### 2015-12-22 Master.Rnw
Master file for 1. round review version of the paper

\documentclass[12pt,english]{article}

\usepackage{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{babel}
\usepackage{url}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage[authoryear]{natbib}
\usepackage[unicode=true]{hyperref}
\usepackage[format=hang, justification=raggedright]{subfig}
\usepackage{array}
\usepackage{multirow}
%\usepackage{authblk}
\usepackage{lscape}
\usepackage{float}
\usepackage{longtable}
\usepackage[noae]{Sweave}
\usepackage{setspace}
\doublespacing
\geometry{verbose,tmargin=2cm,bmargin=2cm,lmargin=3cm,rmargin=3cm}

\begin{document}
\SweaveOpts{concordance=TRUE}

\title{Income and Food Price Transmission\footnote{Revised version of ``Poverty and Price Transmission''. Submitted to \emph{Agricultural Economics} for review}}

\author{Christian Elleby}
\date{2015}

\maketitle
\begin{abstract}
The paper analyzes how international-to-domestic food price transmission
depends on a country\textquoteright s income or expenditure level.
According to conventional wisdom, international food price spikes
have the largest poverty impact on low income countries. However,
the paper presents empirical evidence of an inverse-U shaped relationship
between the degree of international agricultural price changes passed
on to domestic consumers and the country\textquoteright s per capita
income or expenditure level. This implies that retail food prices
in middle income countries respond the strongest to international
commodity price spikes, whereas those in low income countries respond
less, all else equal. The specification of the empirical model is
based on a simple theoretical model of a country\textquoteright s
aggregate food price level. Its predictions are tested on monthly
price index data from 140 countries covering the six year period 2005-11
and results are found to be generally significant and robust. Around
one third to one half of the countries in the sample are to the left
of the turning point, on the increasing part of the parabola, in most
specifications.\end{abstract}

\SweaveInput{intro.Rnw}

\SweaveInput{litrev.Rnw}

\SweaveInput{model.Rnw}

\SweaveInput{analysis.Rnw}

\SweaveInput{conclusion.Rnw}

\bibliographystyle{chicago}
\bibliography{poverty_price_transmission}

\newpage

\appendix

\SweaveInput{figures.Rnw}

\newpage

\SweaveInput{tables.Rnw}

\end{document}

### 2015-12-22 analysis.Rnw
Analysis part of 1. round rev. v. of paper

% !Rnw root = Master.Rnw

\SweaveInput{data_prep.Rnw}
\SweaveInput{regressions.Rnw}

\section{Empirical analysis\label{sec:analysis}}
This section assesses the model's predictions regarding the relationship between income and price transmission. The model in section \ref{sec:model} predicts an inverse-U shaped relationship between income and price transmission but it does not pin down the actual turning point, which is an empirical problem. Our goal in this section is thus to a) establish whether there is a statistically significant quadratic relationship between income or, alternatively, expenditure and food price transmission and b) to determine the income/expenditure level corresponding to the turning point. Section \ref{sec:data} introduces the data and provides some context. Results from the statistical analysis are presented in section \ref{sec:analysis1} and \ref{sec:analysis2}.

\subsection{Data on domestic and international food prices\label{sec:data}}
The domestic price index data analyzed below is sourced from the International Labor Organization (ILO).\footnote{See \href{http://laborsta.ilo.org/}{http://laborsta.ilo.org/}} ILO maintains the most comprehensive database of domestic food price indexes (FPIs) and consumer price indexes (CPIs) which is used to monitor the cost of living and real wages. The ILO food price database is not widely used, for good reasons, despite a broad country coverage and series going back to the 1970s in many cases. A major issue with this data is that the historical series are not recalculated whenever a new base year is introduced. The same applies when other changes are made such as the inclusion or exclusion of certain items such as beverages or tobacco from the index. This means that many of the series have discrete level shifts associated with these changes. Another issue with the data is that some countries report several series. In some cases one series replace another after a period of overlap. Finally, many of the series contain missing values. To ``tidy up'' the raw data I carried out the following operations. First, I restricted the sample to the period 2005:1-2010:12. Secondly, overlapping series were spliced and in cases where multiple complete series were available from the same country, a single one was selected. Thirdly, series with missing values were discarded. Finally, and most critical for the analysis below, level shifts associated with base year changes etc. were identified and removed.\footnote{Concretely I added the change in the break point from the previous value to the break point and all points succeeding it. The value of the adjusted series at the break point is thus the same as the one preceding it.} In total 140 countries have complete price series for the 2005-10 period considered. An important last thing to know about this data is that most of the low income countries' price indexes derive from prices collected from one or two major markets in the capital. Weights and selected items are derived from household expenditure surveys. Differences in consumption patterns around the world of course translate into differences in weights and in the composition of the food baskets themselves. This is probably the largest drawback of using index data in a cross section analysis of price transmission. That being said, even domestic prices of individual food commodities such as wheat or rice would not refer to the same good due to quality and varietal differences, so it is not the case that focusing on a few selected food items consumed by all countries would solve the problem.

There has been two recent international price spikes; one in 2007-8 and the other in 2010-11. Between January 2007 and June 2008 FAO's food and grain price indexes, which are normalized with respect to the 2002-4 average, rose by 91 and 127 points to index 226 and 268, respectively, corresponding to relative price increases of 68 and 90 percent. From the first peak in mid-2008 and to their their respective troughs in February and September 2009 the two indexes fell by 82 and 113 points before rising back up to index 240 and 261 in February and April 2011, respectively.

The boxplot in figure \ref{fig:infl} depicts the empirical distribution of the sample countries' headline (CPI) inflation, their food (FPI) inflation as well as the change in their FPI relative to the CPI, termed real FPI growth, over the two time periods 2007:1-2008:6 and 2010:1-2010:12.

\begin{center}
[Figure \ref{fig:infl} around here] \\
\par\end{center}

As mentioned in section \ref{sec:model}, the nominal food price level is not only driven by exogenous world market shocks, it also includes a stochastic trend which it shares with the non-food price level, representing underlying general domestic inflation. By dividing the FPI with the CPI, we obtain an FPI which is purged of headline inflation. The median country experienced an annualized rate of food and headline inflation of \Sexpr{round(median(infl$foodinfl78),1)} and \Sexpr{round(median(infl$cpiinfl78),1)} percent, respectively, in 2007-8. Real FPI growth, which is approximately the difference between the two, was \Sexpr{round(median(infl$rfpigr78),1)} percent for the median country in that period. Price increases were generally lower in the second period. The median country experienced food and headline inflation rates \Sexpr{round(median(infl$foodinfl1011),1)} and \Sexpr{round(median(infl$cpiinfl1011),1)} percent implying a modest real FPI growth of \Sexpr{round(median(infl$rfpigr1011),1)} percent. In both periods the variation in inflation rates is substantial. The standard deviation and interquartile range of 2007-8 food inflation is \Sexpr{round(sd(infl$foodinfl78),1)} and \Sexpr{round(IQR(infl$foodinfl78),1)} percentage points, respectively, versus \Sexpr{round(sd(infl$foodinfl1011),1)} and \Sexpr{round(IQR(infl$foodinfl1011),1)} percentage points in 2010-11. Although median real FPI growth is lower in the second period the variation is similar in both periods (standard deviations of \Sexpr{round(sd(infl$rfpigr78),1)} and \Sexpr{round(sd(infl$rfpigr1011),1)} and interquartile ranges of \Sexpr{round(IQR(infl$rfpigr78),1)} and \Sexpr{round(IQR(infl$rfpigr1011),1)}, respectively). The three remaining panels contain alternative representations the same data. What they show us that there is some systematic variation in the food prices to be explained. As the positively sloped lowess lines indicate, a country hit by a large nominal and real FPI shock in 2007-8 was more likely to be hit by a relatively large shock again in 2010-11 and vice versa. This suggests that there are indeed characteristics which influences a country's response to a world market shock, an assertion that will be explored further in section \ref{sec:analysis1} and \ref{sec:analysis2}.

\subsection{Determinants of domestic food price changes\label{sec:analysis1}}

Table \ref{tab:2007-8-regressions} and \ref{tab:2010-11-regressions} delve deeper into the  relationship between income and domestic food price changes at the time of the 2007-08 and 2010-11 international price spikes, respectively. Each of the columns refer to a regression of FPI inflation or real FPI growth on a set of income related predictor variables. The models can all be written
\begin{equation}
\pi_{i}^{dom}=\delta_{0} + \delta_{1}x_{i} + \delta_{2}x_{i}^{2} + \delta_{3}z_{i} + u_{i}\label{eq:infl_model}
\end{equation}
where $\pi_{i}^{dom}$ refers to either country $i$'s FPI inflation or its real FPI growth in either the 18 month period 2007:1-2008:6 or the 12 month period 2010:1-2010:12, $x_{i}$ refers to a subset of the explanatory variables considered, $z_{i}$ is a vector of control variables and $u_{i}$ is the error term. The covariates, $x_{i}$, include the log of country $i$'s (PPP adjusted) per capita GDP [$\log(GDP)$], its consumption [$\log(cons)$] and food expenditure level [$\log(food)$]. The control variables in $z_{i}$ are country $i$'s percentage dietary energy supply derived from cereals, roots and tubers ($cshare$), its cereal import dependency ratio ($cdep$) and their interaction.\footnote{Cereal dietary energy supply and cereal dependency data are from FAOSTAT and refer to the 2004-06 average. Data on GDP and expenditure is sourced from the World Bank's website. It is the International Comparison Project (ICP) which is behind the expenditure estimates which refer to year 2005.} Summary statistics on these variables are provided in table \ref{tab:sumstat}. Of primarily interest is the quadratic relationship between domestic price changes $\pi_{i}^{dom}$ and expenditure, $x_{i}$. The first 12 rows of table \ref{tab:tp} summarize information about the turning points based on different estimated versions of model \ref{eq:infl_model}.

\begin{center}
[Table \ref{tab:sumstat} around here]
\par\end{center}

Countries that are largely dependent on grain imports to cover their domestic consumption needs cannot use export restrictions as way of stabilizing domestic prices. Because of this vulnerability to international food price fluctuations we would expect to find stronger domestic price increases during international food price spikes among countries with large import dependency ratios. The analysis in section \ref{sec:model} suggests that international price shocks are transmitted the strongest to middle income countries so we would expect food price inflation in these two periods to follow an inverse-U shaped relationship with per capita income and consumption expenditure. Poor households and populations also spend a larger share of their food budget on staples which makes them more vulnerable to price spikes on grains and other internationally traded agricultural commodities. Therefore, if we did account for this, the estimated relationship between income and domestic food price changes could be spurious. By including the $cshare$ variable in the regression this omitted variable bias should be reduced.

\begin{center}
[Table \ref{tab:2007-8-regressions} around here]
\par\end{center}

<<echo=FALSE>>=
<<dom_infl_0708>>
tp3 = -coef(m3)["log(food)"]/(2*coef(m3)["I(log(food)^2)"])
tp6 = -coef(m6)["log(food)"]/(2*coef(m6)["I(log(food)^2)"])
tp4 = -coef(m4)["log(GDP)"]/(2*coef(m4)["I(log(GDP)^2)"])
tp5 = -coef(m5)["log(cons)"]/(2*coef(m5)["I(log(cons)^2)"])
@

Table \ref{tab:2007-8-regressions} considers domestic food price changes during the 2007:1-2008:6 period for the countries in the sample for which there is data on income, expenditure, dietary energy supply and cereal import dependency. Column (1) relates a country's annualized FPI inflation rate in that period to its per capita GDP level. What we see is that average food inflation during the 2007-8 food crisis is increasing in a country's per capita income level up until a certain point after which it starts to decrease. The same applies for consumption and food expenditure in column (2) and (3), respectively. Only the latter set of coefficients are statistically significant though. Specifically, column (3) suggests that among countries with a 2005 per capita food expenditure level of \Sexpr{round(exp(tp3),1)} {[}$\exp(\Sexpr{round(tp3,2)})${]} PPP adjusted dollars or less there is a positive estimated relationship between expected FPI inflation and food expenditure. After that point the predicted marginal effect turns negative. \Sexpr{nrow(data_0708[which(data_0708$food< exp(tp3) & data_0708$food >0), ])} out of the \Sexpr{nrow(data_0708[which(data_0708$food>0), ])} countries in the sample with data on per capita expenditure levels spend less than that on food.

Column (6) focuses on the relationship between average food expenditure and expected real FPI growth in the 2007-8 period. Again there is a significant quadratic relationship between the two and it is remarkably similar to the specification based on FPI inflation in column (3). For countries with per capita food expenditures less than \Sexpr{round(exp(tp6),1)} PPP adjusted dollars (\Sexpr{nrow(data_0708[which(data_0708$food< exp(tp6) & data_0708$food >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$food>0), ])} countries) there is a positive estimated relationship between food expenditure and real FPI growth. For countries with food expenditures higher than that, the relationship becomes negative.

<<echo=FALSE>>=
<<data_0708>>
gdpsort = data_0708[order(data_0708$GDP), ]
@

The GDP and consumption coefficients in column (4) and (5) are also statistically significant and the conclusion from these models is similar to one based on (3) and (6). The GDP model in column (4) predicts that for countries with a per capita GDP below \Sexpr{round(exp(tp4),1)} PPP adjusted dollars (\Sexpr{nrow(data_0708[which(data_0708$GDP< exp(tp4) & data_0708$GDP >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$GDP>0), ])} countries) there is a positive relationship between real FPI increases and income and vice versa. The consumption expenditure model (5) predicts a turning point of \Sexpr{round(exp(tp5),1)} PPP adjusted dollars where \Sexpr{nrow(data_0708[which(data_0708$cons< exp(tp5) & data_0708$cons >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$cons>0), ])} countries have lower average consumption than that. Number \Sexpr{nrow(data_0708[which(data_0708$GDP< exp(tp4) & data_0708$GDP >0), ])} country in the sample in terms of GDP is Nicaragua. Countries in the lower middle income group  therefore generally experienced the largest food increases during the 2007-8 food crisis, whereas low and high income countries were affected less, once we control for differences in consumption and trade patterns.

Countries with higher cereal consumption shares and import dependency ratios experienced larger food price increases as expected although $cshare$ and $cdep$ are not individually significant in most specifications. The negative coefficient on the interaction term means that the effect of $cdep$ on food price growth is decreasing in $cshare$ and vice-versa, but this mitigating effect is negligible.

\begin{center}
[Table \ref{tab:2010-11-regressions} around here]
\par\end{center}

Table \ref{tab:2010-11-regressions} summarizes the results from a set of regressions based on the same explanatory variables as those from table \ref{tab:2007-8-regressions}. The difference is that the dependent variables are domestic FPI inflation and real FPI growth during the 2010:1-2010:12 period. As can be seen, the results are similar to those in table \ref{tab:2007-8-regressions}, cf. also table \ref{tab:tp} .

\subsection{Price transmission\label{sec:analysis2}}
Although the results in section \ref{sec:analysis1} provide insights into the cross sectional variation in domestic price changes during the two recent international food price spikes they have at least two shortcomings. First, none of the estimated relationships represent price transmission as such. Secondly, the estimates are based on two specific time periods, which means that we ignore much the information contained in the individual price series. To accommodate these two shortcomings I estimate country specific price transmission coefficients and regress these on the same income and expenditure predictors as in section \ref{sec:analysis1}.

<<echo=FALSE>>=
<<ts_pt>>
tp1 = -coef(m1)["log(GDP)"]/(2*coef(m1)["I(log(GDP)^2)"])
tp2 = -coef(m2)["log(cons)"]/(2*coef(m2)["I(log(cons)^2)"])
tp3 = -coef(m3)["log(food)"]/(2*coef(m3)["I(log(food)^2)"])
tp4 = -coef(m4)["log(GDP)"]/(2*coef(m4)["I(log(GDP)^2)"])
tp5 = -coef(m5)["log(cons)"]/(2*coef(m5)["I(log(cons)^2)"])
tp6 = -coef(m6)["log(food)"]/(2*coef(m6)["I(log(food)^2)"])
@

Specifically, for each country, I regress its monthly domestic food price changes on its lagged domestic food price changes as well as on contemporaneous and lagged international food price changes. That is, as in the the \cite{IMF2011} study, I estimate country specific autoregressive distributive lag (ADL) models of the type
\begin{equation}
\pi_{t}^{dom}=\delta_{0}+\sum_{s=1}^{S_{y}}\rho_{s}\pi_{t-s}^{dom}+\sum_{s=0}^{S_{f}}\lambda_{s}\pi_{t-s}^{int}+\epsilon_{t}\label{eq:time_series_model}
\end{equation}
where $\pi_{t}^{dom}=\Delta p_{t}$ represents monthly FPI inflation or real FPI growth at time $t$ (both approximated by the first differences to the log of FPI or real FPI) and $\pi_{t}^{int}$ is the FAO food price index at time $t$, also in log differences. Using first differences instead of price levels is a way to avoid the spurious regression problem associated with trending variables. In order to keep things simple, maximum lag lengths are uniformly set to $S_{y}=S_{f}=6$.

Next I calculate the long run effects of changes in the the FAO food price index on domestic food inflation and real FPI growth, the first of which is a standard measure of food price transmission. The long run multiplier is given by 
\begin{equation}
\theta=\frac{\sum_{s=0}^{S_{f}}\lambda_{s}}{1-\sum_{s=1}^{S_{y}}\rho_{s}}.
\end{equation}
Finally, I regress the price transmission estimates, $\hat{\theta}$, on the same explanatory variables as in table \ref{tab:2007-8-regressions} and \ref{tab:2010-11-regressions}. That is, I estimate the model
\begin{equation}
\hat{\theta}_{i} =\delta_{0} + \delta_{1}x_{i} + \delta_{2}x_{i}^{2} + \delta_{3}z_{i} + u_{i}\label{eq:ptreg}.
\end{equation}
The results from this exercise are summarized in table \ref{tab:lri-regressions} and \ref{tab:tp}. As table \ref{tab:lri-regressions} shows, we find the same inverse-U shaped estimated relationships between income/expenditure and price transmission as we did with income/expenditure and domestic price changes in table \ref{tab:2007-8-regressions} and \ref{tab:2010-11-regressions}. For the specifications with significant income and expenditure coefficients, i.e. (1)-(5), the estimated turning points again imply that roughly one third of the countries are on the increasing part of the curve, cf. \ref{tab:tp}.

\begin{center}
[Table \ref{tab:lri-regressions} around here]
\par\end{center}

<<echo=False>>=
<<panel_pt>>
tp1 = -m1$coefficients[8]/(2*m1$coefficients[9])
tp2 = -m2$coefficients[8]/(2*m2$coefficients[9])
tp3 = -m3$coefficients[8]/(2*m3$coefficients[9])
tp4 = -m4$coefficients[8]/(2*m4$coefficients[9])
tp5 = -m5$coefficients[8]/(2*m5$coefficients[9])
tp6 = -m6$coefficients[8]/(2*m6$coefficients[9])
@

The findings discussed above all point to the conclusion that middle income countries are the most vulnerable to international food price spikes. However, the estimates in table \ref{tab:2007-8-regressions}-\ref{tab:lri-regressions} are based on standard OLS and are potentially subject to unobserved heterogeneity bias. As a last robustness check I therefore exploit the panel structure of the dataset to control for time invariant unobserved heterogeneity which potentially confounds the income price transmission relationship. Results in table \ref{tab:panel-regressions} are based on within estimation of the model
\begin{equation}
\pi_{it}^{dom}=\delta_{0}+\sum_{s=1}^{6}\rho_{s}\pi_{it-s}^{dom}+\lambda_{1}\pi_{it-6}^{int}+\lambda_{2}\pi_{it-6}^{int}\cdot x_{i} +\lambda_{3}\pi_{it-6}^{int}\cdot x_{i}^{2}+u_{it}\label{eq:panel_model},
\end{equation}
where the variables, $\pi_{it}^{dom}$, $\pi_{it}^{int}$ and $x_{i}$ are the same as in \ref{eq:time_series_model} and \ref{eq:ptreg}. For simplicity, only a single lag of the FAO price index is included in the model and the lag length is chosen to be 6. The parameters of interest are $\lambda_{2}$ and $\lambda_{3}$ in front of the two interaction terms. A positive $\lambda_{2}$ and a negative $\lambda_{3}$ would, once more, imply an inverse-U shaped relationship between income/expenditure and price transmission as predicted by the simple model from section \ref{sec:model}. Again, in order to avoid the spurious regression problem associated with unit root processes the model is based on inflation rates rather than price levels.

\begin{center}
[Table \ref{tab:panel-regressions} around here]
\par\end{center}

Model (1),(3), (4) and (5) in table \ref{tab:panel-regressions} lead to significant quadratic expenditure price transmission relationships and the turning points based on these specifications are similar but somewhat higher than the ones from table \ref{tab:2007-8-regressions}-\ref{tab:lri-regressions}, with up to around 50 percent the countries on the increasing part of the income-price transmission curve, c.f table \ref{tab:tp}.

\begin{center}
[Table \ref{tab:tp} around here]
\par\end{center}

### 2015-12-22 model.Rnw
Model sec. of 1. round rev. v. of paper

% !Rnw root = Master.Rnw

\section{A model of aggregate food price transmission\label{sec:model}}
As the last section pointed out, the small literature on aggregate food price transmission is based on informal economic reasoning, rather than an actual economic model. In this section I will present one such model and derive some testable predictions.

A country's consumer price index (CPI) can be expressed as a weighted average
\begin{equation}
P =\alpha P_{f}+(1-\alpha)P_{n}\label{eq:cpi},
\end{equation}
where $P_{f}$ and $P_{n}$ denote the food and non-food price indexes, respectively, and $\alpha$ is the fixed food expenditure share. I have omitted time indexes in order to avoid notational clutter in the equations that follow. However, all price indexes are considered stochastic processes. Moreover, the two components $P_{f}$ and $P_{n}$ must share a common stochastic trend representing underlying inflation. The food price index (FPI), which is of primary interest, can be written
\begin{equation}
P_{f} =\beta P_{c}+(1-\beta)P_{m}\label{eq:fpi},
\end{equation}
where $P_{c}$ and $P_{m}$ denote the farm cost and marketing components of the FPI, respectively, and $\beta$ the farm cost or commodity share. By marketing costs I refer to all other costs than those of the raw agricultural commodities included in the final consumer food price such as those associated with distribution, storage and promotion. Finally, the commodity component of the domestic FPI can be decomposed as 
\begin{equation}
P_{c} =\gamma P_{w}+(1-\gamma)P_{d}\label{eq:compi},
\end{equation}
where $P_{w}$ and $P_{d}$ denote the tradable and purely domestic (i.e. non-tradable) subcomponents of the commodity price index, respectively, and $\gamma$ is the share of tradable food commodities consumed domestically. The defining difference between tradable and non-tradable food commodities is that the former have prices which are determined on the world market whereas the latter prices are determined by domestic supply and demand.

The three share parameters, $\alpha$, $\beta$ and $\gamma$ represent weights on price indexes that are progressively embedded into each other. By substituting (\ref{eq:compi}) into (\ref{eq:fpi}) we get
\begin{equation}
P_{f} = \beta \gamma P_{w}+R\label{eq:fpi1},
\end{equation}
where $R=\beta(1-\gamma)P_{d}+(1-\beta)P_{m}$ is a term which does not depend on world market prices directly. With these definitions a first order approximation to transmission of international agricultural commodity price changes to domestic consumer food prices can be written
\begin{equation}
\theta=\frac{\partial P_{f}}{\partial P_{w}}=\beta\gamma,\label{eq:pt}
\end{equation}
i.e., price transmission depends multiplicatively on the share of tradable commodities in the domestic food commodity price index, $\gamma$, and the commodity cost share in the FPI, $\beta$. Expression (\ref{eq:pt}) is based on the assumption that there is no knock-on effect on domestic domestic prices when world market prices change, as would be appropriate in the short term. If we relax this assumption and allow domestic prices to respond to international price changes through substitution effects we would have to add the term $\beta(1-\gamma)\frac{\mathrm{d}P_{d}}{\mathrm{d}P_{w}}$. For simplicity, I will ignore this second order effect in the analysis below.

In order to analyze how price transmission depends on income we need to make some assumptions about the two cost shares, $\gamma$ and $\beta$. Specifically, I shall assume that
\begin{equation}
\beta=\beta(y),\quad\beta'(y)<0,\quad\beta''(y)>0\label{com_share}
\end{equation}
and
\begin{equation}
\gamma=\gamma(y),\quad\gamma'(y)>0,\quad\gamma''(y)<0,\label{trade_share}
\end{equation}
where $y$ denotes per capita food expenditure or, more generally, income. That is, the wealthier the country, the lower the commodity cost share in the FPI and the higher the share of tradables in the commodity price index. The left panel of figure \ref{fig:pt} illustrates two such share functions with these specific properties. Assumptions (\ref{com_share})-(\ref{trade_share}) imply that price transmission, $\theta$, becomes a concave function of income, a feature which is illustrated in the right panel of figure \ref{fig:pt}.

\begin{center}
[Figure \ref{fig:pt} around here] \\
\par\end{center}

To sum up, there are reasons to expect a domestic FPI reaction to price changes on the international agricultural commodity markets which depends non-monotonically on income or expenditure. Specifically, the simple model sketched above predicts an inverse-U shaped relationship between income and price transmission. This prediction is based, of course, on the assumptions regarding the relationship between income and the two key share parameters, $\beta$ and $\gamma$, discussed above. Therefore, to round off this section I will briefly discuss what we know about these.

\subsection{Farm shares, tradability and income\label{sec:model_assumptions}}
Regarding $\beta$, although there is no database of international farm cost shares in food expenditure, there are some time series estimates available from a few countries. Estimates from the US and the EU do, in particular, seem to support the conclusion that the farm share declines as countries grow richer. \cite{Schnepf2009}, for example, uses USDA data to show that the farm-value added share of consumer food expenditures has fallen from 41 percent in 1950 to 18.5 percent in 2006, meaning that the value share of marketing services has increased by 22.5 percentage points over the 56 year period. In the same period US per capita GDP increased by 240 percent. Similarly, analysts working for the European Commission has calculated that within EU25, the agriculture-value added share in the food supply chain has dropped from 31 percent in 1995 to 24 percent in 2005 \citep{Commission2009}. In this period (EU28) per capita GDP grew by 24 percent according to EUROSTAT. \citet[][p. 797]{Yang2015} comment that ``\emph{... the share of domestic margin services in consumed food is larger in high-income countries \cite{Abbott2012}, ... implying that consumers in high-income countries are more insulated from shocks of international agricultural prices.}''

Regarding the $\gamma$ parameter, representing the cost share of tradable commodities in the food commodity price index, the main issue is conceptual rather than empirical. The problem is that tradability is not an observable feature of a good. A food good might not be traded due to policy distortions such as trade taxes and nontariff barriers but there are also other factors including transport costs, quality differences and local taste preferences which reduce the amount of trade. In the literature goods are typically classified according to the actual trade flows though. \cite{Anderson2009}, for example, divides the agricultural goods produced in country into three groups; import-competing (M), exportable (X), or nontradable (H) and a good is classified as nontradable if three conditions are met. First, there must be no policies in place distorting trade flows. Second, the share of production exported must be less than 2.5 percent and, third, the share of consumption imported must also be less than 2.5 percent. M and X goods are both tradable in the sense that their undistorted prices are determined on the world market (pp. 582-84). Figure \ref{fig:gamma} plots the estimated value shares of tradable agricultural goods relative to production based on the \cite{Anderson2008} database. As the lowess line indicates, there is a positive relationship between the estimated $\gamma$'s and income up to a certain income level, after which all goods produced become tradable.

\begin{center}
[Figure \ref{fig:gamma} around here] \\
\par\end{center}

Finally, regarding $\alpha$, the food share in total expenditure, although I do not make any assumptions about its dependence on income, there is, of course, a well established relationship between the two known as Engel's law.

Although the focus in this as well as in the next section is on international to domestic price transmission or pass-through, it is worth mentioning that the simple model presented above delivers another important prediction; namely that LICs are especially vulnerable to domestic agricultural supply shocks. This is because domestic food prices mainly represent non-tradable food goods and because the cost of these goods mainly represent farm gate prices. I will, however, for the sake of brevity, leave it to future work to test this proposition.

### 2015-12-22 intro.Rnw
Intro to 1. v. of paper

% !Rnw root = Master.Rnw

\SweaveInput{data_prep.Rnw}

\section{Introduction}
Poor households spend a large proportion of their budget on food and are therefore vulnerable to increases in the domestic price of food \citep{Mellor1978}. Spurred by the increased volatility on the international commodity markets the last decade or so and more widely available household survey data, a large number of recent studies have quantified the poverty impact of an international food price spike for specific countries as well as for the world as a whole.\footnote{E.g. \cite{Attanasio2013,Balagtas2014,Coxhead2012,Dessus2008,Ferreira2013,Hoyos2011,Ivanic2012,Janvry2010,Mghenyi2011,Vu2011,Wodon2010}. There is also a special issue of \emph{Agricultural Economics} from 2008 devoted to this topic.}

At the aggregate level, food insecurity resulting from international grain price spikes is mainly seen as a problem for Low Income Countries (LICs) \citep[e.g.][]{Ivanic2008}. However, a country's vulnerability to international food price spikes not only depends on its income distribution around the poverty level. Another fundamental determinant of aggregate vulnerability is the extent to which international price spikes are transmitted to domestic consumer markets. As I argue in the paper, price transmission varies systematically with a country's income level. In a nutshell, low income households are the most vulnerable to domestic food price increases but domestic food prices in Middle Income Countries (MICs) rather than LICs or High Income Countries (HICs) are the most responsive to international food price changes.

<<echo=FALSE>>=
<<infl>>
n=nrow(infl)
@

This conclusion stems from an empirical price transmission analysis based on a large dataset of monthly consumer price indexes from \Sexpr{n} countries covering the period January 2005 to December 2010. My main finding is a significant inverse-U shaped relationship between income or expenditure and price transmission, implying that food prices in MICs, ceteris paribus, respond the strongest to international price changes. This result is robust to the specific choice of estimator and dependent variable and it also holds for the actual domestic price changes in connection with the food crises of 2007-08 and 2010-11.

The specification of the regression function is derived from a simple theoretical model of a country's food price index (FPI). The model makes two fairly uncontroversial assumptions about the farm share of consumer food prices and share of tradable commodities in the food basket; 1) that the farm share of consumer food prices is decreasing in income and 2) that the  share of tradable commodities is increasing in income. These two assumptions combined imply that the model predicts a domestic FPI response to a change in a price index of internationally traded food goods which follows an inverse-U shaped relationship with per capita income or expenditure. Two additional findings from the analysis are 1) domestic agricultural supply shocks are expected to have a larger impact on food prices in LICs than in MICs and 2) LICs which rely heavily on on imports of agricultural commodities are the most vulnerable to an international price spike.

My contribution to the literature consists of a theory based analysis of international agricultural commodity to domestic consumer food price transmission, with an emphasis on its dependence on average income. Most of the papers from the small existing literature on the cross sectional pattern of food price transmission do not base their analysis on model predictions. Rather they take a more agnostic approach and consider factors which simple economic reasoning suggests influence pass-through rates (see section \ref{sec:litrev}). There are, however, a couple of recent papers from the poverty impact literature which consider the role of marketing margins specifically. Two very similar papers by \cite{Dawe2014} and \cite{Minot2015} analyze how assumptions about marketing margins on food goods affect the poverty impact from a price increase in the \cite{Deaton1989} framework. Both papers conclude that, assumptions about marketing margins matters for the poverty impact of price changes. Specifically, assuming equal percentage changes for both farm and consumer prices (proportional margin) leads to an exaggerated or upwards biased impact of higher food prices compared to a situation with fixed margins, which might be a more appropriate assumption. \cite{Yang2015} is another recent paper analyzing how assumptions about marketing margins affects price pass-through and subsequently welfare in a CGE framework. Their conclusion is that also CGE models tend to be overestimate price pass-through particularly for high income countries, when domestic margin services are not explicitly taken into account.

The paper also contributes to a broader policy discussion of the appropriate response to international food crises and its poverty impact. Thanks to the work of Andy Sumner and others, it has become clear that most of the world's poor do not actually live in low income countries \citep{Sumner2012, Kanbur2012}. Nearly three quarters of the world's poor now live in MICs according to World Bank estimates.\footnote{See \url{http://www.worldbank.org/en/country/mic/overview}. MICs are defined as having a 2011 per capita gross national income of US\$1,026 to \$12,475, where the conversion from local currency to US dollars is based on the Atlas method.} This is in a sense a worrying fact given the finding that food prices in MICs are more responsive to international price changes than those in LICs, ceteris paribus. On the other hand, MIC governments are better able than LICs to protect its households from a significant fall in real income.

The paper now proceeds as follows. Section \ref{sec:litrev} provides a brief review of the relevant price transmission literature. Section \ref{sec:model} lays out the simple theory model of country's food price index and its response to an international price spike. Section \ref{sec:analysis} presents the data and tests the model's predictions regarding income and price transmission and, finally, section \ref{sec:conclusion} offers some concluding remarks.

### 2015-12-22 tables.Rnw
Tables file for 1. round rev. v. of paper

% !Rnw root = Master.Rnw

\SweaveInput{regressions.Rnw}

\section{Tables}

<<echo=FALSE, results=tex>>=
library(stargazer)
sumdat = merge(infl, icp_data[,-5], by = "iso3c", all.x=TRUE)
sumdat = merge(sumdat, staples, by = "iso3c", all.x=TRUE)
sumdat = merge(sumdat, cdep, by = "iso3c", all.x=TRUE)
colnames(sumdat)=c("iso", "FPI inflation 07-08", "CPI inflation 07-08", "Real FPI growth 07-08", "FPI inflation 10-11", "CPI inflation 10-11", "Real FPI growth 10-11", "GDP", "cons", "food", "cshare", "cdep")

stargazer(sumdat, digits=1, title='Summary statistics', label = "tab:sumstat", 
          notes=c("Notes: 07-08 and 10-11 refer to the periods 2007:1-2008:6 and 2010:1-2010:12,", 
                  "respectively. `GDP, `cons' and `food' refer to per capita income, -consumption",
                  "and -food expenditure respectively in the year 2005. `cshare' refers to a coun-",
                  "try's percentage dietary energy supply derived from cereals, roots and tubers.", 
                  "`cdep' refers to a country's cereal import dependency ratio, both with respect",
                  "to the years 2004-06. Source: ILO, World Bank, FAO. Own calculations"),
          notes.append=FALSE,
          notes.align="l")
@

<<results=tex,echo=F>>=
<<dom_infl_0708>>
stargazer(m1, m2, m3, m4, m5, m6, title='Income related determinants of domestic food price changes 2007:1-2008:6', label = "tab:2007-8-regressions", dep.var.labels=c("FPI inflation","Real FPI growth"), no.space=TRUE, se = list(rse1, rse2, rse3, rse4, rse5, rse6), omit.stat = c("adj.rsq","f", "ser"), font.size = "footnotesize", covariate.labels=c("$log(GDP)$", "$log(GDP)^2$", "$log(cons)$", "$log(cons)^2$", "$log(food)$", "$log(food)^2$", "$cshare$", "$cdep$", "$cshare \\times cdep$"), 
          notes=c("Robust standard errors are reported in parentheses underneath the coefficients.",
                  "Significance at p = 0.1, p = 0.05 and p = 0.01 are denoted by `*', `**' and `***'", 
                  "respectively"),
          notes.append=FALSE,
          notes.align="l",
          notes.label="Notes:")
@

<<results=tex,echo=F>>=
<<dom_infl_1011>>
stargazer(m1, m2, m3, m4, m5, m6, title='Income related determinants of domestic food price changes 2010:1-2010:12', label = "tab:2010-11-regressions", dep.var.labels=c("FPI inflation","Real FPI growth"), no.space=TRUE, se = list(rse1, rse2, rse3, rse4, rse5, rse6), omit.stat = c("adj.rsq","f", "ser"), font.size = "footnotesize", covariate.labels=c("$log(GDP)$", "$log(GDP)^2$", "$log(cons)$", "$log(cons)^2$", "$log(food)$", "$log(food)^2$", "$cshare$", "$cdep$", "$cshare \\times cdep$"), 
          notes=c("Robust standard errors are reported in parentheses underneath the coefficients.",
                  "Significance at p = 0.1, p = 0.05 and p = 0.01 are denoted by `*', `**' and `***'", 
                  "respectively"),
          notes.append=FALSE,
          notes.align="l",
          notes.label="Notes:")
@

<<results=tex, echo=FALSE>>=
<<ts_pt>>
stargazer(m1, m2, m3, m4, m5, m6, title='Long run multiplier (LRM) regressions', label = "tab:lri-regressions", dep.var.labels=c("LRM, FPI inflation","LRM, real FPI growth"), no.space=TRUE, se = list(rse1, rse2, rse3, rse4, rse5, rse6), omit.stat = c("adj.rsq","f", "ser"), font.size = "footnotesize", covariate.labels=c("$log(GDP)$", "$log(GDP)^2$", "$log(cons)$", "$log(cons)^2$", "$log(food)$", "$log(food)^2$", "$cshare$", "$cdep$", "$cshare \\times cdep$"),
          notes=c("Robust standard errors are reported in parentheses underneath the coefficients.",
                  "Significance at p = 0.1, p = 0.05 and p = 0.01 are denoted by `*', `**' and `***'", 
                  "respectively"),
          notes.append=FALSE,
          notes.align="l",
          notes.label="Notes:")
@

<<results=tex, echo=FALSE>>=
<<panel_pt>>
stargazer(m1, m2, m3, m4, m5, m6, 
          title='Panel regressions', 
          label = "tab:panel-regressions", 
          font.size = "footnotesize", 
          no.space=TRUE, 
          se = list(rse1, rse2, rse3, rse4, rse5, rse6), 
          omit.stat = c("adj.rsq","f"), 
          dep.var.labels=c("$\\Delta log(FPI_{t})$","$\\Delta log(rFPI_{t})$"), 
          omit = 1:12, 
          covariate.labels=c("$\\Delta log(FAO_{t-6})$", "$\\Delta log(FAO_{t-6}) \\times log(GDP)$","$\\Delta log(FAO_{t-6}) \\times log(GDP)^2$", "$\\Delta log(FAO_{t-6}) \\times log(cons)$", "$\\Delta log(FAO_{t-6}) \\times log(cons)^2$", "$\\Delta log(FAO_{t-6}) \\times log(food)$", "$\\Delta log(FAO_{t-6}) \\times log(food)^2$"), 
          notes=c("Robust standard errors are reported in parentheses underneath the coef-",
                  "ficients. Significance at p = 0.1, p = 0.05 and p = 0.01 are denoted by `*',", 
                  " `**' and `***' respectively"),
          notes.append=FALSE,
          notes.align="l",
          notes.label="Notes:")
@

<<tpmat, echo=FALSE>>=
tp = matrix(NA, nrow=6, ncol=4)

<<dom_infl_0708>>
tp[1,1] = -m1$coefficients[2]/(2*m1$coefficients[3])
tp[2,1] = -m2$coefficients[2]/(2*m2$coefficients[3])
tp[3,1] = -m3$coefficients[2]/(2*m3$coefficients[3])
tp[4,1] = -m4$coefficients[2]/(2*m4$coefficients[3])
tp[5,1] = -m5$coefficients[2]/(2*m5$coefficients[3])
tp[6,1] = -m6$coefficients[2]/(2*m6$coefficients[3])

<<dom_infl_1011>>
tp[1,2] = -m1$coefficients[2]/(2*m1$coefficients[3])
tp[2,2] = -m2$coefficients[2]/(2*m2$coefficients[3])
tp[3,2] = -m3$coefficients[2]/(2*m3$coefficients[3])
tp[4,2] = -m4$coefficients[2]/(2*m4$coefficients[3])
tp[5,2] = -m5$coefficients[2]/(2*m5$coefficients[3])
tp[6,2] = -m6$coefficients[2]/(2*m6$coefficients[3])

<<ts_pt>>
tp[1,3] = -m1$coefficients[2]/(2*m1$coefficients[3])
tp[2,3] = -m2$coefficients[2]/(2*m2$coefficients[3])
tp[3,3] = -m3$coefficients[2]/(2*m3$coefficients[3])
tp[4,3] = -m4$coefficients[2]/(2*m4$coefficients[3])
tp[5,3] = -m5$coefficients[2]/(2*m5$coefficients[3])
tp[6,3] = -m6$coefficients[2]/(2*m6$coefficients[3])

<<panel_pt>>
tp[1,4] = -m1$coefficients[8]/(2*m1$coefficients[9])
tp[2,4] = -m2$coefficients[8]/(2*m2$coefficients[9])
tp[3,4] = -m3$coefficients[8]/(2*m3$coefficients[9])
tp[4,4] = -m4$coefficients[8]/(2*m4$coefficients[9])
tp[5,4] = -m5$coefficients[8]/(2*m5$coefficients[9])
tp[6,4] = -m6$coefficients[8]/(2*m6$coefficients[9])

tp = exp(tp)
@

\begin{table}
\caption{Estimated turning points}
\label{tab:tp}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|}
\hline 
Model & dep. var & indep. var & tp (\$, PPP) & significance & \# of countries\tabularnewline
\hline 
\hline 
(\ref{eq:infl_model}) & FPI infl. 07-08  & GDP & \Sexpr{round(tp[1,1],0)} & no & \Sexpr{nrow(data_0708[which(data_0708$GDP< tp[1,1] & data_0708$GDP >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$GDP>0), ])}  (\Sexpr{round(nrow(data_0708[which(data_0708$GDP< tp[1,1] & data_0708$GDP >0), ])/nrow(data_0708[which(data_0708$GDP>0), ])*100,0)}\%)\tabularnewline
\hline 
(\ref{eq:infl_model}) & FPI infl. 07-08 & cons & \Sexpr{round(tp[2,1],0)} & no & \Sexpr{nrow(data_0708[which(data_0708$cons< tp[2,1] & data_0708$cons >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$cons>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$cons< tp[2,1] & data_0708$cons >0), ])/nrow(data_0708[which(data_0708$cons>0), ])*100,0)}\%)\tabularnewline
\hline 
(\ref{eq:infl_model}) & FPI infl. 07-08 & food & \Sexpr{round(tp[3,1],0)}  & yes & \Sexpr{nrow(data_0708[which(data_0708$food< tp[3,1] & data_0708$food >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$food>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$food< tp[3,1] & data_0708$food >0), ])/nrow(data_0708[which(data_0708$food>0), ])*100,0)}\%)\tabularnewline
\hline 
(\ref{eq:infl_model}) & rFPI gr. 07-08 & GDP & \Sexpr{round(tp[4,1],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$GDP< tp[4,1] & data_0708$GDP >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$GDP>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$GDP< tp[4,1] & data_0708$GDP >0), ])/nrow(data_0708[which(data_0708$GDP>0), ])*100,0)}\%)\tabularnewline
\hline 
(\ref{eq:infl_model}) & rFPI gr. 07-08 & cons & \Sexpr{round(tp[5,1],0)}  & yes & \Sexpr{nrow(data_0708[which(data_0708$cons< tp[5,1] & data_0708$cons >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$cons>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$cons< tp[5,1] & data_0708$cons >0), ])/nrow(data_0708[which(data_0708$cons>0), ])*100,0)}\%)\tabularnewline
\hline 
(\ref{eq:infl_model}) & rFPI gr. 07-08 & food & \Sexpr{round(tp[6,1],0)}  & yes & \Sexpr{nrow(data_0708[which(data_0708$food< tp[6,1] & data_0708$food >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$food>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$food< tp[6,1] & data_0708$food >0), ])/nrow(data_0708[which(data_0708$food>0), ])*100,0)}\%)\tabularnewline
\hline 
(\ref{eq:infl_model}) & FPI infl. 10-11 & GDP & \Sexpr{round(tp[1,2],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$GDP< tp[1,2] & data_0708$GDP >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$GDP>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$GDP< tp[1,2] & data_0708$GDP >0), ])/nrow(data_0708[which(data_0708$GDP>0), ])*100,0)}\%)\tabularnewline
\hline 
(\ref{eq:infl_model}) & FPI infl. 10-11 & cons & \Sexpr{round(tp[2,2],0)} & no &  \Sexpr{nrow(data_0708[which(data_0708$cons< tp[2,2] & data_0708$cons >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$cons>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$cons< tp[2,2] & data_0708$cons >0), ])/nrow(data_0708[which(data_0708$cons>0), ])*100,0)}\%)\tabularnewline
\hline 
(\ref{eq:infl_model}) & FPI infl. 10-11 & food & \Sexpr{round(tp[3,2],0)} & yes &  \Sexpr{nrow(data_0708[which(data_0708$food< tp[3,2] & data_0708$food >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$food>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$food< tp[3,2] & data_0708$food >0), ])/nrow(data_0708[which(data_0708$food>0), ])*100,0)}\%)\tabularnewline
\hline 
(\ref{eq:infl_model}) & rFPI gr. 10-11 & GDP & \Sexpr{round(tp[4,2],0)} & yes &  \Sexpr{nrow(data_0708[which(data_0708$GDP< tp[4,2] & data_0708$GDP >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$GDP>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$GDP< tp[4,2] & data_0708$GDP >0), ])/nrow(data_0708[which(data_0708$GDP>0), ])*100,0)}\%)\tabularnewline
\hline 
(\ref{eq:infl_model}) & rFPI gr. 10-11 & cons & \Sexpr{round(tp[5,2],0)} & yes &  \Sexpr{nrow(data_0708[which(data_0708$cons< tp[5,2] & data_0708$cons >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$cons>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$cons< tp[5,2] & data_0708$cons >0), ])/nrow(data_0708[which(data_0708$cons>0), ])*100,0)}\%)\tabularnewline
\hline 
(\ref{eq:infl_model}) & rFPI gr. 10-11 & food & \Sexpr{round(tp[6,2],0)} & yes &  \Sexpr{nrow(data_0708[which(data_0708$food< tp[6,2] & data_0708$food >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$food>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$food< tp[6,2] & data_0708$food >0), ])/nrow(data_0708[which(data_0708$food>0), ])*100,0)}\%)\tabularnewline
\hline 
(\ref{eq:ptreg}) & LR multiplier & GDP & \Sexpr{round(tp[1,3],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$GDP< tp[1,3] & data_0708$GDP >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$GDP>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$GDP< tp[1,3] & data_0708$GDP >0), ])/nrow(data_0708[which(data_0708$GDP>0), ])*100,0)}\%)\tabularnewline
\hline 
(\ref{eq:ptreg}) & LR multiplier & cons & \Sexpr{round(tp[2,3],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$cons< tp[2,3] & data_0708$cons >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$cons>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$cons< tp[2,3] & data_0708$cons >0), ])/nrow(data_0708[which(data_0708$cons>0), ])*100,0)}\%)\tabularnewline
\hline 
(\ref{eq:ptreg}) & LR multiplier & food & \Sexpr{round(tp[3,3],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$food< tp[3,3] & data_0708$food >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$food>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$food< tp[3,3] & data_0708$food >0), ])/nrow(data_0708[which(data_0708$food>0), ])*100,0)}\%)\tabularnewline
\hline 
(\ref{eq:ptreg}) & LR multiplier & GDP & \Sexpr{round(tp[4,3],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$GDP< tp[4,3] & data_0708$GDP >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$GDP>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$GDP< tp[4,3] & data_0708$GDP >0), ])/nrow(data_0708[which(data_0708$GDP>0), ])*100,0)}\%)\tabularnewline
\hline
(\ref{eq:ptreg}) & LR multiplier & cons & \Sexpr{round(tp[5,3],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$cons< tp[5,3] & data_0708$cons >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$cons>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$cons< tp[5,3] & data_0708$cons >0), ])/nrow(data_0708[which(data_0708$cons>0), ])*100,0)}\%)\tabularnewline
\hline 
(\ref{eq:ptreg}) & LR multiplier & food & \Sexpr{round(tp[6,3],0)} & no & \Sexpr{nrow(data_0708[which(data_0708$food< tp[6,3] & data_0708$food >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$food>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$food< tp[6,3] & data_0708$food >0), ])/nrow(data_0708[which(data_0708$food>0), ])*100,0)}\%)\tabularnewline
\hline 
(\ref{eq:panel_model}) & mth FPI infl & GDP & \Sexpr{round(tp[1,4],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$GDP< tp[1,4] & data_0708$GDP >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$GDP>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$GDP< tp[1,4] & data_0708$GDP >0), ])/nrow(data_0708[which(data_0708$GDP>0), ])*100,0)}\%)\tabularnewline
\hline 
(\ref{eq:panel_model}) & mth FPI infl & cons & \Sexpr{round(tp[2,4],0)} & no & \Sexpr{nrow(data_0708[which(data_0708$cons< tp[2,4] & data_0708$cons >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$cons>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$cons< tp[2,4] & data_0708$cons >0), ])/nrow(data_0708[which(data_0708$cons>0), ])*100,0)}\%)\tabularnewline
\hline 
(\ref{eq:panel_model}) & mth FPI infl & food & \Sexpr{round(tp[3,4],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$food< tp[3,4] & data_0708$food >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$food>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$food< tp[3,4] & data_0708$food >0), ])/nrow(data_0708[which(data_0708$food>0), ])*100,0)}\%)\tabularnewline
\hline 
(\ref{eq:panel_model}) & mth rFPI gr. & GDP & \Sexpr{round(tp[4,4],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$GDP< tp[4,4] & data_0708$GDP >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$GDP>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$GDP< tp[4,4] & data_0708$GDP >0), ])/nrow(data_0708[which(data_0708$GDP>0), ])*100,0)}\%)\tabularnewline
\hline 
(\ref{eq:panel_model}) & mth rFPI gr. & cons & \Sexpr{round(tp[5,4],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$cons< tp[5,4] & data_0708$cons >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$cons>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$cons< tp[5,4] & data_0708$cons >0), ])/nrow(data_0708[which(data_0708$cons>0), ])*100,0)}\%)\tabularnewline
\hline 
(\ref{eq:panel_model}) & mth rFPI gr. & food & \Sexpr{round(tp[6,4],0)} & no & \Sexpr{nrow(data_0708[which(data_0708$food< tp[6,4] & data_0708$food >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$food>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$food< tp[6,4] & data_0708$food >0), ])/nrow(data_0708[which(data_0708$food>0), ])*100,0)}\%)\tabularnewline
\hline 
\multicolumn{6}{l}{Notes: The numbers in the fourth column refer to the estimated income or expenditure}\tabularnewline
\multicolumn{6}{l}{turning point (tp). The fifth column indicates whether the estimated quadratic term}\tabularnewline
\multicolumn{6}{l}{coefficients are individually significant at the 10 percent level. The last column shows}\tabularnewline
\multicolumn{6}{l}{the number of countries in the sample with a per capita income or expenditure levels}\tabularnewline 
\multicolumn{6}{l}{below the estimated turning point.}
\end{tabular}
\end{center}
\end {table}

### 2015-12-22 litrev.Rnw
Litt. rev. sec. of 1. round rev. v. of paper

% !Rnw root = Master.Rnw

\section{Review of the price transmission literature \label{sec:litrev}}
This section introduces the agricultural price transmission literature. Results from a few key studies of particular relevance will be discussed in some detail.

There is a vast empirical literature studying the price relationships between spatially separated agricultural markets \citep[see][]{Fackler2001}. Within this field we can distinguish between two strands of studies. First, there are studies assessing the connectedness or degree of integration characterizing different markets within the same country or region.\footnote{\citet{Ravallion1986}, \citet{Baulch1997}, \citet{Goodwin2001} and \citet{Barrett2002} are some of the seminal papers.} Recent papers focusing on the recent period include\citet{Abidoye2014},  \citet{Baquedano2011}, \citet{Baquedano2014}, \citet{Cudjoe2010} and \citet{Esposti2013}. A common thread among these is the use of actual commodity prices in an attempt to test whether the law of one price holds or, more generally, whether the domestic and world food markets are integrated.

The other strand of studies, which are more closely related to this paper, compares the response of the domestic aggregate food price level to a world market price shock, rather than the comovement of individual agricultural commodity prices, across a large number of countries. Notable papers from this literature include \citet{IMF2011}, \citet{Ianchovichina2014} and \citet{Lee2013}. Below, I will briefly summarize the methodology and main findings for each of these.

\cite{IMF2011} contains an analysis of food price price transmission based on a dataset of monthly FPIs from 31 advanced economies and 47 emerging and developing economies over the period 2000–11. Specifically, they estimate country-by-country autoregressive distributive lag (ADL) models
\begin{equation}
\pi_{t}^{dom}=\delta_{0}+\sum_{s=1}^{12}\rho_{s}\pi_{t-s}^{dom}+\sum_{s=0}^{12}\lambda_{s}\pi_{t-s}^{int}+\epsilon_{t},\label{eq:IMF_model}
\end{equation}
where $\pi_{t}^{dom}$ and $\pi_{t}^{int}$ represent monthly domestic and international food inflation, respectively, at time $t$. Their measure of country specific price transmission is the long run multiplier based on the estimated equation (\ref{eq:IMF_model}). What they find is that developing countries experience higher pass-through of international food inflation to domestic food inflation than advanced economies. Specifically, the estimated median long term pass-through of a 1 percent increase in international food inflation is 0.18 percent in developed countries and 0.34 percent in developing countries. There is no formal analysis of the cross sectional variation in pass-through rates but the authors argue that incomplete pass-through is explained by a significant local component in the production of food, including marketing margins, excise taxes, and customs duties. Moreover, the composition of the individual domestic FPIs differ from the composition of the IMF international food price index used to calculate international food inflation.

\cite{Ianchovichina2014} estimates transmission of yearly world market food inflation to domestic food inflation for eighteen Middle East and North Africa (MENA) using price data covering the period from December 2006 to October 2011. Their baseline model is equation (\ref{eq:IMF_model}) augmented with lagged growth rates of the domestic currency/US dollar exchange rate. The augmented model also allows for asymmetric responses to positive and negative international price and exchange rate changes. Average estimated pass-through coefficients for the MENA countries (based on the World Bank international food price index) are 0.25 and the majority of the countries are reported to have  food price pass-through coefficients in the order of 0.2-0.4. A variance decomposition exercise suggests global food inflation is an important source of domestic food inflation in the region with international price changes accounting for 40-50 percent of the domestic variation in many cases.

\citet{Lee2013} uses annual data and a panel regression approach to estimate the model
\begin{equation}
\pi_{it}^{dom}=\beta_{0}+EV_{it}\beta_{1}+IV_{it}\beta_{2}+\epsilon_{t},\label{eq:ADB_model}
\end{equation}
where $\pi_{it}^{dom}$ represents annual domestic food price inflation in country $i$ at time $t$. $EV_{it}$ is a vector of ``external variables'' including current and lagged international food inflation where the international food price level is proxied by the FAO price index, as well as a set of intra- and extra-regional  food inflation averages. $IV_{it}$ is a vector of ``internal variables'' representing domestic supply and demand factors. Two variables in this category of particular interest to the present paper are a country's current per capita GDP and GDP growth rate in US dollars. Surprisingly, the authors find a negative relationship between a country's per capita GDP growth rate and its food inflation rate. The relationship between domestic food inflation and the per capita GDP level is found to be positive albeit only marginally significant.

### 2015-12-22 conclusion.Rnw
Concl. sec of 1. round rev. v. of paper

% !Rnw root = Master.Rnw

\section{Conclusion\label{sec:conclusion}}
This paper contributes to the literature on aggregate food price transmission with three findings. First, domestic food prices in middle income countries react the strongest to international agricultural commodity price changes once we control for differences in consumption and trade dependency patterns between high and low income countries. Secondly, domestic price increases were the largest in middle income countries during the two recent international price spikes in 2007-08 and 2010-11, all else equal. Thirdly, there is a simple economic logic behind this inverse-U shaped income-price transmission relationship. These findings are worrying due to two other stylized facts. The first is that middle income countries are now home to around three quarters of the world’s poor. The second is that poor households spend a large portion of their income on food and are therefore vulnerable to sudden increases in the price of food goods. There is, however, also a positive side to this finding. If, as some authors claim, middle income countries, in principle, have the domestic financial resources to eradicate extreme poverty \citep{Sumner2013}, then they should also possess the necessary resources to implement social safety nets which shield purchasing power of the poor from occasional price spikes on the international markets.

The analysis has at least two major policy implications. First, when the next surge occurs on the international agricultural markets, the international donor community should direct its resources towords countries in which poor households are particularly vulnerable to international food price spikes and where the government is unable to to provide sufficient assistance, rather than the LICs in general. This could be LICs which, for some reason, has to import most of its food goods. Secondly, the strong focus in the media and among agricultural economists on international food crises is misguided in the sense that LICs, in general, are primarily vulnerable to domestic rather than international shocks.

### 2015-12-22 figures.Rnw
Figures file for 1. round rev. v. of paper

% !Rnw root = Master.Rnw

\SweaveInput{data_prep.Rnw}

\section{Figures}
<<pt, echo=FALSE, include=FALSE, fig=TRUE, height=6, width=8>>=
a = 1.3
b = 1.2
par(mfrow=c(1,2))
curve(1-a^(-x), 0, 10, lty=1, xlab = "Income", ylab = "", ylim = c(0,1))
curve(b^(-x), add = T, col = "blue", lty=2)
title(main = "Shares", cex.main = 1)
legend("topright", c(expression(beta(y)), expression(gamma(y))),lty = c(2,1), col = c("blue", "black"))
curve((1-a^(-x))*b^(-x), 0, 10, xlab = "Income", ylab = "")
title(main = "Price transmission", cex.main = 1)
legend("topright", c(expression(theta(y))), lty = 1, col = "black")
@

<<gamma, echo=FALSE, include=FALSE, fig=TRUE, height=6, width=8>>=
library(ggplot2)
library("foreign")

dat = read.csv("dist_trade.csv", sep = ";")
dat$gamma = (dat$vop_covx+dat$vop_covm)/dat$vop_covt*100
ggplot(dat, aes(log(gdppcp00), gamma)) + 
  geom_point() + 
  geom_smooth() +
  theme_bw() +
  xlab("Log of GDP") +
  ylab(expression(paste("Estimated ", gamma))) +
  theme(panel.grid.major = element_line(colour = "grey"))
@

<<infl, echo=FALSE, include=FALSE, fig=TRUE, height=6, width=8>>=
library(gridExtra)

p1 = ggplot(aes(y = value, x = factor(time, labels = c("2007:1-2008:6", "2010:1-2010:12")), 
                fill = factor(variable, labels = c("FPI inflation", "CPI inflation", "Real FPI growth"))), data = infl_long1) + 
  geom_boxplot(outlier.shape = 1) +
  theme_bw() +
  scale_fill_grey(start = 0, end = 1) +
  theme(legend.position=c(.75,.775)) +
  labs(fill = "") +
  ylab("Annualized growth rates (%)") + 
  xlab("")

p2 = ggplot(aes(x=foodinfl78, y=foodinfl1011), data = infl) + 
  geom_point(shape=1) +
  geom_smooth(method=loess) +
  theme_bw() +
  ylab("FPI inflation 2010-11") + 
  xlab("FPI inflation 2007-08") +
  geom_text(aes(label=ifelse(foodinfl78>40,as.character(iso3c),'')),hjust=1.2,just=-1.2, 
            size=3) +
  geom_text(aes(label=ifelse(foodinfl1011>20,as.character(iso3c),'')),hjust=1.2,just=-1.2,
            size=3)

p3 = ggplot(aes(x=cpiinfl78, y=cpiinfl1011), data = infl) +
  geom_point(shape=1) +
  geom_smooth(method=loess) +
  theme_bw() +
  ylab("CPI inflation 2010-11") + 
  xlab("CPI inflation 2007-08") +
  geom_text(aes(label=ifelse(cpiinfl78>30,as.character(iso3c),'')),hjust=1.2,just=-1.2, 
            size=3) +
  geom_text(aes(label=ifelse(cpiinfl1011>15,as.character(iso3c),'')),hjust=1.2,just=-1.2,
            size=3)

p4 = ggplot(aes(x=rfpigr78, y=rfpigr1011), data = infl) + 
  geom_point(shape=1) +
  geom_smooth(method=loess) +
  theme_bw() +
  ylab("Real FPI growth 2010-11") + 
  xlab("Real FPI Growth 2007-08") +
  geom_text(aes(label=ifelse(rfpigr78>12,as.character(iso3c),'')),hjust=1.2,just=-1.2, 
            size=3) +
  geom_text(aes(label=ifelse(rfpigr1011>10,as.character(iso3c),'')),hjust=1.2,just=-1.2,
            size=3)

grid.arrange(p1, p2, p3, p4, ncol=2)
@

\begin{figure}[H]
\begin{center}
\includegraphics[width=1\textwidth]{Master-pt.pdf}
\caption{The effect of income on price transmission}
\label{fig:pt}
\end{center}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=1\textwidth]{Master-gamma.pdf}
\caption{Estimates of $\gamma$. Source: \cite{Anderson2008}. Own calculations.}
\label{fig:gamma}
\end{center}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=1\textwidth]{Master-infl.pdf}
\caption{Domestic price changes 2007-8 and 2010-11. Source: ILO, own calculations.}
\label{fig:infl}
\end{center}
\end{figure}

### 2015-12-22 regressions.Rnw
regression file for 1. round rev. v. of paper

% !Rnw root = Master.Rnw

\SweaveInput{data_prep.Rnw}

<<dom_infl_0708,echo=F>>=
library(sandwich)

m1 = lm(foodinfl78~ log(GDP)+I(log(GDP)^2)+ staple_share*cdep, 
       data = data_0708)
cov = vcovHC(m1, type = "HC1")
rse1 = sqrt(diag(cov))

m2 = lm(foodinfl78~log(cons)+I(log(cons)^2) + staple_share*cdep, 
       data = data_0708)
cov = vcovHC(m2, type = "HC1")
rse2 = sqrt(diag(cov))

m3 = lm(foodinfl78~log(food)+I(log(food)^2) + staple_share*cdep, 
       data = data_0708)
cov = vcovHC(m3, type = "HC1")
rse3 = sqrt(diag(cov))

m4 = lm(rfpigr78~log(GDP)+I(log(GDP)^2) + staple_share*cdep, 
       data = data_0708)
cov = vcovHC(m4, type = "HC1")
rse4 = sqrt(diag(cov))

m5 = lm(rfpigr78~log(cons)+I(log(cons)^2) + staple_share*cdep,
       data = data_0708)
cov = vcovHC(m5, type = "HC1")
rse5 = sqrt(diag(cov))

m6 = lm(rfpigr78~log(food)+I(log(food)^2) + staple_share*cdep, 
       data = data_0708)
cov = vcovHC(m6, type = "HC1")
rse6 = sqrt(diag(cov))
@

<<dom_infl_1011, echo=F>>=

m1 = lm(foodinfl1011~log(GDP)+I(log(GDP)^2) + staple_share*cdep, 
       data = data_1011)
cov = vcovHC(m1, type = "HC1")
rse1 = sqrt(diag(cov))

m2 = lm(foodinfl1011~log(cons)+I(log(cons)^2) + staple_share*cdep, 
       data = data_1011)
cov = vcovHC(m2, type = "HC1")
rse2 = sqrt(diag(cov))

m3 = lm(foodinfl1011~log(food)+I(log(food)^2) + staple_share*cdep, 
       data = data_1011)
cov = vcovHC(m3, type = "HC1")
rse3 = sqrt(diag(cov))

m4 = lm(rfpigr1011~log(GDP)+I(log(GDP)^2) + staple_share*cdep, 
       data = data_1011)
cov = vcovHC(m4, type = "HC1")
rse4 = sqrt(diag(cov))

m5 = lm(rfpigr1011~log(cons)+I(log(cons)^2) + staple_share*cdep,
       data = data_1011)
cov = vcovHC(m5, type = "HC1")
rse5 = sqrt(diag(cov))

m6 = lm(rfpigr1011~log(food)+I(log(food)^2) + staple_share*cdep, 
       data = data_1011)
cov = vcovHC(m6, type = "HC1")
rse6 = sqrt(diag(cov))
@

<<ts_pt, echo=FALSE>>=
library(dynlm)

ptmat = matrix(, nrow = nrow(data_0708), ncol = 2)
data$id = as.numeric(data$iso3c)
for (i in 1:max(data$id)) {
  dat = data[data$id==i,]
  ptmat[i,1] = as.character(dat[1,"iso3c"])
  dat = ts(dat, start=c(2005, 1), end=c(2010, 12), frequency=12)
  m = dynlm(d(log(afpi))~ d(log(fpi)) + 
              L(d(log(fpi)),1:6) +
              L(d(log(afpi)),1:6), data = dat)
  ptmat[i,2] = sum(m$coefficients[2:8])/(1-sum(m$coefficients[9:14]))
}
ptdat = as.data.frame(ptmat)
names(ptdat) = c("iso3c", "lri")
ptdat = merge(ptdat, data_0708, by = "iso3c", all.x=TRUE)
ptdat$lri = as.numeric(ptdat$lri)

m1 = lm(lri~log(GDP) + I(log(GDP)^2)+staple_share*cdep, data = ptdat)
cov = vcovHC(m1, type = "HC1")
rse1 = sqrt(diag(cov))

m2 = lm(lri~log(cons) + I(log(cons)^2) + staple_share*cdep, 
       data = ptdat)
cov = vcovHC(m2, type = "HC1")
rse2 = sqrt(diag(cov))

m3 = lm(lri~log(food) + I(log(food)^2) + staple_share*cdep, 
       data = ptdat)
cov = vcovHC(m3, type = "HC1")
rse3 = sqrt(diag(cov))

ptmat = matrix(, nrow = nrow(data_0708), ncol = 2)
data$id = as.numeric(data$iso3c)
for (i in 1:max(data$id)) {
  dat = data[data$id==i,]
  ptmat[i,1] = as.character(dat[1,"iso3c"])
  dat = ts(dat, start=c(2005, 1), end=c(2010, 12), frequency=12)
  m = dynlm(d(log(rafpi))~ d(log(fpi)) + 
              L(d(log(fpi)),1:6) +
              L(d(log(rafpi)),1:6), data = dat)
  ptmat[i,2] = sum(m$coefficients[2:8])/(1-sum(m$coefficients[9:14]))
}
ptdat1 = as.data.frame(ptmat)
names(ptdat1) = c("iso3c", "lri1")
ptdat = merge(ptdat, ptdat1, by = "iso3c", all.x=TRUE)
ptdat$lri1 = as.numeric(ptdat$lri1)

m4 = lm(lri1~log(GDP)+I(log(GDP)^2) + staple_share*cdep, 
       data = ptdat)
cov = vcovHC(m4, type = "HC1")
rse4 = sqrt(diag(cov))

m5 = lm(lri1~log(cons) + I(log(cons)^2) + staple_share*cdep,
       data = ptdat)
cov = vcovHC(m5, type = "HC1")
rse5 = sqrt(diag(cov))

m6 = lm(lri1~log(food) + I(log(food)^2) + staple_share*cdep, 
       data = ptdat)
cov = vcovHC(m6, type = "HC1")
rse6 = sqrt(diag(cov))
@

<<panel_pt, echo=F>>=
library(plm)

m1 = plm(diff(log(afpi)) ~ lag(diff(log(afpi)),1:6)+
           lag(diff(log(fpi)),6)*log(GDP) +
           lag(diff(log(fpi)),6)*I(log(GDP)^2),
        data = data, model = "within")

cov = vcovHC(m1, method = "arellano", type = "HC1", cluster = "group")
rse1 = sqrt(diag(cov))

m2 = plm(diff(log(afpi)) ~ lag(diff(log(afpi)),1:6)+
           lag(diff(log(fpi)),6)*log(cons) +
           lag(diff(log(fpi)),6)*I(log(cons)^2),
        data = data, model = "within")

cov = vcovHC(m2, method = "arellano", type = "HC1", cluster = "group")
rse2 = sqrt(diag(cov))

m3 = plm(diff(log(afpi)) ~ lag(diff(log(afpi)),1:6)+
           lag(diff(log(fpi)),6)*log(food) +
           lag(diff(log(fpi)),6)*I(log(food)^2),
        data = data, model = "within")

cov = vcovHC(m3, method = "arellano", type = "HC1", cluster = "group")
rse3 = sqrt(diag(cov))

m4 = plm(diff(log(rafpi)) ~ lag(diff(log(rafpi)),1:6)+
           lag(diff(log(fpi)),6)*log(GDP) +
           lag(diff(log(fpi)),6)*I(log(GDP)^2),
        data = data, model = "within")

cov = vcovHC(m4, method = "arellano", type = "HC1", cluster = "group")
rse4 = sqrt(diag(cov))

m5 = plm(diff(log(rafpi)) ~ lag(diff(log(rafpi)),1:6)+
           lag(diff(log(fpi)),6)*log(cons) +
           lag(diff(log(fpi)),6)*I(log(cons)^2),
        data = data, model = "within")

cov = vcovHC(m5, method = "arellano", type = "HC1", cluster = "group")
rse5 = sqrt(diag(cov))

m6 = plm(diff(log(rafpi)) ~ lag(diff(log(rafpi)),1:6)+
           lag(diff(log(fpi)),6)*log(food) +
           lag(diff(log(fpi)),6)*I(log(food)^2),
        data = data, model = "within")

cov = vcovHC(m6, method = "arellano", type = "HC1", cluster = "group")
rse6 = sqrt(diag(cov))
@

### 2015-12-22 data_prep.Rnw
Data prep. file for 1. round rev. v. of paper

% !Rnw root = Master.Rnw

<<datasets, echo=F>>=
library(readstata13)
library(zoo)
library(plm)
library(foreign)
library(reshape)

ilo_data = read.dta13("ilodata_small_clean.dta")
icp_data = read.dta13("icp_exp_2005.dta")
icp_data$cfood = scale(icp_data$food)
wb_pov = read.dta13("wb_pov.dta")
wb_pov$cpovgap2 = scale(wb_pov$pov_gap_2)
sids = read.dta13("sids_list_s.dta")
fao_fpi = read.csv("fao_fpi.csv", sep = ";")
load("staples.Rda")
load("cdep.Rda")
infl_0708 =  read.dta13("infl_0708.dta")
infl_1011 =  read.dta13("infl_1011.dta")

df = ilo_data[,c("id", "year", "month", "acpi", "afpi", "rafpi", "iso3c")]
df1 = merge(df, icp_data, by = "iso3c", all.x=TRUE)
df1 = merge(df1, wb_pov, by = "iso3c", all.x=TRUE)
df1 = merge(df1, sids, by = "iso3c", all.x=TRUE)
df1 = merge(df1, staples, by = "iso3c", all.x=TRUE)
df1 = merge(df1, cdep, by = "iso3c", all.x=TRUE)

df1$tt = as.yearmon(paste(df1$year, df1$month, sep = "-"))
df1 = df1[order(df1$id, df1$tt),]
df2 = cbind(df1, fao_fpi)
df2["sidsdum"][is.na(df2["sidsdum"])] <- 0
data = plm.data(df2, index = c("id", "tt"))

quar_gdp = quantile(icp_data$GDP, c(.25, .5, .75), na.rm = T)
fao_0708 = (fao_fpi[42,2]/fao_fpi[25,2]-1)*100
fao_1011 = (fao_fpi[72,2]/fao_fpi[61,2]-1)*100

infl_0708 = rename(infl_0708, c(id="iso3c"))
data_0708 = merge(infl_0708, icp_data, by = "iso3c", all.x=TRUE)
data_0708 = merge(data_0708, wb_pov, by = "iso3c", all.x=TRUE)
data_0708 = merge(data_0708, sids, by = "iso3c", all.x=TRUE)
data_0708 = merge(data_0708, staples, by = "iso3c", all.x=TRUE)
data_0708 = merge(data_0708, cdep, by = "iso3c", all.x=TRUE)
data_0708["sidsdum"][is.na(data_0708["sidsdum"])] <- 0
data_0708 = cbind(data_0708, fao_0708)

infl_1011 = rename(infl_1011, c(id="iso3c"))
data_1011 = merge(infl_1011, icp_data, by = "iso3c", all.x=TRUE)
data_1011 = merge(data_1011, wb_pov, by = "iso3c", all.x=TRUE)
data_1011 = merge(data_1011, sids, by = "iso3c", all.x=TRUE)
data_1011 = merge(data_1011, staples, by = "iso3c", all.x=TRUE)
data_1011 = merge(data_1011, cdep, by = "iso3c", all.x=TRUE)
data_1011["sidsdum"][is.na(data_1011["sidsdum"])] <- 0
data_1011 = cbind(data_1011, fao_1011)

infl = merge(infl_0708, infl_1011, by = "iso3c")
infl_long = reshape(infl,varying=c("foodinfl78","foodinfl1011","cpiinfl78", "cpiinfl1011", 
                                  "rfpigr78", "rfpigr1011"), direction="long", 
                   idvar="iso3c", sep="")
infl_long1 = melt(infl_long, id = c("iso3c", "time"))
@

### 2016-01-17 litrev.Rnw
Revised litrev.

@@
 \section{Review of the price transmission literature \label{sec:litrev}}
-This section introduces the agricultural price transmission literature. Results from a few key studies of particular relevance will be discussed in some detail.
+This section reviews five studies from the agricultural price transmission literature of particular relevance, namely \citet{IMF2011}, \citet{Ianchovichina2014}, \citet{Lee2013}, \cite{Kalkuhl2014} and \cite{greb12}.
+Similar to this paper, each of these studies estimate price transmission rates for a large number of countries.
+The review discusses the methodological approach and identifies factors that appears to be important determinants of cross-country differences in price transmission rates, findings which will form the basis of my own empirical analysis in section \ref{sec:analysis}.
+The review will show, however, that our knowledge of global patterns of price transmission, in particular the determinants of these patterns, is currently very limited.
+But before I go into the particularities of each study I will say a few words about their common economic and econometric framework.
 
-There is a vast empirical literature studying the price relationships between spatially separated agricultural markets \citep[see][]{Fackler2001}.
-Within this field we can distinguish between two strands of studies.
-First, there are studies assessing the connectedness or degree of integration characterizing different markets within the same country or region.\footnote{\citet{Ravallion1986}, \citet{Baulch1997}, \citet{Goodwin2001} and \citet{Barrett2002} are some of the seminal papers.} Recent papers focusing on the recent period include\citet{Abidoye2014},  \citet{Baquedano2011}, \citet{Baquedano2014}, \citet{Cudjoe2010} and \citet{Esposti2013}.
-A common thread among these is the use of actual commodity prices in an attempt to test whether the law of one price holds or, more generally, whether the domestic and world food markets are integrated.
+\citet{Fackler2001} provide a thorough discussion of the economics of price transmission.
+It seems to be based always on some notion of market integration which simply means that that, due to arbitrage, prices in spatially separated market cannot be out of equilibrium for too long.
+Spatial arbitrage implies that prices of a homogenous good in any two locations differ by, at most, the cost of moving the good from the low to the high cost market.
+This condition is also referred to as the (weak) law of one price.
+There is also an aggregate version of the law of one price stated in terms of price indexes, rather than individual commodity prices, known as Purchasing Power Parity, or PPP \citep[][p.
+977-78]{Fackler2001}.
+Section \ref{sec:model} will further develop the PPP approach to price transmission in particular the role of income.
 
-The other strand of studies, which are more closely related to this paper, compares the response of the domestic aggregate food price level to a world market price shock, rather than the comovement of individual agricultural commodity prices, across a large number of countries.
-Notable papers from this literature include \citet{IMF2011}, \citet{Ianchovichina2014} and \citet{Lee2013}.
-Below, I will briefly summarize the methodology and main findings for each of these.
+There are essentially two econometric approaches to price transmission; one based on single equation models and one based on vector models.
+In the single equation framework there is always a split between exogenous explanatory variable and an endogenous dependent variable.
+In practice, this implies that changes to the international price is assumed to affect the domestic price but not the other way around.
+Vector models, on the other hand, allow for joint dependence and are therefore more general.
 
-\cite{IMF2011} contains an analysis of food price price transmission based on a dataset of monthly FPIs from 31 advanced economies and 47 emerging and developing economies over the period 2000–11.
-Specifically, they estimate country-by-country autoregressive distributive lag (ADL) models
+The single equation framework to price transmission analysis is based on the autoregressive distributive lag, ADL($p$,$q$,$n$), model
 \begin{equation}
-\pi_{t}^{dom}=\delta_{0}+\sum_{s=1}^{12}\rho_{s}\pi_{t-s}^{dom}+\sum_{s=0}^{12}\lambda_{s}\pi_{t-s}^{int}+\epsilon_{t},\label{eq:IMF_model}
+p_{t}^{dom}=\sum_{s=1}^{p}\rho_{s}p_{t-s}^{dom}+\sum_{s=0}^{q}\lambda_{s}p_{t-s}^{int}+\sum_{s=0}^{n}c'_{s}x_{t-s}+\epsilon_{t},\label{eq:ADL_model}
 \end{equation}
-where $\pi_{t}^{dom}$ and $\pi_{t}^{int}$ represent monthly domestic and international food inflation, respectively, at time $t$.
-Their measure of country specific price transmission is the long run multiplier based on the estimated equation (\ref{eq:IMF_model}).
-What they find is that developing countries experience higher pass-through of international food inflation to domestic food inflation than advanced economies.
-Specifically, the estimated median long term pass-through of a 1 percent increase in international food inflation is 0.18 percent in developed countries and 0.34 percent in developing countries.
-There is no formal analysis of the cross sectional variation in pass-through rates but the authors argue that incomplete pass-through is explained by a significant local component in the production of food, including marketing margins, excise taxes, and customs duties.
-Moreover, the composition of the individual domestic FPIs differ from the composition of the IMF international food price index used to calculate international food inflation.
+where $p_{t}^{dom}$ and $p_{t}^{int}$ represent monthly domestic and international prices, respectively and $x_{t}$ is a $K$ dimensional vector of control variables (including a constant).
+The flexibility of the ADL model makes it a workhorse in applied time series analysis.
+One especially attractive feature is that it can be parameterized to allow for estimation of cointegration relationships \citep[see][]{pesaran98}.
+If international and domestic price levels are assumed or found to be unit root processes but do not cointegrate, then the model is estimated in inflation rates rater than price levels.
+A commonly used measure of price transmission in this case is the long run price multiplier
+\begin{equation}
+\theta=\frac{\sum_{s=0}^{q}\lambda_{s}}{1-\sum_{s=1}^{p}\rho_{s}},\label{eq:LR_multiplier}
+\end{equation}
+measuring the long run impact or pass-through of an international price change on the domestic food price level.
+The short term impact of a permanent international price change can be defined as the numerator of (\ref{eq:LR_multiplier}).
 
-\cite{Ianchovichina2014} estimates transmission of yearly world market food inflation to domestic food inflation for eighteen Middle East and North Africa (MENA) using price data covering the period from December 2006 to October 2011.
-Their baseline model is equation (\ref{eq:IMF_model}) augmented with lagged growth rates of the domestic currency/US dollar exchange rate.
-The augmented model also allows for asymmetric responses to positive and negative international price and exchange rate changes.
+The ADL model is appropriate when the causal impact is unidirectional.
+This could be the case when the domestic price is from a small country relative to the world market but, more often, this assumption is invoked when the domestic price variable is a Food Price Index (FPI), i.e.
+the food component of the Consumer Price Index (CPI).
+
+The other major framework for empirical price transmission analysis is based on the cointegrated VAR model, in particular the parametrization known as the vector error correction model (VECM).
+Simple VECMs are used extensively to characterize the degree of market integration and to test the law of one price when both the international and the domestic price relate to commodity prices.
+An attractive feature of the VECM is that it separates the dynamics of the prices in question into a component characterizing the long run equilibrium, i.e.
+the cointegrating relationship, and a short run component characterizing the dynamics bringing the two prices back into equilibrium.\footnote{For further detail see \citet{greb12} and the references therein.
+\citet{Pippenger2008} discuss the pitfalls one must avoid when testing the law of one price in commodity markets.}
+
+The first of the studies considered, \cite{IMF2011}, contains an analysis of food price price transmission based on a dataset of monthly FPI inflation from 31 advanced economies and 47 emerging and developing economies over the period 2000 - 2011.
+Specifically, for each of the countries they estimate a 12-lag ADL models à la equation (\ref{eq:ADL_model}), without any control variables.
+Their international price variable is the IMF food price index in log differences.
+
+A main result in \cite{IMF2011} is that emerging and developing countries generally experience higher pass-through of international food inflation to domestic food inflation than advanced economies.
+Specifically, the estimated median long term pass-through of a 1 percent increase in international food inflation is 0.18 percent in developed countries and 0.34 percent in emerging and developing countries.
+Curiously, \citet{IMF2011} offers no explanation of this key finding.
+However, as \citet{Abbott2009} points out a) developing countries have more basic diets which involve a greater share of staples, and b) processing and distribution margins are typically lower than in developed countries, meaning they can shrink less as costs increase.
+Both of these factors, he argues, caused a stronger domestic food inflation response among developing countries than in developed countries to the high world prices in 2007-08.
+
+\citet{Ianchovichina2014} takes the analysis in \citet{IMF2011} a bit further by including lagged growth rates of the domestic currency/US\$ exchange rate as a control variable.
+Their model also allows for asymmetric responses to positive and negative international price and exchange rate changes.
+Their data covers eighteen Middle East and North Africa (MENA) countries over the period December 2006 to October 2011.
 Average estimated pass-through coefficients for the MENA countries (based on the World Bank international food price index) are 0.25 and the majority of the countries are reported to have  food price pass-through coefficients in the order of 0.2-0.4.
 A variance decomposition exercise suggests global food inflation is an important source of domestic food inflation in the region with international price changes accounting for 40-50 percent of the domestic variation in many cases.
+\citet{Ianchovichina2014} do use formal statistical methods to explain the cross-country variation in estimated pass-through rates.
+Instead they speculate how different types of market interventions and structural differences across countries might contribute to the observed differences in estimated pass-through rates.
 
-\citet{Lee2013} uses annual data and a panel regression approach to estimate the model
+\citet{Lee2013} use annual data from 72 countries over the period 2000 - 2011 and a panel regression approach to estimate the model
 \begin{equation}
@@
 \end{equation}
-where $\pi_{it}^{dom}$ represents annual domestic food price inflation in country $i$ at time $t$.
-$EV_{it}$ is a vector of ``external variables'' including current and lagged international food inflation where the international food price level is proxied by the FAO price index, as well as a set of intra- and extra-regional  food inflation averages.
-$IV_{it}$ is a vector of ``internal variables'' representing domestic supply and demand factors.
-Two variables in this category of particular interest to the present paper are a country's current per capita GDP and GDP growth rate in US dollars.
+where $\pi_{it}^{dom}$ represents annual domestic FPI inflation in country $i$ at time $t$.
+$EV_{it}$ and $IV_{it}$ are vectors of so called ``external'' and  ``internal'' variables.
+The set of external variables consists of current and lagged international food inflation where the international food price level is proxied by the FAO price index, as well as a set of intra- and extra-regional food inflation averages.
+The internal variables represent domestic supply and demand factors as well as various other country characteristics.
+These include per capita GDP levels and growth rates, money supply and population growth rates, indexes of political stability and the degree of free trade, changes in the domestic currency/US\$ exchange rate and an index of food production and, finally, the share of food imports in merchandise imports and the change in this share.
+Note that the  \citet{Lee2013} model is a version of the general model (\ref{eq:ADL_model}) without any autoregressive terms and it can be thought of as a Phillips curve type relationship for the food sector.
+Also note that, similarly with the former two studies, \citet{Lee2013} do not model the determinants of price transmission but rather the determinants of domestic FPI inflation.
+
+We would expect negative coefficients to variables representing supply shocks such as changes in the food production index.
+The share of food in merchandise imports and the change in this share are also considered supply variables.
+An increase in the food import share should, all else equal, increase domestic supply and thereby lower food inflation.
+The problem is if the rise in food imports is a result of a confounding domestic demand or supply shock.
+It is less obvious what the expected sign should be on the food import share in levels i.e.
+whether countries that have relatively large food import shares should have higher or lower domestic food inflation rates than countries with low food import shares.
+As it turns out, the coefficient to food production variable is insignificant in all specifications whereas the coefficient to the food import share in levels and changes are found to be consistently negative and positive, respectively.
+That is, countries with large food import shares have, on average, lower food inflation whereas countries experiencing large increases in their food import shares have higher food inflation on average.
+The latter result could, however, reflect that food imports increase when domestic domestic food prices are on the rise, as a response to food inflation rather than a cause of it.
+
+Coefficients to demand variables are expected to be positive.
+GDP growth and money supply are included in this category whereas per capity GDP is meant as a control variable for the money supply-inflation relationship.
 Surprisingly, the authors find a negative relationship between a country's per capita GDP growth rate and its food inflation rate.
-The relationship between domestic food inflation and the per capita GDP level is found to be positive albeit only marginally significant.
+The relationship between domestic food inflation and the per capita GDP level is found to be positive albeit only marginally significant in some specifications.
+
+A depreciation of the domestic currency/US\$ exchange rate increases domestic food inflation significantly across all specifications, as expected, whereas an increase in the money supply has no effect on food inflation.
+More politically stable countries tend to have lower food inflation whereas countries with a higher free trade index score are found to have higher domestic food inflation.
+
+The study by \citet{Kalkuhl2014} makes use of a comprehensive database of domestic and international food commodity prices.
+For each country he estimates equation (\ref{eq:ADL_model}) separately for each domestic and corresponding international food price.
+In addition he estimates models for each country where the dependent price variable is the food inflation rate.
+His control variables consist of the domestic currency/US\$ exchange rate, an international oil price and and a set of monthly dummies.
+His baseline model assumes lag lengths of $q=p=n=3$.
+Oil prices are included due to their influence on domestic production and trade costs.
+\citet{Kalkuhl2014} does not tabulate the estimated impacts on domestic consumer food prices but, according to his figure 3, the short term impact is generally less that 0.2 for most countries.
+The impact on domestic grain prices is somewhat higher.
+\citet{Kalkuhl2014} does not report results related to the impact of oil prices and exchange rates on domestic food prices.
+
+The final study by \citet{greb12} is different from the three others in two respects.
+First, their price transmission estimates are based on the VECM rather than the ADL model.
+Secondly, in addition to the price transmission regressions, they carry out a meta regression analysis of the determinants of price tranmission.
+For the sake of brevity I will focus solely on their meta regression analysis.
+
+\citet{greb12} consider 8 variables covering geographical, infrastructural, institutional and commodity specific factors which are believed to influence the size of pass through rates.\footnote{These are a Landlocked dummy, Trade openness, a grain state trading enterprise dummy, an Ease of trade indicator from the World Bank, Doing Business survey, the Logistics performance index from the World Bank, the Net cereal import ratio (export – import, 3 year average 2009-2011) to domestic consumption and a retail price dummy.} The authors then regress their own price transmission estimates, as well as estimates from the literature, on these explanatory variables.
+What they find is that these covariates explain more of the variation in the estimated speed of adjustment parameters (the $\alpha$'s) than in the cointegration parameters (the $\beta$'s).
+Another major finding is that more of these covariates have significant and plausible partial effects on the $\alpha$'s than on the $\beta$'s.
+However, none of the estimated partial effects are robust to choices of subsample and, as the the author conclude themselves: \emph{``...
+this} [meta-regression (ed.)] \emph{analysis fails to generate compelling results.''}
+
+In summary, the review of the four studies suggests that international commodity price to domestic FPI pass through rates vary significantly across countries and that developing countries seem to experience more pass-through than developed countries.
+\citet{Lee2013} additioanlly found the domestic currency/US\$ exchange rate, the GDP growth rate, the share of food imports and the degree of political stability to be significant determinants of yearly food inflation rates.
+Whether there are country characteristics that can explain the cross country variation in price transmission rates, on the other hand, is still an open question.

### 2016-01-17 cs_infl_regressions.R
Regression scripts

@@ -1,0 +1,233 @@
+library(car)
+library(lmtest)
+library(MASS)
+library(robustbase)
+library(sandwich)
+library(stats)
+library(plyr)
+library(dplyr)
+library(leaps)
+library(relaimpo)
+library(psych)
+
+##################################################################
+### Cross section inflation regressions
+##################################################################
+
+load("ctr_dat_0708.Rda")
+load("ctr_dat_1011.Rda")
+
+## 2007-8 inflation regressions
+data = ctr.dat.0708
+infl.reg.0708 = list()
+infl.fstat = list()
+infl.fstat.p = list()
+
+m = lm(fpigr.0708~log(gdp)+
+         I(log(gdp)^2)+
+         cshare+
+         cdep+
+         cer +
+         lldc+
+         to+
+         dtf+
+         lpi,
+       data = filter(data, fpigr.0708<3.5*sd(fpigr.0708)))
+cov = vcovHC(m, type = "HC1")
+rse = sqrt(diag(cov))
+
+infl.reg.0708[[1]] = m
+infl.reg.0708[[2]] = rse
+
+f = linearHypothesis(m, c("cshare=0", "lldc=0", "dtf=0", "lpi=0"), 
+                     vcov = vcovHC(m, type = "HC1"))
+infl.fstat[[1]] = f[2,3]
+infl.fstat.p[[1]] = f[2,4]
+
+m = lm(fpigr.0708~log(gdp)+
+         I(log(gdp)^2)+
+         cdep+
+         cer+
+         to,
+       data = filter(data, fpigr.0708<3.5*sd(fpigr.0708)))
+cov = vcovHC(m, type = "HC1")
+rse = sqrt(diag(cov))
+
+infl.reg.0708[[3]] = m
+infl.reg.0708[[4]] = rse
+
+m = lm(fpigr.0708~log(cons)+
+         I(log(cons)^2)+
+         cshare+
+         cdep+
+         cer +
+         lldc+
+         to+
+         dtf+
+         lpi,
+       data = filter(data, fpigr.0708<3.5*sd(fpigr.0708)))
+cov = vcovHC(m, type = "HC1")
+rse = sqrt(diag(cov))
+
+infl.reg.0708[[5]] = m
+infl.reg.0708[[6]] = rse
+
[…]

### 2016-01-17 cs_pt_reg.R

@@ -1,0 +1,240 @@
+
+library(car)
+library(dplyr)
+library(sandwich)
+library(lmtest)
+
+##################################################################
+### Cross section price transmission regressions
+##################################################################
+
+load("ctr_dat_0708.Rda")
+load("coef_ecm_fpi.Rda")
+load("coef_ecm_rfpi.Rda")
+load("coef_bew_fpi_l1.Rda")
+
+data = merge(select(ctr.dat.0708, -(2:4)), select(coef.ecm.fpi, iso3c, lrm, lrm.se))
+# data = merge(select(ctr.dat.0708, -(2:4)), coef.bew.fpi.l1)
+# data = rename(data, lrm.fpi=V1, lrm.fpi.se=V2)
+data = dplyr::rename(data, lrm.fpi=lrm, lrm.fpi.se=lrm.se)
+data = merge(data, select(coef.ecm.rfpi, iso3c, lrm, lrm.se))
+data = dplyr::rename(data, lrm.rfpi=lrm, lrm.rfpi.se=lrm.se)
+
+## 1 lag LR multiplier regressions
+pt.reg.l1 = list()
+pt.fstat = list()
+pt.fstat.p = list()
+
+m = lm(lrm.fpi~log(gdp)+
+         I(log(gdp)^2)+
+         cshare+
+         cdep+
+         cer+
+         lldc+
+         to+
+         dtf+
+         lpi,
+       data = filter(data, lrm.fpi<3.5*sd(lrm.fpi)),
+       weights = 1/lrm.fpi.se)
+cov = vcovHC(m, type = "HC1")
+rse = sqrt(diag(cov))
+
+pt.reg.l1[[1]] = m
+pt.reg.l1[[2]] = rse
+
+f = linearHypothesis(m, c("lldc=0", "to=0", "dtf=0", "lpi=0"), 
+                     vcov = vcovHC(m, type = "HC1"))
+pt.fstat[[1]] = f[2,3]
+pt.fstat.p[[1]] = f[2,4]
+
+
+m = lm(lrm.fpi~log(gdp)+
+         I(log(gdp)^2)+
+         cshare+
+         cdep+
+         cer,
+       data = filter(data, lrm.fpi<3.5*sd(lrm.fpi)),
[…]

### 2016-01-17 panel_reg.R

@@ -1,0 +1,154 @@
+
+library(dplyr)
+library(sandwich)
+library(lmtest)
+library(plm)
+library(foreign)
+
+##################################################################
+### Cross section price transmission regressions
+##################################################################
+
+load("ctr_dat_0708.Rda")
+load("ts_dat.Rda")
+df = merge(ts.dat, ctr.dat.0708, by = "iso3c")
+df = df[!duplicated(df), ]
+data = plm.data(df, index = c("iso3c", "tt"))
+df1 = select(df, -fao.fpi)
+df1 = rename(df1, fao.fpi = r.fao.fpi)
+data1 = plm.data(df1, index = c("iso3c", "tt"))
+# write.dta(data, "panel.dta")
+
+panel.reg = list()
+
+m = plm(diff(log(adj.fpi)) ~ lag(diff(log(adj.fpi)),1) +
+          diff(log(fao.fpi)) +
+          lag(diff(log(fao.fpi)), 1) +
+          diff(log(er)) +
+          lag(diff(log(er)), 1) +
+          I(lag(diff(log(fao.fpi)),1)*log(gdp)) +
+          I(lag(diff(log(fao.fpi)),1)*I(log(gdp)^2)) +
+          I(lag(diff(log(fao.fpi)),1)*cshare) +
+          I(lag(diff(log(fao.fpi)),1)*cdep) +
+          I(lag(diff(log(fao.fpi)),1)*I(cshare*cdep)) +
+          I(lag(diff(log(fao.fpi)),1)*lldc) +
+          I(lag(diff(log(fao.fpi)),1)*to) +
+          I(lag(diff(log(fao.fpi)),1)*dtf) +
+          I(lag(diff(log(fao.fpi)),1)*lpi),
+        data = data, model = "within")
+
+cov = vcovHC(m, method = "arellano", type = "HC1", cluster = "group")
+rse = sqrt(diag(cov))
+
+panel.reg[[1]] = m
+panel.reg[[2]] = rse
+
+m = plm(diff(log(adj.fpi)) ~ lag(diff(log(adj.fpi)),1) +
+          diff(log(fao.fpi)) +
+          lag(diff(log(fao.fpi)), 1) +
+          diff(log(er)) +
+          lag(diff(log(er)), 1) +
[…]

### 2016-01-17 ts_reg.R

@@ -1,0 +1,435 @@
+
+library(dplyr)
+library(dyn)
+library(dynlm)
+
+##################################################################
+### Time series regressions
+##################################################################
+
+load("ts_dat.Rda")
+data = ts.dat
+tbl = table(data$id)
+data$id = as.numeric(data$id)
+
+##################################################################
+
+## Finding the optimal lag lengths. 
+## Model: ADL
+
+# 1) adj.fpi
+p = 6; q=6
+aic.mat = array(NA, dim = c(nrow(tbl), p, q))
+aic.min.mat = matrix(NA, nrow = nrow(tbl), ncol = 2)
+
+for (i in 1:max(data$id)) {
+  dat = filter(data, id==i)
+  dat = select(dat, cpi.brk:rfpi, fao.fpi:er)
+  ts.dat = ts(dat, start=c(2005, 1), end=c(2014, 12), frequency=12)
+  
+  if(is.na(dat$er)[1] == FALSE) {
+    for (j in 1:p) {
+      for (k in 1:q) {
+        m = dynlm(d(log(adj.fpi))~ L(d(log(adj.fpi)),1:j) +
+                    d(log(fao.fpi)) +
+                    L(d(log(fao.fpi)),1:k) + 
+                    d(log(er)) +
+                    L(d(log(er)),1:k), 
+                  data = ts.dat)
+        aic.mat[i,j,k] = extractAIC(m)[2]
+      }
+    }
+  } else {
+    for (j in 1:p) {
+      for (k in 1:q) {
+        m = dynlm(d(log(adj.fpi))~ L(d(log(adj.fpi)),1:j) + 
+                    d(log(fao.fpi)) +
+                    L(d(log(fao.fpi)),1:k), 
+                  data = ts.dat)
+        aic.mat[i,j,k] = extractAIC(m)[2]
+      }
+    }
+  }
+  aic.min.mat[i,] = which(aic.mat[i,,] == min(aic.mat[i,,]), arr.ind = TRUE)
+}
+aic.min.fpi = as.data.frame(aic.min.mat)
+aic.min.fpi$iso3c = names(tbl)
+
+# 2) rfpi
+p = 6; q=6
+aic.mat = array(NA, dim = c(nrow(tbl), p, q))
[…]

### 2016-01-29 cs_infl_regressions.R
Revised version

@@ -1,14 +1,8 @@
 library(car)
 library(lmtest)
-library(MASS)
-library(robustbase)
 library(sandwich)
-library(stats)
-library(plyr)
 library(dplyr)
-library(leaps)
 library(relaimpo)
-library(psych)
 
 ##################################################################
 ### Cross section inflation regressions
@@ -17,11 +11,12 @@ library(psych)
 load("ctr_dat_0708.Rda")
 load("ctr_dat_1011.Rda")
 
-## 2007-8 inflation regressions
+## FPI inflation regressions
+fpi.reg.m = list()
+fpi.reg.se = list()
+
+# 2007-08
 data = ctr.dat.0708
-infl.reg.0708 = list()
-infl.fstat = list()
-infl.fstat.p = list()
 
 m = lm(fpigr.0708~log(gdp)+
          I(log(gdp)^2)+
@@ -36,25 +31,8 @@ m = lm(fpigr.0708~log(gdp)+
 cov = vcovHC(m, type = "HC1")
 rse = sqrt(diag(cov))
 
-infl.reg.0708[[1]] = m
-infl.reg.0708[[2]] = rse
-
-f = linearHypothesis(m, c("cshare=0", "lldc=0", "dtf=0", "lpi=0"), 
-                     vcov = vcovHC(m, type = "HC1"))
-infl.fstat[[1]] = f[2,3]
-infl.fstat.p[[1]] = f[2,4]
-
-m = lm(fpigr.0708~log(gdp)+
-         I(log(gdp)^2)+
-         cdep+
-         cer+
-         to,
-       data = filter(data, fpigr.0708<3.5*sd(fpigr.0708)))
-cov = vcovHC(m, type = "HC1")
-rse = sqrt(diag(cov))
-
-infl.reg.0708[[3]] = m
-infl.reg.0708[[4]] = rse
+fpi.reg.m[[1]] = m
+fpi.reg.se[[1]] = rse
 
 m = lm(fpigr.0708~log(cons)+
          I(log(cons)^2)+
@@ -69,63 +47,11 @@ m = lm(fpigr.0708~log(cons)+
 cov = vcovHC(m, type = "HC1")
 rse = sqrt(diag(cov))
 
-infl.reg.0708[[5]] = m
-infl.reg.0708[[6]] = rse
+fpi.reg.m[[2]] = m
+fpi.reg.se[[2]] = rse
 
-m = lm(rfpigr.0708~log(gdp)+
[…]

### 2016-01-29 analysis.Rnw
Revised version

@@
-% !Rnw root = Master.Rnw
-
-\SweaveInput{data_prep.Rnw}
-\SweaveInput{regressions.Rnw}
+% !Rnw root = master1.Rnw
 
@@
 Our goal in this section is thus to a) establish whether there is a statistically significant quadratic relationship between income or, alternatively, expenditure and food price transmission and b) to determine the income/expenditure level corresponding to the turning point.
-Section \ref{sec:data} introduces the data and provides some context.
+Section \ref{sec:data} introduces the food price data and provides some context.
 Results from the statistical analysis are presented in section \ref{sec:analysis1} and \ref{sec:analysis2}.
 
-\subsection{Data on domestic and international food prices\label{sec:data}}
-The domestic price index data analyzed below is sourced from the International Labor Organization (ILO).\footnote{See \href{http://laborsta.ilo.org/}{http://laborsta.ilo.org/}} ILO maintains the most comprehensive database of domestic food price indexes (FPIs) and consumer price indexes (CPIs) which is used to monitor the cost of living and real wages.
-The ILO food price database is not widely used, for good reasons, despite a broad country coverage and series going back to the 1970s in many cases.
+\subsection{Data\label{sec:data}}
+
+<<echo=FALSE>>=
+load("infl_0708_1011.Rda")
+load("ts_dat.Rda")
+n.full = nrow(table(ts.dat$id))
+n.sub = nrow(infl)
+infl = infl[infl$iso3c != "BLR", ]
+@
+
+\subsubsection{Domestic and international food prices}
+Domestic price data is sourced from the International Labor Organization (ILO).\footnote{See \citet{ilo16}} ILO maintains a comprehensive database of domestic food price indexes (FPIs) and consumer price indexes (CPIs) which is used to monitor the cost of living and real wages.
+There are not many studies which draw on this database, however, despite a broad country coverage and long monthly price series.
 A major issue with this data is that the historical series are not recalculated whenever a new base year is introduced.
@@
 This means that many of the series have discrete level shifts associated with these changes.
-Another issue with the data is that some countries report several series.
+Another issue with the data is that some countries report several series representing different socioeconomic groups or geographical areas.
 In some cases one series replace another after a period of overlap.
 Finally, many of the series contain missing values.
-To ``tidy up'' the raw data I carried out the following operations.
-First, I restricted the sample to the period 2005:1-2010:12.
-Secondly, overlapping series were spliced and in cases where multiple complete series were available from the same country, a single one was selected.
+To ``tidy up'' the raw data I carried out the following steps First, I restricted the sample to the period 2005:1-2014:12.
+Secondly, in cases where multiple complete series were available from the same country, a single one was selected.
 Thirdly, series with missing values were discarded.
-Finally, and most critical for the analysis below, level shifts associated with base year changes etc.
+Fourthly, and most critical for the analysis below, level shifts associated with base year changes etc.
 were identified and removed.\footnote{Concretely I added the change in the break point from the previous value to the break point and all points succeeding it.
-The value of the adjusted series at the break point is thus the same as the one preceding it.} In total 140 countries have complete price series for the 2005-10 period considered.
-An important last thing to know about this data is that most of the low income countries' price indexes derive from prices collected from one or two major markets in the capital.
-Weights and selected items are derived from household expenditure surveys.
+The value of the adjusted series at the break point is thus the same as the one preceding it.} Finally, all the series were normalized with respect to a common base period (Jan 2007).
+In total \Sexpr{n.full} countries have complete price series for the entire ten year period considered whereas, for example, there are \Sexpr{n.sub} countries with complete price series covering the 2005-11 subperiod.
+There is therefore a trade-off between the number of complete price series and the length of the series.
+
+An important detail to know about this data is that many of the low income countries' price indexes derive from prices collected from one or two major markets in the capital.
+Furthermore, weights and selected items are derived from household expenditure surveys which often target certain groups such as the urban middle class.
 Differences in consumption patterns around the world of course translate into differences in weights and in the composition of the food baskets themselves.
 This is probably the largest drawback of using index data in a cross section analysis of price transmission.
-That being said, even domestic prices of individual food commodities such as wheat or rice would not refer to the same good due to quality and varietal differences, so it is not the case that focusing on a few selected food items consumed by all countries would solve the problem.
+That being said, even domestic prices of individual food commodities such as wheat or rice would not refer to the same good due to quality and varietal differences, so it is not the case that focusing on a few selected food items consumed by all countries would necessarily solve the problem.
+
+International food prices are represented by the FAO food price index (FAO FPI) and in some cases its ``real'' version, namely the nominal FAO FPI deflated by the Manufacturing Unit Value (MUV) index published by the World Bank.\footnote{see \citet{fao16} and \citet{worldbank16a}.} Table \ref{tab:ts_data} contain information about all time series data employed in the analysis below.
+
+\subsubsection{Domestic food inflation during the food crises}
 
 There has been two recent international price spikes; one in 2007-8 and the other in 2010-11.
-Between January 2007 and June 2008 FAO's food and grain price indexes, which are normalized with respect to the 2002-4 average, rose by 91 and 127 points to index 226 and 268, respectively, corresponding to relative price increases of 68 and 90 percent.
+The former spike especially is often referred to as ``the food criss'' in the literature but there are also studies that consider the latter price spike \citep[e.g.][]{Headey2008, Headey2013, timmer10, Ivanic2012, Martin2011}.
+Between January 2007 and June 2008 FAO's food and grain price indexes rose by 91 and 127 points to index 226 and 268, respectively, corresponding to relative price increases of 68 and 90 percent.
 From the first peak in mid-2008 and to their their respective troughs in February and September 2009 the two indexes fell by 82 and 113 points before rising back up to index 240 and 261 in February and April 2011, respectively.
@@
 
-As mentioned in section \ref{sec:model}, the nominal food price level is not only driven by exogenous world market shocks, it also includes a stochastic trend which it shares with the non-food price level, representing underlying general domestic inflation.
-By dividing the FPI with the CPI, we obtain an FPI which is purged of headline inflation.
-The median country experienced an annualized rate of food and headline inflation of \Sexpr{round(median(infl$foodinfl78),1)} and \Sexpr{round(median(infl$cpiinfl78),1)} percent, respectively, in 2007-8.
-Real FPI growth, which is approximately the difference between the two, was \Sexpr{round(median(infl$rfpigr78),1)} percent for the median country in that period.
+The median country experienced an annualized rate of food and headline inflation of \Sexpr{round(median(infl$fpigr.0708),1)} and \Sexpr{round(median(infl$cpigr.0708),1)} percent, respectively, in 2007-8.
+Nominal food prices are not only driven by exogenous world market shocks.
+They follow a stochastic trend together with the non-food prices, representing underlying general (core) inflation.
+By dividing the FPI with the CPI, we obtain an FPI which is purged of underlying domestic inflation.
+Real FPI growth, which is approximately the difference between the FPI and CPI inflation rate, was \Sexpr{round(median(infl$rfpigr.0708),1)} percent for the median country in that period.\footnote{A problem with the real FPI is that domestic non-food prices are also affected by international commodity shocks.
+Low or even negative growth in real domestic food prices in these two periods could therefore reflect a response to the concurrent surge in non-food commodities, oil in particular, rather than a puzzling negative response to an international food price shock.}
+
 Price increases were generally lower in the second period.
-The median country experienced food and headline inflation rates \Sexpr{round(median(infl$foodinfl1011),1)} and \Sexpr{round(median(infl$cpiinfl1011),1)} percent implying a modest real FPI growth of \Sexpr{round(median(infl$rfpigr1011),1)} percent.
+The median country experienced food and headline inflation rates \Sexpr{round(median(infl$fpigr.1011),1)} and \Sexpr{round(median(infl$cpigr.1011),1)} percent and a modest real FPI growth of \Sexpr{round(median(infl$rfpigr.1011),1)} percent.
 In both periods the variation in inflation rates is substantial.
-The standard deviation and interquartile range of 2007-8 food inflation is \Sexpr{round(sd(infl$foodinfl78),1)} and \Sexpr{round(IQR(infl$foodinfl78),1)} percentage points, respectively, versus \Sexpr{round(sd(infl$foodinfl1011),1)} and \Sexpr{round(IQR(infl$foodinfl1011),1)} percentage points in 2010-11.
-Although median real FPI growth is lower in the second period the variation is similar in both periods (standard deviations of \Sexpr{round(sd(infl$rfpigr78),1)} and \Sexpr{round(sd(infl$rfpigr1011),1)} and interquartile ranges of \Sexpr{round(IQR(infl$rfpigr78),1)} and \Sexpr{round(IQR(infl$rfpigr1011),1)}, respectively).
-The three remaining panels contain alternative representations the same data.
-What they show us that there is some systematic variation in the food prices to be explained.
-As the positively sloped lowess lines indicate, a country hit by a large nominal and real FPI shock in 2007-8 was more likely to be hit by a relatively large shock again in 2010-11 and vice versa.
+The standard deviation and interquartile range of 2007-8 food inflation is \Sexpr{round(sd(infl$fpigr.0708),1)} and \Sexpr{round(IQR(infl$fpigr.0708),1)} percentage points, respectively, versus \Sexpr{round(sd(infl$fpigr.1011),1)} and \Sexpr{round(IQR(infl$fpigr.1011),1)} percentage points in 2010-11.
+Although median real FPI growth is lower in the second period the variation is similar in both periods (standard deviations of \Sexpr{round(sd(infl$rfpigr.0708),1)} and \Sexpr{round(sd(infl$rfpigr.1011),1)} and interquartile ranges of \Sexpr{round(IQR(infl$rfpigr.0708),1)} and \Sexpr{round(IQR(infl$rfpigr.1011),1)}, respectively).
+The three remaining panels contain alternative representations the same data and there appears to be some systematic variation.
+As the (mostly) positively sloped lowess lines indicate, a country hit by a large nominal and real FPI shock in 2007-8 was more likely to be hit by a relatively large shock again in 2010-11 and vice versa.
 This suggests that there are indeed characteristics which influences a country's response to a world market shock, an assertion that will be explored further in section \ref{sec:analysis1} and \ref{sec:analysis2}.
 
-\subsection{Determinants of domestic food price changes\label{sec:analysis1}}
+\subsubsection{Country characteristics}
+Table \ref{tab:cs_data} lists the variables that, based on the discussion in section \ref{sec:model}, are considered potentially important determinants of price transmission.
+I consider two proxies for a country's income level, namely country $i$'s (PPP adjusted) per capita GDP as well as its household consumption expenditure.
+The other covariates are a country $i$'s percentage dietary energy supply derived from cereals, roots and tubers, its cereal import- and export shares, a dummy indicating whether the country is a landlocked developing country and its trade openness defined as the sum of its imports and exports relative to its GDP.
+In addition there are two indexes included representing how easy it is to trade with the country in question and the quality of its infrastructure.
 
-Table \ref{tab:2007-8-regressions} and \ref{tab:2010-11-regressions} delve deeper into the  relationship between income and domestic food price changes at the time of the 2007-08 and 2010-11 international price spikes, respectively.
-Each of the columns refer to a regression of FPI inflation or real FPI growth on a set of income related predictor variables.
+The analysis in section \ref{sec:model} suggests that international price shocks are transmitted the strongest to middle income countries so we would expect food price inflation in these two periods to follow an inverse-U shaped relationship with per capita income and consumption expenditure.
+
+Some populations consume more basic food goods than others, for cultural reasons, which make their domestic food price level more responsive to price changes on grains and other internationally traded agricultural commodities.
+Furthermore, poor households also spend a larger share of their food budget on staples.
+Therefore, if we did account for this, the estimated relationship between income and domestic food price changes could be spurious.
+
+Net food importers as well as exporters are, all else equal, more connected to the world market and domestic food prices in these countries should therefore react stronger to world market price shocks than prices in countries that are self sufficient in food.
+Countries that are largely dependent on grain imports to cover their domestic consumption needs can lower tariffs in response to an international price shock, but this will only have a substantial effect if tariffs are high to begin with.
+Many developing countries lowered their tariffs during the food crisis but the scope for major reductions was there only for a few countries \citet{imf08, Demeke2009}.
+All in all, we would expect to find a stronger transmission of international food price spikes to domestic price increases among countries with large import dependency ratios.
+Food importers could, of course, resort more often than other countries to domestic price stabilizing policies.
+In this case, there would be no or perhaps even a negative relationship between a country's import share and its price transmission rate.
+Such considerations also apply to the cereal consumption share discussed above.
+
+Net food exporters, on the other hand, can use export restrictions as way of stabilizing domestic prices and some countries often do.
+Major grain exporters such as India, Argentina, Russia and Ukraine, among others, intervened in their export markets during the food crisis \citep{mitra09}.
+The question is whether Such interventions are, however, highly unpopular among the farmers and do not always work.
+
+\subsection{Determinants of domestic food price changes during the 2007-8 and 2010-11 food crises\label{sec:analysis1}}
+This section will discuss whether the cross country distribution of domestic food price changes during the recent international price spikes was as expected.
+That is, did middle income countries with open markets, good infrastructure and large shares of grains in their food consumption experience the greatest food price increases in these two periods?
+A discussion of whether there is a systematic variation in the degree to which domestic food prices react to international food price changes in general is postponed to section \ref{sec:analysis2}.
+
+Table \ref{tab:fpi-regressions} and \ref{tab:rfpi-regressions} consider the cross sectional variation in domestic food price changes during the 2007-8 and 2010-11 food crises.
+Each of the columns refer to a regression of FPI inflation or real FPI growth on a subset of the explanatory variables discussed above.
 The models can all be written
 \begin{equation}
-\pi_{i}^{dom}=\delta_{0} + \delta_{1}x_{i} + \delta_{2}x_{i}^{2} + \delta_{3}z_{i} + u_{i}\label{eq:infl_model}
+\pi_{i}^{dom}=\delta_{0} + \delta_{1}x_{i} + u_{i}\label{eq:infl_model}
 \end{equation}
-where $\pi_{i}^{dom}$ refers to either country $i$'s FPI inflation or its real FPI growth in either the 18 month period 2007:1-2008:6 or the 12 month period 2010:1-2010:12, $x_{i}$ refers to a subset of the explanatory variables considered, $z_{i}$ is a vector of control variables and $u_{i}$ is the error term.
-The covariates, $x_{i}$, include the log of country $i$'s (PPP adjusted) per capita GDP [$\log(GDP)$], its consumption [$\log(cons)$] and food expenditure level [$\log(food)$].
-The control variables in $z_{i}$ are country $i$'s percentage dietary energy supply derived from cereals, roots and tubers ($cshare$), its cereal import dependency ratio ($cdep$) and their interaction.\footnote{Cereal dietary energy supply and cereal dependency data are from FAOSTAT and refer to the 2004-06 average.
-Data on GDP and expenditure is sourced from the World Bank's website.
-It is the International Comparison Project (ICP) which is behind the expenditure estimates which refer to year 2005.} Summary statistics on these variables are provided in table \ref{tab:sumstat}.
-Of primarily interest is the quadratic relationship between domestic price changes $\pi_{i}^{dom}$ and expenditure, $x_{i}$.
-The first 12 rows of table \ref{tab:tp} summarize information about the turning points based on different estimated versions of model \ref{eq:infl_model}.
+where $\pi_{i}^{dom}$ refers to either country $i$'s FPI inflation or its real FPI growth in either the 18 month period 2007:1-2008:6 or the 12 month period 2010:1-2010:12.
+$x_{i}$ is the covariate vector and $u_{i}$ is the error term.
 
 \begin{center}
-[Table \ref{tab:sumstat} around here]
+[Table \ref{tab:fpi-regressions} around here]
 \par\end{center}
 
-Countries that are largely dependent on grain imports to cover their domestic consumption needs cannot use export restrictions as way of stabilizing domestic prices.
-Because of this vulnerability to international food price fluctuations we would expect to find stronger domestic price increases during international food price spikes among countries with large import dependency ratios.
-The analysis in section \ref{sec:model} suggests that international price shocks are transmitted the strongest to middle income countries so we would expect food price inflation in these two periods to follow an inverse-U shaped relationship with per capita income and consumption expenditure.
-Poor households and populations also spend a larger share of their food budget on staples which makes them more vulnerable to price spikes on grains and other internationally traded agricultural commodities.
-Therefore, if we did account for this, the estimated relationship between income and domestic food price changes could be spurious.
-By including the $cshare$ variable in the regression this omitted variable bias should be reduced.
+In table \ref{tab:fpi-regressions}, the non-income related explanatory variables that are significant in half of the specifications or more are a country's cereal consumption share, its cereal import and export shares and landlocked status.
+As expected, countries with large cereal consumption shares experienced significantly higher food inflation than countries with lower cereal consumption shares during the 2010-11 international price spike, when we control for other factor affecting price transmission.
+The estimated effect is negative for the 2007-08 food crisis period but much smaller in magnitude and statistically insignificant.
+This could suggests that vulnerable countries intervened more heavily in their food markets during the first spike than in the second.
+LLDCs experienced significantly lower food inflation in the latter period (around 4 percentage points less on average) but not in the former.
+Again, the lack of significance could be a result of more intervention in the former period, which makes the relationship more noisy.
+Grain exporters experienced more food inflation on average in both periods but the effect is very small.
+Perhaps the most surprising finding is that grain importers experienced lower food inflation on average in both periods.
+As discussed above, this would suggest that large food importers intervene more in their food markets than other countries.
 
 \begin{center}
-[Table \ref{tab:2007-8-regressions} around here]
+[Table \ref{tab:rfpi-regressions} around here]
 \par\end{center}
 
+Table \ref{tab:rfpi-regressions} considers domestic real FPI growth rather than FPI inflation during during the 2007-08 and 2010-11 international price spikes.
+The main differences compared to table \ref{tab:fpi-regressions} is that countries with better infrastucture, i.e.
+a higher LPI index, are found to have higher real FPI growth rates but only statistically significantly so in the former period.
+The $R^2$ is quite high in all specifications, especially for the FPI inflation regressions in table \ref{tab:fpi-regressions}.\footnote{GDP and consumption expenditure contribute up to around half of the $R^2$ according to the LMG metric \citep{lindeman80}.
+These calculations were carried out in \textsf{R} using the \emph{relaimpo} package \citep{groemping06}.}
+
+In order to find out whether the negative coefficient on the Cereal import variable and the insignificance of the Landlocked dummy in 2007-08 is due to policy interventions during the food crisis, I included a set of intervention dummies to the 2007-08 regressions in table \ref{tab:fpi-regressions} and \ref{tab:rfpi-regressions}.
+Results are shown in table \ref{tab:int-regressions}.\footnote{These dummies are based on table 1 in \citet{Demeke2009}.
+``Export restrictions'' = 1 if the country Restricted or banned export, ``Price controls'' = 1 if the country administered price control or restricted private trade, ``Stock releases'' = 1 if the country released stocks (public or imported) at subsidized prices'', ``Tariff reductions'' = 1 if the country reduced tariffs or customs fees on imports and ``Tax reductions'' = 1 if the country suspended or reduced VAT or other taxes, all as of 1.
+December 2008.} As can be seen, this increases the coefficients on the LLDC dummy but not enough to achieve statistical significance at conventional levels.
+Furthermore, the coefficient on the Cereal import variable remains negative.
+Only the Tax reduction dummy is significant in more than one specification.
+The coefficient is positive though, suggesting that tax reductions did not lead to lower food prices increases during the food crisis, rather the opposite.
+
+\subsection{Determinants of price transmission \label{sec:analysis2}}
+Although the results in section \ref{sec:analysis1} provide insights into the cross sectional variation in domestic price changes during the two recent international food price spikes they have at least two shortcomings.
+First, none of the estimated relationships represent price transmission as such as there may be the other and perhaps more important causes of the observed domestic price changes.
+Secondly, the estimates are based on two specific time periods, which means that we ignore much the information contained in the individual price series and it also raises questions about the reliability of the results.
+To accommodate these shortcomings I approach the issue from two different angles.
+First, in line with the small existing literature on global food price transmission patterns, I estimate country specific price transmission coefficients based on a time series model.
+These estimated coefficients are then subsequently regressed on the same income and expenditure predictors as in section \ref{sec:analysis1}.
+Secondly, I estimate the effects of the predictor variables on price transmission directly in a panel model.
+
+\subsubsection{Meta analysis of estimated price transmission rates\label{sec:meta-analysis}}
+This section first presents a set of country specific price transmission estimates.
+Next a meta-regression analysis is carried out to see whether the estimates vary systematically with the proposed determinants of price transmission.
+
+Following the literature, I regress monthly domestic food price changes on its lagged domestic food price changes as well as on contemporaneous and lagged international food price changes and exchange rates \citep[see][]{Ianchovichina2014, IMF2011, Kalkuhl2014}.
+That is, for each country $i$ an autoregressive distributive lag (ADL) models was fitted
+\begin{equation}
+\pi_{i,t}^{dom}=\delta_{i,0}+\sum_{s=1}^{p}\rho_{i,s}\pi_{i,t-s}^{dom}+\sum_{s=0}^{q}\lambda_{i,s}\pi_{t-s}^{int}+\sum_{s=0}^{q}\phi_{i,s}er_{t-s}+\epsilon_{i,t}\label{eq:time_series_model}
+\end{equation}
+where $\pi_{i,t}^{dom}$ represents country $i$'s monthly FPI inflation or real FPI growth at time $t$ (both approximated by the first differences to the log of FPI or real FPI), $\pi_{t}^{int}$ is the FAO food price index at time $t$ and $er_{i,t}$ is the country $i$'s domestic currency/US\$ exchange rate, both in log differences.
+
+The ADL model is appropriate when the causal impact is unidirectional.
+This could be the case when the domestic price is from a small country relative to the world market but, more often, this assumption is invoked when the domestic price variable is a Food Price Index (FPI), i.e.
+the food component of the Consumer Price Index (CPI).
+
+Prior to estimation I tested each of the series for unit roots and cointegration with international prices.\footnote{I tested for stationarity with the KPSS test \citep{kwiatkowski92} and for unit roots with the augmented Dickey-Fuller test \citep{dickey79}.
+I used the Phillips and Ouliaris ``Pz'' test \citep{phillips90} to test for cointegration.
+These tests are implemented in \textsf{R} via the packages \emph{tseries} \citep{trapletti15} and \emph{urca} \citep{pfaff08}, respectively.} Results indicated that the domestic FPI and real FPI series, are nonstationary in levels but stationary in first differences.
+Domestic price levels were also found not to be cointegrated with world prices.
+Consequently, each model was estimated in growth rates rather than levels in order to avoid the problems associated with spurious regressions.
+
+Next, for each country, I calculated the long run multiplier (LRM) associated with a change in the the FAO food price index given by
+\begin{equation}
+\theta_{i}=\frac{\sum_{s=0}^{q}\lambda_{i,s}}{1-\sum_{s=1}^{p}\rho_{i,s}}.
+\end{equation}
+
 <<echo=FALSE>>=
-<<dom_infl_0708>>
-tp3 = -coef(m3)["log(food)"]/(2*coef(m3)["I(log(food)^2)"])
-tp6 = -coef(m6)["log(food)"]/(2*coef(m6)["I(log(food)^2)"])
-tp4 = -coef(m4)["log(GDP)"]/(2*coef(m4)["I(log(GDP)^2)"])
-tp5 = -coef(m5)["log(cons)"]/(2*coef(m5)["I(log(cons)^2)"])
+library(dplyr)
+load("lrm_fpi_l1_12.Rda")
+fpi = select(lrm.fpi.l.1_12, -iso3c)
+load("lrm_rfpi_l1_12.Rda")
+rfpi = select(lrm.rfpi.l.1_12, -iso3c)
+fivenum.fpi = data.frame(sapply(fpi, summary))
+fivenum.rfpi = data.frame(sapply(rfpi, summary))
+n = nrow(fpi)
 @
 
-Table \ref{tab:2007-8-regressions} considers domestic food price changes during the 2007:1-2008:6 period for the countries in the sample for which there is data on income, expenditure, dietary energy supply and cereal import dependency.
-Column (1) relates a country's annualized FPI inflation rate in that period to its per capita GDP level.
-What we see is that average food inflation during the 2007-8 food crisis is increasing in a country's per capita income level up until a certain point after which it starts to decrease.
-The same applies for consumption and food expenditure in column (2) and (3), respectively.
-Only the latter set of coefficients are statistically significant though.
-Specifically, column (3) suggests that among countries with a 2005 per capita food expenditure level of \Sexpr{round(exp(tp3),1)} {[}$\exp(\Sexpr{round(tp3,2)})${]} PPP adjusted dollars or less there is a positive estimated relationship between expected FPI inflation and food expenditure.
-After that point the predicted marginal effect turns negative.
-\Sexpr{nrow(data_0708[which(data_0708$food< exp(tp3) & data_0708$food >0), ])} out of the \Sexpr{nrow(data_0708[which(data_0708$food>0), ])} countries in the sample with data on per capita expenditure levels spend less than that on food.
+The LRM is a common measure of price transmission.
+\cite{IMF2011}, for example, contains an analysis of food price price transmission based on a dataset of monthly FPI inflation from 31 advanced economies and 47 emerging and developing economies over the period 2000 - 2011.
+Their median estimated LRM pass-through of a 1 percent increase in international food inflation is 0.18 percent in developed countries and 0.34 percent in emerging and developing countries.
+Similarly, \citet{Ianchovichina2014} report an average estimated LRM for the eighteen Middle East and North Africa (MENA) countries of 0.25 and the majority of the countries are reported to have  food price pass-through LRMs in the order of 0.2-0.4.
+Both studies use 12 lags in their country specific ADL models, i.e.
+they assume $p=q=12$.\footnote{\cite{IMF2011} do not include exchange rates in their regressions.} My own estimates based on a sample of \Sexpr{n} countries covering the period 2005:1 to 2014:12 are very similar.
+The average and median FPI transmission LRM from a model with 12 lags of all variables are \Sexpr{round(fivenum.fpi["Mean", "V12"], 2)} and \Sexpr{round(fivenum.fpi["Median", "V12"], 2)}, respectively.
+That being said, it is not obvious that so many lags should be included in the models.
+When lag length selection is based on the Akaike Information Criterion (AIC), for example, the optimal number lags in most of the models turn out to be $p=q=1$.\footnote{For each country, I calculated the AIC for each of the $12^2$ combinations of $(p,q)\, \in (1,..,12)$.
+Then I selected the $(p,q)$ combination with the lowest AIC.} But when a common lag length of 1 is used for all countries, average and median FPI transmission LRMs are much lower, namely \Sexpr{round(fivenum.fpi["Mean", "V1"], 2)} and \Sexpr{round(fivenum.fpi["Median", "V1"], 2)}, respectively.\footnote{I also tested for residual autocorrelation with the Durbin-Watson test.
+There was no evidence of residual autocorrelation in either case.}
 
-Column (6) focuses on the relationship between average food expenditure and expected real FPI growth in the 2007-8 period.
-Again there is a significant quadratic relationship between the two and it is remarkably similar to the specification based on FPI inflation in column (3).
-For countries with per capita food expenditures less than \Sexpr{round(exp(tp6),1)} PPP adjusted dollars (\Sexpr{nrow(data_0708[which(data_0708$food< exp(tp6) & data_0708$food >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$food>0), ])} countries) there is a positive estimated relationship between food expenditure and real FPI growth.
-For countries with food expenditures higher than that, the relationship becomes negative.
+Finally, FPI and Real FPI transmission estimates, $\hat{\theta}$, are regressed, on the same explanatory variables as in table \ref{tab:fpi-regressions} and \ref{tab:rfpi-regressions}.
+That is, the following model is fitted
+\begin{equation}
+\hat{\theta}_{i} =\delta_{0} + \delta_{1}x_{i} + u_{i}\label{eq:ptreg},
+\end{equation}
+where the covariates $x$ are the same as in equation (\ref{eq:infl_model}).
+The dependent variable in Equation (\ref{eq:ptreg}) is an estimate rather than an observed value and the the equation therefore represents an estimated dependent variable (EDV) regression \citep[see][]{hanushek74, lewis05}.
+A common approach in this situation is to use Weighted Least Squares (WLS) estimation with weights $w_{i}$ that are inversely proportional to the standard error of $\theta_{i}$ \citep{saxonhouse76}.
+This way more precise estimates are given a higher weight in the regression.
+The main reason though for using WLS rather than OLS is that the latter may suffer from heteroskedasticity which leads to inefficient coefficient estimates inconsistent standard errors.
+The inconsistent OLS standard errors can, however, easily be corrected with a robust standard error estimator.
+\citet{lewis05} uses formal reasoning backed up by a Monte Carlo analysis to argue that OLS with robust standard errors is likely to produce better estimates of parameter uncertainty than WLS based on inverse standard error weights, but the latter is likely to produce more efficient parameter estimates.
+Furthermore, they conclude that OLS with robust standard errors is the best approach unless information about the sampling error of the dependent variable is not only available but also highly reliable.
+
+Table \ref{tab:fpi-pt-regressions} presents the estimates from equation \ref{eq:ptreg} based on OLS with robust standard errors as well as WLS.
+I tested for heteroskedasticity with the Breusch-Pagan test \citep{breusch79} but there was no evidence against the null of homoskedasticity.
+Consequently, ordinary and robust standard errors were very similar.
+The results also turned out to be robust to the choice of lag lengths in connection with the LRM and therefore, only the results based on the single lag LRM are reported.
+
+\begin{center}
+[Table \ref{tab:fpi-pt-regressions} around here]
+\par\end{center}
+
+As can be seen, results are fairly robust to the choice of estimator although the significance levels are somewhat higher in the OLS case.
+As in the domestic food inflation case (table \ref{tab:fpi-regressions} -\ref{tab:int-regressions}) we find that cereal exporters respond stronger to international food price changes whereas cereal importers and cereal consumers respond less to internstional price shocks.
+These effects are, however, very small and/or statistically insignificant.
+LLDCs, curiously, seem to respond stronger on average to international price shocks than other countries.
+GDP and consumption expenditure are also apparently important determinants of food price transmission.
+
+Finally, table \ref{tab:rfpi-pt-regressions} repeats the exercise for the LRM based on the real FPI data. Compared to table \ref{tab:fpi-pt-regressions}, not much changes.
+
+\begin{center}
+[Table \ref{tab:rfpi-pt-regressions} around here]
+\par\end{center}
+
+\subsubsection{A panel model of price transmission determinants}
+The findings discussed above suggest that middle income countries are the most vulnerable to international food price spikes even after controlling for other factors which plausibly affect price transmission.
+However, the estimates in table \ref{tab:fpi-pt-regressions} and \ref{tab:rfpi-pt-regressions} are potentially subject to unobserved heterogeneity bias.
+As a last robustness check I therefore exploit the panel structure of the dataset to control for time invariant unobserved heterogeneity which potentially confounds the relationship between price transmission and its determinants.
+Specifically, the following model is fitted
+\begin{equation}
+\pi_{it}^{dom}=\delta_{0}+
+\sum_{s=1}^{6}\rho_{s}\pi_{it-s}^{dom}+
+\sum_{s=0}^{1}(\lambda_{s}\pi_{t-s}^{int}+\phi_{s}er_{it-s})+
+\delta_{1}\pi_{t}^{int}\cdot x_{i} +
+u_{it}\label{eq:panel_model},
+\end{equation}
+where all variables are the same as before.
+I tried to include more lags of $\pi_{t}^{int}$ and $er_{it}$ but their coefficients were small and generally insignificant.
+The main effect of $\pi^{int}$ on $\pi_{i}^{dom}$ is on impact, i.e.
+at time $t=0$.
+This is the reason why the interactions terms $\pi_{t}^{int}\cdot x_{i} $ also refer to the impact effect.
+
+It is well known that witin estimation of dynamic panel models such as equation (\ref{eq:panel_model}) suffers from Nickell bias.
+The problem is that the demeaned lagged dependent variable is correlated with the demeaned error term which makes the estimator inconsistent \citep{nickell81}.
+This inconsistency, however, is mainly a problem in short panel datasets where the number of cross section obeservations $N$  is much larger than the number of time periods $T$.
+Specifically, \citet{nickell81} demonstrates that the bias is of the order $1/T$ as $N\rightarrow \infty$.
+In the present case where $T>100$ the bias should be negligible.
+Nontheless, as a last robustness check, I also consider the  Anderson-Hsiao (AH) estimator \citep{anderson82} based on the first-difference transformation of the model in log levels with lags 7-12 of the dependent variable acting as instruments.
+
+Estimation results are shown in table \ref{tab:fe-panel-regressions}-\ref{tab:fd-panel-regressions}.\footnote{Note that coefficients on the lagged dependent variable are omitted.} In table \ref{tab:fe-panel-regressions}, the only explanatory varibles that remain significant are those related to income, namely the GDP and Consumption expenditure terms.
+The effect sizes in table \ref{tab:fd-panel-regressions} are similar to those in table \ref{tab:fd-panel-regressions} but the standard errors are generally larger so the significance mostly disappears.
+Only the Cereal export share becomes significant in the FPI regressions.
 
 <<echo=FALSE>>=
-<<data_0708>>
-gdpsort = data_0708[order(data_0708$GDP), ]
+load("ctr_dat_0708.Rda")
+dat = ctr.dat.0708
+r1 = readRDS("fpi_reg_m.rds")
+r2 = readRDS("rfpi_reg_m.rds")
+r3 = readRDS("int_reg_m.rds")
+r4 = readRDS("fpi_pt_reg_m.rds")
+r5 = readRDS("rfpi_pt_reg_m.rds")
+r6 = readRDS("panel_reg_m.rds")
+r7 = readRDS("panel_reg_fd_m.rds")
+
+tp = matrix(NA, nrow=4, ncol=7)
+
+tp[1,1] = -coef(r1[[1]])["log(gdp)"]/(2*coef(r1[[1]])["I(log(gdp)^2)"])
+tp[2,1] = -coef(r1[[2]])["log(cons)"]/(2*coef(r1[[2]])["I(log(cons)^2)"])
+tp[3,1] = -coef(r1[[3]])["log(gdp)"]/(2*coef(r1[[3]])["I(log(gdp)^2)"])
+tp[4,1] = -coef(r1[[4]])["log(cons)"]/(2*coef(r1[[4]])["I(log(cons)^2)"])
+
+tp[1,2] = -coef(r2[[1]])["log(gdp)"]/(2*coef(r2[[1]])["I(log(gdp)^2)"])
+tp[2,2] = -coef(r2[[2]])["log(cons)"]/(2*coef(r2[[2]])["I(log(cons)^2)"])
+tp[3,2] = -coef(r2[[3]])["log(gdp)"]/(2*coef(r2[[3]])["I(log(gdp)^2)"])
+tp[4,2] = -coef(r2[[4]])["log(cons)"]/(2*coef(r2[[4]])["I(log(cons)^2)"])
+
+tp[1,3] = -coef(r3[[1]])["log(gdp)"]/(2*coef(r3[[1]])["I(log(gdp)^2)"])
+tp[2,3] = -coef(r3[[2]])["log(cons)"]/(2*coef(r3[[2]])["I(log(cons)^2)"])
+tp[3,3] = -coef(r3[[3]])["log(gdp)"]/(2*coef(r3[[3]])["I(log(gdp)^2)"])
+tp[4,3] = -coef(r3[[4]])["log(cons)"]/(2*coef(r3[[4]])["I(log(cons)^2)"])
+
+tp[1,4] = -coef(r4[[1]])["log(gdp)"]/(2*coef(r4[[1]])["I(log(gdp)^2)"])
+tp[2,4] = -coef(r4[[2]])["log(cons)"]/(2*coef(r4[[2]])["I(log(cons)^2)"])
+tp[3,4] = -coef(r4[[3]])["log(gdp)"]/(2*coef(r4[[3]])["I(log(gdp)^2)"])
+tp[4,4] = -coef(r4[[4]])["log(cons)"]/(2*coef(r4[[4]])["I(log(cons)^2)"])
+
+tp[1,5] = -coef(r5[[1]])["log(gdp)"]/(2*coef(r5[[1]])["I(log(gdp)^2)"])
+tp[2,5] = -coef(r5[[2]])["log(cons)"]/(2*coef(r5[[2]])["I(log(cons)^2)"])
+tp[3,5] = -coef(r5[[3]])["log(gdp)"]/(2*coef(r5[[3]])["I(log(gdp)^2)"])
+tp[4,5] = -coef(r5[[4]])["log(cons)"]/(2*coef(r5[[4]])["I(log(cons)^2)"])
+
+tp[1,6] = -coef(r6[[1]])["I(lag(diff(log(fao.fpi)), 0) * log(gdp))"]/
+  (2*coef(r6[[1]])["I(lag(diff(log(fao.fpi)), 0) * I(log(gdp)^2))"])
+tp[2,6] = -coef(r6[[2]])["I(lag(diff(log(fao.fpi)), 0) * log(cons))"]/
+  (2*coef(r6[[2]])["I(lag(diff(log(fao.fpi)), 0) * I(log(cons)^2))"])
+tp[3,6] = -coef(r6[[3]])["I(lag(diff(log(fao.fpi)), 0) * log(gdp))"]/
+  (2*coef(r6[[3]])["I(lag(diff(log(fao.fpi)), 0) * I(log(gdp)^2))"])
+tp[4,6] = -coef(r6[[4]])["I(lag(diff(log(fao.fpi)), 0) * log(cons))"]/
+  (2*coef(r6[[4]])["I(lag(diff(log(fao.fpi)), 0) * I(log(cons)^2))"])
+
+tp[1,7] = -coef(r7[[1]])["I(lag(log(fao.fpi), 0) * log(gdp))"]/
+  (2*coef(r7[[1]])["I(lag(log(fao.fpi), 0) * I(log(gdp)^2))"])
+tp[2,7] = -coef(r7[[2]])["I(lag(log(fao.fpi), 0) * log(cons))"]/
+  (2*coef(r7[[2]])["I(lag(log(fao.fpi), 0) * I(log(cons)^2))"])
+tp[3,7] = -coef(r7[[3]])["I(lag(log(fao.fpi), 0) * log(gdp))"]/
+  (2*coef(r7[[3]])["I(lag(log(fao.fpi), 0) * I(log(gdp)^2))"])
+tp[4,7] = -coef(r7[[4]])["I(lag(log(fao.fpi), 0) * log(cons))"]/
+  (2*coef(r7[[4]])["I(lag(log(fao.fpi), 0) * I(log(cons)^2))"])
+
+tp = exp(tp)
+
+nc = matrix(NA, nrow=4, ncol=7)
+nc.pc = matrix(NA, nrow=4, ncol=7)
+n.gdp = nrow(dat[which(dat$gdp>0), ])
+n.cons = nrow(dat[which(dat$cons>0), ])
+
+nc[1,1] = nrow(dat[which(dat$gdp< tp[1,1] & dat$gdp >0), ])
+nc[2,1] = nrow(dat[which(dat$cons< tp[2,1] & dat$cons >0), ])
+nc[3,1] = nrow(dat[which(dat$gdp< tp[3,1] & dat$gdp >0), ])
+nc[4,1] = nrow(dat[which(dat$cons< tp[4,1] & dat$cons >0), ])
+
+nc.pc[1,1] = round(nc[1,1]/n.gdp*100,0)
+nc.pc[2,1] = round(nc[2,1]/n.cons*100,0)
+nc.pc[3,1] = round(nc[3,1]/n.gdp*100,0)
+nc.pc[4,1] = round(nc[4,1]/n.cons*100,0)
+
+nc[1,2] = nrow(dat[which(dat$gdp< tp[1,2] & dat$gdp >0), ])
+nc[2,2] = nrow(dat[which(dat$cons< tp[2,2] & dat$cons >0), ])
+nc[3,2] = nrow(dat[which(dat$gdp< tp[3,2] & dat$gdp >0), ])
+nc[4,2] = nrow(dat[which(dat$cons< tp[4,2] & dat$cons >0), ])
+
+nc.pc[1,2] = round(nc[1,2]/n.gdp*100,0)
+nc.pc[2,2] = round(nc[2,2]/n.cons*100,0)
+nc.pc[3,2] = round(nc[3,2]/n.gdp*100,0)
+nc.pc[4,2] = round(nc[4,2]/n.cons*100,0)
+
+nc[1,3] = nrow(dat[which(dat$gdp< tp[1,3] & dat$gdp >0), ])
+nc[2,3] = nrow(dat[which(dat$cons< tp[2,3] & dat$cons >0), ])
+nc[3,3] = nrow(dat[which(dat$gdp< tp[3,3] & dat$gdp >0), ])
+nc[4,3] = nrow(dat[which(dat$cons< tp[4,3] & dat$cons >0), ])
+
+nc.pc[1,3] = round(nc[1,3]/n.gdp*100,0)
+nc.pc[2,3] = round(nc[2,3]/n.cons*100,0)
+nc.pc[3,3] = round(nc[3,3]/n.gdp*100,0)
+nc.pc[4,3] = round(nc[4,3]/n.cons*100,0)
+
+nc[1,4] = nrow(dat[which(dat$gdp< tp[1,4] & dat$gdp >0), ])
+nc[2,4] = nrow(dat[which(dat$cons< tp[2,4] & dat$cons >0), ])
+nc[3,4] = nrow(dat[which(dat$gdp< tp[3,4] & dat$gdp >0), ])
+nc[4,4] = nrow(dat[which(dat$cons< tp[4,4] & dat$cons >0), ])
+
+nc.pc[1,4] = round(nc[1,4]/n.gdp*100,0)
+nc.pc[2,4] = round(nc[2,4]/n.cons*100,0)
+nc.pc[3,4] = round(nc[3,4]/n.gdp*100,0)
+nc.pc[4,4] = round(nc[4,4]/n.cons*100,0)
+
+nc[1,5] = nrow(dat[which(dat$gdp< tp[1,5] & dat$gdp >0), ])
+nc[2,5] = nrow(dat[which(dat$cons< tp[2,5] & dat$cons >0), ])
+nc[3,5] = nrow(dat[which(dat$gdp< tp[3,5] & dat$gdp >0), ])
+nc[4,5] = nrow(dat[which(dat$cons< tp[4,5] & dat$cons >0), ])
+
+nc.pc[1,5] = round(nc[1,5]/n.gdp*100,0)
+nc.pc[2,5] = round(nc[2,5]/n.cons*100,0)
+nc.pc[3,5] = round(nc[3,5]/n.gdp*100,0)
+nc.pc[4,5] = round(nc[4,5]/n.cons*100,0)
+
+nc[1,6] = nrow(dat[which(dat$gdp< tp[1,6] & dat$gdp >0), ])
+nc[2,6] = nrow(dat[which(dat$cons< tp[2,6] & dat$cons >0), ])
+nc[3,6] = nrow(dat[which(dat$gdp< tp[3,6] & dat$gdp >0), ])
+nc[4,6] = nrow(dat[which(dat$cons< tp[4,6] & dat$cons >0), ])
+
+nc.pc[1,6] = round(nc[1,6]/n.gdp*100,0)
+nc.pc[2,6] = round(nc[2,6]/n.cons*100,0)
+nc.pc[3,6] = round(nc[3,6]/n.gdp*100,0)
+nc.pc[4,6] = round(nc[4,6]/n.cons*100,0)
+
+nc[1,7] = nrow(dat[which(dat$gdp< tp[1,7] & dat$gdp >0), ])
+nc[2,7] = nrow(dat[which(dat$cons< tp[2,7] & dat$cons >0), ])
+nc[3,7] = nrow(dat[which(dat$gdp< tp[3,7] & dat$gdp >0), ])
+nc[4,7] = nrow(dat[which(dat$cons< tp[4,7] & dat$cons >0), ])
+
+nc.pc[1,7] = round(nc[1,7]/n.gdp*100,0)
+nc.pc[2,7] = round(nc[2,7]/n.cons*100,0)
+nc.pc[3,7] = round(nc[3,7]/n.gdp*100,0)
+nc.pc[4,7] = round(nc[4,7]/n.cons*100,0)
+
+nc.pc.colmean = colMeans(nc.pc)
+nc.pc.mean = mean(nc.pc.colmean)
+nc.pc.panel.mean = mean(nc.pc.colmean[6:7])
+nc.pc.nonpanel.mean = mean(nc.pc.colmean[-(6:7)])
 @
 
-The GDP and consumption coefficients in column (4) and (5) are also statistically significant and the conclusion from these models is similar to one based on (3) and (6).
-The GDP model in column (4) predicts that for countries with a per capita GDP below \Sexpr{round(exp(tp4),1)} PPP adjusted dollars (\Sexpr{nrow(data_0708[which(data_0708$GDP< exp(tp4) & data_0708$GDP >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$GDP>0), ])} countries) there is a positive relationship between real FPI increases and income and vice versa.
-The consumption expenditure model (5) predicts a turning point of \Sexpr{round(exp(tp5),1)} PPP adjusted dollars where \Sexpr{nrow(data_0708[which(data_0708$cons< exp(tp5) & data_0708$cons >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$cons>0), ])} countries have lower average consumption than that.
-Number \Sexpr{nrow(data_0708[which(data_0708$GDP< exp(tp4) & data_0708$GDP >0), ])} country in the sample in terms of GDP is Nicaragua.
-Countries in the lower middle income group  therefore generally experienced the largest food increases during the 2007-8 food crisis, whereas low and high income countries were affected less, once we control for differences in consumption and trade patterns.
-
-Countries with higher cereal consumption shares and import dependency ratios experienced larger food price increases as expected although $cshare$ and $cdep$ are not individually significant in most specifications.
-The negative coefficient on the interaction term means that the effect of $cdep$ on food price growth is decreasing in $cshare$ and vice-versa, but this mitigating effect is negligible.
-
-\begin{center}
-[Table \ref{tab:2010-11-regressions} around here]
-\par\end{center}
-
-Table \ref{tab:2010-11-regressions} summarizes the results from a set of regressions based on the same explanatory variables as those from table \ref{tab:2007-8-regressions}.
-The difference is that the dependent variables are domestic FPI inflation and real FPI growth during the 2010:1-2010:12 period.
-As can be seen, the results are similar to those in table \ref{tab:2007-8-regressions}, cf.
-also table \ref{tab:tp} .
-
-\subsection{Price transmission\label{sec:analysis2}}
-Although the results in section \ref{sec:analysis1} provide insights into the cross sectional variation in domestic price changes during the two recent international food price spikes they have at least two shortcomings.
-First, none of the estimated relationships represent price transmission as such.
-Secondly, the estimates are based on two specific time periods, which means that we ignore much the information contained in the individual price series.
-To accommodate these two shortcomings I estimate country specific price transmission coefficients and regress these on the same income and expenditure predictors as in section \ref{sec:analysis1}.
-
-<<echo=FALSE>>=
-<<ts_pt>>
-tp1 = -coef(m1)["log(GDP)"]/(2*coef(m1)["I(log(GDP)^2)"])
-tp2 = -coef(m2)["log(cons)"]/(2*coef(m2)["I(log(cons)^2)"])
-tp3 = -coef(m3)["log(food)"]/(2*coef(m3)["I(log(food)^2)"])
-tp4 = -coef(m4)["log(GDP)"]/(2*coef(m4)["I(log(GDP)^2)"])
-tp5 = -coef(m5)["log(cons)"]/(2*coef(m5)["I(log(cons)^2)"])
-tp6 = -coef(m6)["log(food)"]/(2*coef(m6)["I(log(food)^2)"])
-@
-
-Specifically, for each country, I regress its monthly domestic food price changes on its lagged domestic food price changes as well as on contemporaneous and lagged international food price changes.
-That is, as in the the \cite{IMF2011} study, I estimate country specific autoregressive distributive lag (ADL) models of the type
-\begin{equation}
-\pi_{t}^{dom}=\delta_{0}+\sum_{s=1}^{S_{y}}\rho_{s}\pi_{t-s}^{dom}+\sum_{s=0}^{S_{f}}\lambda_{s}\pi_{t-s}^{int}+\epsilon_{t}\label{eq:time_series_model}
-\end{equation}
-where $\pi_{t}^{dom}=\Delta p_{t}$ represents monthly FPI inflation or real FPI growth at time $t$ (both approximated by the first differences to the log of FPI or real FPI) and $\pi_{t}^{int}$ is the FAO food price index at time $t$, also in log differences.
-Using first differences instead of price levels is a way to avoid the spurious regression problem associated with trending variables.
-In order to keep things simple, maximum lag lengths are uniformly set to $S_{y}=S_{f}=6$.
-
-Next I calculate the long run effects of changes in the the FAO food price index on domestic food inflation and real FPI growth, the first of which is a standard measure of food price transmission.
-The long run multiplier is given by
-\begin{equation}
-\theta=\frac{\sum_{s=0}^{S_{f}}\lambda_{s}}{1-\sum_{s=1}^{S_{y}}\rho_{s}}.
-\end{equation}
-Finally, I regress the price transmission estimates, $\hat{\theta}$, on the same explanatory variables as in table \ref{tab:2007-8-regressions} and \ref{tab:2010-11-regressions}.
-That is, I estimate the model
-\begin{equation}
-\hat{\theta}_{i} =\delta_{0} + \delta_{1}x_{i} + \delta_{2}x_{i}^{2} + \delta_{3}z_{i} + u_{i}\label{eq:ptreg}.
-\end{equation}
-The results from this exercise are summarized in table \ref{tab:lri-regressions} and \ref{tab:tp}.
-As table \ref{tab:lri-regressions} shows, we find the same inverse-U shaped estimated relationships between income/expenditure and price transmission as we did with income/expenditure and domestic price changes in table \ref{tab:2007-8-regressions} and \ref{tab:2010-11-regressions}.
-For the specifications with significant income and expenditure coefficients, i.e.
-(1)-(5), the estimated turning points again imply that roughly one third of the countries are on the increasing part of the curve, cf.
-\ref{tab:tp}.
-
-\begin{center}
-[Table \ref{tab:lri-regressions} around here]
-\par\end{center}
-
-<<echo=False>>=
-<<panel_pt>>
-tp1 = -m1$coefficients[8]/(2*m1$coefficients[9])
-tp2 = -m2$coefficients[8]/(2*m2$coefficients[9])
-tp3 = -m3$coefficients[8]/(2*m3$coefficients[9])
-tp4 = -m4$coefficients[8]/(2*m4$coefficients[9])
-tp5 = -m5$coefficients[8]/(2*m5$coefficients[9])
-tp6 = -m6$coefficients[8]/(2*m6$coefficients[9])
-@
-
-The findings discussed above all point to the conclusion that middle income countries are the most vulnerable to international food price spikes.
-However, the estimates in table \ref{tab:2007-8-regressions}-\ref{tab:lri-regressions} are based on standard OLS and are potentially subject to unobserved heterogeneity bias.
-As a last robustness check I therefore exploit the panel structure of the dataset to control for time invariant unobserved heterogeneity which potentially confounds the income price transmission relationship.
-Results in table \ref{tab:panel-regressions} are based on within estimation of the model
-\begin{equation}
-\pi_{it}^{dom}=\delta_{0}+\sum_{s=1}^{6}\rho_{s}\pi_{it-s}^{dom}+\lambda_{1}\pi_{it-6}^{int}+\lambda_{2}\pi_{it-6}^{int}\cdot x_{i} +\lambda_{3}\pi_{it-6}^{int}\cdot x_{i}^{2}+u_{it}\label{eq:panel_model},
-\end{equation}
-where the variables, $\pi_{it}^{dom}$, $\pi_{it}^{int}$ and $x_{i}$ are the same as in \ref{eq:time_series_model} and \ref{eq:ptreg}.
-For simplicity, only a single lag of the FAO price index is included in the model and the lag length is chosen to be 6.
-The parameters of interest are $\lambda_{2}$ and $\lambda_{3}$ in front of the two interaction terms.
-A positive $\lambda_{2}$ and a negative $\lambda_{3}$ would, once more, imply an inverse-U shaped relationship between income/expenditure and price transmission as predicted by the simple model from section \ref{sec:model}.
-Again, in order to avoid the spurious regression problem associated with unit root processes the model is based on inflation rates rather than price levels.
-
-\begin{center}
-[Table \ref{tab:panel-regressions} around here]
-\par\end{center}
-
-Model (1),(3), (4) and (5) in table \ref{tab:panel-regressions} lead to significant quadratic expenditure price transmission relationships and the turning points based on these specifications are similar but somewhat higher than the ones from table \ref{tab:2007-8-regressions}-\ref{tab:lri-regressions}, with up to around 50 percent the countries on the increasing part of the income-price transmission curve, c.f table \ref{tab:tp}.
-
-\begin{center}
-[Table \ref{tab:tp} around here]
-\par\end{center}
+\subsection{Estimated income turning points}
+Table \ref{tab:tp} summarizes the estimated income turning points derived from the coefficients in table \ref{tab:fpi-regressions}-\ref{tab:fd-panel-regressions}.
+The figures in the last column refer to the number of countries with a per capita GDP or household consumption expenditure less than the estimated turning point shown in the third-last column.
+The average number of countries that are to the left of the turning point relative to the number of countries in the sample with data on these variables, across all specifications, is \Sexpr{round(nc.pc.mean,0)} percent.
+There are, however, important differences between the different estimators.
+Panel estimation, in particular, leads to a significantly higher proportion of countries to the left of the turning point than cross section estimation (\Sexpr{round(nc.pc.panel.mean,0)} vs.
+\Sexpr{round(nc.pc.nonpanel.mean,0)} percent) on average.

### 2016-01-29 tables.Rnw
Revised version. Note that the tp table created a failed message.

@@
-% !Rnw root = Master.Rnw
-
-\SweaveInput{regressions.Rnw}
+% !Rnw root = master1.Rnw
 
@@
 
-<<echo=FALSE, results=tex>>=
+\begin{table}[h]
+\caption{Time series data on prices and exchange rates}
+\label{tab:ts_data}
+\begin{center}
+{\small{}}%
+\begin{tabular}{|c|>{\raggedright}m{5cm}|>{\raggedright}m{3cm}|>{\raggedright}p{3cm}|}
+\hline 
+\textbf{\small{}Variable} & \textbf{\small{}Description} & \textbf{\small{}Source} & \textbf{\small{}Expected effect}\tabularnewline
+\hline 
+{\small{}FPI} & {\small{}Domestic food price index.} & {\small{}ILO, LABORSTA} & -\tabularnewline
+\hline
+{\small{}CPI} & {\small{}Domestic consumer price index.} & {\small{}ILO, LABORSTA} & -\tabularnewline
+\hline 
+{\small{}Real FPI} & {\small{}Domestic real food price index. FPI deflated by the CPI} & {\small{}ILO, LABORSTA} & -\tabularnewline
+\hline 
+{\small{}FAO FPI} & {\small{}FAO food price index.} & {\small{}FAO} & Positive\tabularnewline
+\hline 
+{\small{}Real FAO FPI} & {\small{}Real FAO food price index. Nominal index deflated by the
+World Bank Manufacturing Unit Value index} & {\small{}FAO} & Positive\tabularnewline
+\hline 
+{\small{}Exchange rate} & {\small{}Domestic currency/US\$ exchange rate} & {\small{}IMF} & Positive\tabularnewline
+\hline 
+\end{tabular}{\small \par}
+\end{center}
+\end {table}
+
+\begin{table}[h]
+\caption{Cross sectional data on country characteristics}
+\label{tab:cs_data}
+\begin{center}
+
+{\small{}}%
+\begin{tabular}{|c|>{\raggedright}m{5cm}|>{\raggedright}m{3cm}|>{\raggedright}p{3cm}|}
+\hline 
+\textbf{\small{}Variable} & \textbf{\small{}Description} & \textbf{\small{}Source} & \textbf{\small{}Expected effect}\tabularnewline
+\hline 
+{\small{}GDP} & {\small{}Gross domestic product. Log of per capita GDP (current int\$, PPP)} & {\small{}International Comparison Program 2005. World Bank database} & Inverse-U\tabularnewline
+\hline 
+{\small{}Consumption} & {\small{}Household consumption.
+Log of individual consumption expenditure by households (current int\$, PPP)} & {\small{}International Comparison Program 2005.
+World Bank database} & Inverse-U\tabularnewline
+\hline 
+{\small{}Cereal consumption share} & {\small{}Share of dietary energy supply derived from
+cereals, roots and tubers (\%). 3-year average, 2004-06} & {\small{}FAO, FAOSTAT} & Positive\tabularnewline
+\hline 
+{\small{}Cereal import share} & {\small{}Cereal imports as a share of domestic supply (\%). 3-year average, 2004-06} & {\small{}FAO, FAOSTAT} & Positive\tabularnewline
+\hline 
+{\small{}Cereal export share} & {\small{}Cereal exports as a share of domestic production (\%). 3-year average, 2004-06} & {\small{}FAO, FAOSTAT} & Positive\tabularnewline
+\hline 
+{\small{}LLDC} & {\small{}Dummy. 1 if country is a landlocked developing country} & {\small{}UNCTAD} & Negative\tabularnewline
+\hline 
+{\small{}Trade openness} & {\small{}Sum of exports and imports of goods and services as a share of GDP (\%). 2006-10 average} & {\small{}World Bank Development Indicators} & Positive\tabularnewline
+\hline 
+{\small{}DTF} & {\small{}Distance to Frontier index.
+An economy\textquoteright s distance to frontier measured on a scale from 0 to 100, where 0 represents the lowest performance and 100 the frontier.} & {\small{}Doing Business survey, Trading Across Borders .
+World Bank} & Positive\tabularnewline
+\hline 
+{\small{}LPI} & {\small{} Logistics performance index.
+The LPI measures ease of trade and transport of goods.
+The index ranges from 1 to 5, with a higher score representing better performance.} & {\small{}Logistics Performance Index 2007.
+World Bank database} & Positive\tabularnewline
+\hline 
+\end{tabular}{\small \par}
+\end{center}
+\end {table}
+
+<<results=tex,echo=F>>=
 library(stargazer)
-sumdat = merge(infl, icp_data[,-5], by = "iso3c", all.x=TRUE)
-sumdat = merge(sumdat, staples, by = "iso3c", all.x=TRUE)
-sumdat = merge(sumdat, cdep, by = "iso3c", all.x=TRUE)
-colnames(sumdat)=c("iso", "FPI inflation 07-08", "CPI inflation 07-08", "Real FPI growth 07-08", "FPI inflation 10-11", "CPI inflation 10-11", "Real FPI growth 10-11", "GDP", "cons", "food", "cshare", "cdep")
+reg.list = readRDS("fpi_reg_m.rds")
+se.list = readRDS("fpi_reg_se.rds")
 
-stargazer(sumdat, digits=1, title='Summary statistics', label = "tab:sumstat", 
-          notes=c("Notes: 07-08 and 10-11 refer to the periods 2007:1-2008:6 and 2010:1-2010:12,", 
-                  "respectively. `GDP, `cons' and `food' refer to per capita income, -consumption",
-                  "and -food expenditure respectively in the year 2005. `cshare' refers to a coun-",
-                  "try's percentage dietary energy supply derived from cereals, roots and tubers.", 
-                  "`cdep' refers to a country's cereal import dependency ratio, both with respect",
-                  "to the years 2004-06. Source: ILO, World Bank, FAO. Own calculations"),
-          notes.append=FALSE,
-          notes.align="l")
+stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
+          title='Determinants of domestic FPI inflation 2007-8 and 2010-11', 
+          label = "tab:fpi-regressions", 
+          dep.var.labels=c("FPI inflation 07-08", "FPI inflation 10-11"), 
+          no.space=TRUE,
+          se = list(se.list[[1]], se.list[[2]], se.list[[3]], se.list[[4]]), 
+          omit.stat = c("adj.rsq","f", "ser"), 
+          font.size = "footnotesize", 
+          covariate.labels=c("log of GDP",
+                             "log of GDP squared", 
+                             "log of consumption", 
+                             "log of consumption squared", 
+                             "Cereal consumption share", 
+                             "Cereal import share", 
+                             "Cereal export share",
+                             "LLDC",
+                             "State trading enterprice",
+                             "Trade openness", 
+                             "DTF", 
+                             "LPI"))
 @
@@
 <<results=tex,echo=F>>=
-<<dom_infl_0708>>
-stargazer(m1, m2, m3, m4, m5, m6, title='Income related determinants of domestic food price changes 2007:1-2008:6', label = "tab:2007-8-regressions", dep.var.labels=c("FPI inflation","Real FPI growth"), no.space=TRUE, se = list(rse1, rse2, rse3, rse4, rse5, rse6), omit.stat = c("adj.rsq","f", "ser"), font.size = "footnotesize", covariate.labels=c("$log(GDP)$", "$log(GDP)^2$", "$log(cons)$", "$log(cons)^2$", "$log(food)$", "$log(food)^2$", "$cshare$", "$cdep$", "$cshare \\times cdep$"),
-          notes=c("Robust standard errors are reported in parentheses underneath the coefficients.",
-                  "Significance at p = 0.1, p = 0.05 and p = 0.01 are denoted by `*', `**' and `***'", 
-                  "respectively"),
-          notes.append=FALSE,
-          notes.align="l",
-          notes.label="Notes:")
+reg.list = readRDS("rfpi_reg_m.rds")
+se.list = readRDS("rfpi_reg_se.rds")
+
+stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
+          title='Determinants of domestic real FPI growth 2007-8 and 2010-11', 
+          label = "tab:rfpi-regressions", 
+          dep.var.labels=c("Real FPI growth 07-08", "Real FPI growth 10-11"), 
+          no.space=TRUE,
+          se = list(se.list[[1]], se.list[[2]], se.list[[3]], se.list[[4]]), 
+          omit.stat = c("adj.rsq","f", "ser"), 
+          font.size = "footnotesize", 
+          covariate.labels=c("log of GDP",
+                             "log of GDP squared", 
+                             "log of consumption", 
+                             "log of consumption squared", 
+                             "Cereal consumption share", 
+                             "Cereal import share", 
+                             "Cereal export share",
+                             "LLDC",
+                             "State trading enterprice", 
+                             "Trade openness", 
+                             "DTF", 
+                             "LPI"))
 @
@@
 <<results=tex,echo=F>>=
-<<dom_infl_1011>>
-stargazer(m1, m2, m3, m4, m5, m6, title='Income related determinants of domestic food price changes 2010:1-2010:12', label = "tab:2010-11-regressions", dep.var.labels=c("FPI inflation","Real FPI growth"), no.space=TRUE, se = list(rse1, rse2, rse3, rse4, rse5, rse6), omit.stat = c("adj.rsq","f", "ser"), font.size = "footnotesize", covariate.labels=c("$log(GDP)$", "$log(GDP)^2$", "$log(cons)$", "$log(cons)^2$", "$log(food)$", "$log(food)^2$", "$cshare$", "$cdep$", "$cshare \\times cdep$"),
-          notes=c("Robust standard errors are reported in parentheses underneath the coefficients.",
-                  "Significance at p = 0.1, p = 0.05 and p = 0.01 are denoted by `*', `**' and `***'", 
-                  "respectively"),
-          notes.append=FALSE,
-          notes.align="l",
-          notes.label="Notes:")
+reg.list = readRDS("int_reg_m.rds")
+se.list = readRDS("int_reg_se.rds")
+
+stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
+          title='Determinants of domestic FPI inflation 2007-8', 
+          label = "tab:int-regressions", 
+          dep.var.labels=c("FPI inflation 07-08", "Real FPI growth 07-08"), 
+          no.space=TRUE,
+          se = list(se.list[[1]], se.list[[2]], se.list[[3]], se.list[[4]]), 
+          omit.stat = c("adj.rsq","f", "ser"), 
+          font.size = "footnotesize", 
+          covariate.labels=c("log of GDP",
+                             "log of GDP squared", 
+                             "log of consumption", 
+                             "log of consumption squared", 
+                             "Cereal consumption share", 
+                             "Cereal import share", 
+                             "Cereal export share",
+                             "LLDC", 
+                             "State trading enterprice",
+                             "Trade openness", 
+                             "DTF", 
+                             "LPI",
+                             "Export restrictions",
+                             "Price controls",
+                             "Stock releases",
+                             "Tariff reductions",
+                             "Tax reductions"))
+@
+
+<<results=tex,echo=F>>=
+library(stargazer)
+reg.list = readRDS("fpi_pt_reg_m.rds")
+se.list = readRDS("fpi_pt_reg_se.rds")
+
+stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
+          title='Determinants of long run FPI transmission rates', 
+          label = "tab:fpi-pt-regressions", 
+          dep.var.labels=c("FPI transmission LRM"), 
+          column.labels = c("OLS", "WLS"),
+          column.separate = c(2, 2),
+          no.space=TRUE,
+          se = list(se.list[[1]], se.list[[2]], se.list[[3]], se.list[[4]]), 
+          omit.stat = c("adj.rsq","f", "ser"), 
+          font.size = "footnotesize", 
+          covariate.labels=c("log of GDP",
+                             "log of GDP squared", 
+                             "log of consumption", 
+                             "log of consumption squared",
+                             "Cereal consumption share", 
+                             "Cereal import share", 
+                             "Cereal export share",
+                             "LLDC", 
+                             "Trade openness", 
+                             "DTF", 
+                             "LPI"))
+@
+
+<<results=tex,echo=F>>=
+reg.list = readRDS("rfpi_pt_reg_m.rds")
+se.list = readRDS("rfpi_pt_reg_se.rds")
+
+stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
+          title='Determinants of long run real FPI transmission rates', 
+          label = "tab:rfpi-pt-regressions", 
+          dep.var.labels=c("Real FPI transmission LRM"), 
+          column.labels = c("OLS", "WLS"),
+          column.separate = c(2, 2),
+          no.space=TRUE,
+          se = list(se.list[[1]], se.list[[2]], se.list[[3]], se.list[[4]]), 
+          omit.stat = c("adj.rsq","f", "ser"), 
+          font.size = "footnotesize", 
+          covariate.labels=c("log of GDP",
+                             "log of GDP squared", 
+                             "log of consumption", 
+                             "log of consumption squared",
+                             "Cereal consumption share", 
+                             "Cereal import share", 
+                             "Cereal export share",
+                             "LLDC", 
+                             "Trade openness", 
+                             "DTF", 
+                             "LPI"))
 @
@@
 <<results=tex, echo=FALSE>>=
-<<ts_pt>>
-stargazer(m1, m2, m3, m4, m5, m6, title='Long run multiplier (LRM) regressions', label = "tab:lri-regressions", dep.var.labels=c("LRM, FPI inflation","LRM, real FPI growth"), no.space=TRUE, se = list(rse1, rse2, rse3, rse4, rse5, rse6), omit.stat = c("adj.rsq","f", "ser"), font.size = "footnotesize", covariate.labels=c("$log(GDP)$", "$log(GDP)^2$", "$log(cons)$", "$log(cons)^2$", "$log(food)$", "$log(food)^2$", "$cshare$", "$cdep$", "$cshare \\times cdep$"),
-          notes=c("Robust standard errors are reported in parentheses underneath the coefficients.",
-                  "Significance at p = 0.1, p = 0.05 and p = 0.01 are denoted by `*', `**' and `***'", 
-                  "respectively"),
-          notes.append=FALSE,
-          notes.align="l",
-          notes.label="Notes:")
+library(plm)
+reg.list = readRDS("panel_reg_m.rds")
+
+stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
+          title='Dynamic panel regressions (within (FE) estimator)', 
+          label = "tab:fe-panel-regressions", 
+          dep.var.labels=c("FPI inflation", "Real FPI growth"), 
+          no.space=TRUE,
+          omit = c("adj.fpi", "rfpi"),
+          omit.stat = c("adj.rsq","f", "ser"), 
+          font.size = "footnotesize",
+          covariate.labels=c("$\\pi_{t}^{i}$",
+                             "$\\pi_{t-1}^{i}$",
+                             "$\\Delta \\log(er_{t})$",
+                             "$\\Delta \\log(er_{t-1})$",
+                             "$\\pi_{t}^{i}\\times \\text{log of GDP}$",
+                             "$\\pi_{t}^{i}\\times \\text{log of GDP squared}$", 
+                             "$\\pi_{t}^{i}\\times \\text{log of consumption}$", 
+                             "$\\pi_{t}^{i}\\times \\text{log of consumption squared}$", 
+                             "$\\pi_{t}^{i}\\times \\text{Cereal consumption share}$", 
+                             "$\\pi_{t}^{i}\\times \\text{Cereal import share}$", 
+                             "$\\pi_{t}^{i}\\times \\text{Cereal export share}$",
+                             "$\\pi_{t}^{i}\\times \\text{LLDC}$", 
+                             "$\\pi_{t}^{i}\\times \\text{Trade openness}$", 
+                             "$\\pi_{t}^{i}\\times \\text{DTF}$", 
+                             "$\\pi_{t}^{i}\\times \\text{LPI}$"))
 @
@@
 <<results=tex, echo=FALSE>>=
-<<panel_pt>>
-stargazer(m1, m2, m3, m4, m5, m6, 
-          title='Panel regressions', 
-          label = "tab:panel-regressions", 
+library(plm)
+reg.list = readRDS("panel_reg_fd_m.rds")
+
+stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
+          title='Dynamic panel regressions (Anderson-Hsiao (AH) estimator)', 
+          label = "tab:fd-panel-regressions", 
+          dep.var.labels=c("FPI inflation", "Real FPI growth"), 
+          no.space=TRUE,
+          omit = c("adj.fpi", "rfpi"),
+          omit.stat = c("adj.rsq","f", "ser"), 
           font.size = "footnotesize",
-          no.space=TRUE, 
-          se = list(rse1, rse2, rse3, rse4, rse5, rse6), 
-          omit.stat = c("adj.rsq","f"), 
-          dep.var.labels=c("$\\Delta log(FPI_{t})$","$\\Delta log(rFPI_{t})$"), 
-          omit = 1:12, 
-          covariate.labels=c("$\\Delta log(FAO_{t-6})$", "$\\Delta log(FAO_{t-6}) \\times log(GDP)$","$\\Delta log(FAO_{t-6}) \\times log(GDP)^2$", "$\\Delta log(FAO_{t-6}) \\times log(cons)$", "$\\Delta log(FAO_{t-6}) \\times log(cons)^2$", "$\\Delta log(FAO_{t-6}) \\times log(food)$", "$\\Delta log(FAO_{t-6}) \\times log(food)^2$"),
-          notes=c("Robust standard errors are reported in parentheses underneath the coef-",
-                  "ficients. Significance at p = 0.1, p = 0.05 and p = 0.01 are denoted by `*',", 
-                  " `**' and `***' respectively"),
-          notes.append=FALSE,
-          notes.align="l",
-          notes.label="Notes:")
+          covariate.labels=c("$\\pi_{t}^{i}$",
+                             "$\\pi_{t-1}^{i}$",
+                             "$\\Delta \\log(er_{t})$",
+                             "$\\Delta \\log(er_{t-1})$",
+                             "$\\pi_{t}^{i}\\times \\text{log of GDP}$",
+                             "$\\pi_{t}^{i}\\times \\text{log of GDP squared}$", 
+                             "$\\pi_{t}^{i}\\times \\text{log of consumption}$", 
+                             "$\\pi_{t}^{i}\\times \\text{log of consumption squared}$", 
+                             "$\\pi_{t}^{i}\\times \\text{Cereal consumption share}$", 
+                             "$\\pi_{t}^{i}\\times \\text{Cereal import share}$", 
+                             "$\\pi_{t}^{i}\\times \\text{Cereal export share}$",
+                             "$\\pi_{t}^{i}\\times \\text{LLDC}$", 
+                             "$\\pi_{t}^{i}\\times \\text{Trade openness}$", 
+                             "$\\pi_{t}^{i}\\times \\text{DTF}$", 
+                             "$\\pi_{t}^{i}\\times \\text{LPI}$"))
 @
 
-<<tpmat, echo=FALSE>>=
-tp = matrix(NA, nrow=6, ncol=4)
+<<echo=FALSE>>=
 
-<<dom_infl_0708>>
-tp[1,1] = -m1$coefficients[2]/(2*m1$coefficients[3])
-tp[2,1] = -m2$coefficients[2]/(2*m2$coefficients[3])
-tp[3,1] = -m3$coefficients[2]/(2*m3$coefficients[3])
-tp[4,1] = -m4$coefficients[2]/(2*m4$coefficients[3])
-tp[5,1] = -m5$coefficients[2]/(2*m5$coefficients[3])
-tp[6,1] = -m6$coefficients[2]/(2*m6$coefficients[3])
+## Remember to change variable names!!!!!
 
-<<dom_infl_1011>>
-tp[1,2] = -m1$coefficients[2]/(2*m1$coefficients[3])
-tp[2,2] = -m2$coefficients[2]/(2*m2$coefficients[3])
-tp[3,2] = -m3$coefficients[2]/(2*m3$coefficients[3])
-tp[4,2] = -m4$coefficients[2]/(2*m4$coefficients[3])
-tp[5,2] = -m5$coefficients[2]/(2*m5$coefficients[3])
-tp[6,2] = -m6$coefficients[2]/(2*m6$coefficients[3])
+load("ctr_dat_0708.Rda")
+dat = ctr.dat.0708
+r1 = readRDS("fpi_reg_m.rds")
+r2 = readRDS("rfpi_reg_m.rds")
+r3 = readRDS("int_reg_m.rds")
+r4 = readRDS("fpi_pt_reg_m.rds")
+r5 = readRDS("rfpi_pt_reg_m.rds")
+r6 = readRDS("panel_reg_m.rds")
+r7 = readRDS("panel_reg_fd_m.rds")
 
-<<ts_pt>>
-tp[1,3] = -m1$coefficients[2]/(2*m1$coefficients[3])
-tp[2,3] = -m2$coefficients[2]/(2*m2$coefficients[3])
-tp[3,3] = -m3$coefficients[2]/(2*m3$coefficients[3])
-tp[4,3] = -m4$coefficients[2]/(2*m4$coefficients[3])
-tp[5,3] = -m5$coefficients[2]/(2*m5$coefficients[3])
-tp[6,3] = -m6$coefficients[2]/(2*m6$coefficients[3])
+tp = matrix(NA, nrow=4, ncol=7)
 
-<<panel_pt>>
-tp[1,4] = -m1$coefficients[8]/(2*m1$coefficients[9])
-tp[2,4] = -m2$coefficients[8]/(2*m2$coefficients[9])
-tp[3,4] = -m3$coefficients[8]/(2*m3$coefficients[9])
-tp[4,4] = -m4$coefficients[8]/(2*m4$coefficients[9])
-tp[5,4] = -m5$coefficients[8]/(2*m5$coefficients[9])
-tp[6,4] = -m6$coefficients[8]/(2*m6$coefficients[9])
+tp[1,1] = -coef(r1[[1]])["log(gdp)"]/(2*coef(r1[[1]])["I(log(gdp)^2)"])
+tp[2,1] = -coef(r1[[2]])["log(cons)"]/(2*coef(r1[[2]])["I(log(cons)^2)"])
+tp[3,1] = -coef(r1[[3]])["log(gdp)"]/(2*coef(r1[[3]])["I(log(gdp)^2)"])
+tp[4,1] = -coef(r1[[4]])["log(cons)"]/(2*coef(r1[[4]])["I(log(cons)^2)"])
+
+tp[1,2] = -coef(r2[[1]])["log(gdp)"]/(2*coef(r2[[1]])["I(log(gdp)^2)"])
+tp[2,2] = -coef(r2[[2]])["log(cons)"]/(2*coef(r2[[2]])["I(log(cons)^2)"])
+tp[3,2] = -coef(r2[[3]])["log(gdp)"]/(2*coef(r2[[3]])["I(log(gdp)^2)"])
+tp[4,2] = -coef(r2[[4]])["log(cons)"]/(2*coef(r2[[4]])["I(log(cons)^2)"])
+
+tp[1,3] = -coef(r3[[1]])["log(gdp)"]/(2*coef(r3[[1]])["I(log(gdp)^2)"])
+tp[2,3] = -coef(r3[[2]])["log(cons)"]/(2*coef(r3[[2]])["I(log(cons)^2)"])
+tp[3,3] = -coef(r3[[3]])["log(gdp)"]/(2*coef(r3[[3]])["I(log(gdp)^2)"])
+tp[4,3] = -coef(r3[[4]])["log(cons)"]/(2*coef(r3[[4]])["I(log(cons)^2)"])
+
+tp[1,4] = -coef(r4[[1]])["log(gdp)"]/(2*coef(r4[[1]])["I(log(gdp)^2)"])
+tp[2,4] = -coef(r4[[2]])["log(cons)"]/(2*coef(r4[[2]])["I(log(cons)^2)"])
+tp[3,4] = -coef(r4[[3]])["log(gdp)"]/(2*coef(r4[[3]])["I(log(gdp)^2)"])
+tp[4,4] = -coef(r4[[4]])["log(cons)"]/(2*coef(r4[[4]])["I(log(cons)^2)"])
+
+tp[1,5] = -coef(r5[[1]])["log(gdp)"]/(2*coef(r5[[1]])["I(log(gdp)^2)"])
+tp[2,5] = -coef(r5[[2]])["log(cons)"]/(2*coef(r5[[2]])["I(log(cons)^2)"])
+tp[3,5] = -coef(r5[[3]])["log(gdp)"]/(2*coef(r5[[3]])["I(log(gdp)^2)"])
+tp[4,5] = -coef(r5[[4]])["log(cons)"]/(2*coef(r5[[4]])["I(log(cons)^2)"])
+
+tp[1,6] = -coef(r6[[1]])["I(lag(diff(log(fao.fpi)), 0) * log(gdp))"]/
+  (2*coef(r6[[1]])["I(lag(diff(log(fao.fpi)), 0) * I(log(gdp)^2))"])
+tp[2,6] = -coef(r6[[2]])["I(lag(diff(log(fao.fpi)), 0) * log(cons))"]/
+  (2*coef(r6[[2]])["I(lag(diff(log(fao.fpi)), 0) * I(log(cons)^2))"])
+tp[3,6] = -coef(r6[[3]])["I(lag(diff(log(fao.fpi)), 0) * log(gdp))"]/
+  (2*coef(r6[[3]])["I(lag(diff(log(fao.fpi)), 0) * I(log(gdp)^2))"])
+tp[4,6] = -coef(r6[[4]])["I(lag(diff(log(fao.fpi)), 0) * log(cons))"]/
+  (2*coef(r6[[4]])["I(lag(diff(log(fao.fpi)), 0) * I(log(cons)^2))"])
+
+tp[1,7] = -coef(r7[[1]])["I(lag(log(fao.fpi), 0) * log(gdp))"]/
+  (2*coef(r7[[1]])["I(lag(log(fao.fpi), 0) * I(log(gdp)^2))"])
+tp[2,7] = -coef(r7[[2]])["I(lag(log(fao.fpi), 0) * log(cons))"]/
+  (2*coef(r7[[2]])["I(lag(log(fao.fpi), 0) * I(log(cons)^2))"])
+tp[3,7] = -coef(r7[[3]])["I(lag(log(fao.fpi), 0) * log(gdp))"]/
+  (2*coef(r7[[3]])["I(lag(log(fao.fpi), 0) * I(log(gdp)^2))"])
+tp[4,7] = -coef(r7[[4]])["I(lag(log(fao.fpi), 0) * log(cons))"]/
+  (2*coef(r7[[4]])["I(lag(log(fao.fpi), 0) * I(log(cons)^2))"])
 
 tp = exp(tp)
+
+nc = matrix(NA, nrow=4, ncol=7)
+nc.pc = matrix(NA, nrow=4, ncol=7)
+n.gdp = nrow(dat[which(dat$gdp>0), ])
+n.cons = nrow(dat[which(dat$cons>0), ])
+
+nc[1,1] = nrow(dat[which(dat$gdp< tp[1,1] & dat$gdp >0), ])
+nc[2,1] = nrow(dat[which(dat$cons< tp[2,1] & dat$cons >0), ])
+nc[3,1] = nrow(dat[which(dat$gdp< tp[3,1] & dat$gdp >0), ])
+nc[4,1] = nrow(dat[which(dat$cons< tp[4,1] & dat$cons >0), ])
+
+nc.pc[1,1] = round(nc[1,1]/n.gdp*100,0)
+nc.pc[2,1] = round(nc[2,1]/n.cons*100,0)
+nc.pc[3,1] = round(nc[3,1]/n.gdp*100,0)
+nc.pc[4,1] = round(nc[4,1]/n.cons*100,0)
+
+nc[1,2] = nrow(dat[which(dat$gdp< tp[1,2] & dat$gdp >0), ])
+nc[2,2] = nrow(dat[which(dat$cons< tp[2,2] & dat$cons >0), ])
+nc[3,2] = nrow(dat[which(dat$gdp< tp[3,2] & dat$gdp >0), ])
+nc[4,2] = nrow(dat[which(dat$cons< tp[4,2] & dat$cons >0), ])
+
+nc.pc[1,2] = round(nc[1,2]/n.gdp*100,0)
+nc.pc[2,2] = round(nc[2,2]/n.cons*100,0)
+nc.pc[3,2] = round(nc[3,2]/n.gdp*100,0)
+nc.pc[4,2] = round(nc[4,2]/n.cons*100,0)
+
+nc[1,3] = nrow(dat[which(dat$gdp< tp[1,3] & dat$gdp >0), ])
+nc[2,3] = nrow(dat[which(dat$cons< tp[2,3] & dat$cons >0), ])
+nc[3,3] = nrow(dat[which(dat$gdp< tp[3,3] & dat$gdp >0), ])
+nc[4,3] = nrow(dat[which(dat$cons< tp[4,3] & dat$cons >0), ])
+
+nc.pc[1,3] = round(nc[1,3]/n.gdp*100,0)
+nc.pc[2,3] = round(nc[2,3]/n.cons*100,0)
+nc.pc[3,3] = round(nc[3,3]/n.gdp*100,0)
+nc.pc[4,3] = round(nc[4,3]/n.cons*100,0)
+
+nc[1,4] = nrow(dat[which(dat$gdp< tp[1,4] & dat$gdp >0), ])
+nc[2,4] = nrow(dat[which(dat$cons< tp[2,4] & dat$cons >0), ])
+nc[3,4] = nrow(dat[which(dat$gdp< tp[3,4] & dat$gdp >0), ])
+nc[4,4] = nrow(dat[which(dat$cons< tp[4,4] & dat$cons >0), ])
+
+nc.pc[1,4] = round(nc[1,4]/n.gdp*100,0)
+nc.pc[2,4] = round(nc[2,4]/n.cons*100,0)
+nc.pc[3,4] = round(nc[3,4]/n.gdp*100,0)
+nc.pc[4,4] = round(nc[4,4]/n.cons*100,0)
+
+nc[1,5] = nrow(dat[which(dat$gdp< tp[1,5] & dat$gdp >0), ])
+nc[2,5] = nrow(dat[which(dat$cons< tp[2,5] & dat$cons >0), ])
+nc[3,5] = nrow(dat[which(dat$gdp< tp[3,5] & dat$gdp >0), ])
+nc[4,5] = nrow(dat[which(dat$cons< tp[4,5] & dat$cons >0), ])
+
+nc.pc[1,5] = round(nc[1,5]/n.gdp*100,0)
+nc.pc[2,5] = round(nc[2,5]/n.cons*100,0)
+nc.pc[3,5] = round(nc[3,5]/n.gdp*100,0)
+nc.pc[4,5] = round(nc[4,5]/n.cons*100,0)
+
+nc[1,6] = nrow(dat[which(dat$gdp< tp[1,6] & dat$gdp >0), ])
+nc[2,6] = nrow(dat[which(dat$cons< tp[2,6] & dat$cons >0), ])
+nc[3,6] = nrow(dat[which(dat$gdp< tp[3,6] & dat$gdp >0), ])
+nc[4,6] = nrow(dat[which(dat$cons< tp[4,6] & dat$cons >0), ])
+
+nc.pc[1,6] = round(nc[1,6]/n.gdp*100,0)
+nc.pc[2,6] = round(nc[2,6]/n.cons*100,0)
+nc.pc[3,6] = round(nc[3,6]/n.gdp*100,0)
+nc.pc[4,6] = round(nc[4,6]/n.cons*100,0)
+
+nc[1,7] = nrow(dat[which(dat$gdp< tp[1,7] & dat$gdp >0), ])
+nc[2,7] = nrow(dat[which(dat$cons< tp[2,7] & dat$cons >0), ])
+nc[3,7] = nrow(dat[which(dat$gdp< tp[3,7] & dat$gdp >0), ])
+nc[4,7] = nrow(dat[which(dat$cons< tp[4,7] & dat$cons >0), ])
+
+nc.pc[1,7] = round(nc[1,7]/n.gdp*100,0)
+nc.pc[2,7] = round(nc[2,7]/n.cons*100,0)
+nc.pc[3,7] = round(nc[3,7]/n.gdp*100,0)
+nc.pc[4,7] = round(nc[4,7]/n.cons*100,0)
+
+nc.pc.colmean = colMeans(nc.pc)
+nc.pc.mean = mean(nc.pc.colmean)
+nc.pc.panel.mean = mean(nc.pc.colmean[6:7])
+nc.pc.nonpanel.mean = mean(nc.pc.colmean[-(6:7)])
 @
@@
 \label{tab:tp}
-\begin{center}
-\begin{tabular}{|c|c|c|c|c|c|}
+\small
+\centering
+\begin{tabular}{|c|c|c|c|c|c|c|}
 \hline 
-Model & dep. var & indep. var & tp (\$, PPP) & significance & \# of countries\tabularnewline
+Table & Estimator & dep. var & indep. var & tp (\$, PPP) & sign. & \# ctrs left of tp.\tabularnewline
 \hline 
 \hline
-(\ref{eq:infl_model}) & FPI infl.
-07-08  & GDP & \Sexpr{round(tp[1,1],0)} & no & \Sexpr{nrow(data_0708[which(data_0708$GDP< tp[1,1] & data_0708$GDP >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$GDP>0), ])}  (\Sexpr{round(nrow(data_0708[which(data_0708$GDP< tp[1,1] & data_0708$GDP >0), ])/nrow(data_0708[which(data_0708$GDP>0), ])*100,0)}\%)\tabularnewline
+\ref{tab:fpi-regressions} & OLS & FPI infl. 07-08  & GDP & \Sexpr{round(tp[1,1],0)} & yes & \Sexpr{nc[1,1]}/\Sexpr{n.gdp} (\Sexpr{nc.pc[1,1]} \%) \tabularnewline
 \hline 
-(\ref{eq:infl_model}) & FPI infl.
-07-08 & cons & \Sexpr{round(tp[2,1],0)} & no & \Sexpr{nrow(data_0708[which(data_0708$cons< tp[2,1] & data_0708$cons >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$cons>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$cons< tp[2,1] & data_0708$cons >0), ])/nrow(data_0708[which(data_0708$cons>0), ])*100,0)}\%)\tabularnewline
+\ref{tab:fpi-regressions} & OLS & FPI infl. 07-08  & Cons & \Sexpr{round(tp[2,1],0)} & yes & \Sexpr{nc[2,1]}/\Sexpr{n.cons} (\Sexpr{nc.pc[2,1]} \%) \tabularnewline
 \hline 
-(\ref{eq:infl_model}) & FPI infl.
-07-08 & food & \Sexpr{round(tp[3,1],0)}  & yes & \Sexpr{nrow(data_0708[which(data_0708$food< tp[3,1] & data_0708$food >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$food>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$food< tp[3,1] & data_0708$food >0), ])/nrow(data_0708[which(data_0708$food>0), ])*100,0)}\%)\tabularnewline
+\ref{tab:fpi-regressions} & OLS & FPI infl. 10-11  & GDP & \Sexpr{round(tp[3,1],0)} & no & \Sexpr{nc[3,1]}/\Sexpr{n.gdp} (\Sexpr{nc.pc[3,1]} \%) \tabularnewline
 \hline 
-(\ref{eq:infl_model}) & rFPI gr.
-07-08 & GDP & \Sexpr{round(tp[4,1],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$GDP< tp[4,1] & data_0708$GDP >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$GDP>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$GDP< tp[4,1] & data_0708$GDP >0), ])/nrow(data_0708[which(data_0708$GDP>0), ])*100,0)}\%)\tabularnewline
+\ref{tab:fpi-regressions} & OLS & FPI infl. 10-11  & Cons & \Sexpr{round(tp[4,1],0)} & yes & \Sexpr{nc[4,1]}/\Sexpr{n.cons} (\Sexpr{nc.pc[4,1]} \%) \tabularnewline
 \hline 
-(\ref{eq:infl_model}) & rFPI gr.
-07-08 & cons & \Sexpr{round(tp[5,1],0)}  & yes & \Sexpr{nrow(data_0708[which(data_0708$cons< tp[5,1] & data_0708$cons >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$cons>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$cons< tp[5,1] & data_0708$cons >0), ])/nrow(data_0708[which(data_0708$cons>0), ])*100,0)}\%)\tabularnewline
+\ref{tab:rfpi-regressions} & OLS & rFPI growth 07-08  & GDP & \Sexpr{round(tp[1,2],0)} & yes & \Sexpr{nc[1,2]}/\Sexpr{n.gdp} (\Sexpr{nc.pc[1,2]} \%) \tabularnewline
 \hline 
-(\ref{eq:infl_model}) & rFPI gr.
-07-08 & food & \Sexpr{round(tp[6,1],0)}  & yes & \Sexpr{nrow(data_0708[which(data_0708$food< tp[6,1] & data_0708$food >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$food>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$food< tp[6,1] & data_0708$food >0), ])/nrow(data_0708[which(data_0708$food>0), ])*100,0)}\%)\tabularnewline
+\ref{tab:rfpi-regressions} & OLS & rFPI growth 07-08  & Cons & \Sexpr{round(tp[2,2],0)} & yes & \Sexpr{nc[2,2]}/\Sexpr{n.cons} (\Sexpr{nc.pc[2,2]} \%) \tabularnewline
 \hline 
-(\ref{eq:infl_model}) & FPI infl.
-10-11 & GDP & \Sexpr{round(tp[1,2],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$GDP< tp[1,2] & data_0708$GDP >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$GDP>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$GDP< tp[1,2] & data_0708$GDP >0), ])/nrow(data_0708[which(data_0708$GDP>0), ])*100,0)}\%)\tabularnewline
+\ref{tab:rfpi-regressions} & OLS & rFPI growth 10-11  & GDP & \Sexpr{round(tp[3,2],0)} & yes & \Sexpr{nc[3,2]}/\Sexpr{n.gdp} (\Sexpr{nc.pc[3,2]} \%) \tabularnewline
 \hline 
-(\ref{eq:infl_model}) & FPI infl.
-10-11 & cons & \Sexpr{round(tp[2,2],0)} & no &  \Sexpr{nrow(data_0708[which(data_0708$cons< tp[2,2] & data_0708$cons >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$cons>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$cons< tp[2,2] & data_0708$cons >0), ])/nrow(data_0708[which(data_0708$cons>0), ])*100,0)}\%)\tabularnewline
+\ref{tab:rfpi-regressions} & OLS & rFPI growth 10-11  & Cons & \Sexpr{round(tp[4,2],0)} & yes & \Sexpr{nc[4,2]}/\Sexpr{n.cons} (\Sexpr{nc.pc[4,2]} \%) \tabularnewline
 \hline 
-(\ref{eq:infl_model}) & FPI infl.
-10-11 & food & \Sexpr{round(tp[3,2],0)} & yes &  \Sexpr{nrow(data_0708[which(data_0708$food< tp[3,2] & data_0708$food >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$food>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$food< tp[3,2] & data_0708$food >0), ])/nrow(data_0708[which(data_0708$food>0), ])*100,0)}\%)\tabularnewline
+\ref{tab:int-regressions} & OLS & FPI infl. 07-08  & GDP & \Sexpr{round(tp[1,3],0)} & yes & \Sexpr{nc[1,3]}/\Sexpr{n.gdp} (\Sexpr{nc.pc[1,3]} \%) \tabularnewline
 \hline 
-(\ref{eq:infl_model}) & rFPI gr.
-10-11 & GDP & \Sexpr{round(tp[4,2],0)} & yes &  \Sexpr{nrow(data_0708[which(data_0708$GDP< tp[4,2] & data_0708$GDP >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$GDP>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$GDP< tp[4,2] & data_0708$GDP >0), ])/nrow(data_0708[which(data_0708$GDP>0), ])*100,0)}\%)\tabularnewline
+\ref{tab:int-regressions} & OLS & FPI infl. 07-08  & Cons & \Sexpr{round(tp[2,3],0)} & yes & \Sexpr{nc[2,3]}/\Sexpr{n.cons} (\Sexpr{nc.pc[2,3]} \%) \tabularnewline
 \hline 
-(\ref{eq:infl_model}) & rFPI gr.
-10-11 & cons & \Sexpr{round(tp[5,2],0)} & yes &  \Sexpr{nrow(data_0708[which(data_0708$cons< tp[5,2] & data_0708$cons >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$cons>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$cons< tp[5,2] & data_0708$cons >0), ])/nrow(data_0708[which(data_0708$cons>0), ])*100,0)}\%)\tabularnewline
+\ref{tab:int-regressions} & OLS & FPI infl. 07-08  & GDP & \Sexpr{round(tp[3,3],0)} & yes & \Sexpr{nc[3,3]}/\Sexpr{n.gdp} (\Sexpr{nc.pc[3,3]} \%) \tabularnewline
 \hline 
-(\ref{eq:infl_model}) & rFPI gr.
-10-11 & food & \Sexpr{round(tp[6,2],0)} & yes &  \Sexpr{nrow(data_0708[which(data_0708$food< tp[6,2] & data_0708$food >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$food>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$food< tp[6,2] & data_0708$food >0), ])/nrow(data_0708[which(data_0708$food>0), ])*100,0)}\%)\tabularnewline
+\ref{tab:int-regressions} & OLS & FPI infl. 07-08  & Cons & \Sexpr{round(tp[4,3],0)} & yes & \Sexpr{nc[4,3]}/\Sexpr{n.cons} (\Sexpr{nc.pc[4,3]} \%) \tabularnewline
 \hline 
-(\ref{eq:ptreg}) & LR multiplier & GDP & \Sexpr{round(tp[1,3],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$GDP< tp[1,3] & data_0708$GDP >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$GDP>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$GDP< tp[1,3] & data_0708$GDP >0), ])/nrow(data_0708[which(data_0708$GDP>0), ])*100,0)}\%)\tabularnewline
+\ref{tab:fpi-pt-regressions} & OLS & FPI LRM  & GDP & \Sexpr{round(tp[1,4],0)} & yes & \Sexpr{nc[1,4]}/\Sexpr{n.gdp} (\Sexpr{nc.pc[1,4]} \%) \tabularnewline
 \hline 
-(\ref{eq:ptreg}) & LR multiplier & cons & \Sexpr{round(tp[2,3],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$cons< tp[2,3] & data_0708$cons >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$cons>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$cons< tp[2,3] & data_0708$cons >0), ])/nrow(data_0708[which(data_0708$cons>0), ])*100,0)}\%)\tabularnewline
+\ref{tab:fpi-pt-regressions} & OLS & FPI LRM  & Cons & \Sexpr{round(tp[2,4],0)} & yes & \Sexpr{nc[2,4]}/\Sexpr{n.cons} (\Sexpr{nc.pc[2,4]} \%) \tabularnewline
 \hline 
-(\ref{eq:ptreg}) & LR multiplier & food & \Sexpr{round(tp[3,3],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$food< tp[3,3] & data_0708$food >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$food>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$food< tp[3,3] & data_0708$food >0), ])/nrow(data_0708[which(data_0708$food>0), ])*100,0)}\%)\tabularnewline
+\ref{tab:fpi-pt-regressions} & WLS & FPI LRM  & GDP & \Sexpr{round(tp[3,4],0)} & no & \Sexpr{nc[3,4]}/\Sexpr{n.gdp} (\Sexpr{nc.pc[3,4]} \%) \tabularnewline
 \hline 
-(\ref{eq:ptreg}) & LR multiplier & GDP & \Sexpr{round(tp[4,3],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$GDP< tp[4,3] & data_0708$GDP >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$GDP>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$GDP< tp[4,3] & data_0708$GDP >0), ])/nrow(data_0708[which(data_0708$GDP>0), ])*100,0)}\%)\tabularnewline
+\ref{tab:fpi-pt-regressions} & WLS & FPI LRM  & Cons & \Sexpr{round(tp[4,4],0)} & yes & \Sexpr{nc[4,4]}/\Sexpr{n.cons} (\Sexpr{nc.pc[4,4]} \%) \tabularnewline
 \hline
-(\ref{eq:ptreg}) & LR multiplier & cons & \Sexpr{round(tp[5,3],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$cons< tp[5,3] & data_0708$cons >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$cons>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$cons< tp[5,3] & data_0708$cons >0), ])/nrow(data_0708[which(data_0708$cons>0), ])*100,0)}\%)\tabularnewline
+\ref{tab:rfpi-pt-regressions} & OLS & rFPI LRM  & GDP & \Sexpr{round(tp[1,5],0)} & yes & \Sexpr{nc[1,5]}/\Sexpr{n.gdp} (\Sexpr{nc.pc[1,5]} \%) \tabularnewline
 \hline 
-(\ref{eq:ptreg}) & LR multiplier & food & \Sexpr{round(tp[6,3],0)} & no & \Sexpr{nrow(data_0708[which(data_0708$food< tp[6,3] & data_0708$food >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$food>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$food< tp[6,3] & data_0708$food >0), ])/nrow(data_0708[which(data_0708$food>0), ])*100,0)}\%)\tabularnewline
+\ref{tab:rfpi-pt-regressions} & OLS & rFPI LRM  & Cons & \Sexpr{round(tp[2,5],0)} & yes & \Sexpr{nc[2,5]}/\Sexpr{n.cons} (\Sexpr{nc.pc[2,5]} \%) \tabularnewline
 \hline 
-(\ref{eq:panel_model}) & mth FPI infl & GDP & \Sexpr{round(tp[1,4],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$GDP< tp[1,4] & data_0708$GDP >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$GDP>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$GDP< tp[1,4] & data_0708$GDP >0), ])/nrow(data_0708[which(data_0708$GDP>0), ])*100,0)}\%)\tabularnewline
+\ref{tab:rfpi-pt-regressions} & WLS & rFPI LRM  & GDP & \Sexpr{round(tp[3,5],0)} & no & \Sexpr{nc[3,5]}/\Sexpr{n.gdp} (\Sexpr{nc.pc[3,5]} \%) \tabularnewline
 \hline 
-(\ref{eq:panel_model}) & mth FPI infl & cons & \Sexpr{round(tp[2,4],0)} & no & \Sexpr{nrow(data_0708[which(data_0708$cons< tp[2,4] & data_0708$cons >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$cons>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$cons< tp[2,4] & data_0708$cons >0), ])/nrow(data_0708[which(data_0708$cons>0), ])*100,0)}\%)\tabularnewline
+\ref{tab:rfpi-pt-regressions} & WLS & rFPI LRM  & Cons & \Sexpr{round(tp[4,5],0)} & no & \Sexpr{nc[4,5]}/\Sexpr{n.cons} (\Sexpr{nc.pc[4,5]} \%) \tabularnewline
 \hline
-(\ref{eq:panel_model}) & mth FPI infl & food & \Sexpr{round(tp[3,4],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$food< tp[3,4] & data_0708$food >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$food>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$food< tp[3,4] & data_0708$food >0), ])/nrow(data_0708[which(data_0708$food>0), ])*100,0)}\%)\tabularnewline
+\ref{tab:fe-panel-regressions} & FE & FPI infl.  & GDP & \Sexpr{round(tp[1,6],0)} & yes & \Sexpr{nc[1,6]}/\Sexpr{n.gdp} (\Sexpr{nc.pc[1,6]} \%) \tabularnewline
 \hline 
-(\ref{eq:panel_model}) & mth rFPI gr.
-& GDP & \Sexpr{round(tp[4,4],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$GDP< tp[4,4] & data_0708$GDP >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$GDP>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$GDP< tp[4,4] & data_0708$GDP >0), ])/nrow(data_0708[which(data_0708$GDP>0), ])*100,0)}\%)\tabularnewline
+\ref{tab:fe-panel-regressions} & FE & FPI infl.  & Cons & \Sexpr{round(tp[2,6],0)} & yes & \Sexpr{nc[2,6]}/\Sexpr{n.cons} (\Sexpr{nc.pc[2,6]} \%) \tabularnewline
 \hline 
-(\ref{eq:panel_model}) & mth rFPI gr.
-& cons & \Sexpr{round(tp[5,4],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$cons< tp[5,4] & data_0708$cons >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$cons>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$cons< tp[5,4] & data_0708$cons >0), ])/nrow(data_0708[which(data_0708$cons>0), ])*100,0)}\%)\tabularnewline
+\ref{tab:fe-panel-regressions} & FE & rFPI growth  & GDP & \Sexpr{round(tp[3,6],0)} & yes & \Sexpr{nc[3,6]}/\Sexpr{n.gdp} (\Sexpr{nc.pc[3,6]} \%) \tabularnewline
 \hline 
-(\ref{eq:panel_model}) & mth rFPI gr.
-& food & \Sexpr{round(tp[6,4],0)} & no & \Sexpr{nrow(data_0708[which(data_0708$food< tp[6,4] & data_0708$food >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$food>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$food< tp[6,4] & data_0708$food >0), ])/nrow(data_0708[which(data_0708$food>0), ])*100,0)}\%)\tabularnewline
+\ref{tab:fe-panel-regressions} & FE & rFPI growth  & Cons & \Sexpr{round(tp[4,6],0)} & yes & \Sexpr{nc[4,6]}/\Sexpr{n.cons} (\Sexpr{nc.pc[4,6]} \%) \tabularnewline
 \hline
-\multicolumn{6}{l}{Notes: The numbers in the fourth column refer to the estimated income or expenditure}\tabularnewline
-\multicolumn{6}{l}{turning point (tp). The fifth column indicates whether the estimated quadratic term}\tabularnewline
-\multicolumn{6}{l}{coefficients are individually significant at the 10 percent level. The last column shows}\tabularnewline
-\multicolumn{6}{l}{the number of countries in the sample with a per capita income or expenditure levels}\tabularnewline 
-\multicolumn{6}{l}{below the estimated turning point.}
+\ref{tab:fd-panel-regressions} & AH & FPI infl.  & GDP & \Sexpr{round(tp[1,7],0)} & no & \Sexpr{nc[1,7]}/\Sexpr{n.gdp} (\Sexpr{nc.pc[1,7]} \%) \tabularnewline
+\hline 
+\ref{tab:fd-panel-regressions} & AH & FPI infl.  & Cons & \Sexpr{round(tp[2,7],0)} & no & \Sexpr{nc[2,7]}/\Sexpr{n.cons} (\Sexpr{nc.pc[2,7]} \%) \tabularnewline
+\hline 
+\ref{tab:fd-panel-regressions} & AH & rFPI growth  & GDP & \Sexpr{round(tp[3,7],0)} & no & \Sexpr{nc[3,7]}/\Sexpr{n.gdp} (\Sexpr{nc.pc[3,7]} \%) \tabularnewline
+\hline 
+\ref{tab:fd-panel-regressions} & AH & rFPI growth  & Cons & \Sexpr{round(tp[4,7],0)} & yes & \Sexpr{nc[4,7]}/\Sexpr{n.cons} (\Sexpr{nc.pc[4,7]} \%) \tabularnewline
+\hline 
 \end{tabular}
-\end{center}
 \end {table}

### 2016-01-30 cs_pt_reg.R
Revised version. Regressions are split up into those based on nominal vs. real FPI

@@ -9,23 +9,23 @@ library(lmtest)
 ##################################################################
 
 load("ctr_dat_0708.Rda")
-load("coef_ecm_fpi.Rda")
-load("coef_ecm_rfpi.Rda")
 load("coef_bew_fpi_l1.Rda")
+load("coef_bew_rfpi_l1.Rda")
+load("coef_bew_fpi_l12.Rda")
+load("coef_bew_rfpi_l12.Rda")
 
-data = merge(select(ctr.dat.0708, -(2:4)), select(coef.ecm.fpi, iso3c, lrm, lrm.se))
-# data = merge(select(ctr.dat.0708, -(2:4)), coef.bew.fpi.l1)
-# data = rename(data, lrm.fpi=V1, lrm.fpi.se=V2)
-data = dplyr::rename(data, lrm.fpi=lrm, lrm.fpi.se=lrm.se)
-data = merge(data, select(coef.ecm.rfpi, iso3c, lrm, lrm.se))
-data = dplyr::rename(data, lrm.rfpi=lrm, lrm.rfpi.se=lrm.se)
+## FPI LR multiplier regressions
+fpi.pt.reg.m = list()
+fpi.pt.reg.se = list()
+fpi.bptest.stat = list()
+fpi.bptest.p = list()
 
-## 1 lag LR multiplier regressions
-pt.reg.l1 = list()
-pt.fstat = list()
-pt.fstat.p = list()
+# 1 lag LRM
+data = merge(ctr.dat.0708, coef.bew.fpi.l1)
+data = rename(data, lrm1 = lrm, lrm1.se = lrm.se)
 
-m = lm(lrm.fpi~log(gdp)+
+# OLS
+m = lm(lrm1~log(gdp)+
          I(log(gdp)^2)+
          cshare+
          cdep+
@@ -34,34 +34,17 @@ m = lm(lrm.fpi~log(gdp)+
          to+
          dtf+
          lpi,
-       data = filter(data, lrm.fpi<3.5*sd(lrm.fpi)),
-       weights = 1/lrm.fpi.se)
+       data = filter(data, lrm1<3.5*sd(lrm1)))
 cov = vcovHC(m, type = "HC1")
 rse = sqrt(diag(cov))
 
-pt.reg.l1[[1]] = m
-pt.reg.l1[[2]] = rse
+fpi.pt.reg.m[[1]] = m
[…]

### 2016-01-30 panel_reg.R
Revised version. Regressions split up into FE and AH models.

@@ -6,7 +6,7 @@ library(plm)
 library(foreign)
 
 ##################################################################
-### Cross section price transmission regressions
+### Panel price transmission regressions
 ##################################################################
 
 load("ctr_dat_0708.Rda")
@@ -17,138 +17,245 @@ data = plm.data(df, index = c("iso3c", "tt"))
 df1 = select(df, -fao.fpi)
 df1 = rename(df1, fao.fpi = r.fao.fpi)
 data1 = plm.data(df1, index = c("iso3c", "tt"))
-# write.dta(data, "panel.dta")
 
-panel.reg = list()
+panel.reg.m = list()
+p = 6
 
-m = plm(diff(log(adj.fpi)) ~ lag(diff(log(adj.fpi)),1) +
-          diff(log(fao.fpi)) +
-          lag(diff(log(fao.fpi)), 1) +
-          diff(log(er)) +
-          lag(diff(log(er)), 1) +
-          I(lag(diff(log(fao.fpi)),1)*log(gdp)) +
-          I(lag(diff(log(fao.fpi)),1)*I(log(gdp)^2)) +
-          I(lag(diff(log(fao.fpi)),1)*cshare) +
-          I(lag(diff(log(fao.fpi)),1)*cdep) +
-          I(lag(diff(log(fao.fpi)),1)*I(cshare*cdep)) +
-          I(lag(diff(log(fao.fpi)),1)*lldc) +
-          I(lag(diff(log(fao.fpi)),1)*to) +
-          I(lag(diff(log(fao.fpi)),1)*dtf) +
-          I(lag(diff(log(fao.fpi)),1)*lpi),
-        data = data, model = "within")
+#####################################################
+## Fixed effects (FE) estimator
 
-cov = vcovHC(m, method = "arellano", type = "HC1", cluster = "group")
-rse = sqrt(diag(cov))
+# FPI regressions
+m = plm(diff(log(adj.fpi)) ~ lag(diff(log(adj.fpi)),1:p) +
+          lag(diff(log(fao.fpi)), 0:1) +
+          lag(diff(log(er)), 0:1) +
+          I(lag(diff(log(fao.fpi)),0)*log(gdp)) +
[…]

### 2016-02-29 Master.Rnw
Master file for 2. round version of paper.

@@
 \documentclass[12pt,english]{article}
-
-\usepackage{geometry}
+\usepackage{color,soul}
 \usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
+\usepackage{inputenc}
 \usepackage{babel}
@@
 \usepackage{graphicx}
-\usepackage{setspace}
 \usepackage[authoryear]{natbib}
@@
 \usepackage{multirow}
-%\usepackage{authblk}
 \usepackage{lscape}
@@
 \usepackage{longtable}
+\usepackage{array}
 \usepackage[noae]{Sweave}
@@
 \doublespacing
+\usepackage{geometry}
 \geometry{verbose,tmargin=2cm,bmargin=2cm,lmargin=3cm,rmargin=3cm}
+\usepackage{amsmath}
+\usepackage[flushleft]{threeparttable}
 
@@
 
-\title{Income and Food Price Transmission\footnote{Revised version of ``Poverty and Price Transmission''. Submitted to \emph{Agricultural Economics} for review}}
+\title{Income and Food Price Transmission\footnote{Revised version of MS 2015310 ``Income and Price Transmission''.
+Submitted to \emph{Agricultural Economics} for 2.
+round review.
+I thank two anonymous reviewers for their valuable comments.}}
 
-\author{Christian Elleby}
-\date{2015}
+\date{\vspace{-5ex}}
 
 \maketitle
+
 \begin{abstract}
-The paper analyzes how international-to-domestic food price transmission
-depends on a country\textquoteright s income or expenditure level.
-According to conventional wisdom, international food price spikes
-have the largest poverty impact on low income countries. However,
-the paper presents empirical evidence of an inverse-U shaped relationship
-between the degree of international agricultural price changes passed
-on to domestic consumers and the country\textquoteright s per capita
-income or expenditure level. This implies that retail food prices
-in middle income countries respond the strongest to international
-commodity price spikes, whereas those in low income countries respond
-less, all else equal. The specification of the empirical model is
-based on a simple theoretical model of a country\textquoteright s
-aggregate food price level. Its predictions are tested on monthly
-price index data from 140 countries covering the six year period 2005-11
-and results are found to be generally significant and robust. Around
-one third to one half of the countries in the sample are to the left
-of the turning point, on the increasing part of the parabola, in most
-specifications.\end{abstract}
+This paper examines determinants of international-to-domestic food price transmission using aggregate monthly price data for more than 100 countries covering the ten year period 2005-15.
+Findings indicate that middle-income countries experience the highest rates of food price transmission in the long run.
+Other determinants of food price transmission include a country's cereal import, export and consumption shares, its trade openness, the quality of its infrastructure and government involvement in grain trade.
+A novel model is presented which is able to explain the economic logic behind the observed inverse-U shaped relationship between income and price transmission.
+
+\end{abstract}
+
+\textbf{Keywords:} price transmission, food crisis, food inflation, determinants, regression analysis, income\newpage
 
 \SweaveInput{intro.Rnw}
-
-\SweaveInput{litrev.Rnw}
-
 \SweaveInput{model.Rnw}
-
 \SweaveInput{analysis.Rnw}
-
 \SweaveInput{conclusion.Rnw}
@@
 \bibliographystyle{chicago}
-\bibliography{poverty_price_transmission}
+\bibliography{inc_pt}
 
+\SweaveInput{figures1.Rnw}
 \newpage
-
+\SweaveInput{tables.Rnw}
+\newpage
 \appendix
-
-\SweaveInput{figures.Rnw}
-
-\newpage
-
-\SweaveInput{tables.Rnw}
+\SweaveInput{appendix.Rnw}
 

### 2016-02-29 analysis.Rnw
Analysis section of 2. round version of paper.

--- 2015-12-22 analysis.Rnw
@@
-% !Rnw root = Master.Rnw
-
-\SweaveInput{data_prep.Rnw}
-\SweaveInput{regressions.Rnw}
+% !Rnw root = master.Rnw
 
 \section{Empirical analysis\label{sec:analysis}}
-This section assesses the model's predictions regarding the relationship between income and price transmission.
-The model in section \ref{sec:model} predicts an inverse-U shaped relationship between income and price transmission but it does not pin down the actual turning point, which is an empirical problem.
-Our goal in this section is thus to a) establish whether there is a statistically significant quadratic relationship between income or, alternatively, expenditure and food price transmission and b) to determine the income/expenditure level corresponding to the turning point.
+This section assesses the impact of the determinants discussed above on international-to-domestic aggregate food price transmission.
 Section \ref{sec:data} introduces the data and provides some context.
@@
 
-\subsection{Data on domestic and international food prices\label{sec:data}}
-The domestic price index data analyzed below is sourced from the International Labor Organization (ILO).\footnote{See \href{http://laborsta.ilo.org/}{http://laborsta.ilo.org/}} ILO maintains the most comprehensive database of domestic food price indexes (FPIs) and consumer price indexes (CPIs) which is used to monitor the cost of living and real wages.
-The ILO food price database is not widely used, for good reasons, despite a broad country coverage and series going back to the 1970s in many cases.
-A major issue with this data is that the historical series are not recalculated whenever a new base year is introduced.
+\subsection{Data\label{sec:data}}
+
+<<echo=FALSE>>=
+load("infl_0708_1011.Rda")
+load("ts_dat.Rda")
+n.full = nrow(table(ts.dat$id))
+n.sub = nrow(infl)
+infl = infl[infl$iso3c != "BLR", ]
+@
+
+\subsubsection{Domestic and international food prices}
+Data on domestic prices are sourced from the International Labor Organization (ILO).\footnote{See \citet{ilo16}} ILO maintains a comprehensive database of domestic food price indexes (FPIs) and consumer price indexes (CPIs) that is used to monitor the cost of living and real wages.
+There are not many studies which draw on this database, however, despite a broad country coverage and long monthly price series.
+A major issue with this data is that historical series are not recalculated whenever a new base year is introduced.
 The same applies when other changes are made such as the inclusion or exclusion of certain items such as beverages or tobacco from the index.
 This means that many of the series have discrete level shifts associated with these changes.
-Another issue with the data is that some countries report several series.
 In some cases one series replace another after a period of overlap.
 Finally, many of the series contain missing values.
-To ``tidy up'' the raw data I carried out the following operations.
-First, I restricted the sample to the period 2005:1-2010:12.
-Secondly, overlapping series were spliced and in cases where multiple complete series were available from the same country, a single one was selected.
+To ``tidy up'' the raw data I carried out the following steps First, I restricted the sample to the period 2005:1-2014:12.
+Secondly, in cases where multiple complete series were available from the same country, a single one was selected.
 Thirdly, series with missing values were discarded.
-Finally, and most critical for the analysis below, level shifts associated with base year changes etc.
+Fourthly, and most critical for the analysis below, level shifts associated with base year changes etc.
 were identified and removed.\footnote{Concretely I added the change in the break point from the previous value to the break point and all points succeeding it.
-The value of the adjusted series at the break point is thus the same as the one preceding it.} In total 140 countries have complete price series for the 2005-10 period considered.
-An important last thing to know about this data is that most of the low income countries' price indexes derive from prices collected from one or two major markets in the capital.
-Weights and selected items are derived from household expenditure surveys.
+The value of the adjusted series at the break point is thus the same as the one preceding it.} Finally, all the series were normalized with respect to a common base period (Jan 2007).
+In total \Sexpr{n.full} countries have complete price series for the entire ten year period considered whereas, for example, there are \Sexpr{n.sub} countries with complete price series covering the 2005-11 subperiod.
+There is therefore a trade-off between the number of complete price series and the length of the series.
+
+An important detail to know about this data is that many of the low income countries' price indexes derive from prices collected from a few major markets in the capital.
+Furthermore, weights and selected items are derived from household expenditure surveys which often target certain groups such as the urban middle class.
 Differences in consumption patterns around the world of course translate into differences in weights and in the composition of the food baskets themselves.
 This is probably the largest drawback of using index data in a cross section analysis of price transmission.
-That being said, even domestic prices of individual food commodities such as wheat or rice would not refer to the same good due to quality and varietal differences, so it is not the case that focusing on a few selected food items consumed by all countries would solve the problem.
+That being said, even domestic prices of individual food commodities such as wheat or rice would not refer to the same good due to quality and varietal differences, so it is not the case that focusing on a few selected food items consumed by all countries would necessarily solve the problem.
+
+International food prices are represented by the FAO food price index (FAO FPI) and in some cases its ``real'' version, namely the nominal FAO FPI deflated by the Manufactures Unit Value (MUV) index published by the World Bank.\footnote{see \citet{fao16} and \citet{worldbank16a}.}
+
+\subsubsection{Domestic food inflation during the food crises}
 
 There has been two recent international price spikes; one in 2007-8 and the other in 2010-11.
-Between January 2007 and June 2008 FAO's food and grain price indexes, which are normalized with respect to the 2002-4 average, rose by 91 and 127 points to index 226 and 268, respectively, corresponding to relative price increases of 68 and 90 percent.
-From the first peak in mid-2008 and to their their respective troughs in February and September 2009 the two indexes fell by 82 and 113 points before rising back up to index 240 and 261 in February and April 2011, respectively.
+The former spike especially is often referred to as ``the food crisis'' in the literature \citep[e.g.][]{Headey2008, Headey2013, timmer10, Ivanic2012, Martin2011}.
+Between January 2007 and June 2008 the FAO FPI rose by 91 points to index 226, corresponding to a 68 percent relative price increase.
+From the first peak in mid-2008 and to the trough in February 2009 the FAO FPI fell by 82 points before rising back up to index 240 in February 2011.
 
-The boxplot in figure \ref{fig:infl} depicts the empirical distribution of the sample countries' headline (CPI) inflation, their food (FPI) inflation as well as the change in their FPI relative to the CPI, termed real FPI growth, over the two time periods 2007:1-2008:6 and 2010:1-2010:12.
+The boxplot in figure \ref{fig:infl} depicts the empirical distribution of the sample countries' headline (CPI) inflation, their food (FPI) inflation as well as the change in their FPI relative to the CPI, termed real FPI (RFPI) growth, over the two time periods 2007:1-2008:6 and 2010:1-2010:12.
 
@@
 
-As mentioned in section \ref{sec:model}, the nominal food price level is not only driven by exogenous world market shocks, it also includes a stochastic trend which it shares with the non-food price level, representing underlying general domestic inflation.
-By dividing the FPI with the CPI, we obtain an FPI which is purged of headline inflation.
-The median country experienced an annualized rate of food and headline inflation of \Sexpr{round(median(infl$foodinfl78),1)} and \Sexpr{round(median(infl$cpiinfl78),1)} percent, respectively, in 2007-8.
-Real FPI growth, which is approximately the difference between the two, was \Sexpr{round(median(infl$rfpigr78),1)} percent for the median country in that period.
-Price increases were generally lower in the second period.
-The median country experienced food and headline inflation rates \Sexpr{round(median(infl$foodinfl1011),1)} and \Sexpr{round(median(infl$cpiinfl1011),1)} percent implying a modest real FPI growth of \Sexpr{round(median(infl$rfpigr1011),1)} percent.
-In both periods the variation in inflation rates is substantial.
-The standard deviation and interquartile range of 2007-8 food inflation is \Sexpr{round(sd(infl$foodinfl78),1)} and \Sexpr{round(IQR(infl$foodinfl78),1)} percentage points, respectively, versus \Sexpr{round(sd(infl$foodinfl1011),1)} and \Sexpr{round(IQR(infl$foodinfl1011),1)} percentage points in 2010-11.
-Although median real FPI growth is lower in the second period the variation is similar in both periods (standard deviations of \Sexpr{round(sd(infl$rfpigr78),1)} and \Sexpr{round(sd(infl$rfpigr1011),1)} and interquartile ranges of \Sexpr{round(IQR(infl$rfpigr78),1)} and \Sexpr{round(IQR(infl$rfpigr1011),1)}, respectively).
-The three remaining panels contain alternative representations the same data.
-What they show us that there is some systematic variation in the food prices to be explained.
-As the positively sloped lowess lines indicate, a country hit by a large nominal and real FPI shock in 2007-8 was more likely to be hit by a relatively large shock again in 2010-11 and vice versa.
+The median country experienced an annualized rate of food and headline inflation of \Sexpr{round(median(infl$fpigr.0708),1)} and \Sexpr{round(median(infl$cpigr.0708),1)} percent, respectively, in 2007-08.
+Nominal food prices are not only driven by exogenous world market shocks; they follow a stochastic trend together with the non-food prices, representing underlying general (core) inflation.
+Dividing with the CPI is a simple way of purging the FPI of underlying domestic inflation.
+RFPI growth, which is approximately the difference between the FPI and CPI inflation rate, was \Sexpr{round(median(infl$rfpigr.0708),1)} percent for the median country in 2007-08.\footnote{A problem with the real FPI is that domestic non-food prices are also affected by international commodity shocks.
+Low or even negative growth in real domestic food prices in these two periods could therefore reflect a response to the concurrent surge in non-food commodities, oil in particular, rather than a puzzling negative response to an international food price shock.} Domestic price increases were generally lower in 2010-11.
+The median country experienced food and headline inflation rates \Sexpr{round(median(infl$fpigr.1011),1)} and \Sexpr{round(median(infl$cpigr.1011),1)} percent and a modest RFPI growth of \Sexpr{round(median(infl$rfpigr.1011),1)} percent.
+There is, however, a substantial variation in the inflation rates in both periods.
+
+The three remaining panels of figure \ref{fig:infl} contain alternative representations the same data.
+There appears to be some systematic variation; as the (mostly) positively sloped lowess lines indicate, a country hit by a large nominal and real FPI shock in 2007-8 was more likely to be hit by a relatively large shock again in 2010-11 and vice versa.
 This suggests that there are indeed characteristics which influences a country's response to a world market shock, an assertion that will be explored further in section \ref{sec:analysis1} and \ref{sec:analysis2}.
 
-\subsection{Determinants of domestic food price changes\label{sec:analysis1}}
+\subsubsection{Price transmission determinants}
+Table \ref{tab:cs_data} in the appendix lists the variables that, based on the discussion in section \ref{sec:model}, are considered potentially important determinants of price transmission.
+A country's PPP adjusted per capita GDP is considered a proxy for its income level.
+Other covariates considered are a country's percentage dietary energy supply derived from cereals, roots and tubers, its cereal import- and export shares, a dummy indicating whether the country is a landlocked developing country (LLDC), its trade openness defined as the sum of its imports and exports relative to its GDP and a dummy indicating whether there is a grain state trading enterprise (STE) in the country.
+In addition there are two indexes included representing the regulatory performance of the country in question and the quality of its infrastructure.
+These two indexes are abbreviated DTF (Distance to Frontier) and LPI (Logistics Performance Index), respectively.\footnote{The covariates are similar to those of \citet{greb12}.}
 
-Table \ref{tab:2007-8-regressions} and \ref{tab:2010-11-regressions} delve deeper into the  relationship between income and domestic food price changes at the time of the 2007-08 and 2010-11 international price spikes, respectively.
-Each of the columns refer to a regression of FPI inflation or real FPI growth on a set of income related predictor variables.
-The models can all be written
-\begin{equation}
-\pi_{i}^{dom}=\delta_{0} + \delta_{1}x_{i} + \delta_{2}x_{i}^{2} + \delta_{3}z_{i} + u_{i}\label{eq:infl_model}
-\end{equation}
-where $\pi_{i}^{dom}$ refers to either country $i$'s FPI inflation or its real FPI growth in either the 18 month period 2007:1-2008:6 or the 12 month period 2010:1-2010:12, $x_{i}$ refers to a subset of the explanatory variables considered, $z_{i}$ is a vector of control variables and $u_{i}$ is the error term.
-The covariates, $x_{i}$, include the log of country $i$'s (PPP adjusted) per capita GDP [$\log(GDP)$], its consumption [$\log(cons)$] and food expenditure level [$\log(food)$].
-The control variables in $z_{i}$ are country $i$'s percentage dietary energy supply derived from cereals, roots and tubers ($cshare$), its cereal import dependency ratio ($cdep$) and their interaction.\footnote{Cereal dietary energy supply and cereal dependency data are from FAOSTAT and refer to the 2004-06 average.
-Data on GDP and expenditure is sourced from the World Bank's website.
-It is the International Comparison Project (ICP) which is behind the expenditure estimates which refer to year 2005.} Summary statistics on these variables are provided in table \ref{tab:sumstat}.
-Of primarily interest is the quadratic relationship between domestic price changes $\pi_{i}^{dom}$ and expenditure, $x_{i}$.
-The first 12 rows of table \ref{tab:tp} summarize information about the turning points based on different estimated versions of model \ref{eq:infl_model}.
+\subsection{Determinants of domestic food price changes during the food crisis\label{sec:analysis1}}
+Table \ref{tab:07-08-fc-regressions} considers the cross sectional variation in domestic food price changes during the 2007-8 food crises.
+Each of the columns refer to a regression of FPI inflation or real FPI growth on the explanatory variables discussed above as well as the percentage change in a country's domestic currency/US dollar exchange rate during the food crisis.
+All variables except for GDP and the dummies have been standardized in order to facilitate comparisons of effect sizes.
 
 \begin{center}
-[Table \ref{tab:sumstat} around here]
+[Table \ref{tab:07-08-fc-regressions} around here]
 \par\end{center}
 
-Countries that are largely dependent on grain imports to cover their domestic consumption needs cannot use export restrictions as way of stabilizing domestic prices.
-Because of this vulnerability to international food price fluctuations we would expect to find stronger domestic price increases during international food price spikes among countries with large import dependency ratios.
-The analysis in section \ref{sec:model} suggests that international price shocks are transmitted the strongest to middle income countries so we would expect food price inflation in these two periods to follow an inverse-U shaped relationship with per capita income and consumption expenditure.
-Poor households and populations also spend a larger share of their food budget on staples which makes them more vulnerable to price spikes on grains and other internationally traded agricultural commodities.
-Therefore, if we did account for this, the estimated relationship between income and domestic food price changes could be spurious.
-By including the $cshare$ variable in the regression this omitted variable bias should be reduced.
+Columns (1) and (3) are based on regression models that include the full set of explanatory variables.
+Income has a significant inverse-U shaped effect as predicted.
+The other significant covariates are a country's cereal import and export shares, its trade openness, the STE dummy and its LPI.
+Grain exporters experienced more food inflation on average whereas importers experienced less.
+The coefficient to the STE dummy is negative and large in magnitude suggesting that countries with grain state trading enterprises were able to control domestic food prices better than other countries.
+Two of the significant variables have unexpected signs, namely a country's cereal import share and its quality of infrastructure and trade as summarized by its LPI.
+It is not clear why large grain importers experienced lower food inflation during the food crisis.
+It could be though, that large food importers intervene more in their food markets than other countries.
+This interpretation is consistent with the finding by \citet{Lee2013} that large food importers have lower food inflation in general but more research is needed before we can say anything conclusive about this.
+Better infrastructure should imply more integrated markets and stronger price transmission but countries with a high LPI score experienced lower food inflation on average during the food crisis.
+The effect on real FPI growth is insignificant though, which would suggest that these countries also experienced lower non-food inflation.
+Countries with depreciating currencies experienced significantly stronger FPI inflation as expected but the effect on real FPI growth is insignificant.
+The latter finding is not surprising given that exchange rate changes affect domestic food- as well as non-food prices.
+
+A concern is that some of the variables are highly correlated which leads to large standard errors and insignificant effects.
+Therefore, the subset of covariates included in columns (2) and (4) are those which minimizes the Akaike information criterion (AIC) as determined by stepwise regression.\footnote{This was carried out using the \emph{MASS} package in \textsf{R} \citep{venables02}} Stepwise regressions have been criticized for leading to models where the remaining terms appear to be more important than they actually are.
+However, as can be seen, no substantial changes result from this exercise.
+
+What effect did the government interventions have on domestic food inflation during the food crisis?
+Table \ref{tab:int-regressions} in the appendix reports results from a set of regressions similar to those in table \ref{tab:07-08-fc-regressions} except that a set of dummies are included indicating whether or not the government intervened in various ways during the food crisis.\footnote{These dummies are based on table 1 in \citet{Demeke2009}.
+``Export'' = 1 if the country Restricted or banned export, ``Price'' = 1 if the country administered price control or restricted private trade, ``Stock'' = 1 if the country released stocks (public or imported) at subsidized prices'', ``Tariff'' = 1 if the country reduced tariffs or customs fees on imports and ``Tax'' = 1 if the country suspended or reduced VAT or other taxes, all as of 1.
+December 2008.\label{int_note}} Some of the coefficients to these intervention dummies are substantial but they are all statistically insignificant.
+Table \ref{tab:1011-regressions} in the appendix contains results related to the 2010-11 international price spike.
+Again, estimated effects are similar to those in  table \ref{tab:07-08-fc-regressions} but most are statistically insignificant.
+In fact,  the cereal consumption share is the only robustly significant predictor of domestic food inflation in 2010-11.
+
+\subsection{Determinants of price transmission \label{sec:analysis2}}
+Although the results in section \ref{sec:analysis1} provide interesting insights into the cross sectional variation in domestic price changes during the two recent international food price spikes they have at least two shortcomings.
+First, none of the estimated relationships represent price transmission as such as there may be the other and perhaps more important causes of the observed domestic price changes.
+Secondly, the estimates are based on a specific short time period, which means that we ignore much the information contained in the individual price series and it also raises questions about the reliability of the results.
+To accommodate these shortcomings I approach the issue from two different angles.
+First, in line with the small existing literature on global food price transmission patterns, I estimate a set of long run price transmission rates based on country specific time series models.
+These estimates are then subsequently regressed on the same country characteristics as in section \ref{sec:analysis1}.
+Secondly, I estimate the effects of the predictor variables on short run price transmission directly in a panel model.
+
+\subsubsection{Meta analysis of estimated long run price transmission rates\label{sec:meta-analysis}}
+This section carries out a meta-regression analysis to see whether a country's estimated long run price transmission rate varies systematically with the proposed determinants of price transmission.
+
+Following the literature, monthly domestic food price changes are regressed on lagged domestic food price changes as well as on contemporaneous and lagged international food price changes and exchange rates \citep[see][]{Ianchovichina2014, IMF2011, Kalkuhl2014}.
+That is, for each country $i$ an autoregressive distributive lag (ADL) models is fitted
+\begin{equation}
+\pi_{i,t}^{dom}=\delta_{i,0}+\sum_{s=1}^{p}\rho_{i,s}\pi_{i,t-s}^{dom}+\sum_{s=0}^{q}(\lambda_{i,s}\pi_{t-s}^{int}+\phi_{i,s}er_{t-s})+\epsilon_{i,t}\label{eq:time_series_model}
+\end{equation}
+where $\pi_{i,t}^{dom}$ represents country $i$'s monthly FPI inflation or RFPI growth at time $t$ (both approximated by the first differences to the log of its FPI or RFPI), $\pi_{t}^{int}$ denotes the FAO FPI or, in case of the RFPI growth model, the FAO RFPI at time $t$ and $er_{i,t}$ is the country $i$'s domestic currency/US\$ exchange rate, both in log differences.
+
+The ADL model is appropriate when the causal impact is unidirectional.
+This could be the case when the domestic price is from a small country relative to the world market but, more often, this assumption is invoked when the domestic price variable is a Food Price Index (FPI), i.e.
+the food component of the Consumer Price Index (CPI).
+
+Prior to estimation I tested each of the series for unit roots and cointegration with international prices.\footnote{I tested for stationarity with the KPSS test \citep{kwiatkowski92} and for unit roots with the augmented Dickey-Fuller test \citep{dickey79}.
+I used the Phillips and Ouliaris ``Pz'' test \citep{phillips90} to test for cointegration.
+These tests are implemented in \textsf{R} via the packages \emph{tseries} \citep{trapletti15} and \emph{urca} \citep{pfaff08}, respectively.} Results indicated that the domestic FPI and real FPI series, are nonstationary in levels but stationary in first differences.
+Domestic price levels were also found not to be cointegrated with world prices.
+Consequently, each model was estimated in growth rates rather than levels in order to avoid spurious regressions.
+
+Next, for each country, I calculated the long run multiplier (LRM) associated with a change in the the FAO food price index given by
+\begin{equation}
+\theta_{i}=\frac{\sum_{s=0}^{q}\lambda_{i,s}}{1-\sum_{s=1}^{p}\rho_{i,s}}.
+\end{equation}
+
+<<echo=FALSE>>=
+library(dplyr)
+load("lrm_fpi_l1_12.Rda")
+fpi = select(lrm.fpi.l.1_12, -iso3c)
+load("lrm_rfpi_l1_12.Rda")
+rfpi = select(lrm.rfpi.l.1_12, -iso3c)
+fivenum.fpi = data.frame(sapply(fpi, summary))
+fivenum.rfpi = data.frame(sapply(rfpi, summary))
+n = nrow(fpi)
+@
+
+The LRM is a common measure of price transmission.
+\cite{IMF2011}, for example, contains an analysis of food price price transmission based on a dataset of monthly FPI inflation from 31 advanced economies and 47 emerging and developing economies over the period 2000 - 2011.
+Their median estimated LRM pass-through of a 1 percent increase in international food inflation is 0.18 percent in developed countries and 0.34 percent in emerging and developing countries.
+Similarly, \citet{Ianchovichina2014} report an average estimated LRM for the eighteen Middle East and North Africa (MENA) countries of 0.25 and the majority of the countries are reported to have  food price pass-through LRMs in the order of 0.2-0.4.
+Both studies use 12 lags in their country specific ADL models, i.e.
+they assume $p=q=12$.\footnote{\cite{IMF2011} do not include exchange rates in their regressions.} My own estimates based on a sample of \Sexpr{n} countries covering the period 2005:1 to 2014:12 are very similar.
+The average and median FPI LRM from a model with 12 lags of all variables are \Sexpr{round(fivenum.fpi["Mean", "V12"], 2)} and \Sexpr{round(fivenum.fpi["Median", "V12"], 2)}, respectively.
+That being said, it is not obvious that so many lags should be included in the models.
+When lag length selection is based on the Akaike Information Criterion (AIC), for example, the optimal number lags in most of the models turn out to be $p=q=1$.\footnote{For each country, I calculated the AIC for each of the $12^2$ combinations of $(p,q)\, \in (1,..,12)$.
+Then I selected the $(p,q)$ combination with the lowest AIC.} But when a common lag length of 1 is used for all countries, average and median FPI transmission LRMs are much lower, namely \Sexpr{round(fivenum.fpi["Mean", "V1"], 2)} and \Sexpr{round(fivenum.fpi["Median", "V1"], 2)}, respectively, whereas the average of the standard errors of the LRM estimates very similar.\footnote{I also tested for residual autocorrelation with the Durbin-Watson test.
+There was no evidence of residual autocorrelation in either case.
+Standard errors of the LRMs are based on the so called Bewley transformation of the ADL model instrumented with lags of the dependent variable.} As a compromise between the long lag lengths used in the literature which might cause problems with overfitting and the short optimal lag lengths implied by the AIC criterion which leads to LRM estimates that are lower than commonly found in the literature, the country specific models are based on common lag lengths of $p=q=6$.
+This choice leads to average long run FPI and RFPI transmission rates of \Sexpr{round(fivenum.fpi["Mean", "V6"], 2)} and \Sexpr{round(fivenum.rfpi["Mean", "V6"], 2)}, respectively.
+One interpretation of the very low RFPI LRM average is that shocks to a country's FPI is passed on almost fully to its non-FPI in the long run, a phenomenon known as propagation in the literature \citep[e.g.][]{pedersen10}.
+
+The long run transmission estimates, $\hat{\theta}$, are regressed, on the same explanatory variables as in table \ref{tab:07-08-fc-regressions}.
+However, since the dependent variable is an estimate rather than an observed value and the model now represents an estimated dependent variable (EDV) regression \citep[see][]{hanushek74, lewis05}.
+A common approach in this situation is to use Weighted Least Squares (WLS) estimation with weights $w_{i}$ that are inversely proportional to the standard error of $\theta_{i}$ \citep{saxonhouse76}.
+This way more precise estimates are given a higher weight in the regression.
+The main reason though for using WLS rather than OLS is that the latter may suffer from heteroskedasticity which leads to inefficient coefficient estimates inconsistent standard errors.
+The inconsistent OLS standard errors can, however, easily be corrected with a robust standard error estimator.
+\citet{lewis05} uses formal reasoning backed up by a Monte Carlo analysis to argue that OLS with robust standard errors is likely to produce more reliable estimates than WLS based on inverse standard error weights, and that the latter is likely to produce downward biased standard errors.
+
+Table \ref{tab:fpi-pt-regressions} presents estimates of FPI transmission determinants based on OLS [columns (1)-(2)] as well as WLS [columns (3)-(4)], both with robust standard errors.
+Again, all variables except per capita GDP and the dummies have been standardized and, as in table \ref{tab:fpi-pt-regressions}, column (1) and (3) represent regressions that are based on the full set of covariates whereas columns (2) and (4) represent regressions with the lowest AIC.
 
 \begin{center}
-[Table \ref{tab:2007-8-regressions} around here]
+[Table \ref{tab:fpi-pt-regressions} around here]
 \par\end{center}
 
-<<echo=FALSE>>=
-<<dom_infl_0708>>
-tp3 = -coef(m3)["log(food)"]/(2*coef(m3)["I(log(food)^2)"])
-tp6 = -coef(m6)["log(food)"]/(2*coef(m6)["I(log(food)^2)"])
-tp4 = -coef(m4)["log(GDP)"]/(2*coef(m4)["I(log(GDP)^2)"])
-tp5 = -coef(m5)["log(cons)"]/(2*coef(m5)["I(log(cons)^2)"])
-@
+First, as can be seen, results are robust to the choice of estimator.
+Secondly, although the results in table \ref{tab:fpi-pt-regressions} relate to determinants of FPI transmission rather than FPI inflation as in table \ref{tab:07-08-fc-regressions}, the estimated coefficients and significance patterns are similar across the two sets of regressions.
+A large grain consumption and grain import share is associated with lower price transmission and so is having a grain state trading enterprise.
+Large grain exporters, on the other hand, experience larger price transmission on average and so do more open countries.
+The signs of some of these coefficients could reflect systematic differences in the degree of government intervention in the domestic food market as discussed in section \ref{sec:analysis1}.
+Income again has a highly significant inverse-U shaped affect of FPI transmission and, oddly, countries with better infrastructure are once more found to experience lower price transmission on average.
+A possible explanation for the latter effect could be, that the LPI is simply another proxy for income.
 
-Table \ref{tab:2007-8-regressions} considers domestic food price changes during the 2007:1-2008:6 period for the countries in the sample for which there is data on income, expenditure, dietary energy supply and cereal import dependency.
-Column (1) relates a country's annualized FPI inflation rate in that period to its per capita GDP level.
-What we see is that average food inflation during the 2007-8 food crisis is increasing in a country's per capita income level up until a certain point after which it starts to decrease.
-The same applies for consumption and food expenditure in column (2) and (3), respectively.
-Only the latter set of coefficients are statistically significant though.
-Specifically, column (3) suggests that among countries with a 2005 per capita food expenditure level of \Sexpr{round(exp(tp3),1)} {[}$\exp(\Sexpr{round(tp3,2)})${]} PPP adjusted dollars or less there is a positive estimated relationship between expected FPI inflation and food expenditure.
-After that point the predicted marginal effect turns negative.
-\Sexpr{nrow(data_0708[which(data_0708$food< exp(tp3) & data_0708$food >0), ])} out of the \Sexpr{nrow(data_0708[which(data_0708$food>0), ])} countries in the sample with data on per capita expenditure levels spend less than that on food.
+Table \ref{tab:rfpi-pt-regressions} in the appendix contain regression results based on estimated long run RFPI transmission rates from country specific time series models with 6 lags.
+The signs of the coefficients are similar to those in table \ref{tab:fpi-pt-regressions} but most effects are statistically insignificant.
+Similarly, table \ref{tab:fpi-pt1-regressions} and \ref{tab:rfpi-pt1-regressions} in the appendix contain results based on FPI and real FPI transmission rates calculated from time series models with a single lag.
+These also point to only a few significant predictors of price transmission.
+A plausible reason is that the LRM estimates based on models with a single lag are relatively more noisy and therefore more difficult to explain.
 
-Column (6) focuses on the relationship between average food expenditure and expected real FPI growth in the 2007-8 period.
-Again there is a significant quadratic relationship between the two and it is remarkably similar to the specification based on FPI inflation in column (3).
-For countries with per capita food expenditures less than \Sexpr{round(exp(tp6),1)} PPP adjusted dollars (\Sexpr{nrow(data_0708[which(data_0708$food< exp(tp6) & data_0708$food >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$food>0), ])} countries) there is a positive estimated relationship between food expenditure and real FPI growth.
-For countries with food expenditures higher than that, the relationship becomes negative.
+\subsubsection{A panel model of short run price transmission determinants}
+The regression results in table \ref{tab:07-08-fc-regressions} and \ref{tab:fpi-pt-regressions} are based on a small number of observations which decreases the precision of the estimates.
+Furthermore, the estimates are potentially subject to unobserved heterogeneity bias.
+The last set of regressions therefore exploits the panel structure of the dataset to control for time invariant unobserved heterogeneity which potentially confounds the estimated relationships.
+Specifically the following model is fitted on monthly price data covering the full 2005-15 period
+\begin{equation}
+\pi_{it}^{dom}=\beta_{0}+
+\sum_{s=1}^{6}\rho_{s}\pi_{it-s}^{dom}+
+\sum_{s=0}^{1}(\lambda_{s}\pi_{t-s}^{int}+\phi_{s}er_{it-s})+
+\beta_{1}\pi_{t-1}^{int}\cdot x_{i} +
+u_{it}\label{eq:panel_model},
+\end{equation}
+where $x_{i}$ is the vector of country characteristics.
 
-<<echo=FALSE>>=
-<<data_0708>>
-gdpsort = data_0708[order(data_0708$GDP), ]
-@
+It is well known that dynamic panel models such as equation (\ref{eq:panel_model}) suffers from (Nickell) bias due to correlation between the lagged dependent variable and the error term.
+However, as demonstrated by \citet{nickell81}, the bias of the within estimator decreases at a rate $1/T$, i.e.
+endogeneity is mainly a problem in short panel datasets with only a few time periods $T$.
+In the present case where $T>100$, the bias should be negligible.
 
-The GDP and consumption coefficients in column (4) and (5) are also statistically significant and the conclusion from these models is similar to one based on (3) and (6).
-The GDP model in column (4) predicts that for countries with a per capita GDP below \Sexpr{round(exp(tp4),1)} PPP adjusted dollars (\Sexpr{nrow(data_0708[which(data_0708$GDP< exp(tp4) & data_0708$GDP >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$GDP>0), ])} countries) there is a positive relationship between real FPI increases and income and vice versa.
-The consumption expenditure model (5) predicts a turning point of \Sexpr{round(exp(tp5),1)} PPP adjusted dollars where \Sexpr{nrow(data_0708[which(data_0708$cons< exp(tp5) & data_0708$cons >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$cons>0), ])} countries have lower average consumption than that.
-Number \Sexpr{nrow(data_0708[which(data_0708$GDP< exp(tp4) & data_0708$GDP >0), ])} country in the sample in terms of GDP is Nicaragua.
-Countries in the lower middle income group  therefore generally experienced the largest food increases during the 2007-8 food crisis, whereas low and high income countries were affected less, once we control for differences in consumption and trade patterns.
-
-Countries with higher cereal consumption shares and import dependency ratios experienced larger food price increases as expected although $cshare$ and $cdep$ are not individually significant in most specifications.
-The negative coefficient on the interaction term means that the effect of $cdep$ on food price growth is decreasing in $cshare$ and vice-versa, but this mitigating effect is negligible.
-
-\begin{center}
-[Table \ref{tab:2010-11-regressions} around here]
-\par\end{center}
-
-Table \ref{tab:2010-11-regressions} summarizes the results from a set of regressions based on the same explanatory variables as those from table \ref{tab:2007-8-regressions}.
-The difference is that the dependent variables are domestic FPI inflation and real FPI growth during the 2010:1-2010:12 period.
-As can be seen, the results are similar to those in table \ref{tab:2007-8-regressions}, cf.
-also table \ref{tab:tp} .
-
-\subsection{Price transmission\label{sec:analysis2}}
-Although the results in section \ref{sec:analysis1} provide insights into the cross sectional variation in domestic price changes during the two recent international food price spikes they have at least two shortcomings.
-First, none of the estimated relationships represent price transmission as such.
-Secondly, the estimates are based on two specific time periods, which means that we ignore much the information contained in the individual price series.
-To accommodate these two shortcomings I estimate country specific price transmission coefficients and regress these on the same income and expenditure predictors as in section \ref{sec:analysis1}.
-
-<<echo=FALSE>>=
-<<ts_pt>>
-tp1 = -coef(m1)["log(GDP)"]/(2*coef(m1)["I(log(GDP)^2)"])
-tp2 = -coef(m2)["log(cons)"]/(2*coef(m2)["I(log(cons)^2)"])
-tp3 = -coef(m3)["log(food)"]/(2*coef(m3)["I(log(food)^2)"])
-tp4 = -coef(m4)["log(GDP)"]/(2*coef(m4)["I(log(GDP)^2)"])
-tp5 = -coef(m5)["log(cons)"]/(2*coef(m5)["I(log(cons)^2)"])
-tp6 = -coef(m6)["log(food)"]/(2*coef(m6)["I(log(food)^2)"])
-@
-
-Specifically, for each country, I regress its monthly domestic food price changes on its lagged domestic food price changes as well as on contemporaneous and lagged international food price changes.
-That is, as in the the \cite{IMF2011} study, I estimate country specific autoregressive distributive lag (ADL) models of the type
-\begin{equation}
-\pi_{t}^{dom}=\delta_{0}+\sum_{s=1}^{S_{y}}\rho_{s}\pi_{t-s}^{dom}+\sum_{s=0}^{S_{f}}\lambda_{s}\pi_{t-s}^{int}+\epsilon_{t}\label{eq:time_series_model}
-\end{equation}
-where $\pi_{t}^{dom}=\Delta p_{t}$ represents monthly FPI inflation or real FPI growth at time $t$ (both approximated by the first differences to the log of FPI or real FPI) and $\pi_{t}^{int}$ is the FAO food price index at time $t$, also in log differences.
-Using first differences instead of price levels is a way to avoid the spurious regression problem associated with trending variables.
-In order to keep things simple, maximum lag lengths are uniformly set to $S_{y}=S_{f}=6$.
-
-Next I calculate the long run effects of changes in the the FAO food price index on domestic food inflation and real FPI growth, the first of which is a standard measure of food price transmission.
-The long run multiplier is given by
-\begin{equation}
-\theta=\frac{\sum_{s=0}^{S_{f}}\lambda_{s}}{1-\sum_{s=1}^{S_{y}}\rho_{s}}.
-\end{equation}
-Finally, I regress the price transmission estimates, $\hat{\theta}$, on the same explanatory variables as in table \ref{tab:2007-8-regressions} and \ref{tab:2010-11-regressions}.
-That is, I estimate the model
-\begin{equation}
-\hat{\theta}_{i} =\delta_{0} + \delta_{1}x_{i} + \delta_{2}x_{i}^{2} + \delta_{3}z_{i} + u_{i}\label{eq:ptreg}.
-\end{equation}
-The results from this exercise are summarized in table \ref{tab:lri-regressions} and \ref{tab:tp}.
-As table \ref{tab:lri-regressions} shows, we find the same inverse-U shaped estimated relationships between income/expenditure and price transmission as we did with income/expenditure and domestic price changes in table \ref{tab:2007-8-regressions} and \ref{tab:2010-11-regressions}.
-For the specifications with significant income and expenditure coefficients, i.e.
-(1)-(5), the estimated turning points again imply that roughly one third of the countries are on the increasing part of the curve, cf.
-\ref{tab:tp}.
-
-\begin{center}
-[Table \ref{tab:lri-regressions} around here]
-\par\end{center}
-
-<<echo=False>>=
-<<panel_pt>>
-tp1 = -m1$coefficients[8]/(2*m1$coefficients[9])
-tp2 = -m2$coefficients[8]/(2*m2$coefficients[9])
-tp3 = -m3$coefficients[8]/(2*m3$coefficients[9])
-tp4 = -m4$coefficients[8]/(2*m4$coefficients[9])
-tp5 = -m5$coefficients[8]/(2*m5$coefficients[9])
-tp6 = -m6$coefficients[8]/(2*m6$coefficients[9])
-@
-
-The findings discussed above all point to the conclusion that middle income countries are the most vulnerable to international food price spikes.
-However, the estimates in table \ref{tab:2007-8-regressions}-\ref{tab:lri-regressions} are based on standard OLS and are potentially subject to unobserved heterogeneity bias.
-As a last robustness check I therefore exploit the panel structure of the dataset to control for time invariant unobserved heterogeneity which potentially confounds the income price transmission relationship.
-Results in table \ref{tab:panel-regressions} are based on within estimation of the model
-\begin{equation}
-\pi_{it}^{dom}=\delta_{0}+\sum_{s=1}^{6}\rho_{s}\pi_{it-s}^{dom}+\lambda_{1}\pi_{it-6}^{int}+\lambda_{2}\pi_{it-6}^{int}\cdot x_{i} +\lambda_{3}\pi_{it-6}^{int}\cdot x_{i}^{2}+u_{it}\label{eq:panel_model},
-\end{equation}
-where the variables, $\pi_{it}^{dom}$, $\pi_{it}^{int}$ and $x_{i}$ are the same as in \ref{eq:time_series_model} and \ref{eq:ptreg}.
-For simplicity, only a single lag of the FAO price index is included in the model and the lag length is chosen to be 6.
-The parameters of interest are $\lambda_{2}$ and $\lambda_{3}$ in front of the two interaction terms.
-A positive $\lambda_{2}$ and a negative $\lambda_{3}$ would, once more, imply an inverse-U shaped relationship between income/expenditure and price transmission as predicted by the simple model from section \ref{sec:model}.
-Again, in order to avoid the spurious regression problem associated with unit root processes the model is based on inflation rates rather than price levels.
+Estimation results based on the within estimator are shown in table \ref{tab:panel-regressions}.\footnote{Note that coefficients on the lagged dependent variable are omitted.}
 
@@
 
-Model (1),(3), (4) and (5) in table \ref{tab:panel-regressions} lead to significant quadratic expenditure price transmission relationships and the turning points based on these specifications are similar but somewhat higher than the ones from table \ref{tab:2007-8-regressions}-\ref{tab:lri-regressions}, with up to around 50 percent the countries on the increasing part of the income-price transmission curve, c.f table \ref{tab:tp}.
+Regarding the specification of the model, the main effect of $\pi_{t-s}^{int}$ on $\pi_{it}^{dom}$ was found to be at lag $s=1$ and this is also the reason why the interactions terms $\pi_{t-1}^{int}\cdot x_{i} $ refer to the one period lagged effect.
+Consequently, the set of coefficients to the interaction terms represent factors affecting short run price transmission.\footnote{Results were found to be robust to the exact number of included lags of $\pi_{it}^{dom}$.}
 
-\begin{center}
-[Table \ref{tab:tp} around here]
-\par\end{center}
+As can be seen, the signs of the estimated coefficients on the interaction terms are mostly consistent with the estimates in table \ref{tab:07-08-fc-regressions} and \ref{tab:fpi-pt-regressions}.
+Income has a significant inverse-U shaped effect on short run FPI transmission whereas the grain consumption ratio and landlockedness have significant effects, positive and negative, respectively, on short run RFPI transmission.
+Finally, countries with grain state trading enterprises are found to have significantly lower short run (nominal as well as real) food price transmission.
+
+<<echo=false>>=
+load("tp_ctr_mean.Rda")
+load("tp_ctr_mean1.Rda")
+@
+
+\subsection{Estimated income-price transmission turning points}
+Table \ref{tab:tp} summarizes the estimated income turning points derived from the significant coefficients in table \ref{tab:07-08-fc-regressions}-\ref{tab:panel-regressions}.
+The ratios in the last column refer to the number of countries with a per capita GDP less than the estimated turning point shown in the second-last column, relative to the total number of countries in the sample with data on income.
+The average share of countries that are to the left of the turning point, across all specifications, is \Sexpr{round(nc.pc.mean[1],0)} percent.
+There are, however, important differences between the different estimators.
+Panel estimation, in particular, leads to a significantly higher proportion of countries to the left of the turning point than cross section estimation (\Sexpr{round(nc.pc.mean[2],0)} vs.
+\Sexpr{round(nc.pc.mean[3],0)} percent for the non-panel estimators) on average.
+In constant 2011 international dollars, the average income level in 2014 among the group of low-, lower middle- and upper middle income countries was \$1,517, \$5,752 and \$13,719, respectively \citep{worldbank16b}.\footnote{According to the World Bank classification scheme.
+MICs are defined as having a 2011 per capita gross national income of US\$1,026 to \$12,475, where the conversion from local currency to US dollars is based on the Atlas method.} In comparison, the income level leading to the highest predicted price transmission, averaged across all specifications, is \$\Sexpr{format(round(tp.mean[1],0), big.mark= ",", scientific= FALSE)}.
+Countries from the lower end of the high-income spectrum experience the highest short run food price transmission whereas lower middle income countries experience the highest long run food price transmission according to these estimates.

### 2016-03-02 conclusion.Rnw
Conclusion section of 2. round version of paper.

@@
-% !Rnw root = Master.Rnw
+% !Rnw root = master.Rnw
 
 \section{Conclusion\label{sec:conclusion}}
-This paper contributes to the literature on aggregate food price transmission with three findings.
-First, domestic food prices in middle income countries react the strongest to international agricultural commodity price changes once we control for differences in consumption and trade dependency patterns between high and low income countries.
-Secondly, domestic price increases were the largest in middle income countries during the two recent international price spikes in 2007-08 and 2010-11, all else equal.
-Thirdly, there is a simple economic logic behind this inverse-U shaped income-price transmission relationship.
-These findings are worrying due to two other stylized facts.
-The first is that middle income countries are now home to around three quarters of the world’s poor.
-The second is that poor households spend a large portion of their income on food and are therefore vulnerable to sudden increases in the price of food goods.
-There is, however, also a positive side to this finding.
-If, as some authors claim, middle income countries, in principle, have the domestic financial resources to eradicate extreme poverty \citep{Sumner2013}, then they should also possess the necessary resources to implement social safety nets which shield purchasing power of the poor from occasional price spikes on the international markets.
+This paper contributes to the literature on aggregate food price transmission with an economic analysis and a number of empirical findings.
+Specifically, the paper examines how factors related to trade, consumption, income, geography and infrastructure affects a country's domestic food price response to international agricultural commodity price fluctuations.
+A novelty of the economic analysis, is that it predicts an inverse-U shaped relationship between a country's income level and its domestic food price responsiveness to an international shock, i.e.
+the degree of price transmission.
 
-The analysis has at least two major policy implications.
-First, when the next surge occurs on the international agricultural markets, the international donor community should direct its resources towords countries in which poor households are particularly vulnerable to international food price spikes and where the government is unable to to provide sufficient assistance, rather than the LICs in general.
-This could be LICs which, for some reason, has to import most of its food goods.
-Secondly, the strong focus in the media and among agricultural economists on international food crises is misguided in the sense that LICs, in general, are primarily vulnerable to domestic rather than international shocks.
+Three different approaches involving cross-sectional, time series and panel econometrics are used to quantify the importance of these country characteristics on the degree of price transmission.
+The first approach consists of a cross country regression analysis of domestic food price changes during the 2007-08 food crisis.
+Results show that countries with large cereal import shares experienced significantly lower food inflation on average during the food crisis than other countries whereas cereal exporters and more open countries in general experienced more food inflation on average.
+Various types of government interventions intended to stabilize domestic food prices during the food crisis did not have a significant effect, but countries with an agricultural state trading enterprise (STE) experienced significantly lower domestic food price increases in 2007-08.
+Finally, per capita income is found to have a significant inverse-U shaped effect on domestic food inflation during the food crisis.
+
+The second approach employs country specific time series models to estimate the long run relationship between international and domestic food prices.
+These price transmission estimates are then subsequently regressed on the set of explanatory variables representing different country characteristics.
+As it turns out, the regression results based on estimated price transmission rates are very similar to those based on domestic food inflation during the food crisis.
+The only major difference is that countries with large cereal consumption shares are found to have significantly lower price transmission on average.
+As with the negative coefficient to the cereal import share, this is a surprising finding which could indicate that governments in these countries intervene more in order to compensate for the exposure to world market fluctuations.
+
+In the third and final approach the determinants of short run price transmission are estimated directly with a dynamic panel model.
+Results indicate that short run food price transmission is affected by the many of the same factors as as long run transmission and in the same manner.
+I conclude that countries from the middle of the income spectrum generally have the highest food price transmission rates once we control for factors that are commonly believed to affect price transmission.
+This is a worrying finding in a way, given the fact that nearly three quarters of the world's poor now live in middle-income countries \citep{worldbank16, Sumner2012, Kanbur2012}.
+
+The policy implications of these findings is that focus should be on low-income households in middle-income countries during an international food price surge, since these appear to be the most vulnerable in such situations.
+Poor households in low-income countries, on the other hand, are especially vulnerable domestic shocks according to the economic model.
+This prediction is left open for future research though.

### 2016-03-02 cs_infl_regressions.R
Revised inflation regression script. Used for 2. round rev. of paper

@@ -3,258 +3,302 @@ library(lmtest)
 library(sandwich)
 library(dplyr)
 library(relaimpo)
+library(MASS)
 
 ##################################################################
 ### Cross section inflation regressions
 ##################################################################
 
-load("ctr_dat_0708.Rda")
-load("ctr_dat_1011.Rda")
+load("infl_ctr_dat_0708.Rda")
+load("infl_ctr_dat_1011.Rda")
 
-## FPI inflation regressions
-fpi.reg.m = list()
-fpi.reg.se = list()
+boxplot(ctr.dat.0708$fpigr.0708)
+boxplot(ctr.dat.0708$fpigr.0708[ctr.dat.0708$fpigr.0708<3.5*sd(ctr.dat.0708$fpigr.0708)])
 
+boxplot(ctr.dat.0708$rfpigr.0708)
+boxplot(ctr.dat.0708$rfpigr.0708[ctr.dat.0708$rfpigr.0708<3.5*sd(ctr.dat.0708$rfpigr.0708)])
+
+boxplot(ctr.dat.1011$fpigr.1011)
+boxplot(ctr.dat.1011$fpigr.1011[ctr.dat.1011$fpigr.1011<3.5*sd(ctr.dat.1011$fpigr.1011)])
+
+boxplot(ctr.dat.1011$rfpigr.1011)
+boxplot(ctr.dat.1011$rfpigr.1011[ctr.dat.1011$rfpigr.1011<3.5*sd(ctr.dat.1011$rfpigr.1011)])
+
+## Food crisis regressions
 # 2007-08
 data = ctr.dat.0708
+data1 = na.omit(dplyr::select(data, fpigr.0708, rfpigr.0708, ergr.0708,
+                              gdp.const.int, cshare, cir, cer, 
+                              lldc, ste, to, dtf, lpi.infra))
 
-m = lm(fpigr.0708~log(gdp)+
-         I(log(gdp)^2)+
-         cshare+
-         cdep+
-         cer +
+fc.reg0708.m = list()
+fc.reg0708.se = list()
+
+m = lm(scale(fpigr.0708)~scale(ergr.0708) +
+         log(gdp.const.int)+
+         I(log(gdp.const.int)^2)+
+         scale(cshare)+
+         scale(cir)+
+         scale(cer) +
          lldc+
-         to+
-         dtf+
-         lpi,
-       data = filter(data, fpigr.0708<3.5*sd(fpigr.0708)))
+         ste +
[…]

### 2016-03-02 cs_pt_reg.R
Revised LRM regression script. Used for 2. round rev. of paper

@@ -3,16 +3,20 @@ library(car)
 library(dplyr)
 library(sandwich)
 library(lmtest)
+library(MASS)
 
 ##################################################################
 ### Cross section price transmission regressions
 ##################################################################
 
-load("ctr_dat_0708.Rda")
+load("infl_ctr_dat_0708.Rda")
+load("infl_ctr_dat_1011.Rda")
 load("coef_bew_fpi_l1.Rda")
 load("coef_bew_rfpi_l1.Rda")
-load("coef_bew_fpi_l12.Rda")
-load("coef_bew_rfpi_l12.Rda")
+load("imp_se_fpi_l1_12.Rda")
+load("imp_se_rfpi_l1_12.Rda")
+load("coef_bew_fpi_l6.Rda")
+load("coef_bew_rfpi_l6.Rda")
 
 ## FPI LR multiplier regressions
 fpi.pt.reg.m = list()
@@ -20,21 +24,25 @@ fpi.pt.reg.se = list()
 fpi.bptest.stat = list()
 fpi.bptest.p = list()
 
-# 1 lag LRM
-data = merge(ctr.dat.0708, coef.bew.fpi.l1)
-data = rename(data, lrm1 = lrm, lrm1.se = lrm.se)
+# data = merge(ctr.dat.0708, coef.bew.fpi.l1)
+data = merge(ctr.dat.0708, coef.bew.fpi.l6)
+data = dplyr::rename(data, lrm1 = lrm, lrm1.se = lrm.se)
+data1 = na.omit(dplyr::select(data, lrm1, lrm1.se,
+                              gdp.const.int, cshare, cir, cer, 
+                              lldc, ste, to, dtf, lpi.infra))
 
 # OLS
-m = lm(lrm1~log(gdp)+
-         I(log(gdp)^2)+
-         cshare+
-         cdep+
-         cer+
+m = lm(scale(lrm1)~log(gdp.const.int)+
+         I(log(gdp.const.int)^2)+
+         scale(cshare)+
+         scale(cir)+
+         scale(cer) +
          lldc+
-         to+
-         dtf+
-         lpi,
-       data = filter(data, lrm1<3.5*sd(lrm1)))
+         ste +
+         scale(to)+
+         scale(dtf)+
+         scale(lpi.infra),
[…]

### 2016-03-02 data_prep.R
Revised version of data prep. file. Not used in 2. round rev. version of paper

@@ -1,257 +1,169 @@
-library(plyr)
+#####################################################################
+#### Creating cross section dataset
+#####################################################################
+
 library(dplyr)
 library(zoo)
-library(plm)
 library(foreign)
 library(reshape)
 library(reshape2)
-library(ggplot2)
-library(data.table)
-library(lattice)
-library(readstata13)
 library(countrycode)
+library(data.table)
+library(tempdisagg)
 
-### Data files used
-## Domestic price indexes and exchange rates
-# laborstaB9.csv
-# laborstaBA.csv
-# er.csv
-## Covariates
+### Raw Data
 # landlocked.csv
 # trade_openness.csv
 # ste.csv
 # dtf.csv
 # lpi.csv
-# psd_grain.csv
+# cereal_production.csv
+# cereal_supply.csv
+# cereal_import.csv
+# cereal_export.csv
 # gdp_05.cs
 # cons_05.csv
-# fao_fpi.csv
-
-#####################################################################
 
 str = "C:/Users/nqw235/Google Drev/Research/pov_price_transmission/"
 
-### The ILO data
-temp = paste(str, "laborstaB9.csv", sep = "")
-B9 = read.csv(temp, sep = ";")
-temp = paste(str, "laborstaBA.csv", sep = "")
-BA = read.csv(temp, sep = ";")
-
-## Reshape to long format
-B9$id = paste0(B9$CODE.COUNTRY, B9$CODE.AREA)
-B9[B9$id == "LKLK1" & B9$YEAR == 2002, ]
-B9 = B9[-2659, ] # coding mistake
-B9l = reshape(B9, varying=c(9:21), direction="long", v.names="cpi", 
-              idvar=c("id", "YEAR"),
-              sep="", timevar = "month")
-B9l = B9l[order(B9l$id, B9l$YEAR, B9l$month), ]
-B9l$SRC_ID = NULL
-B9l = rename(B9l, c("NOTES" = "cpi.notes"))
-
-BA$id = paste0(BA$CODE.COUNTRY, BA$CODE.AREA)
-BA[BA$id == "LKLK1" & BA$YEAR == 2002, ]
-BA = BA[-2629, ] # coding mistake
[…]

### 2016-03-02 data_prep.Rnw
Revised data preparation script. Used for 2. round version of paper

@@
 <<datasets, echo=F>>=
-library(readstata13)
+library(plyr)
+library(dplyr)
 library(zoo)
@@
 library(reshape)
+library(reshape2)
+library(ggplot2)
+library(data.table)
+library(lattice)
+library(readstata13)
+library(countrycode)
 
-ilo_data = read.dta13("ilodata_small_clean.dta")
-icp_data = read.dta13("icp_exp_2005.dta")
-icp_data$cfood = scale(icp_data$food)
-wb_pov = read.dta13("wb_pov.dta")
-wb_pov$cpovgap2 = scale(wb_pov$pov_gap_2)
-sids = read.dta13("sids_list_s.dta")
-fao_fpi = read.csv("fao_fpi.csv", sep = ";")
-load("staples.Rda")
-load("cdep.Rda")
-infl_0708 =  read.dta13("infl_0708.dta")
-infl_1011 =  read.dta13("infl_1011.dta")
+### Data files used
+# laborstaB9.csv
+# laborstaBA.csv
+
+#####################################################################
+
+str = "C:/Users/nqw235/Google Drev/Research/pov_price_transmission/"
+
+### The ILO data
+temp = paste(str, "laborstaB9.csv", sep = "")
+B9 = read.csv(temp, sep = ";")
+temp = paste(str, "laborstaBA.csv", sep = "")
+BA = read.csv(temp, sep = ";")
+
+## Reshape to long format
+B9$id = paste0(B9$CODE.COUNTRY, B9$CODE.AREA)
+B9[B9$id == "LKLK1" & B9$YEAR == 2002, ]
+B9 = B9[-2659, ] # coding mistake
+B9l = reshape(B9, varying=c(9:21), direction="long", v.names="cpi", 
+              idvar=c("id", "YEAR"),
+              sep="", timevar = "month")
+B9l = B9l[order(B9l$id, B9l$YEAR, B9l$month), ]
+B9l$SRC_ID = NULL
+B9l = rename(B9l, c("NOTES" = "cpi.notes"))
+
+BA$id = paste0(BA$CODE.COUNTRY, BA$CODE.AREA)
+BA[BA$id == "LKLK1" & BA$YEAR == 2002, ]
+BA = BA[-2629, ] # coding mistake
+BAl = reshape(BA, varying=c(9:21), direction="long", v.names="fpi", 
+              idvar=c("id", "YEAR"),
+              sep="", timevar = "month")
+BAl = BAl[order(BAl$id, BAl$YEAR, BAl$month), ]
+BAl$SRC_ID = NULL
+BAl = rename(BAl, c("NOTES" = "fpi.notes"))
+
+## Merge CPI and FPI datasets
+ilo = merge(B9l, BAl, all=TRUE)
+ilo = ilo[order(ilo$id, ilo$YEAR, ilo$month), ]
+ilo = rename(ilo, c("YEAR" = "year", "COUNTRY"="country", 
+                    "CODE.COUNTRY"="c.code", "CODE.AREA"="a.code"))
+
+ilo = ilo[ilo$month != 13, ]
+ilo$tt = as.yearmon(paste(ilo$year, ilo$month, sep = "-"))
+
+rm(B9, BA, B9l, BAl)
+
+## Identifying countries with multiple price series and select single series
+ilo$id = factor(ilo$id)
+str(ilo$id)
+summary(tabulate(ilo$id))
+table(ilo$id)
+table(table(ilo$id))
+ilo$c.code = factor(ilo$c.code)
+str(ilo$c.code)
+table(table(ilo$c.code))
+tbl = table(ilo$c.code)
+# subset of countries with > 1 price serie
+dat = droplevels(ilo[ilo$c.code %in% names(tbl)[tbl > 180],,drop=FALSE])
+table(dat$c.code)
+length(table(dat$c.code))
+
+# visual inspection of the series
+c = "MM"
+datsub = ilo[ilo$c.code == c, ]
+dat.m = melt(datsub,measure.vars=c('cpi','fpi'))
+ggplot(dat.m) +
+ geom_line(aes(x=as.numeric(tt),y=value,group=id,color=id)) +
+  facet_grid(~variable)
+
+# drop/keep (ir)relevant series
+ilo = ilo[ilo$id != "BRBR1", ]
+ilo = ilo[ilo$c.code != "BS", ]
+ilo = ilo[ilo$c.code != "CL", ]
+ilo = ilo[ilo$c.code != "CV", ]
+ilo = ilo[ilo$id != "ETET1", ]
+ilo = ilo[!(ilo$id %in% c("ININ4", "ININ6", "ININ7")), ]
+ilo = ilo[ilo$c.code != "KE", ]
+ilo = ilo[ilo$c.code != "LB", ]
+ilo = ilo[!(ilo$id %in% c("MGMG1", "MGMG2")), ]
+ilo = ilo[ilo$c.code != "MN", ]
+ilo = ilo[ilo$id != "MZMZ1", ]
+ilo = ilo[ilo$id != "NINI1", ]
+ilo = ilo[ilo$id != "OMOM1", ]
+ilo = ilo[ilo$id != "PAPA1", ]
+ilo = ilo[ilo$id != "RW", ]
+ilo = ilo[ilo$id != "SASA2", ]
+ilo = ilo[ilo$id != "SLSL1", ]
+ilo = ilo[ilo$id != "SN", ]
+ilo = ilo[ilo$id != "SYSY1", ]
+ilo = ilo[ilo$c.code != "VE", ]
+# Additional countries with multiple series
+ilo = ilo[ilo$id != "ZMZM2", ]
+ilo = ilo[ilo$c.code != "KN", ]
+ilo = ilo[ilo$c.code != "SZ", ]
+ilo = ilo[ilo$id != "MMMM1", ]
+
+ilo$id = factor(ilo$id)
+str(ilo$id)
+ilo$c.code = factor(ilo$c.code)
+str(ilo$c.code)
+rm(dat, dat.m, datsub)
+ilo$id = NULL
+ilo = rename(ilo, c("c.code" = "id"))
+
+## Selecting the gross sample of countries
+x = 180 -(3*12)
+x/12 # number of years required
+dat = droplevels(ilo[ilo$id %in% names(tbl)[tbl >= x],,drop=FALSE])
+dat$id = factor(dat$id)
+str(dat$id)
+# require comple series
+dat = dat %>% dplyr::filter(ave(!is.na(cpi), id, FUN = all))
+dat = dat %>% dplyr::filter(ave(!is.na(fpi), id, FUN = all))
+dat$id = factor(dat$id)
+str(dat$id)
+ilo = dat
+rm(dat)
+
+## Identifying and correcting for technical breaks, coding errors etc.
+DT = data.table(ilo)
+DT[, temp1 := 0]
+setkey(DT, id)
+DT[DT[, .I[1], by = key(DT)]$V1, temp1 := 2] 
+DT[DT[, .I[.N], by = key(DT)]$V1, temp1 := 1]
+DF = data.frame(DT)
+DF = within(DF, lag.cpi.notes <- c(NA, head(as.character(cpi.notes), -1)))
+DF = within(DF, lag.cpi <- c(NA, head((cpi), -1)))
+
+DF = within(DF, lag.fpi.notes <- c(NA, head(as.character(fpi.notes), -1)))
+DF = within(DF, lag.fpi <- c(NA, head((fpi), -1)))
+
+DF$cpi.brk = factor(ifelse(as.character(DF$cpi.notes)!=as.character(DF$lag.cpi.notes) & 
+                             is.na(DF$cpi.notes) == FALSE &
+                             is.na(DF$lag.cpi.notes) == FALSE &
+                             DF$temp1 !=2 &
+                             abs(DF$cpi/DF$lag.cpi-1) > 0.1, 
+                           1,0))
+DF$fpi.brk = factor(ifelse(as.character(DF$fpi.notes)!=as.character(DF$lag.fpi.notes) & 
+                             is.na(DF$fpi.notes) == FALSE &
+                             is.na(DF$lag.fpi.notes) == FALSE &
+                             DF$temp1 !=2 &
+                             abs(DF$fpi/DF$lag.fpi-1) > 0.1, 
+                           1,0))
+summary(DF$cpi.brk)
+summary(DF$fpi.brk)
+
+View(DF[which((DF$cpi.brk) == 1),])
+View(DF[which((DF$fpi.brk) == 1),])
+
+DF$adj.cpi = ilo$cpi
+DF$adj.fpi = ilo$fpi
+
+# fix coding errors
+DF[which(DF$id == "MD" & DF$year == 2007 & DF$month == 1), "cpi"] = 1195.14931/10
+
+# fix the structural breaks
+for (c in levels(DF$id)) 
+  {
+  temp = filter(DF, id == c)
+  temp$d.cpi = c(NA, diff(temp$cpi))
+  temp$x = ifelse(temp$cpi.brk == 1, temp$d.cpi, 0)
+  temp = within(temp, cs.cpi <- cumsum(x))
+  temp$adj.cpi = temp$cpi-temp$cs.cpi
+
+  temp$d.fpi = c(NA, diff(temp$fpi))
+  temp$y = ifelse(temp$fpi.brk == 1, temp$d.fpi, 0)
+  temp = within(temp, cs.fpi <- cumsum(y))
+  temp$adj.fpi = temp$fpi-temp$cs.fpi
+
+  DF[which(DF$id == c), c("adj.cpi", "adj.fpi")] = temp[, c("adj.cpi", "adj.fpi")]
+  }
+
+# visual inspection of the series with structural breaks
+c = "AM"
+datsub = DF[DF$id == c, ]
+dat.m = melt(datsub,measure.vars=c('cpi','fpi'))
+ggplot(dat.m) +
+ geom_line(aes(x=as.numeric(tt),y=value,group=id,color=id)) +
+  facet_grid(~variable)
+
+dat.m = melt(datsub,measure.vars=c('adj.cpi','adj.fpi'))
+ggplot(dat.m) +
+ geom_line(aes(x=as.numeric(tt),y=value,group=id,color=id)) +
+  facet_grid(~variable)
+
+View(DF[which((DF$id) == c),])
+
+## finishing touches
+ilo = select(DF, country:tt, cpi.brk, fpi.brk, adj.cpi, adj.fpi)
+rm(dat.m, datsub, DF, DT, temp)
+
+ilo = mutate(ilo, rfpi = adj.fpi/adj.cpi*100)
+
+# visual inspection of the series
+xyplot(rfpi ~ tt | id,
+       data = ilo[which(as.numeric(ilo$id)>=97 & as.numeric(ilo$id)<109), ],
+       type = "l",
+       col = c("black", "red"))
+
+# standardising country names
+ilo$iso3c = factor(countrycode(ilo$id, "iso2c", "iso3c", warn = TRUE))
+ilo = ilo[ilo$id != "T1", ]
+str(ilo$iso3c)
+ilo$id = factor(ilo$id)
+str(ilo$id)
+
+#####################################################################
+
+### Covariates
+temp = paste(str, "icp_exp_2005.dta", sep = "")
+icp_data = read.dta13(temp)
+
+temp = paste(str, "wb_pov.dta", sep = "")
+wb_pov = read.dta13(temp)
+
+temp = paste(str, "fao_fpi.csv", sep = "")
+fao_fpi = read.csv(temp, sep = ";")
+
+temp = paste(str, "staples.Rda", sep = "")
+load(temp)
+
+temp = paste(str, "cdep.Rda", sep = "")
+load(temp)
+
+temp = paste(str, "infl_0708.dta", sep = "")
+infl_0708 =  read.dta13(temp)
+
+temp = paste(str, "infl_1011.dta", sep = "")
+infl_1011 =  read.dta13(temp)
 
@@
 df1 = merge(df1, wb_pov, by = "iso3c", all.x=TRUE)
-df1 = merge(df1, sids, by = "iso3c", all.x=TRUE)
 df1 = merge(df1, staples, by = "iso3c", all.x=TRUE)
@@
 df2 = cbind(df1, fao_fpi)
-df2["sidsdum"][is.na(df2["sidsdum"])] <- 0
 data = plm.data(df2, index = c("id", "tt"))
@@
 data_0708 = merge(data_0708, wb_pov, by = "iso3c", all.x=TRUE)
-data_0708 = merge(data_0708, sids, by = "iso3c", all.x=TRUE)
 data_0708 = merge(data_0708, staples, by = "iso3c", all.x=TRUE)
 data_0708 = merge(data_0708, cdep, by = "iso3c", all.x=TRUE)
-data_0708["sidsdum"][is.na(data_0708["sidsdum"])] <- 0
 data_0708 = cbind(data_0708, fao_0708)
@@
 data_1011 = merge(data_1011, wb_pov, by = "iso3c", all.x=TRUE)
-data_1011 = merge(data_1011, sids, by = "iso3c", all.x=TRUE)
 data_1011 = merge(data_1011, staples, by = "iso3c", all.x=TRUE)
 data_1011 = merge(data_1011, cdep, by = "iso3c", all.x=TRUE)
-data_1011["sidsdum"][is.na(data_1011["sidsdum"])] <- 0
 data_1011 = cbind(data_1011, fao_1011)

### 2016-03-02 figures.Rnw
Revised figures script. Used for 2. round rev. version of paper

@@
 
-dat = read.csv("dist_trade.csv", sep = ";")
+## Import data into R.
+str = "C:/Users/nqw235/Google Drev/Research/pov_price_transmission/"
+temp = paste(str, "dist_trade.csv", sep = "")
+
+dat = read.csv(temp, sep = ";")
 dat$gamma = (dat$vop_covx+dat$vop_covm)/dat$vop_covt*100
@@
   xlab("FPI inflation 2007-08") +
-  geom_text(aes(label=ifelse(foodinfl78>40,as.character(iso3c),'')),hjust=1.2,just=-1.2, 
+  geom_text(aes(label=ifelse(foodinfl78>40,as.character(iso3c),'')),hjust=1.2,vjust=-1.2, 
             size=3) +
-  geom_text(aes(label=ifelse(foodinfl1011>20,as.character(iso3c),'')),hjust=1.2,just=-1.2,
+  geom_text(aes(label=ifelse(foodinfl1011>20,as.character(iso3c),'')),hjust=1.2,vjust=-1.2,
             size=3)
@@
   xlab("CPI inflation 2007-08") +
-  geom_text(aes(label=ifelse(cpiinfl78>30,as.character(iso3c),'')),hjust=1.2,just=-1.2, 
+  geom_text(aes(label=ifelse(cpiinfl78>30,as.character(iso3c),'')),hjust=1.2,vjust=-1.2, 
             size=3) +
-  geom_text(aes(label=ifelse(cpiinfl1011>15,as.character(iso3c),'')),hjust=1.2,just=-1.2,
+  geom_text(aes(label=ifelse(cpiinfl1011>15,as.character(iso3c),'')),hjust=1.2,vjust=-1.2,
             size=3)
@@
   xlab("Real FPI Growth 2007-08") +
-  geom_text(aes(label=ifelse(rfpigr78>12,as.character(iso3c),'')),hjust=1.2,just=-1.2, 
+  geom_text(aes(label=ifelse(rfpigr78>12,as.character(iso3c),'')),hjust=1.2,vjust=-1.2, 
             size=3) +
-  geom_text(aes(label=ifelse(rfpigr1011>10,as.character(iso3c),'')),hjust=1.2,just=-1.2,
+  geom_text(aes(label=ifelse(rfpigr1011>10,as.character(iso3c),'')),hjust=1.2,vjust=-1.2,
             size=3)

### 2016-03-02 intro.Rnw
Revised table script for 2. round revision version of paper

@@
-% !Rnw root = Master.Rnw
-
-\SweaveInput{data_prep.Rnw}
+% !Rnw root = master.Rnw
 
 \section{Introduction}
-Poor households spend a large proportion of their budget on food and are therefore vulnerable to increases in the domestic price of food \citep{Mellor1978}.
-Spurred by the increased volatility on the international commodity markets the last decade or so and more widely available household survey data, a large number of recent studies have quantified the poverty impact of an international food price spike for specific countries as well as for the world as a whole.\footnote{E.g.
-\cite{Attanasio2013,Balagtas2014,Coxhead2012,Dessus2008,Ferreira2013,Hoyos2011,Ivanic2012,Janvry2010,Mghenyi2011,Vu2011,Wodon2010}.
-There is also a special issue of \emph{Agricultural Economics} from 2008 devoted to this topic.}
+In response to the substantial increase in international food prices during the first decade of the new millennium, a large number of recent studies have quantified global and local poverty impacts of international price shocks.\footnote{E.g.
+\cite{Attanasio2013,Balagtas2014,Coxhead2012,Ferreira2013,Hoyos2011,Ivanic2012,Janvry2010,Mghenyi2011,Vu2011,Wodon2010}.
+In addition, a supplement to volume 39 of \emph{Agricultural Economics} is entirely devoted to this topic \citep[see][]{masters08}.} A major determinant of a country's vulnerability to such shocks is the responsiveness of its domestic food prices \citep[e.g.][]{Hoyos2011}.
+But what are the main determinants of the domestic price responsiveness to international shocks?
+Are they related to trade, consumption, geography or something else?
+As the title of the paper suggests, a country's income level seems to be an important determinant but other factors also play a role.
+In particular, countries which control their international trade through an agricultural state trading enterprise are found to experience lower food price transmission on average in the short run as well as in the long run.
 
-At the aggregate level, food insecurity resulting from international grain price spikes is mainly seen as a problem for Low Income Countries (LICs) \citep[e.g.][]{Ivanic2008}.
-However, a country's vulnerability to international food price spikes not only depends on its income distribution around the poverty level.
-Another fundamental determinant of aggregate vulnerability is the extent to which international price spikes are transmitted to domestic consumer markets.
-As I argue in the paper, price transmission varies systematically with a country's income level.
-In a nutshell, low income households are the most vulnerable to domestic food price increases but domestic food prices in Middle Income Countries (MICs) rather than LICs or High Income Countries (HICs) are the most responsive to international food price changes.
+The literature on domestic price impacts of international commodity price shocks is vast and the topic has experienced a revival in the wake of the 2007-08 food crisis \citep[see][]{Fackler2001, kouyate16}.
+However, since the majority of price transmission studies focus on a single country we know little about the determinants, other than distance and borders, of price transmission differences across countries.
+Moreover, most price transmission studies focus on individual commodity prices rather than food prices in general.
+This paper therefore contributes with a cross country analysis of price transmission determinants that is based on aggregate food price data, i.e.
+food price indexes, rather than prices of individual agricultural commodities.
 
-<<echo=FALSE>>=
-<<infl>>
-n=nrow(infl)
-@
+Regarding the choice between individual or aggregate prices, each approach has its strengths and weaknesses \citep{headey10}.
+Commodity specific analyses have a stronger microeconomic foundation in the law of one price but results do not generalize to other food goods or even to the same food good traded in other domestic markets.
+Transmission estimates based on aggregate price data, on the other hand, give a more comprehensive picture and therefore better reflect domestic welfare costs of international food price fluctuations.
+A potential weakness though of the latter approach, is that a country's food price index might not reflect the composition of poor people's food baskets which is often based mainly on grain.
+Another weakness is that the composition of a country's food price index differs substantially from that of the international food price indexes and as well as from other countries' price indexes, which makes it more difficult to compare transmission rates across countries.
+These idiosyncrasies do not, however, preclude us from studying broad patterns in price transmission rates.
 
-This conclusion stems from an empirical price transmission analysis based on a large dataset of monthly consumer price indexes from \Sexpr{n} countries covering the period January 2005 to December 2010.
-My main finding is a significant inverse-U shaped relationship between income or expenditure and price transmission, implying that food prices in MICs, ceteris paribus, respond the strongest to international price changes.
-This result is robust to the specific choice of estimator and dependent variable and it also holds for the actual domestic price changes in connection with the food crises of 2007-08 and 2010-11.
+The empirical analysis is based on insights derived from a simple economic model that is more suitable for aggregate price data than the law of one price workhorse model.
+A key prediction is that a country's income level should have an independent effect on price transmission, in addition to factors related to trade openness, geography and so on.
+The model predicts a novel inverse-U shaped relationship between a country's income level and its price transmission rate.
+This income-price transmission relationship turns out to be a robust feature of the data.
 
-The specification of the regression function is derived from a simple theoretical model of a country's food price index (FPI).
-The model makes two fairly uncontroversial assumptions about the farm share of consumer food prices and share of tradable commodities in the food basket; 1) that the farm share of consumer food prices is decreasing in income and 2) that the  share of tradable commodities is increasing in income.
-These two assumptions combined imply that the model predicts a domestic FPI response to a change in a price index of internationally traded food goods which follows an inverse-U shaped relationship with per capita income or expenditure.
-Two additional findings from the analysis are 1) domestic agricultural supply shocks are expected to have a larger impact on food prices in LICs than in MICs and 2) LICs which rely heavily on on imports of agricultural commodities are the most vulnerable to an international price spike.
+There are also other significant predictors of the domestic responsiveness to international food price shocks.
+Countries which import a large share of their grains and countries whose diets are mainly based on cereals are, surprisingly, found to have lower price transmission on average whereas the opposite holds for countries which export most of their grain.
+Countries that are more open in general in terms of their trade to GDP ratio also experience more price transmission whereas countries with an agricultural state trading enterprise experience lower price transmission.
+These country characteristics also significantly explain differences in domestic food inflation during the 2007-08 food crisis.
 
-My contribution to the literature consists of a theory based analysis of international agricultural commodity to domestic consumer food price transmission, with an emphasis on its dependence on average income.
-Most of the papers from the small existing literature on the cross sectional pattern of food price transmission do not base their analysis on model predictions.
-Rather they take a more agnostic approach and consider factors which simple economic reasoning suggests influence pass-through rates (see section \ref{sec:litrev}).
-There are, however, a couple of recent papers from the poverty impact literature which consider the role of marketing margins specifically.
-Two very similar papers by \cite{Dawe2014} and \cite{Minot2015} analyze how assumptions about marketing margins on food goods affect the poverty impact from a price increase in the \cite{Deaton1989} framework.
-Both papers conclude that, assumptions about marketing margins matters for the poverty impact of price changes.
-Specifically, assuming equal percentage changes for both farm and consumer prices (proportional margin) leads to an exaggerated or upwards biased impact of higher food prices compared to a situation with fixed margins, which might be a more appropriate assumption.
-\cite{Yang2015} is another recent paper analyzing how assumptions about marketing margins affects price pass-through and subsequently welfare in a CGE framework.
-Their conclusion is that also CGE models tend to be overestimate price pass-through particularly for high income countries, when domestic margin services are not explicitly taken into account.
-
-The paper also contributes to a broader policy discussion of the appropriate response to international food crises and its poverty impact.
-Thanks to the work of Andy Sumner and others, it has become clear that most of the world's poor do not actually live in low income countries \citep{Sumner2012, Kanbur2012}.
-Nearly three quarters of the world's poor now live in MICs according to World Bank estimates.\footnote{See \url{http://www.worldbank.org/en/country/mic/overview}.
-MICs are defined as having a 2011 per capita gross national income of US\$1,026 to \$12,475, where the conversion from local currency to US dollars is based on the Atlas method.} This is in a sense a worrying fact given the finding that food prices in MICs are more responsive to international price changes than those in LICs, ceteris paribus.
-On the other hand, MIC governments are better able than LICs to protect its households from a significant fall in real income.
+What are the policy implications of these findings?
+One obvious implication is that we may have to revise somewhat our perception of vulnerability and country focus in connection with international food crises.
+Donors and international aid organizations tend to a focus mostly on low-income food deficit countries but, according to the estimates in this paper, it seems that poor households living in middle-income countries are actually the most vulnerable to an international food price spike.
+Moreover, it does not seem to be the case that international price changes in general are passed on at a higher rate to countries that import most of their grains and this was not the case either during the 2007-08 food crisis.
 
 The paper now proceeds as follows.
-Section \ref{sec:litrev} provides a brief review of the relevant price transmission literature.
-Section \ref{sec:model} lays out the simple theory model of country's food price index and its response to an international price spike.
-Section \ref{sec:analysis} presents the data and tests the model's predictions regarding income and price transmission and, finally, section \ref{sec:conclusion} offers some concluding remarks.
+Section \ref{sec:model} discusses the determinants of price transmission and lays out a model of a country's food price index.
+Section \ref{sec:analysis} presents the empirical analysis and, finally, section \ref{sec:conclusion} offers some concluding remarks.

### 2016-03-02 tables.Rnw
Revised table script for 2. round revision version of paper

--- 2015-12-22 tables.Rnw
@@
-% !Rnw root = Master.Rnw
+% !Rnw root = master.Rnw
 
-\SweaveInput{regressions.Rnw}
+\section*{Tables}
 
-\section{Tables}
+<<results=tex,echo=F>>=
+library(stargazer)
+reg.list = readRDS("fc_reg0708_m.rds")
+se.list = readRDS("fc_reg0708_se.rds")
 
-<<echo=FALSE, results=tex>>=
-library(stargazer)
-sumdat = merge(infl, icp_data[,-5], by = "iso3c", all.x=TRUE)
-sumdat = merge(sumdat, staples, by = "iso3c", all.x=TRUE)
-sumdat = merge(sumdat, cdep, by = "iso3c", all.x=TRUE)
-colnames(sumdat)=c("iso", "FPI inflation 07-08", "CPI inflation 07-08", "Real FPI growth 07-08", "FPI inflation 10-11", "CPI inflation 10-11", "Real FPI growth 10-11", "GDP", "cons", "food", "cshare", "cdep")
-
-stargazer(sumdat, digits=1, title='Summary statistics', label = "tab:sumstat", 
-          notes=c("Notes: 07-08 and 10-11 refer to the periods 2007:1-2008:6 and 2010:1-2010:12,", 
-                  "respectively. `GDP, `cons' and `food' refer to per capita income, -consumption",
-                  "and -food expenditure respectively in the year 2005. `cshare' refers to a coun-",
-                  "try's percentage dietary energy supply derived from cereals, roots and tubers.", 
-                  "`cdep' refers to a country's cereal import dependency ratio, both with respect",
-                  "to the years 2004-06. Source: ILO, World Bank, FAO. Own calculations"),
+stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
+          title='Determinants of domestic food price changes 2007-8', 
+          label = "tab:07-08-fc-regressions", 
+          dep.var.labels=c("FPI inflation", "RFPI growth"),
+          no.space=TRUE,
+          se = list(se.list[[1]], se.list[[2]], se.list[[3]], se.list[[4]]), 
+          keep.stat = c("rsq", "n", "aic"),
+          font.size = "small", 
+          covariate.labels=c("$\\Delta ER$",
+                             "$GDP$",
+                             "$GDP^{2}$", 
+                             "$CCONS$", 
+                             "$CIMP$", 
+                             "$CEXP$",
+                             "$LLDC$", 
+                             "$STE$",
+                             "$TO$", 
+                             "$DTF$", 
+                             "$LPI$"),
+          notes=c("*, ** and *** signify significance at the 10, 5 and 1 percent",
+                  "levels, respectively. Robust standard errors in parentheses;", 
+                  "samples exclude outliers defined as observations $\\pm$ 3.5 stan-", 
+                  "dard deviations from the mean of the dependent variable.", 
+                  "A description of the variables and their sources can be found", 
+                  "in table \\ref{tab:cs_data} and section \\ref{sec:analysis}."),
           notes.append=FALSE,
@@
 <<results=tex,echo=F>>=
-<<dom_infl_0708>>
-stargazer(m1, m2, m3, m4, m5, m6, title='Income related determinants of domestic food price changes 2007:1-2008:6', label = "tab:2007-8-regressions", dep.var.labels=c("FPI inflation","Real FPI growth"), no.space=TRUE, se = list(rse1, rse2, rse3, rse4, rse5, rse6), omit.stat = c("adj.rsq","f", "ser"), font.size = "footnotesize", covariate.labels=c("$log(GDP)$", "$log(GDP)^2$", "$log(cons)$", "$log(cons)^2$", "$log(food)$", "$log(food)^2$", "$cshare$", "$cdep$", "$cshare \\times cdep$"),
-          notes=c("Robust standard errors are reported in parentheses underneath the coefficients.",
-                  "Significance at p = 0.1, p = 0.05 and p = 0.01 are denoted by `*', `**' and `***'", 
-                  "respectively"),
+library(stargazer)
+reg.list = readRDS("fpi_pt_reg_m.rds")
+se.list = readRDS("fpi_pt_reg_se.rds")
+bp.stat.list = readRDS("fpi_bptest_stat.rds")
+bp.p.list = readRDS("fpi_bptest_p.rds")
+
+stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
+          title='Determinants of long run FPI transmission rates', 
+          label = "tab:fpi-pt-regressions", 
+          dep.var.labels=c("FPI LRM (6 lags)"), 
+          column.labels = c("OLS", "WLS"),
+          column.separate = c(2, 2),
+          no.space=TRUE,
+          se = list(se.list[[1]], se.list[[2]], se.list[[3]], se.list[[4]]), 
+          omit.stat = c("adj.rsq","f", "ser"), 
+          font.size = "small", 
+          covariate.labels=c("$GDP$",
+                             "$GDP^{2}$", 
+                             "$CCONS$", 
+                             "$CIMP$", 
+                             "$CEXP$",
+                             "$LLDC$", 
+                             "$STE$",
+                             "$TO$", 
+                             "$DTF$", 
+                             "$LPI$"),
+          notes=c("*, ** and *** signify significance at the 10, 5 and 1 percent",
+                  "levels, respectively. Robust standard errors in parentheses;", 
+                  "samples exclude outliers defined as observations $\\pm$ 3.5 stan-", 
+                  "dard deviations from the mean of the dependent variable.", 
+                  "A description of the variables and their sources can be found", 
+                  "in table \\ref{tab:cs_data} and section \\ref{sec:analysis}."),
           notes.append=FALSE,
-          notes.align="l",
-          notes.label="Notes:")
-@
-
-<<results=tex,echo=F>>=
-<<dom_infl_1011>>
-stargazer(m1, m2, m3, m4, m5, m6, title='Income related determinants of domestic food price changes 2010:1-2010:12', label = "tab:2010-11-regressions", dep.var.labels=c("FPI inflation","Real FPI growth"), no.space=TRUE, se = list(rse1, rse2, rse3, rse4, rse5, rse6), omit.stat = c("adj.rsq","f", "ser"), font.size = "footnotesize", covariate.labels=c("$log(GDP)$", "$log(GDP)^2$", "$log(cons)$", "$log(cons)^2$", "$log(food)$", "$log(food)^2$", "$cshare$", "$cdep$", "$cshare \\times cdep$"),
-          notes=c("Robust standard errors are reported in parentheses underneath the coefficients.",
-                  "Significance at p = 0.1, p = 0.05 and p = 0.01 are denoted by `*', `**' and `***'", 
-                  "respectively"),
-          notes.append=FALSE,
-          notes.align="l",
-          notes.label="Notes:")
+          notes.align="l")
 @
@@
 <<results=tex, echo=FALSE>>=
-<<ts_pt>>
-stargazer(m1, m2, m3, m4, m5, m6, title='Long run multiplier (LRM) regressions', label = "tab:lri-regressions", dep.var.labels=c("LRM, FPI inflation","LRM, real FPI growth"), no.space=TRUE, se = list(rse1, rse2, rse3, rse4, rse5, rse6), omit.stat = c("adj.rsq","f", "ser"), font.size = "footnotesize", covariate.labels=c("$log(GDP)$", "$log(GDP)^2$", "$log(cons)$", "$log(cons)^2$", "$log(food)$", "$log(food)^2$", "$cshare$", "$cdep$", "$cshare \\times cdep$"),
-          notes=c("Robust standard errors are reported in parentheses underneath the coefficients.",
-                  "Significance at p = 0.1, p = 0.05 and p = 0.01 are denoted by `*', `**' and `***'", 
-                  "respectively"),
+library(plm)
+reg.list = readRDS("panel_reg_m.rds")
+se.list = readRDS("panel_reg_se.rds")
+
+stargazer(reg.list[[1]], reg.list[[3]],
+          title='Determinants of short run FPI transmission rates', 
+          label = "tab:panel-regressions", 
+          dep.var.labels=c("FPI inflation", "RFPI growth"),
+          single.row = TRUE,
+          se = list(se.list[[1]], se.list[[3]]),
+          omit = c("adj.fpi", "rfpi"),
+          omit.stat = c("adj.rsq","f", "ser"), 
+          font.size = "small",
+          covariate.labels=c("$\\pi_{t}^{i}$",
+                             "$\\pi_{t-1}^{i}$",
+                             "$\\Delta ER_{t}$",
+                             "$\\Delta ER_{t-1}$",
+                             "$\\pi_{t-1}^{i}\\times GDP$",
+                             "$\\pi_{t-1}^{i}\\times GDP^2$", 
+                             "$\\pi_{t-1}^{i}\\times CCONS$", 
+                             "$\\pi_{t-1}^{i}\\times CIMP$", 
+                             "$\\pi_{t-1}^{i}\\times CEXP$",
+                             "$\\pi_{t-1}^{i}\\times LLDC$", 
+                             "$\\pi_{t-1}^{i}\\times STE$", 
+                             "$\\pi_{t-1}^{i}\\times TO$", 
+                             "$\\pi_{t-1}^{i}\\times DTF$", 
+                             "$\\pi_{t-1}^{i}\\times LPI$"),
+          notes=c("*, ** and *** signify significance at the 10, 5 and 1",
+                  "percent levels, respectively. Robust clustered standard", 
+                  "errors in parentheses. A description of the variables and", 
+                  "their sources can be found in table \\ref{tab:cs_data} and section \\ref{sec:analysis}."),
           notes.append=FALSE,
-          notes.align="l",
-          notes.label="Notes:")
+          notes.align="l")
 @
 
-<<results=tex, echo=FALSE>>=
-<<panel_pt>>
-stargazer(m1, m2, m3, m4, m5, m6, 
-          title='Panel regressions', 
-          label = "tab:panel-regressions", 
-          font.size = "footnotesize", 
-          no.space=TRUE, 
-          se = list(rse1, rse2, rse3, rse4, rse5, rse6), 
-          omit.stat = c("adj.rsq","f"), 
-          dep.var.labels=c("$\\Delta log(FPI_{t})$","$\\Delta log(rFPI_{t})$"), 
-          omit = 1:12, 
-          covariate.labels=c("$\\Delta log(FAO_{t-6})$", "$\\Delta log(FAO_{t-6}) \\times log(GDP)$","$\\Delta log(FAO_{t-6}) \\times log(GDP)^2$", "$\\Delta log(FAO_{t-6}) \\times log(cons)$", "$\\Delta log(FAO_{t-6}) \\times log(cons)^2$", "$\\Delta log(FAO_{t-6}) \\times log(food)$", "$\\Delta log(FAO_{t-6}) \\times log(food)^2$"),
-          notes=c("Robust standard errors are reported in parentheses underneath the coef-",
-                  "ficients. Significance at p = 0.1, p = 0.05 and p = 0.01 are denoted by `*',", 
-                  " `**' and `***' respectively"),
-          notes.append=FALSE,
-          notes.align="l",
-          notes.label="Notes:")
-@
+<<echo=FALSE>>=
+load("infl_ctr_dat_0708.Rda")
+dat = ctr.dat.0708
+r1 = readRDS("fc_reg0708_m.rds")
+r2 = readRDS("fpi_pt_reg_m.rds")
+r3 = readRDS("panel_reg_m.rds")
 
-<<tpmat, echo=FALSE>>=
-tp = matrix(NA, nrow=6, ncol=4)
-
-<<dom_infl_0708>>
-tp[1,1] = -m1$coefficients[2]/(2*m1$coefficients[3])
-tp[2,1] = -m2$coefficients[2]/(2*m2$coefficients[3])
-tp[3,1] = -m3$coefficients[2]/(2*m3$coefficients[3])
-tp[4,1] = -m4$coefficients[2]/(2*m4$coefficients[3])
-tp[5,1] = -m5$coefficients[2]/(2*m5$coefficients[3])
-tp[6,1] = -m6$coefficients[2]/(2*m6$coefficients[3])
-
-<<dom_infl_1011>>
-tp[1,2] = -m1$coefficients[2]/(2*m1$coefficients[3])
-tp[2,2] = -m2$coefficients[2]/(2*m2$coefficients[3])
-tp[3,2] = -m3$coefficients[2]/(2*m3$coefficients[3])
-tp[4,2] = -m4$coefficients[2]/(2*m4$coefficients[3])
-tp[5,2] = -m5$coefficients[2]/(2*m5$coefficients[3])
-tp[6,2] = -m6$coefficients[2]/(2*m6$coefficients[3])
-
-<<ts_pt>>
-tp[1,3] = -m1$coefficients[2]/(2*m1$coefficients[3])
-tp[2,3] = -m2$coefficients[2]/(2*m2$coefficients[3])
-tp[3,3] = -m3$coefficients[2]/(2*m3$coefficients[3])
-tp[4,3] = -m4$coefficients[2]/(2*m4$coefficients[3])
-tp[5,3] = -m5$coefficients[2]/(2*m5$coefficients[3])
-tp[6,3] = -m6$coefficients[2]/(2*m6$coefficients[3])
-
-<<panel_pt>>
-tp[1,4] = -m1$coefficients[8]/(2*m1$coefficients[9])
-tp[2,4] = -m2$coefficients[8]/(2*m2$coefficients[9])
-tp[3,4] = -m3$coefficients[8]/(2*m3$coefficients[9])
-tp[4,4] = -m4$coefficients[8]/(2*m4$coefficients[9])
-tp[5,4] = -m5$coefficients[8]/(2*m5$coefficients[9])
-tp[6,4] = -m6$coefficients[8]/(2*m6$coefficients[9])
+tp = matrix(NA, nrow=6, ncol=1)
+tp[1,1] = -coef(r1[[1]])["log(gdp.const.int)"]/(2*coef(r1[[1]])["I(log(gdp.const.int)^2)"])
+tp[2,1] = -coef(r1[[3]])["log(gdp.const.int)"]/(2*coef(r1[[3]])["I(log(gdp.const.int)^2)"])
+tp[3,1] = -coef(r2[[1]])["log(gdp.const.int)"]/(2*coef(r2[[1]])["I(log(gdp.const.int)^2)"])
+tp[4,1] = -coef(r2[[3]])["log(gdp.const.int)"]/(2*coef(r2[[3]])["I(log(gdp.const.int)^2)"])
+tp[5,1] = -coef(r3[[1]])["I(lag(diff(log(fao.fpi)), 1) * log(gdp.const.int))"]/
+  (2*coef(r3[[1]])["I(lag(diff(log(fao.fpi)), 1) * I(log(gdp.const.int)^2))"])
+tp[6,1] = -coef(r3[[3]])["I(lag(diff(log(fao.fpi)), 1) * log(gdp.const.int))"]/
+  (2*coef(r3[[3]])["I(lag(diff(log(fao.fpi)), 1) * I(log(gdp.const.int)^2))"])
 
 tp = exp(tp)
+n.gdp = length(which(!is.na(dat$gdp.const.int)))
+
+nc = matrix(NA, nrow=6, ncol=1)
+nc[1,1] = nrow(dat[which(dat$gdp.const.int< tp[1,1] & !is.na(dat$gdp.const.int)), ])
+nc[2,1] = nrow(dat[which(dat$gdp.const.int< tp[2,1] & !is.na(dat$gdp.const.int)), ])
+nc[3,1] = nrow(dat[which(dat$gdp.const.int< tp[3,1] & !is.na(dat$gdp.const.int)), ])
+nc[4,1] = nrow(dat[which(dat$gdp.const.int< tp[4,1] & !is.na(dat$gdp.const.int)), ])
+nc[5,1] = nrow(dat[which(dat$gdp.const.int< tp[5,1] & !is.na(dat$gdp.const.int)), ])
+nc[6,1] = nrow(dat[which(dat$gdp.const.int< tp[6,1] & !is.na(dat$gdp.const.int)), ])
+
+nc.pc.mean =  matrix(NA, nrow=3, ncol=1)
+nc.pc.mean[1] = mean(nc[1:5])/n.gdp
+nc.pc.mean[2] = mean(nc[5])/n.gdp
+nc.pc.mean[3] = mean(nc[1:4])/n.gdp
+nc.pc.mean = nc.pc.mean*100
+
+tp.mean =  matrix(NA, nrow=3, ncol=1)
+tp.mean[1] = mean(tp[1:5])
+tp.mean[2] = mean(tp[5])
+tp.mean[3] = mean(tp[1:4])
+save(nc.pc.mean, file = "tp_ctr_mean.Rda")
+save(tp.mean, file = "tp_ctr_mean1.Rda")
 @
@@
 \begin{table}
-\caption{Estimated turning points}
+\centering
+\begin{threeparttable}
+\caption{Estimated turning points (TPs)}
 \label{tab:tp}
-\begin{center}
+\small
 \begin{tabular}{|c|c|c|c|c|c|}
 \hline 
-Model & dep. var & indep. var & tp (\$, PPP) & significance & \# of countries\tabularnewline
+Table & Estimator & Dependent variable & TP (\$, PPP) & Left of TP\tabularnewline
 \hline 
 \hline
-(\ref{eq:infl_model}) & FPI infl.
-07-08  & GDP & \Sexpr{round(tp[1,1],0)} & no & \Sexpr{nrow(data_0708[which(data_0708$GDP< tp[1,1] & data_0708$GDP >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$GDP>0), ])}  (\Sexpr{round(nrow(data_0708[which(data_0708$GDP< tp[1,1] & data_0708$GDP >0), ])/nrow(data_0708[which(data_0708$GDP>0), ])*100,0)}\%)\tabularnewline
+\ref{tab:07-08-fc-regressions} & OLS & FPI infl. 07-08  &  \Sexpr{round(tp[1,1],0)} & \Sexpr{nc[1,1]}/\Sexpr{n.gdp} \tabularnewline
 \hline 
-(\ref{eq:infl_model}) & FPI infl.
-07-08 & cons & \Sexpr{round(tp[2,1],0)} & no & \Sexpr{nrow(data_0708[which(data_0708$cons< tp[2,1] & data_0708$cons >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$cons>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$cons< tp[2,1] & data_0708$cons >0), ])/nrow(data_0708[which(data_0708$cons>0), ])*100,0)}\%)\tabularnewline
+\ref{tab:07-08-fc-regressions} & OLS & RFPI growth 07-08  &  \Sexpr{round(tp[2,1],0)} & \Sexpr{nc[2,1]}/\Sexpr{n.gdp} \tabularnewline
 \hline 
-(\ref{eq:infl_model}) & FPI infl.
-07-08 & food & \Sexpr{round(tp[3,1],0)}  & yes & \Sexpr{nrow(data_0708[which(data_0708$food< tp[3,1] & data_0708$food >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$food>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$food< tp[3,1] & data_0708$food >0), ])/nrow(data_0708[which(data_0708$food>0), ])*100,0)}\%)\tabularnewline
+\ref{tab:fpi-pt-regressions} & OLS & FPI LRM  &  \Sexpr{round(tp[3,1],0)} & \Sexpr{nc[3,1]}/\Sexpr{n.gdp} \tabularnewline
 \hline 
-(\ref{eq:infl_model}) & rFPI gr.
-07-08 & GDP & \Sexpr{round(tp[4,1],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$GDP< tp[4,1] & data_0708$GDP >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$GDP>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$GDP< tp[4,1] & data_0708$GDP >0), ])/nrow(data_0708[which(data_0708$GDP>0), ])*100,0)}\%)\tabularnewline
+\ref{tab:fpi-pt-regressions} & WLS & FPI LRM  &  \Sexpr{round(tp[4,1],0)} & \Sexpr{nc[4,1]}/\Sexpr{n.gdp} \tabularnewline
 \hline 
-(\ref{eq:infl_model}) & rFPI gr.
-07-08 & cons & \Sexpr{round(tp[5,1],0)}  & yes & \Sexpr{nrow(data_0708[which(data_0708$cons< tp[5,1] & data_0708$cons >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$cons>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$cons< tp[5,1] & data_0708$cons >0), ])/nrow(data_0708[which(data_0708$cons>0), ])*100,0)}\%)\tabularnewline
+\ref{tab:panel-regressions} & FE & $\Delta \log(FPI)_{t}$  &  \Sexpr{round(tp[5,1],0)} & \Sexpr{nc[5,1]}/\Sexpr{n.gdp} \tabularnewline
 \hline
-(\ref{eq:infl_model}) & rFPI gr.
-07-08 & food & \Sexpr{round(tp[6,1],0)}  & yes & \Sexpr{nrow(data_0708[which(data_0708$food< tp[6,1] & data_0708$food >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$food>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$food< tp[6,1] & data_0708$food >0), ])/nrow(data_0708[which(data_0708$food>0), ])*100,0)}\%)\tabularnewline
-\hline 
-(\ref{eq:infl_model}) & FPI infl.
-10-11 & GDP & \Sexpr{round(tp[1,2],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$GDP< tp[1,2] & data_0708$GDP >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$GDP>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$GDP< tp[1,2] & data_0708$GDP >0), ])/nrow(data_0708[which(data_0708$GDP>0), ])*100,0)}\%)\tabularnewline
-\hline 
-(\ref{eq:infl_model}) & FPI infl.
-10-11 & cons & \Sexpr{round(tp[2,2],0)} & no &  \Sexpr{nrow(data_0708[which(data_0708$cons< tp[2,2] & data_0708$cons >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$cons>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$cons< tp[2,2] & data_0708$cons >0), ])/nrow(data_0708[which(data_0708$cons>0), ])*100,0)}\%)\tabularnewline
-\hline 
-(\ref{eq:infl_model}) & FPI infl.
-10-11 & food & \Sexpr{round(tp[3,2],0)} & yes &  \Sexpr{nrow(data_0708[which(data_0708$food< tp[3,2] & data_0708$food >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$food>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$food< tp[3,2] & data_0708$food >0), ])/nrow(data_0708[which(data_0708$food>0), ])*100,0)}\%)\tabularnewline
-\hline 
-(\ref{eq:infl_model}) & rFPI gr.
-10-11 & GDP & \Sexpr{round(tp[4,2],0)} & yes &  \Sexpr{nrow(data_0708[which(data_0708$GDP< tp[4,2] & data_0708$GDP >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$GDP>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$GDP< tp[4,2] & data_0708$GDP >0), ])/nrow(data_0708[which(data_0708$GDP>0), ])*100,0)}\%)\tabularnewline
-\hline 
-(\ref{eq:infl_model}) & rFPI gr.
-10-11 & cons & \Sexpr{round(tp[5,2],0)} & yes &  \Sexpr{nrow(data_0708[which(data_0708$cons< tp[5,2] & data_0708$cons >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$cons>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$cons< tp[5,2] & data_0708$cons >0), ])/nrow(data_0708[which(data_0708$cons>0), ])*100,0)}\%)\tabularnewline
-\hline 
-(\ref{eq:infl_model}) & rFPI gr.
-10-11 & food & \Sexpr{round(tp[6,2],0)} & yes &  \Sexpr{nrow(data_0708[which(data_0708$food< tp[6,2] & data_0708$food >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$food>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$food< tp[6,2] & data_0708$food >0), ])/nrow(data_0708[which(data_0708$food>0), ])*100,0)}\%)\tabularnewline
-\hline 
-(\ref{eq:ptreg}) & LR multiplier & GDP & \Sexpr{round(tp[1,3],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$GDP< tp[1,3] & data_0708$GDP >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$GDP>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$GDP< tp[1,3] & data_0708$GDP >0), ])/nrow(data_0708[which(data_0708$GDP>0), ])*100,0)}\%)\tabularnewline
-\hline 
-(\ref{eq:ptreg}) & LR multiplier & cons & \Sexpr{round(tp[2,3],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$cons< tp[2,3] & data_0708$cons >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$cons>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$cons< tp[2,3] & data_0708$cons >0), ])/nrow(data_0708[which(data_0708$cons>0), ])*100,0)}\%)\tabularnewline
-\hline 
-(\ref{eq:ptreg}) & LR multiplier & food & \Sexpr{round(tp[3,3],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$food< tp[3,3] & data_0708$food >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$food>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$food< tp[3,3] & data_0708$food >0), ])/nrow(data_0708[which(data_0708$food>0), ])*100,0)}\%)\tabularnewline
-\hline 
-(\ref{eq:ptreg}) & LR multiplier & GDP & \Sexpr{round(tp[4,3],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$GDP< tp[4,3] & data_0708$GDP >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$GDP>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$GDP< tp[4,3] & data_0708$GDP >0), ])/nrow(data_0708[which(data_0708$GDP>0), ])*100,0)}\%)\tabularnewline
-\hline
-(\ref{eq:ptreg}) & LR multiplier & cons & \Sexpr{round(tp[5,3],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$cons< tp[5,3] & data_0708$cons >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$cons>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$cons< tp[5,3] & data_0708$cons >0), ])/nrow(data_0708[which(data_0708$cons>0), ])*100,0)}\%)\tabularnewline
-\hline 
-(\ref{eq:ptreg}) & LR multiplier & food & \Sexpr{round(tp[6,3],0)} & no & \Sexpr{nrow(data_0708[which(data_0708$food< tp[6,3] & data_0708$food >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$food>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$food< tp[6,3] & data_0708$food >0), ])/nrow(data_0708[which(data_0708$food>0), ])*100,0)}\%)\tabularnewline
-\hline 
-(\ref{eq:panel_model}) & mth FPI infl & GDP & \Sexpr{round(tp[1,4],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$GDP< tp[1,4] & data_0708$GDP >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$GDP>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$GDP< tp[1,4] & data_0708$GDP >0), ])/nrow(data_0708[which(data_0708$GDP>0), ])*100,0)}\%)\tabularnewline
-\hline 
-(\ref{eq:panel_model}) & mth FPI infl & cons & \Sexpr{round(tp[2,4],0)} & no & \Sexpr{nrow(data_0708[which(data_0708$cons< tp[2,4] & data_0708$cons >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$cons>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$cons< tp[2,4] & data_0708$cons >0), ])/nrow(data_0708[which(data_0708$cons>0), ])*100,0)}\%)\tabularnewline
-\hline 
-(\ref{eq:panel_model}) & mth FPI infl & food & \Sexpr{round(tp[3,4],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$food< tp[3,4] & data_0708$food >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$food>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$food< tp[3,4] & data_0708$food >0), ])/nrow(data_0708[which(data_0708$food>0), ])*100,0)}\%)\tabularnewline
-\hline 
-(\ref{eq:panel_model}) & mth rFPI gr.
-& GDP & \Sexpr{round(tp[4,4],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$GDP< tp[4,4] & data_0708$GDP >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$GDP>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$GDP< tp[4,4] & data_0708$GDP >0), ])/nrow(data_0708[which(data_0708$GDP>0), ])*100,0)}\%)\tabularnewline
-\hline 
-(\ref{eq:panel_model}) & mth rFPI gr.
-& cons & \Sexpr{round(tp[5,4],0)} & yes & \Sexpr{nrow(data_0708[which(data_0708$cons< tp[5,4] & data_0708$cons >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$cons>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$cons< tp[5,4] & data_0708$cons >0), ])/nrow(data_0708[which(data_0708$cons>0), ])*100,0)}\%)\tabularnewline
-\hline 
-(\ref{eq:panel_model}) & mth rFPI gr.
-& food & \Sexpr{round(tp[6,4],0)} & no & \Sexpr{nrow(data_0708[which(data_0708$food< tp[6,4] & data_0708$food >0), ])}/\Sexpr{nrow(data_0708[which(data_0708$food>0), ])} (\Sexpr{round(nrow(data_0708[which(data_0708$food< tp[6,4] & data_0708$food >0), ])/nrow(data_0708[which(data_0708$food>0), ])*100,0)}\%)\tabularnewline
-\hline 
-\multicolumn{6}{l}{Notes: The numbers in the fourth column refer to the estimated income or expenditure}\tabularnewline
-\multicolumn{6}{l}{turning point (tp). The fifth column indicates whether the estimated quadratic term}\tabularnewline
-\multicolumn{6}{l}{coefficients are individually significant at the 10 percent level. The last column shows}\tabularnewline
-\multicolumn{6}{l}{the number of countries in the sample with a per capita income or expenditure levels}\tabularnewline 
-\multicolumn{6}{l}{below the estimated turning point.}
 \end{tabular}
-\end{center}
-\end {table}
+\begin{tablenotes}
+\item \emph{Note:} The numbers in the fourth column are the estimated turning points based on coefficients from the regression specified in the first three columns.
+I.e.
+the per capita income level corresponding to the highest level of predicted domestic food inflation or price transmission.
+The final column calculates how many countries that have per capita income less than the level corresponding to the estimated  turning point, relative to the total number of countries in the sample with data on income.
+\end{tablenotes}
+\end{threeparttable}
+\end{table}

### 2016-03-02 model.Rnw
revised model section for 2. round revision version of paper

@@
-% !Rnw root = Master.Rnw
+% !Rnw root = master.Rnw
 
-\section{A model of aggregate food price transmission\label{sec:model}}
-As the last section pointed out, the small literature on aggregate food price transmission is based on informal economic reasoning, rather than an actual economic model.
-In this section I will present one such model and derive some testable predictions.
+\section{Determinants of food price transmission\label{sec:model}}
+What are the main determinants of a country's responsiveness to an international food price shock?
+I.e., which factors would we expect to influence the relationship between international and domestic food prices?
+To answer this question this section first reviews literature.
+Next, a simple model is presented of a country's food price level and its responsiveness to international shocks based on insights from the literature review.
+
+\subsection{Literature review}
+There are only a few studies carrying out actual cross section analyses of price transmission determinants.
+There are, however, several studies discussing this informally, especially in relation to the 2007-08 food crisis.
+Arguments and main findings from a selection of these studies are summarized below.
+
+\citet{Hoyos2011} point out that domestic food inflation during the food crisis varied a lot across countries.\footnote{Domestic food inflation is defined as the percentage change in a country's Food Price Index (FPI) which is a component of the Consumer Price Index (CPI).} Furthermore, domestic food inflation was lower on average than international food inflation.
+\citeauthor{Hoyos2011} offer 4 reasons why was is the case: (i) weak price transmission perhaps due to poor infrastructure, (ii) differences in the composition of food baskets and the fact that domestic food baskets contain nontradables, (iii) lack of competition in the domestic food markets and (iv) government interventions for example in the form of price controls.
+
+\citet{Abbott2009} discusses the causes and consequences of the food crisis.
+Regarding the cross country differences in food inflation rates, he makes three observations: i) developing countries have more basic diets which involve a greater share of staples; ii) processing and distribution margins are typically lower than in developed countries, which means that they can shrink less as costs increase.
+Both factors, he argues, caused a stronger domestic food inflation response among developing countries than in developed countries to the high world prices in 2007-08.
+And, iii) a country's exchange rate regime has a large impact on the domestic price impact of an international shock; countries which peg their currency to the dollar are more exposed to international price fluctuations than countries whose currencies follow the euro.
+
+\citet{Lee2013} quantify the importance of various ``external'' and ``internal'' variables on domestic food inflation based on data from 72 countries over the period 2000 - 2011.
+The set of external variables consists of current and lagged international food inflation, as well as a set of intra- and extra-regional food inflation averages.
+The internal variables represent domestic supply and demand factors as well as various other country characteristics.
+What they find is that  countries with large food import shares have lower food inflation whereas countries experiencing large increases in their food import shares have higher food inflation on average.
+The latter result could, however, reflect that food imports increase when domestic domestic food prices are on the rise, as a response to food inflation rather than a cause of it.
+Surprisingly, the authors also find a negative relationship between a country's per capita GDP growth rate (a demand variable) and its food inflation rate.
+The relationship between domestic food inflation and the per capita GDP level is found to be positive albeit only marginally significant in some specifications.
+A depreciation of the domestic currency/US\$ exchange rate increases domestic food inflation significantly across all specifications, as expected, whereas an increase in the money supply has no effect on food inflation.
+More politically stable countries tend to have lower food inflation whereas countries with a higher free trade index score are found to have higher domestic food inflation.
+
+Two studies quantify the determinants of price transmission and market integration based on individual agricultural commodity prices rather than food price indexes.
+\citet{kouyate16} carry out a meta-regression analysis of the effect of distance on agricultural commodity price transmission for a large number of market pairs.
+Their results show that geographical distance and separation by borders reduces the likelihood of price cointegration and weakens the speed of speed of price transmission significantly.
+These findings are consistent with the well established stylized facts from the trade literature, that borders and distance reduces trade.
+
+Finally, \citet{greb12} consider 8 variables covering geographical, infrastructural, institutional and commodity specific factors which are believed to influence agricultural price transmission.
+The authors regress their own price transmission estimates, as well as estimates obtained from the literature, on these country characteristics.
+What they find is that the covariates explain more of the variation in the estimated speed of adjustment parameters (the $\alpha$'s) than in the cointegration parameters (the $\beta$'s).
+Another major finding is that more of these covariates have significant and plausible partial effects on the $\alpha$'s than on the $\beta$'s.
+However, none of the determinants have a robust robust impact on price transmission.
+
+\subsection{A model of a country's food price level}
+This section presents a simple model of a country's food price level that is inspired by the arguments and findings discussed above, related to determinants of domestic food inflation and price transmission.
+The benefit of a working with a formal economic model is that it imposes some structure on the empirical analysis.
+In fact, despite its simplicity, it is possible to derive several predictions from the model regarding the determinants of price transmission.
 
@@
 \begin{equation}
-P =\alpha P_{f}+(1-\alpha)P_{n}\label{eq:cpi},
+P =a P_{f}+(1-a)P_{n}\label{eq:cpi},
 \end{equation}
-where $P_{f}$ and $P_{n}$ denote the food and non-food price indexes, respectively, and $\alpha$ is the fixed food expenditure share.
-I have omitted time indexes in order to avoid notational clutter in the equations that follow.
-However, all price indexes are considered stochastic processes.
-Moreover, the two components $P_{f}$ and $P_{n}$ must share a common stochastic trend representing underlying inflation.
+where $P_{f}$ and $P_{n}$ denote the food and non-food price indexes, respectively, and $a$ is the food expenditure share.
+Time indexes are omitted in order to avoid notational clutter.
 The food price index (FPI), which is of primary interest, can be written
 \begin{equation}
-P_{f} =\beta P_{c}+(1-\beta)P_{m}\label{eq:fpi},
+P_{f} =b P_{c}+(1-b)P_{m}\label{eq:fpi},
 \end{equation}
-where $P_{c}$ and $P_{m}$ denote the farm cost and marketing components of the FPI, respectively, and $\beta$ the farm cost or commodity share.
-By marketing costs I refer to all other costs than those of the raw agricultural commodities included in the final consumer food price such as those associated with distribution, storage and promotion.
-Finally, the commodity component of the domestic FPI can be decomposed as
+where $P_{c}$ and $P_{m}$ denote the farm cost and marketing components of the FPI, respectively, and $b$ is the farm cost or commodity share.\footnote{By marketing costs I refer to all other costs than those of the raw agricultural commodities included in the final consumer food price such as those associated with distribution, storage and promotion.}  Finally, the commodity component of the domestic FPI can be decomposed as
 \begin{equation}
-P_{c} =\gamma P_{w}+(1-\gamma)P_{d}\label{eq:compi},
+P_{c} =c P_{w}+(1-c)P_{d}\label{eq:compi},
 \end{equation}
-where $P_{w}$ and $P_{d}$ denote the tradable and purely domestic (i.e.
-non-tradable) subcomponents of the commodity price index, respectively, and $\gamma$ is the share of tradable food commodities consumed domestically.
-The defining difference between tradable and non-tradable food commodities is that the former have prices which are determined on the world market whereas the latter prices are determined by domestic supply and demand.
+where $P_{w}$ and $P_{d}$ denote the tradable and purely domestic subcomponents of the commodity price index, respectively, and $c$ is the share of tradable food commodities in $P_{c}$.
+The defining difference between tradable and domestic food commodities is that the former have prices which are determined on the world market whereas the latter prices are determined by domestic supply and demand.
 
-The three share parameters, $\alpha$, $\beta$ and $\gamma$ represent weights on price indexes that are progressively embedded into each other.
-By substituting (\ref{eq:compi}) into (\ref{eq:fpi}) we get
+The three share parameters, $a$, $b$ and $c$ represent weights on price indexes that are progressively embedded into each other. By substituting (\ref{eq:compi}) into (\ref{eq:fpi}) we get
 \begin{equation}
-P_{f} = \beta \gamma P_{w}+R\label{eq:fpi1},
+P_{f} = b c P_{w}+R\label{eq:fpi1},
 \end{equation}
-where $R=\beta(1-\gamma)P_{d}+(1-\beta)P_{m}$ is a term which does not depend on world market prices directly.
-With these definitions a first order approximation to transmission of international agricultural commodity price changes to domestic consumer food prices can be written
+where $R=b(1-c)P_{d}+(1-b)P_{m}$ is a term which does not depend on world market prices directly.
+With these definitions a first-order approximation to transmission of international agricultural commodity price changes to domestic consumer food prices can be written
 \begin{equation}
-\theta=\frac{\partial P_{f}}{\partial P_{w}}=\beta\gamma,\label{eq:pt}
+\theta\equiv\frac{\partial P_{f}}{\partial P_{w}}=bc,\label{eq:pt}
 \end{equation}
-i.e., price transmission depends multiplicatively on the share of tradable commodities in the domestic food commodity price index, $\gamma$, and the commodity cost share in the FPI, $\beta$.
+i.e., price transmission depends multiplicatively on the share of tradable commodities in the commodity price index, $c$, and the commodity cost share in the FPI, $b$.
 Expression (\ref{eq:pt}) is based on the assumption that there is no knock-on effect on domestic domestic prices when world market prices change, as would be appropriate in the short term.
-If we relax this assumption and allow domestic prices to respond to international price changes through substitution effects we would have to add the term $\beta(1-\gamma)\frac{\mathrm{d}P_{d}}{\mathrm{d}P_{w}}$.
+If we relax this assumption and allow domestic prices to respond to international price changes through substitution effects we would have to add the term $b(1-c)\frac{\mathrm{d}P_{d}}{\mathrm{d}P_{w}}$.
 For simplicity, I will ignore this second order effect in the analysis below.
 
-In order to analyze how price transmission depends on income we need to make some assumptions about the two cost shares, $\gamma$ and $\beta$. Specifically, I shall assume that
+By imposing some structure on the two cost shares, $b$ and $c$ we will be able to analyze determinants of a country's price transmission.
+In particular, we can write these shares as functions of certain country characteristics:
 \begin{equation}
-\beta=\beta(y),\quad\beta'(y)<0,\quad\beta''(y)>0\label{com_share}
+b=b(x_{b}), \quad c=c(x_{c}).
 \end{equation}
-and
-\begin{equation}
-\gamma=\gamma(y),\quad\gamma'(y)>0,\quad\gamma''(y)<0,\label{trade_share}
-\end{equation}
-where $y$ denotes per capita food expenditure or, more generally, income.
-That is, the wealthier the country, the lower the commodity cost share in the FPI and the higher the share of tradables in the commodity price index.
-The left panel of figure \ref{fig:pt} illustrates two such share functions with these specific properties.
-Assumptions (\ref{com_share})-(\ref{trade_share}) imply that price transmission, $\theta$, becomes a concave function of income, a feature which is illustrated in the right panel of figure \ref{fig:pt}.
+But what are these characteristics? As I will argue below, a country's income level should feature prominently among them.
 
-\begin{center}
-[Figure \ref{fig:pt} around here] \\
-\par\end{center}
+We can reasonably assume that the wealthier the country, the lower the commodity cost share in the FPI, c.f.
+\citet{Abbott2009}.
+Formally, if $y$ denotes income then $y\in x_{b}$ and $\frac{\partial b}{\partial y} <0$.
+Furthermore, if we believe that countries become more integrated with the world market as they develop, not only due to better infrastructure and more open markets, but also due to for example higher quality standards associated with the adoption of modern production and management practices, then it would be appropriate to also assume $y\in x_{c}$ and $\frac{\partial c}{\partial y} >0$.\footnote{We must also impose the following set of restrictions, $\lim_{y\rightarrow \infty} b(y)=0$, $\lim_{y\rightarrow \infty} c(y)=1$, $\lim_{y\rightarrow 0^{+}} b(y)=1$ and $\lim_{y\rightarrow 0~{+}} c(y)=0$, since otherwise these functions could not represent cost shares.} These assumptions in combination imply an inverse-U shaped relationship between price transmission, $\theta$, and income.
 
-To sum up, there are reasons to expect a domestic FPI reaction to price changes on the international agricultural commodity markets which depends non-monotonically on income or expenditure.
+Income is not the only determinant of price transmission of course.
+A country's openness to trade, the quality of its infrastructure and the composition of its food consumption set are all factors which we would expect to influence the transmission of world market price fluctuations to domestic food markets.
+Economies that are more open in general, perhaps for historical reasons, should have a higher tradables share and therefore experience more price transmission.
+Similarly with countries that, due to climatic or geographical reasons, are forced to import most of their food supply.
+On the other hand, there are countries that are self-sufficient due to poor infrastructure, perhaps in combination with landlockedness, resulting in prohibitive transport costs.
+The tradables share in these countries should be very low and so should the price transmission rate.
+Consumption patterns are influenced by cultural as well as geographical factors.
+In particular, some countries consume more grains than others and the commodity component in the food price index should therefore be larger.
+Consequently, domestic aggregate food prices are more exposed international shocks in such countries.
+Finally, there are policies that are capable of affecting both shares.
+Trade policies in particular have a direct effect on the tradability of a good.
+
+To sum up, there are reasons to expect a domestic FPI reaction to price fluctuations on the international agricultural commodity markets which depends non-monotonically on income.
 Specifically, the simple model sketched above predicts an inverse-U shaped relationship between income and price transmission.
-This prediction is based, of course, on the assumptions regarding the relationship between income and the two key share parameters, $\beta$ and $\gamma$, discussed above.
-Therefore, to round off this section I will briefly discuss what we know about these.
-
-\subsection{Farm shares, tradability and income\label{sec:model_assumptions}}
-Regarding $\beta$, although there is no database of international farm cost shares in food expenditure, there are some time series estimates available from a few countries.
-Estimates from the US and the EU do, in particular, seem to support the conclusion that the farm share declines as countries grow richer.
-\cite{Schnepf2009}, for example, uses USDA data to show that the farm-value added share of consumer food expenditures has fallen from 41 percent in 1950 to 18.5 percent in 2006, meaning that the value share of marketing services has increased by 22.5 percentage points over the 56 year period.
-In the same period US per capita GDP increased by 240 percent.
-Similarly, analysts working for the European Commission has calculated that within EU25, the agriculture-value added share in the food supply chain has dropped from 31 percent in 1995 to 24 percent in 2005 \citep{Commission2009}.
-In this period (EU28) per capita GDP grew by 24 percent according to EUROSTAT.
-\citet[][p.
-797]{Yang2015} comment that ``\emph{...
-the share of domestic margin services in consumed food is larger in high-income countries \cite{Abbott2012}, ...
-implying that consumers in high-income countries are more insulated from shocks of international agricultural prices.}''
-
-Regarding the $\gamma$ parameter, representing the cost share of tradable commodities in the food commodity price index, the main issue is conceptual rather than empirical.
-The problem is that tradability is not an observable feature of a good.
-A food good might not be traded due to policy distortions such as trade taxes and nontariff barriers but there are also other factors including transport costs, quality differences and local taste preferences which reduce the amount of trade.
-In the literature goods are typically classified according to the actual trade flows though.
-\cite{Anderson2009}, for example, divides the agricultural goods produced in country into three groups; import-competing (M), exportable (X), or nontradable (H) and a good is classified as nontradable if three conditions are met.
-First, there must be no policies in place distorting trade flows.
-Second, the share of production exported must be less than 2.5 percent and, third, the share of consumption imported must also be less than 2.5 percent.
-M and X goods are both tradable in the sense that their undistorted prices are determined on the world market (pp.
-582-84).
-Figure \ref{fig:gamma} plots the estimated value shares of tradable agricultural goods relative to production based on the \cite{Anderson2008} database.
-As the lowess line indicates, there is a positive relationship between the estimated $\gamma$'s and income up to a certain income level, after which all goods produced become tradable.
-
-\begin{center}
-[Figure \ref{fig:gamma} around here] \\
-\par\end{center}
-
-Finally, regarding $\alpha$, the food share in total expenditure, although I do not make any assumptions about its dependence on income, there is, of course, a well established relationship between the two known as Engel's law.
-
-Although the focus in this as well as in the next section is on international to domestic price transmission or pass-through, it is worth mentioning that the simple model presented above delivers another important prediction; namely that LICs are especially vulnerable to domestic agricultural supply shocks.
-This is because domestic food prices mainly represent non-tradable food goods and because the cost of these goods mainly represent farm gate prices.
-I will, however, for the sake of brevity, leave it to future work to test this proposition.
+Factors related to trade openness, consumption, geography and the quality of infrastructure also, arguably, affect price transmission.

### 2016-03-02 panel_reg.R
Revised panel regression script for 2. round revision version of paper

@@ -3,259 +3,157 @@ library(dplyr)
 library(sandwich)
 library(lmtest)
 library(plm)
-library(foreign)
 
 ##################################################################
 ### Panel price transmission regressions
 ##################################################################
 
-load("ctr_dat_0708.Rda")
+load("infl_ctr_dat_0708.Rda")
 load("ts_dat.Rda")
 df = merge(ts.dat, ctr.dat.0708, by = "iso3c")
-df = df[!duplicated(df), ]
-data = plm.data(df, index = c("iso3c", "tt"))
-df1 = select(df, -fao.fpi)
-df1 = rename(df1, fao.fpi = r.fao.fpi)
-data1 = plm.data(df1, index = c("iso3c", "tt"))
+data = plm.data(df, index = c("iso3c", "ff"))
+df1 = dplyr::select(df, -fao.fpi)
+df1 = dplyr::rename(df1, fao.fpi = r.fao.fpi)
+data1 = plm.data(df1, index = c("iso3c", "ff"))
 
 panel.reg.m = list()
+panel.reg.se = list()
 p = 6
 
 #####################################################
 ## Fixed effects (FE) estimator
 
 # FPI regressions
-m = plm(diff(log(adj.fpi)) ~ lag(diff(log(adj.fpi)),1:p) +
-          lag(diff(log(fao.fpi)), 0:1) +
-          lag(diff(log(er)), 0:1) +
-          I(lag(diff(log(fao.fpi)),0)*log(gdp)) +
-          I(lag(diff(log(fao.fpi)),0)*I(log(gdp)^2)) +
-          I(lag(diff(log(fao.fpi)),0)*cshare) +
-          I(lag(diff(log(fao.fpi)),0)*cdep) +
-          I(lag(diff(log(fao.fpi)),0)*lldc) +
-          I(lag(diff(log(fao.fpi)),0)*to) +
-          I(lag(diff(log(fao.fpi)),0)*dtf) +
-          I(lag(diff(log(fao.fpi)),0)*lpi) |
-          lag(diff(log(fao.fpi)), 0:1) +
-          lag(diff(log(er)), 0:1) +
-          lag(diff(log(adj.fpi)),1:p) +
-          I(lag(diff(log(fao.fpi)),0)*log(gdp)) +
[…]

### 2016-03-02 ts_reg.R
revised time series regression script for 2. round rev. version of paper

@@ -2,6 +2,7 @@
 library(dplyr)
 library(dyn)
 library(dynlm)
+library(lmtest)
 
 ##################################################################
 ### Time series regressions
@@ -18,13 +19,13 @@ data$id = as.numeric(data$id)
 ## Model: ADL
 
 # 1) adj.fpi
-p = 6; q=6
+p = 12; q=12
 aic.mat = array(NA, dim = c(nrow(tbl), p, q))
 aic.min.mat = matrix(NA, nrow = nrow(tbl), ncol = 2)
 
 for (i in 1:max(data$id)) {
   dat = filter(data, id==i)
-  dat = select(dat, cpi.brk:rfpi, fao.fpi:er)
+  dat = dplyr::select(dat, cpi.brk:rfpi, fao.fpi:er, POILAPSP)
   ts.dat = ts(dat, start=c(2005, 1), end=c(2014, 12), frequency=12)
   
   if(is.na(dat$er)[1] == FALSE) {
@@ -34,7 +35,9 @@ for (i in 1:max(data$id)) {
                     d(log(fao.fpi)) +
                     L(d(log(fao.fpi)),1:k) + 
                     d(log(er)) +
-                    L(d(log(er)),1:k), 
+                    L(d(log(er)),1:k) +
+                    d(log(POILAPSP)) +
+                    L(d(log(POILAPSP)),1:k), 
                   data = ts.dat)
         aic.mat[i,j,k] = extractAIC(m)[2]
       }
@@ -44,7 +47,9 @@ for (i in 1:max(data$id)) {
       for (k in 1:q) {
         m = dynlm(d(log(adj.fpi))~ L(d(log(adj.fpi)),1:j) + 
                     d(log(fao.fpi)) +
-                    L(d(log(fao.fpi)),1:k), 
+                    L(d(log(fao.fpi)),1:k) +
+                    d(log(POILAPSP)) +
+                    L(d(log(POILAPSP)),1:k), 
                   data = ts.dat)
         aic.mat[i,j,k] = extractAIC(m)[2]
       }
@@ -56,13 +61,13 @@ aic.min.fpi = as.data.frame(aic.min.mat)
 aic.min.fpi$iso3c = names(tbl)
 
 # 2) rfpi
-p = 6; q=6
+p = 12; q=12
 aic.mat = array(NA, dim = c(nrow(tbl), p, q))
[…]

### 2016-03-02 appendix.Rnw
appendix for 2. round rev. version of paper

% !Rnw root = master.Rnw

\section*{Appendix}

\begin{table}[h]
\caption{Variable names, description and sources}
\label{tab:cs_data}
\begin{center}

{\small{}}%
\begin{tabular}{|c|>{\raggedright}m{8cm}|>{\raggedright}m{3cm}|}
\hline 
\textbf{\small{}Variable} & \textbf{\small{}Description} & \textbf{\small{}Source} \tabularnewline
\hline
{\small{}FPI} & {\small{}Domestic food price index.} & {\small{}ILO, LABORSTA} \tabularnewline
\hline 
{\small{}RFPI} & {\small{}Domestic real food price index. FPI deflated by the CPI} & {\small{}ILO, LABORSTA, own calculations} \tabularnewline
\hline 
{\small{}FAO FPI} & {\small{}FAO food price index.} & {\small{}FAO} \tabularnewline
\hline 
{\small{}FAO RFPI} & {\small{}Real FAO food price index. Nominal index deflated by the
World Bank Manufactures Unit Value index} & {\small{}FAO} \tabularnewline
\hline 
{\small{}ER} & {\small{}Exchange rate. Log of the domestic currency/US\$ exchange rate} & {\small{}IMF} \tabularnewline
\hline 
{\small{}GDP} & {\small{}GDP. Log of per capita GDP (constant 2011 international \$, PPP)} & {\small{}World Bank database} \tabularnewline
\hline 
{\small{}CCONS} & {\small{}Cereal consumption share. Share of dietary energy supply derived from
cereals, roots and tubers (\%). 3-year average, 2004-06} & {\small{}FAO, FAOSTAT} \tabularnewline
\hline 
{\small{}CIMP} & {\small{}Cereal import share. Cereal imports as a share of domestic supply (\%). 3-year average, 2004-06} & {\small{}FAO, FAOSTAT} \tabularnewline
\hline 
{\small{}CEXP} & {\small{}Cereal export share. Cereal exports as a share of domestic production (\%). 3-year average, 2004-06} & {\small{}FAO, FAOSTAT} \tabularnewline
\hline 
{\small{}LLDC} & {\small{}Dummy. 1 if country is a landlocked developing country} & {\small{}UNCTAD} \tabularnewline
\hline
{\small{}STE} & {\small{}State trading enterprise dummy. 1 if country has a grain state trading enterprise} & {\small{}\citet{greb12}} \tabularnewline
\hline
{\small{}TO} & {\small{}Trade openness. Sum of exports and imports of goods and services as a share of GDP (\%). 2006-10 average} & {\small{}World Bank Development Indicators} \tabularnewline
\hline 
{\small{}DTF} & {\small{}Distance to Frontier index. An economy\textquoteright s distance to frontier measured on a scale from 0 to 100, where 0 represents the lowest performance and 100 the frontier.} & {\small{}Doing Business survey, Trading Across Borders. World Bank} \tabularnewline
\hline 
{\small{}LPI} & {\small{} Logistics performance index: Quality of trade and transport-related infrastructure. The index ranges from 1 to 5, with a higher score representing better performance.} & {\small{}Logistics Performance Index 2007. World Bank database} \tabularnewline
\hline 
\end{tabular}{\small \par}
\end{center}
\end {table}

<<results=tex,echo=F>>=
library(stargazer)
reg.list = readRDS("int_reg_m.rds")
se.list = readRDS("int_reg_se.rds")

stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
          title='Determinants of domestic food price changes 2007-8', 
          label = "tab:int-regressions",
          dep.var.labels=c("FPI inflation", "RFPI growth"),
          no.space=TRUE,
          se = list(se.list[[1]], se.list[[2]], se.list[[3]], se.list[[4]]), 
          omit.stat = c("adj.rsq","f", "ser"), 
          font.size = "footnotesize", 
          covariate.labels=c("$\\Delta ER$",
                             "$GDP$",
                             "$GDP^{2}$", 
                             "$CCONS$", 
                             "$CIMP$", 
                             "$CEXP$",
                             "$LLDC$", 
                             "$STE$",
                             "$TO$", 
                             "$DTF$", 
                             "$LPI$",
                             "$Export$",
                             "$Price$",
                             "$Stock$",
                             "$Tariff$",
                             "$Tax$"),
          notes=c("*, ** and *** signify significance at the 10, 5 and 1 percent",
                  "levels, respectively. Robust standard errors in parentheses;", 
                  "samples exclude outliers defined as observations $\\pm$ 3.5 stan-", 
                  "dard deviations from the mean of the dependent variable.", 
                  "A description of the variables and their sources can be found", 
                  "in table \\ref{tab:cs_data} and section \\ref{sec:analysis}. Footnote \\ref{int_note} contains a description", 
                  "of the 5 intervention dummies."),
          notes.append=FALSE,
          notes.align="l")
@

<<results=tex,echo=F>>=
# library(stargazer)
reg.list = readRDS("fc_reg1011_m.rds")
se.list = readRDS("fc_reg1011_se.rds")

stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
          title='Determinants of domestic food price changes 2010-11', 
          label = "tab:1011-regressions",
          dep.var.labels=c("FPI inflation", "RFPI grwth"),
          no.space=TRUE,
          se = list(se.list[[1]], se.list[[2]], se.list[[3]], se.list[[4]]), 
          omit.stat = c("adj.rsq","f", "ser"), 
          font.size = "small", 
         covariate.labels=c("$\\Delta ER$",
                             "$GDP$",
                             "$GDP^{2}$", 
                             "$CCONS$", 
                             "$CIMP$", 
                             "$CEXP$",
                             "$LLDC$", 
                             "$STE$",
                             "$TO$", 
                             "$DTF$", 
                             "$LPI$"),
          notes=c("*, ** and *** signify significance at the 10, 5 and 1 percent",
                  "levels, respectively. Robust standard errors in parentheses;", 
                  "samples exclude outliers defined as observations $\\pm$ 3.5 stan-", 
                  "dard deviations from the mean of the dependent variable.", 
                  "A description of the variables and their sources can be found", 
                  "in table \\ref{tab:cs_data} and section \\ref{sec:analysis}."),
          notes.append=FALSE,
          notes.align="l")
@

<<results=tex,echo=F>>=
reg.list = readRDS("rfpi_pt_reg_m.rds")
se.list = readRDS("rfpi_pt_reg_se.rds")

stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
          title='Determinants of long run real FPI transmission rates', 
          label = "tab:rfpi-pt-regressions", 
          dep.var.labels=c("RFPI LRM (6 lags)"), 
          column.labels = c("OLS", "WLS"),
          column.separate = c(2, 2),
          no.space=TRUE,
          se = list(se.list[[1]], se.list[[2]], se.list[[3]], se.list[[4]]), 
          omit.stat = c("adj.rsq","f", "ser"), 
          font.size = "small", 
          covariate.labels=c("$GDP$",
                             "$GDP^{2}$", 
                             "$CCONS$", 
                             "$CIMP$", 
                             "$CEXP$",
                             "$LLDC$", 
                             "$STE$",
                             "$TO$", 
                             "$DTF$", 
                             "$LPI$"),
          notes=c("*, ** and *** signify significance at the 10, 5 and 1 percent",
                  "levels, respectively. Robust standard errors in parentheses;", 
                  "samples exclude outliers defined as observations $\\pm$ 3.5 stan-", 
                  "dard deviations from the mean of the dependent variable.", 
                  "A description of the variables and their sources can be found", 
                  "in table \\ref{tab:cs_data} and section \\ref{sec:analysis}."),
          notes.append=FALSE,
          notes.align="l")
@

<<results=tex,echo=F>>=
reg.list = readRDS("fpi_pt_reg_m1.rds")
se.list = readRDS("fpi_pt_reg_se1.rds")

stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
          title='Determinants of long run FPI transmission rates', 
          label = "tab:fpi-pt1-regressions", 
          dep.var.labels=c("FPI LRM (single lag)"), 
          column.labels = c("OLS", "WLS"),
          column.separate = c(2, 2),
          no.space=TRUE,
          se = list(se.list[[1]], se.list[[2]], se.list[[3]], se.list[[4]]), 
          omit.stat = c("adj.rsq","f", "ser"), 
          font.size = "small", 
          covariate.labels=c("$GDP$",
                             "$GDP^{2}$", 
                             "$CCONS$", 
                             "$CIMP$", 
                             "$CEXP$",
                             "$LLDC$", 
                             "$STE$",
                             "$TO$", 
                             "$DTF$", 
                             "$LPI$"),
          notes=c("*, ** and *** signify significance at the 10, 5 and 1 percent",
                  "levels, respectively. Robust standard errors in parentheses;", 
                  "samples exclude outliers defined as observations $\\pm$ 3.5 stan-", 
                  "dard deviations from the mean of the dependent variable.", 
                  "A description of the variables and their sources can be found", 
                  "in table \\ref{tab:cs_data} and section \\ref{sec:analysis}."),
          notes.append=FALSE,
          notes.align="l")
@

<<results=tex,echo=F>>=
reg.list = readRDS("rfpi_pt_reg_m1.rds")
se.list = readRDS("rfpi_pt_reg_se1.rds")

stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
          title='Determinants of long run real FPI transmission rates', 
          label = "tab:rfpi-pt1-regressions", 
          dep.var.labels=c("RFPI LRM (single lag)"), 
          column.labels = c("OLS", "WLS"),
          column.separate = c(2, 2),
          no.space=TRUE,
          se = list(se.list[[1]], se.list[[2]], se.list[[3]], se.list[[4]]), 
          omit.stat = c("adj.rsq","f", "ser"), 
          font.size = "small", 
          covariate.labels=c("$GDP$",
                             "$GDP^{2}$", 
                             "$CCONS$", 
                             "$CIMP$", 
                             "$CEXP$",
                             "$LLDC$", 
                             "$STE$",
                             "$TO$", 
                             "$DTF$", 
                             "$LPI$"),
         notes=c("*, ** and *** signify significance at the 10, 5 and 1 percent",
                  "levels, respectively. Robust standard errors in parentheses;", 
                  "samples exclude outliers defined as observations $\\pm$ 3.5 stan-", 
                  "dard deviations from the mean of the dependent variable.", 
                  "A description of the variables and their sources can be found", 
                  "in table \\ref{tab:cs_data} and section \\ref{sec:analysis}."),
          notes.append=FALSE,
          notes.align="l")
@

<<results=tex, echo=FALSE>>=
# library(plm)
# reg.list = readRDS("panel_reg_m.rds")
# 
# stargazer(reg.list[[2]], reg.list[[4]],
#           title='Determinants of short run FPI transmission rates. Anderson-Hsiao estimator', 
#           label = "tab:panel-regressions1", 
#           dep.var.labels=c("FPI inflation", "RFPI growth"),
#           single.row = TRUE,
#           omit = c("adj.fpi", "rfpi"),
#           omit.stat = c("adj.rsq","f", "ser"), 
#           font.size = "small",
#           covariate.labels=c("$\\pi_{t}^{i}$",
#                              "$\\pi_{t-1}^{i}$",
#                              "$\\Delta ER_{t}$",
#                              "$\\Delta ER_{t-1}$",
#                              "$\\pi_{t}^{i}\\times GDP$",
#                              "$\\pi_{t}^{i}\\times GDP^2$", 
#                              "$\\pi_{t}^{i}\\times CCONS$", 
#                              "$\\pi_{t}^{i}\\times CIMP$", 
#                              "$\\pi_{t}^{i}\\times CEXP$",
#                              "$\\pi_{t}^{i}\\times LLDC$", 
#                              "$\\pi_{t}^{i}\\times STE$", 
#                              "$\\pi_{t}^{i}\\times TO$", 
#                              "$\\pi_{t}^{i}\\times DTF$", 
#                              "$\\pi_{t}^{i}\\times LPI$"),
#           notes=c("*, ** and *** signify significance at the 10, 5 and 1",
#                   "percent levels, respectively. Standard errors in pa-", 
#                   "rentheses. See table \\ref{tab:cs_data} and section \\ref{sec:analysis} for a description",
#                   "of the variables and their sources."),
#           notes.append=FALSE,
#           notes.align="l")
@

### 2016-03-02 price_data_prep.R
price data preparation script for 2. round rev. version of paper

@@ -1,0 +1,493 @@
+#####################################################################
+#### Creating the domestic price dataset 
+#####################################################################
+
+library(foreign)
+library(dplyr)
+library(reshape)
+library(zoo)
+library(data.table)
+library(ggplot2)
+library(lattice)
+library(countrycode)
+
+### Raw data:
+# laborstaB9.csv
+# laborstaBA.csv
+
+#####################################################################
+#### Reading and merging the raw price data
+
+str = "C:/Users/nqw235/Google Drev/Research/pov_price_transmission/"
+
+### The ILO data
+temp = paste(str, "laborstaB9.csv", sep = "")
+B9 = read.csv(temp, sep = ";")
+temp = paste(str, "laborstaBA.csv", sep = "")
+BA = read.csv(temp, sep = ";")
+
+## Reshape to long format
+B9$id = paste0(B9$CODE.COUNTRY, B9$CODE.AREA)
+B9[B9$id == "LKLK1" & B9$YEAR == 2002, ]
+B9 = B9[-2659, ] # coding mistake
+B9l = reshape(B9, varying=c(9:21), direction="long", v.names="cpi", 
+              idvar=c("id", "YEAR"),
+              sep="", timevar = "month")
+B9l = B9l[order(B9l$id, B9l$YEAR, B9l$month), ]
+B9l$SRC_ID = NULL
+B9l = rename(B9l, c("NOTES" = "cpi.notes"))
+
+BA$id = paste0(BA$CODE.COUNTRY, BA$CODE.AREA)
+BA[BA$id == "LKLK1" & BA$YEAR == 2002, ]
+BA = BA[-2629, ] # coding mistake
+BAl = reshape(BA, varying=c(9:21), direction="long", v.names="fpi", 
+              idvar=c("id", "YEAR"),
+              sep="", timevar = "month")
+BAl = BAl[order(BAl$id, BAl$YEAR, BAl$month), ]
+BAl$SRC_ID = NULL
+BAl = rename(BAl, c("NOTES" = "fpi.notes"))
+
+## Merge CPI and FPI datasets
+ilo = merge(B9l, BAl, all=TRUE)
[…]

### 2016-03-02 reply_reviewer_1.Rnw
reply to editor and reviwers in connection with 2. round rev. version of paper

\documentclass[12pt,english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{babel}
\usepackage{setspace}
\doublespacing
\usepackage{geometry}
\geometry{verbose,tmargin=2cm,bmargin=2cm,lmargin=3cm,rmargin=3cm}
\usepackage{color,soul}
\usepackage[authoryear]{natbib}

\begin{document}
\SweaveOpts{concordance=TRUE}

\title{Reply to reviewer R1}
\date{\vspace{-5ex}}
\maketitle

\subsection*{Reviewer R1}
\subsubsection*{Some Formal Aspects:}
\begin{itemize}
\item \emph{Please include Abstract and Keywords on the first page.}

Done.

\item \emph{In the beginning of the paper, it would also be good to mention why this study is important. For example, why is it important to look at how price on the agricultural products transmitted differently to the different countries?}

The introduction now has more emphasis on the paper's contribution to the price transmission literature which mostly consists of single country, single commodity case studies. I also stress the policy implications of the main findings.

\item \emph{A clarification in the introduction on how authors chose the LIC, MIC or HIC groups and what database was used for such segregation (I assume World Bank database?) and actual country names by group (Appendix?) if other than standard WB methodology would be helpful.}

I have included a footnote in the analysis pointing out that I use the World Bank income classification scheme for the country groupings.

\item \emph{Literature Review section should represent literature related to the problem and suggesting why this question is relevant and what effect it has. Provide more theoretical
background as well to set the discussion in appropriate context.}

In light of this criticism I have thoroughly revised the literature review section to make it clear that its purpose is to identify factors which the literature considers important determinants of price transmission. Regarding the theoretical background, I now discuss in greater detail than before how various country characteristics are believed to affect price transmission.

\item \emph{Not clear how table 6 is constructed. Please spend some time describing the table 6 in more details.}

I have changed the table notes so that it should be clear what the numbers represent. I have also revised the main text.

\end{itemize}

\subsubsection*{Minor Comments:}
\begin{itemize}

\item \emph{Instead of rewriting equation (11), author can make just a reference to equation (1) and describe the adjustments that were made.}

I have skipped the old equation (1) instead.

\item \emph{Table 5 provides results for the panel regression. What period is this for?}

The panel regressions use monthly price data from the entire 2005-15 period. I have corrected some mistakes in the old table and revised the table notes.

\item \emph{Typos: p.2, two lines above last paragraph, double on;}

Corrected. I have rewritten the introduction to accommodate comments from both reviewers.

\item \emph{P.7, 5th line from the bottom, double domestic}

Corrected.

\item \emph{P.12, 2nd line from top, double their}

Corrected.

\item \emph{P. 16, the IMF}

Corrected. I have rewritten section 2 and 3 to accommodate comments from both reviewers.

\item \emph{P. 17 should be tables 4 and 6.}

Corrected. Note that there are only 4 tables in the revised main text.

\end{itemize}

\end{document}

### 2016-03-02 reply_reviewer_R5.Rnw
reply to editor and reviwers in connection with 2. round rev. version of paper

\documentclass[12pt,english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{babel}
\usepackage{setspace}
\doublespacing
\usepackage{geometry}
\geometry{verbose,tmargin=2cm,bmargin=2cm,lmargin=3cm,rmargin=3cm}
\usepackage{color,soul}
\usepackage[authoryear]{natbib}

\begin{document}
\SweaveOpts{concordance=TRUE}

\title{Reply to reviewer R5}
\date{\vspace{-5ex}}
\maketitle

\subsection*{Reviewer R5}
\subsubsection*{Major issues:}

\begin{itemize}

\item \emph{The focus and research question of the paper should be broadened and not focus too narrowly on the role of income as explanation for price transmission. Transportation costs (infrastructure) are also relevant; the role of cereals' share in the diet and cereal import dependency is considered in the regression model but not in the theoretical section 3; cereals export share is, however, neglected (which should be considered together with imports). My recommendation would be to have a broader conceptual framework section 3 which discusses in general the determinants of price transmission and the related literature (see also the paper of Greb et al. (2012) on a meta-regression on co-integration). The theoretical model can be shortened and integrated in this section or put in the appendix.}

I have broadened the research question and the analysis as requested. The paper now connects stronger to the \citet{Lee2013} and \citet{greb12} papers. I am grateful, by the way, for bringing the \citet{greb12} study to my attention. However, the existence of an inverse-U shaped income-price transmission relationship is a novel finding and I therefore find that an extra focus on this is justified. Instead of relegating the conceptual model to the appendix I have incorporated the additional determinants into the framework. I have also shortened somewhat the description of the model.

\item \emph{The choice of 6 lags for the regression of (11) is problematic with only 72 observations. The lag length should be based on some criterion (be it economic theory or information criterion, which could be applied for all countries and the choice could be based on the median optimal lag-length).}

Thanks for pointing out this. As I discuss in the revised section 3, lag length selection based on the AIC leads to a median optimal lag length of 1. However, the long run multipliers (LRMs) based on ADL models with a single lag turn out very low on average; much lower than is commonly found in the literature. Also, the LRMs based on a single lag are very imprecise. The reported LRMs based on 6 lags, therefore represent a compromise between the 12 lag models found in the literature and 1 lag models implied by the information criterion. Results based on time series models with a single lag are reported in the appendix.

\item \emph{The consideration of food consumption expenditures (and its square) as explanatory variable in the regression models lack any theoretical argument and should therefore be removed; the author has only argued for considering income (GDP or consumption expenditure as proxy), but I cannot see why food expenditure should influence price transmission.}

In light of the request that I should broaden the analysis and consider other determinants than income, I have narrowed the income-related explanatory variables down to one, namely per-capita GDP in constant 2011 international \$.

\item \emph{International prices and domestic prices are reported in different currencies. Shouldn't they be converted into the same currency or shouldn't alternatively exchange rates be integrated as explanatory variable to reduce standard errors or avoid potential omitted variable biases?}

I have included exchange rate changes as an explanatory variable. Thanks for the suggestion.

\item \emph{The regression of the estimated long run multipliers in Eq. (13) lacks a critical discussion and reference to techniques of meta-analysis (see, e.g. Stanley and Doucouliagos). Contrary to prices, the estimated long-run multiplier is not a random draw from a population. Therefore, some theoretical justification is needed for applying asymptotic theory. Also, it is neglected whether the long-run multipliers are statistically significant or not (inverse standard errors or sample variance might be integrated as weights for the meta regression).}

Again, thanks for pointing out this to me; I did not think of this before. In my defense, none of the other papers from the price transmission literature, including Kalkuhl (2014), mention this issue. Nonetheless, the revised section 3 discusses this problem at some length. The discussion ends up with a regression table comparing OLS and WLS estimates, both with robust standard errors.

\item \emph{The regression of Eq. (14) suffers from dynamic panel bias (Nickell 1981). As T=72, the bias might be small, but this issues should be discussed and/or system or difference GMM regression provided as robustness check.}

Given the extended period 2005-15 (see comment below), T=120 and the bias should be even smaller. However, not mentioning the Nickell bias in the first version of the paper was a mistake (which is now corrected). I tried to estimate the model with GMM but interaction terms are not permitted in standard statistical software packages such as STATA or R.

\item \emph{Also related to (14): The choice of the lag-length for the endogenous variable is ad-hoc; the 6-month lag for the international price index is ad-hoc and needs justification. Most time-series models show that transmission is highest in the same month or on the following month, suggesting the choice of 0 or 1 lag. The choice of the lag-length could also be supported by using information criteria.}

Again, thanks for pointing this out. You are right, transmission turns out to be highest for the first lag so I have revised the model accordingly. There is also some discussion of this in the revised section 3.

\item \emph{On the conclusion part: It is true that most of the global poor live in middle income countries. The regression analyses of the paper, however, neglected the size of the poor population. Time-series models on individual countries, for example, find very low transmission (or no statistically significant transmission) to India and China (Kalkuhl 2014). These two countries also intervene regularly on domestic markets to stabilize prices. Thus, the implications for poor people worldwide are much more nuanced than suggested by the inverse u-shaped income-relationship.}

It is true, as you point out, that the two giants China and India intervene heavily in their grain markets and that grain prices there remained comparatively stable during the 2007-08 food crisis. However, to my knowledge, there is little evidence suggesting that aggregate food prices in those countries are more stable and less responsive to international shocks than those in other countries. Besides, Kalkuhl (2014) estimates a grain price transmission elasticity of .42 for China based on the FAO FPI, which is fairly high. The number for India is .19 which is somewhat lower but still much higher than many low-income countries' price transmission rate. I would like to have gone into a discussion of this issue in the paper but I am up against the word limit. In order to quantify the effects of the policy interventions during the food crisis I included a set of intervention dummies in the food crisis regressions (table 6 in the appendix). However, none of the intervention dummies turned out significant. Besides, there is a potential endogeneity issue with these policy dummies which is why I do not emphasize these results.

\end{itemize}

\subsubsection*{Minor issues:}

\begin{itemize}

\item \emph{Is the international food price index by FAO deflated? If not, it should be done as also deflated ('real') domestic food price indices are used.}

It was not before. I now use the deflated FAO FPI in connection with the Real FPI (RFPI) regressions.

\item \emph{There are several typos in the manuscript and the author should proof-read it carefully again}

I have tried to weed out all the typos in the revised manuscript.

\item \emph{Why has only the timer period of 2005 to 2012 been considered and price data for 2013-2014 neglected?}

I had no good reason for doing this. The actual reason is that I started working on the paper several years ago and I did not update the data as time passed. However, the revised manuscript is based on price data covering the 10 year period 2005-15 (not including 2015 data).

\item \emph{On the problem of changing base years of the domestic food price indices. The price series of LABORSTAT contain for many countries comments when base years have changed; thus, it is possible to re-index most price series to a consistent base year (leaving aside the issue of changing baskets or weights, which should best be addressed by an additional dummy variables which is included in each regression, or by just dropping the observation following a change of the basket).}

I use January 2007 as the base period for all countries in the revised version. The problem with the strategy of addressing base year changes and other structural breaks with dummies is that it only works with time series models. For the food inflation regression in table 1, the dependent variable needs to represent actual food inflation rather than a structural break. This is why I chose to fix the breaks the way I do.

\end{itemize}

\bibliographystyle{chicago}
\bibliography{inc_pt}

\end{document}

### 2016-03-02 reply_rev1.Rnw
reply to editor and reviwers in connection with 2. round rev. version of paper

\documentclass[12pt,english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{babel}
\usepackage{setspace}
\doublespacing
\usepackage{geometry}
\geometry{verbose,tmargin=2cm,bmargin=2cm,lmargin=3cm,rmargin=3cm}
\usepackage{color,soul}
\usepackage[authoryear]{natbib}

\begin{document}
\SweaveOpts{concordance=TRUE}

\title{Reply to the editor and to the reviewers}
\date{\vspace{-5ex}}
\maketitle

Dear Dr. Abdulai. I am pleased to resubmit for possible publication a thoroughly revised version of MS 2015310 "Income and Food Price Transmission". I have done my best to accommodate the constructive criticisms by the two reviewers and yourself. You stressed that the introduction needs to focus more on the contribution to the literature. You also stressed that I should broaden the research question and consider additional factors such as infrastructure that are commonly believed to influence price transmission. This second request in particular meant that I had to to rewrite the whole paper, but I think quality of the paper has benefited from this.

Below I respond to the reviewer comments one by one. I also respond to each of the reviewers in separate documents.

\subsection*{Reviewer R1}
\subsubsection*{Some Formal Aspects:}
\begin{itemize}
\item \emph{Please include Abstract and Keywords on the first page.}

Done.

\item \emph{In the beginning of the paper, it would also be good to mention why this study is important. For example, why is it important to look at how price on the agricultural products transmitted differently to the different countries?}

The introduction now has more emphasis on the paper's contribution to the price transmission literature which mostly consists of single country, single commodity case studies. I also stress the policy implications of the main findings.

\item \emph{A clarification in the introduction on how authors chose the LIC, MIC or HIC groups and what database was used for such segregation (I assume World Bank database?) and actual country names by group (Appendix?) if other than standard WB methodology would be helpful.}

I have included a footnote in the analysis pointing out that I use the World Bank income classification scheme for the country groupings.

\item \emph{Literature Review section should represent literature related to the problem and suggesting why this question is relevant and what effect it has. Provide more theoretical
background as well to set the discussion in appropriate context.}

In light of this criticism I have thoroughly revised the literature review section to make it clear that its purpose is to identify factors which the literature considers important determinants of price transmission. Regarding the theoretical background, I now discuss in greater detail than before how various country characteristics are believed to affect price transmission.

\item \emph{Not clear how table 6 is constructed. Please spend some time describing the table 6 in more details.}

I have changed the table notes so that it should be clear what the numbers represent. I have also revised the main text.

\end{itemize}

\subsubsection*{Minor Comments:}
\begin{itemize}

\item \emph{Instead of rewriting equation (11), author can make just a reference to equation (1) and describe the adjustments that were made.}

I have skipped the old equation (1) instead.

\item \emph{Table 5 provides results for the panel regression. What period is this for?}

The panel regressions use monthly price data from the entire 2005-15 period. I have corrected some mistakes in the old table and revised the table notes.

\item \emph{Typos: p.2, two lines above last paragraph, double on;}

Corrected. I have rewritten the introduction to accommodate comments from both reviewers.

\item \emph{P.7, 5th line from the bottom, double domestic}

Corrected.

\item \emph{P.12, 2nd line from top, double their}

Corrected.

\item \emph{P. 16, the IMF}

Corrected. I have rewritten section 2 and 3 to accommodate comments from both reviewers.

\item \emph{P. 17 should be tables 4 and 6.}

Corrected. Note that there are only 4 tables in the revised main text.

\end{itemize}

\subsection*{Reviewer R5}
\subsubsection*{Major issues:}

\begin{itemize}

\item \emph{The focus and research question of the paper should be broadened and not focus too narrowly on the role of income as explanation for price transmission. Transportation costs (infrastructure) are also relevant; the role of cereals' share in the diet and cereal import dependency is considered in the regression model but not in the theoretical section 3; cereals export share is, however, neglected (which should be considered together with imports). My recommendation would be to have a broader conceptual framework section 3 which discusses in general the determinants of price transmission and the related literature (see also the paper of Greb et al. (2012) on a meta-regression on co-integration). The theoretical model can be shortened and integrated in this section or put in the appendix.}

I have broadened the research question and the analysis as requested. The paper now connects stronger to the \citet{Lee2013} and \citet{greb12} papers. I am grateful, by the way, for bringing the \citet{greb12} study to my attention. However, the existence of an inverse-U shaped income-price transmission relationship is a novel finding and I therefore find that an extra focus on this is justified. Instead of relegating the conceptual model to the appendix I have incorporated the additional determinants into the framework. I have also shortened somewhat the description of the model.

\item \emph{The choice of 6 lags for the regression of (11) is problematic with only 72 observations. The lag length should be based on some criterion (be it economic theory or information criterion, which could be applied for all countries and the choice could be based on the median optimal lag-length).}

Thanks for pointing out this. As I discuss in the revised section 3, lag length selection based on the AIC leads to a median optimal lag length of 1. However, the long run multipliers (LRMs) based on ADL models with a single lag turn out very low on average; much lower than is commonly found in the literature. Also, the LRMs based on a single lag are very imprecise. The reported LRMs based on 6 lags, therefore represent a compromise between the 12 lag models found in the literature and 1 lag models implied by the information criterion. Results based on time series models with a single lag are reported in the appendix.

\item \emph{The consideration of food consumption expenditures (and its square) as explanatory variable in the regression models lack any theoretical argument and should therefore be removed; the author has only argued for considering income (GDP or consumption expenditure as proxy), but I cannot see why food expenditure should influence price transmission.}

In light of the request that I should broaden the analysis and consider other determinants than income, I have narrowed the income-related explanatory variables down to one, namely per-capita GDP in constant 2011 international \$.

\item \emph{International prices and domestic prices are reported in different currencies. Shouldn't they be converted into the same currency or shouldn't alternatively exchange rates be integrated as explanatory variable to reduce standard errors or avoid potential omitted variable biases?}

I have included exchange rate changes as an explanatory variable. Thanks for the suggestion.

\item \emph{The regression of the estimated long run multipliers in Eq. (13) lacks a critical discussion and reference to techniques of meta-analysis (see, e.g. Stanley and Doucouliagos). Contrary to prices, the estimated long-run multiplier is not a random draw from a population. Therefore, some theoretical justification is needed for applying asymptotic theory. Also, it is neglected whether the long-run multipliers are statistically significant or not (inverse standard errors or sample variance might be integrated as weights for the meta regression).}

Again, thanks for pointing out this to me; I did not think of this before. In my defense, none of the other papers from the price transmission literature, including Kalkuhl (2014), mention this issue. Nonetheless, the revised section 3 discusses this problem at some length. The discussion ends up with a regression table comparing OLS and WLS estimates, both with robust standard errors.

\item \emph{The regression of Eq. (14) suffers from dynamic panel bias (Nickell 1981). As T=72, the bias might be small, but this issues should be discussed and/or system or difference GMM regression provided as robustness check.}

Given the extended period 2005-15 (see comment below), T=120 and the bias should be even smaller. However, not mentioning the Nickell bias in the first version of the paper was a mistake (which is now corrected). I tried to estimate the model with GMM but interaction terms are not permitted in standard statistical software packages such as STATA or R.

\item \emph{Also related to (14): The choice of the lag-length for the endogenous variable is ad-hoc; the 6-month lag for the international price index is ad-hoc and needs justification. Most time-series models show that transmission is highest in the same month or on the following month, suggesting the choice of 0 or 1 lag. The choice of the lag-length could also be supported by using information criteria.}

Again, thanks for pointing this out. You are right, transmission turns out to be highest for the first lag so I have revised the model accordingly. There is also some discussion of this in the revised section 3.

\item \emph{On the conclusion part: It is true that most of the global poor live in middle income countries. The regression analyses of the paper, however, neglected the size of the poor population. Time-series models on individual countries, for example, find very low transmission (or no statistically significant transmission) to India and China (Kalkuhl 2014). These two countries also intervene regularly on domestic markets to stabilize prices. Thus, the implications for poor people worldwide are much more nuanced than suggested by the inverse u-shaped income-relationship.}

It is true, as you point out, that the two giants China and India intervene heavily in their grain markets and that grain prices there remained comparatively stable during the 2007-08 food crisis. However, to my knowledge, there is little evidence suggesting that aggregate food prices in those countries are more stable and less responsive to international shocks than those in other countries. Besides, Kalkuhl (2014) estimates a grain price transmission elasticity of .42 for China based on the FAO FPI, which is fairly high. The number for India is .19 which is somewhat lower but still much higher than many low-income countries' price transmission rate. I would like to have gone into a discussion of this issue in the paper but I am up against the word limit. In order to quantify the effects of the policy interventions during the food crisis I included a set of intervention dummies in the food crisis regressions (table 6 in the appendix). However, none of the intervention dummies turned out significant. Besides, there is a potential endogeneity issue with these policy dummies which is why I do not emphasize these results.

\end{itemize}

\subsubsection*{Minor issues:}

\begin{itemize}

\item \emph{Is the international food price index by FAO deflated? If not, it should be done as also deflated ('real') domestic food price indices are used.}

It was not before. I now use the deflated FAO FPI in connection with the Real FPI (RFPI) regressions.

\item \emph{There are several typos in the manuscript and the author should proof-read it carefully again}

I have tried to weed out all the typos in the revised manuscript.

\item \emph{Why has only the timer period of 2005 to 2012 been considered and price data for 2013-2014 neglected?}

I had no good reason for doing this. The actual reason is that I started working on the paper several years ago and I did not update the data as time passed. However, the revised manuscript is based on price data covering the 10 year period 2005-15 (not including 2015 data).

\item \emph{On the problem of changing base years of the domestic food price indices. The price series of LABORSTAT contain for many countries comments when base years have changed; thus, it is possible to re-index most price series to a consistent base year (leaving aside the issue of changing baskets or weights, which should best be addressed by an additional dummy variables which is included in each regression, or by just dropping the observation following a change of the basket).}

I use January 2007 as the base period for all countries in the revised version. The problem with the strategy of addressing base year changes and other structural breaks with dummies is that it only works with time series models. For the food inflation regression in table 1, the dependent variable needs to represent actual food inflation rather than a structural break. This is why I chose to fix the breaks the way I do.

\end{itemize}

\bibliographystyle{chicago}
\bibliography{inc_pt}

\end{document}

### 2016-03-02 ts_data_prep.R
time series preparation script for 2. round rev. version of paper

@@ -1,0 +1,184 @@
+#####################################################################
+#### Creating the time series dataset
+#####################################################################
+
+library(dplyr)
+library(zoo)
+library(foreign)
+library(countrycode)
+library(reshape)
+library(reshape2)
+library(stringr)
+
+### Raw Data
+# er.csv
+# fao_fpi.csv
+
+str = "C:/Users/nqw235/Google Drev/Research/pov_price_transmission/"
+
+#####################################################################
+
+## Int. food prices (FAO FPI)
+## Source: FAO
+temp = paste(str, "fao_fpi.csv", sep = "")
+fao.fpi = read.csv(temp, sep = ";", header = TRUE)
+fao.fpi$tt = seq(as.Date("2000/1/1"), by = "month", length.out = nrow(fao.fpi))
+fao.fpi$tt = as.yearmon(fao.fpi$tt)
+fao.fpi$date = NULL
+
+## Int. oil prices (POILAPSP)
+## Source: IMF
+temp = paste(str, "imf_oil.csv", sep = "")
+imf.oil = read.csv(temp, sep = ";", header = TRUE)
+imf.oil$tt = seq(as.Date("1980/1/1"), by = "month", length.out = nrow(imf.oil))
+imf.oil$tt = as.yearmon(imf.oil$tt)
+imf.oil$date = NULL
+
+## Exchange rates (dom. currency/US$)
+## Source: IMF, IFS
+temp = paste(str, "er.csv", sep = "")
+er = read.csv(temp, sep = ";", header = TRUE)
+eu = select(er, Euro.Area)
+eu$tt = seq(as.Date("2000/1/1"), by = "month", length.out = nrow(eu))
+eu$tt = as.yearmon(eu$tt)
+
+er.nam = data.frame(names(er)[-1])
+colnames(er.nam)[1] = "country"
+er.nam$iso3c = factor(countrycode(er.nam$country,  "country.name", "iso3c", warn = TRUE))
+er.nam$x = paste0("ctr_", er.nam$iso3c)
+
+for (i in 1: nrow(er.nam)) {
[…]

### 2016-03-02 ur_test.R
unit root test script for 2. round rev. version of paper

@@ -1,0 +1,120 @@
+
+library(dplyr)
+library(urca)
+library(tseries)
+
+##################################################################
+### Unit root and cointegration tests
+##################################################################
+
+load("ts_dat.Rda")
+data = ts.dat
+tbl = table(data$id)
+data$id = as.numeric(data$id)
+
+## FPI inflation
+test.mat = matrix(NA, nrow = nrow(tbl), ncol = 8)
+for (i in 1: nrow(tbl)) {
+  dat = dplyr::filter(data, id==i)
+  dat = dplyr::select(dat, cpi.brk:rfpi, fao.fpi:er)
+  
+  # KPSS test
+  kpss.lev = kpss.test(log(dat$adj.fpi),  null = "L", lshort = FALSE)
+  kpss.dif = kpss.test(diff(log(dat$adj.fpi)),  null = "L", lshort = FALSE)
+  test.mat[i,1] = kpss.lev[[1]]
+  test.mat[i,2] = kpss.dif[[1]]
+  test.mat[i,3] = kpss.lev[[3]]
+  test.mat[i,4] = kpss.dif[[3]]
+
+  # ADF test
+  adf.lev = adf.test(log(dat$adj.fpi), alternative = "s")
+  adf.dif = adf.test(diff(log(dat$adj.fpi)), alternative = "s")
+  test.mat[i,5] = adf.lev[[1]]
+  test.mat[i,6] = adf.dif[[1]]
+  test.mat[i,7] = adf.lev[[4]]
+  test.mat[i,8] = adf.dif[[4]]
+}
+ur.test = as.data.frame(test.mat)
+ur.test$iso3c = names(tbl)
+
+## real FPI growth
+test.mat = matrix(NA, nrow = nrow(tbl), ncol = 8)
+for (i in 1: nrow(tbl)) {
+  dat = dplyr::filter(data, id==i)
+  dat = dplyr::select(dat, cpi.brk:rfpi, fao.fpi:er)
+  
+  # KPSS test
+  kpss.lev = kpss.test(log(dat$rfpi),  null = "L", lshort = FALSE)
+  kpss.dif = kpss.test(diff(log(dat$rfpi)),  null = "L", lshort = FALSE)
+  test.mat[i,1] = kpss.lev[[1]]
+  test.mat[i,2] = kpss.dif[[1]]
+  test.mat[i,3] = kpss.lev[[3]]
+  test.mat[i,4] = kpss.dif[[3]]
+  
[…]

### 2016-06-21 Master.tex
Initial changes in connection with AE rejection

--- 2016-02-29 analysis.Rnw
@@
+\documentclass[12pt,english]{article}
+\usepackage{color,soul}
+\usepackage[T1]{fontenc}
+\usepackage{inputenc}
+\usepackage{babel}
+\usepackage{url}
+\usepackage{graphicx}
+\usepackage[authoryear]{natbib}
+\usepackage[unicode=true]{hyperref}
+\usepackage[format=hang, justification=raggedright]{subfig}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{lscape}
+\usepackage{float}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage[noae]{Sweave}
+\usepackage{setspace}
+\doublespacing
+\usepackage{geometry}
+\geometry{verbose,tmargin=2cm,bmargin=2cm,lmargin=3cm,rmargin=3cm}
+\usepackage{amsmath}
+\usepackage[flushleft]{threeparttable}
+
+\begin{document}
+\input{Master-concordance}
+
+\title{Income and Food Price Transmission\footnote{Revised version of MS 2015310 ``Income and Price Transmission''.
+Submitted to \emph{Agricultural Economics} for 2.
+round review.
+I thank two anonymous reviewers for their valuable comments.}}
+
+\date{\vspace{-5ex}}
+
+\maketitle
+
+\begin{abstract}
+This paper examines determinants of international-to-domestic food price transmission using aggregate monthly price data for more than 100 countries covering the ten year period 2005-15.
+Findings indicate that middle-income countries experience the highest rates of food price transmission in the long run.
+Other determinants of food price transmission include a country's cereal import, export and consumption shares, its trade openness, the quality of its infrastructure and government involvement in grain trade.
+A novel model is presented which is able to explain the economic logic behind the observed inverse-U shaped relationship between income and price transmission.
+
+\end{abstract}
+
+\textbf{Keywords:} price transmission, food crisis, food inflation, determinants, regression analysis, income\newpage
+
+% !Rnw root = master.Rnw
+
+\section{Introduction}
+In response to the substantial increase in international food prices during the first decade of the new millennium, a large number of recent studies have quantified global and local poverty impacts of international price shocks.\footnote{E.g.
+\cite{Attanasio2013,Balagtas2014,Coxhead2012,Ferreira2013,Hoyos2011,Ivanic2012,Janvry2010,Mghenyi2011,Vu2011,Wodon2010}.
+In addition, a supplement to volume 39 of \emph{Agricultural Economics} is entirely devoted to this topic \citep[see][]{masters08}.} A major determinant of a country's vulnerability to such shocks is the responsiveness of its domestic food prices \citep[e.g.][]{Hoyos2011}.
+But what are the main determinants of the domestic price responsiveness to international shocks?
+Are they related to trade, consumption, geography or something else?
+As the title of the paper suggests, a country's income level seems to be an important determinant but other factors also play a role.
+In particular, countries which control their international trade through an agricultural state trading enterprise are found to experience lower food price transmission on average in the short run as well as in the long run.
+
+The literature on domestic price impacts of international commodity price shocks is vast and the topic has experienced a revival in the wake of the 2007-08 food crisis \citep[see][]{Fackler2001, kouyate16}.
+However, since the majority of price transmission studies focus on a single country we know little about the determinants, other than distance and borders, of price transmission differences across countries.
+Moreover, most price transmission studies focus on individual commodity prices rather than food prices in general.
+This paper therefore contributes with a cross country analysis of price transmission determinants that is based on aggregate food price data, i.e.
+food price indexes, rather than prices of individual agricultural commodities.
+
+Regarding the choice between individual or aggregate prices, each approach has its strengths and weaknesses \citep{headey10}.
+Commodity specific analyses have a stronger microeconomic foundation in the law of one price but results do not generalize to other food goods or even to the same food good traded in other domestic markets.
+Transmission estimates based on aggregate price data, on the other hand, give a more comprehensive picture and therefore better reflect domestic welfare costs of international food price fluctuations.
+A potential weakness though of the latter approach, is that a country's food price index might not reflect the composition of poor people's food baskets which is often based mainly on grain.
+Another weakness is that the composition of a country's food price index differs substantially from that of the international food price indexes and as well as from other countries' price indexes, which makes it more difficult to compare transmission rates across countries.
+These idiosyncrasies do not, however, preclude us from studying broad patterns in price transmission rates.
+
+The empirical analysis is based on insights derived from a simple economic model that is more suitable for aggregate price data than the law of one price workhorse model.
+A key prediction is that a country's income level should have an independent effect on price transmission, in addition to factors related to trade openness, geography and so on.
+The model predicts a novel inverse-U shaped relationship between a country's income level and its price transmission rate.
+This income-price transmission relationship turns out to be a robust feature of the data.
+
+There are also other significant predictors of the domestic responsiveness to international food price shocks.
+Countries which import a large share of their grains and countries whose diets are mainly based on cereals are, surprisingly, found to have lower price transmission on average whereas the opposite holds for countries which export most of their grain.
+Countries that are more open in general in terms of their trade to GDP ratio also experience more price transmission whereas countries with an agricultural state trading enterprise experience lower price transmission.
+These country characteristics also significantly explain differences in domestic food inflation during the 2007-08 food crisis.
+
+What are the policy implications of these findings?
+One obvious implication is that we may have to revise somewhat our perception of vulnerability and country focus in connection with international food crises.
+Donors and international aid organizations tend to a focus mostly on low-income food deficit countries but, according to the estimates in this paper, it seems that poor households living in middle-income countries are actually the most vulnerable to an international food price spike.
+Moreover, it does not seem to be the case that international price changes in general are passed on at a higher rate to countries that import most of their grains and this was not the case either during the 2007-08 food crisis.
+
+The paper now proceeds as follows.
+Section \ref{sec:model} discusses the determinants of price transmission and lays out a model of a country's food price index.
+Section \ref{sec:analysis} presents the empirical analysis and, finally, section \ref{sec:conclusion} offers some concluding remarks.
+% !Rnw root = master.Rnw
+
+\section{Determinants of food price transmission\label{sec:model}}
+What are the main determinants of a country's responsiveness to an international food price shock?
+I.e., which factors would we expect to influence the relationship between international and domestic food prices?
+To answer this question this section first reviews literature.
+Next, a simple model is presented of a country's food price level and its responsiveness to international shocks based on insights from the literature review.
+
+\subsection{Literature review}
+There are only a few studies carrying out actual cross section analyses of price transmission determinants.
+There are, however, several studies discussing this informally, especially in relation to the 2007-08 food crisis.
+Arguments and main findings from a selection of these studies are summarized below.
+
+\citet{Hoyos2011} point out that domestic food inflation during the food crisis varied a lot across countries.\footnote{Domestic food inflation is defined as the percentage change in a country's Food Price Index (FPI) which is a component of the Consumer Price Index (CPI).} Furthermore, domestic food inflation was lower on average than international food inflation.
+\citeauthor{Hoyos2011} offer 4 reasons why was is the case: (i) weak price transmission perhaps due to poor infrastructure, (ii) differences in the composition of food baskets and the fact that domestic food baskets contain nontradables, (iii) lack of competition in the domestic food markets and (iv) government interventions for example in the form of price controls.
+
+\citet{Abbott2009} discusses the causes and consequences of the food crisis.
+Regarding the cross country differences in food inflation rates, he makes three observations: i) developing countries have more basic diets which involve a greater share of staples; ii) processing and distribution margins are typically lower than in developed countries, which means that they can shrink less as costs increase.
+Both factors, he argues, caused a stronger domestic food inflation response among developing countries than in developed countries to the high world prices in 2007-08.
+And, iii) a country's exchange rate regime has a large impact on the domestic price impact of an international shock; countries which peg their currency to the dollar are more exposed to international price fluctuations than countries whose currencies follow the euro.
+
+\citet{Lee2013} quantify the importance of various ``external'' and ``internal'' variables on domestic food inflation based on data from 72 countries over the period 2000 - 2011.
+The set of external variables consists of current and lagged international food inflation, as well as a set of intra- and extra-regional food inflation averages.
+The internal variables represent domestic supply and demand factors as well as various other country characteristics.
+What they find is that  countries with large food import shares have lower food inflation whereas countries experiencing large increases in their food import shares have higher food inflation on average.
+The latter result could, however, reflect that food imports increase when domestic domestic food prices are on the rise, as a response to food inflation rather than a cause of it.
+Surprisingly, the authors also find a negative relationship between a country's per capita GDP growth rate (a demand variable) and its food inflation rate.
+The relationship between domestic food inflation and the per capita GDP level is found to be positive albeit only marginally significant in some specifications.
+A depreciation of the domestic currency/US\$ exchange rate increases domestic food inflation significantly across all specifications, as expected, whereas an increase in the money supply has no effect on food inflation.
+More politically stable countries tend to have lower food inflation whereas countries with a higher free trade index score are found to have higher domestic food inflation.
+
+Two studies quantify the determinants of price transmission and market integration based on individual agricultural commodity prices rather than food price indexes.
+\citet{kouyate16} carry out a meta-regression analysis of the effect of distance on agricultural commodity price transmission for a large number of market pairs.
+Their results show that geographical distance and separation by borders reduces the likelihood of price cointegration and weakens the speed of speed of price transmission significantly.
+These findings are consistent with the well established stylized facts from the trade literature, that borders and distance reduces trade.
+
+Finally, \citet{greb12} consider 8 variables covering geographical, infrastructural, institutional and commodity specific factors which are believed to influence agricultural price transmission.
+The authors regress their own price transmission estimates, as well as estimates obtained from the literature, on these country characteristics.
+What they find is that the covariates explain more of the variation in the estimated speed of adjustment parameters (the $\alpha$'s) than in the cointegration parameters (the $\beta$'s).
+Another major finding is that more of these covariates have significant and plausible partial effects on the $\alpha$'s than on the $\beta$'s.
+However, none of the determinants have a robust robust impact on price transmission.
+
+\subsection{A model of a country's food price level}
+This section presents a simple model of a country's food price level that is inspired by the arguments and findings discussed above, related to determinants of domestic food inflation and price transmission.
+The benefit of a working with a formal economic model is that it imposes some structure on the empirical analysis.
+In fact, despite its simplicity, it is possible to derive several predictions from the model regarding the determinants of price transmission.
+
+A country's consumer price index (CPI) can be expressed as a weighted average
+\begin{equation}
+P =a P_{f}+(1-a)P_{n}\label{eq:cpi},
+\end{equation}
+where $P_{f}$ and $P_{n}$ denote the food and non-food price indexes, respectively, and $a$ is the food expenditure share.
+Time indexes are omitted in order to avoid notational clutter.
+The food price index (FPI), which is of primary interest, can be written
+\begin{equation}
+P_{f} =b P_{c}+(1-b)P_{m}\label{eq:fpi},
+\end{equation}
+where $P_{c}$ and $P_{m}$ denote the farm cost and marketing components of the FPI, respectively, and $b$ is the farm cost or commodity share.\footnote{By marketing costs I refer to all other costs than those of the raw agricultural commodities included in the final consumer food price such as those associated with distribution, storage and promotion.}  Finally, the commodity component of the domestic FPI can be decomposed as
+\begin{equation}
+P_{c} =c P_{w}+(1-c)P_{d}\label{eq:compi},
+\end{equation}
+where $P_{w}$ and $P_{d}$ denote the tradable and purely domestic subcomponents of the commodity price index, respectively, and $c$ is the share of tradable food commodities in $P_{c}$.
+The defining difference between tradable and domestic food commodities is that the former have prices which are determined on the world market whereas the latter prices are determined by domestic supply and demand.
+
+The three share parameters, $a$, $b$ and $c$ represent weights on price indexes that are progressively embedded into each other. By substituting (\ref{eq:compi}) into (\ref{eq:fpi}) we get
+\begin{equation}
+P_{f} = b c P_{w}+R\label{eq:fpi1},
+\end{equation}
+where $R=b(1-c)P_{d}+(1-b)P_{m}$ is a term which does not depend on world market prices directly.
+With these definitions a first-order approximation to transmission of international agricultural commodity price changes to domestic consumer food prices can be written
+\begin{equation}
+\theta\equiv\frac{\partial P_{f}}{\partial P_{w}}=bc,\label{eq:pt}
+\end{equation}
+i.e., price transmission depends multiplicatively on the share of tradable commodities in the commodity price index, $c$, and the commodity cost share in the FPI, $b$.
+Expression (\ref{eq:pt}) is based on the assumption that there is no knock-on effect on domestic domestic prices when world market prices change, as would be appropriate in the short term.
+If we relax this assumption and allow domestic prices to respond to international price changes through substitution effects we would have to add the term $b(1-c)\frac{\mathrm{d}P_{d}}{\mathrm{d}P_{w}}$.
+For simplicity, I will ignore this second order effect in the analysis below.
+
+By imposing some structure on the two cost shares, $b$ and $c$ we will be able to analyze determinants of a country's price transmission.
+In particular, we can write these shares as functions of certain country characteristics:
+\begin{equation}
+b=b(x_{b}), \quad c=c(x_{c}).
+\end{equation}
+But what are these characteristics? As I will argue below, a country's income level should feature prominently among them.
+
+We can reasonably assume that the wealthier the country, the lower the commodity cost share in the FPI, c.f.
+\citet{Abbott2009}.
+Formally, if $y$ denotes income then $y\in x_{b}$ and $\frac{\partial b}{\partial y} <0$.
+Furthermore, if we believe that countries become more integrated with the world market as they develop, not only due to better infrastructure and more open markets, but also due to for example higher quality standards associated with the adoption of modern production and management practices, then it would be appropriate to also assume $y\in x_{c}$ and $\frac{\partial c}{\partial y} >0$.\footnote{We must also impose the following set of restrictions, $\lim_{y\rightarrow \infty} b(y)=0$, $\lim_{y\rightarrow \infty} c(y)=1$, $\lim_{y\rightarrow 0^{+}} b(y)=1$ and $\lim_{y\rightarrow 0~{+}} c(y)=0$, since otherwise these functions could not represent cost shares.} These assumptions in combination imply an inverse-U shaped relationship between price transmission, $\theta$, and income.
+
+Income is not the only determinant of price transmission of course.
+A country's openness to trade, the quality of its infrastructure and the composition of its food consumption set are all factors which we would expect to influence the transmission of world market price fluctuations to domestic food markets.
+Economies that are more open in general, perhaps for historical reasons, should have a higher tradables share and therefore experience more price transmission.
+Similarly with countries that, due to climatic or geographical reasons, are forced to import most of their food supply.
+On the other hand, there are countries that are self-sufficient due to poor infrastructure, perhaps in combination with landlockedness, resulting in prohibitive transport costs.
+The tradables share in these countries should be very low and so should the price transmission rate.
+Consumption patterns are influenced by cultural as well as geographical factors.
+In particular, some countries consume more grains than others and the commodity component in the food price index should therefore be larger.
+Consequently, domestic aggregate food prices are more exposed international shocks in such countries.
+Finally, there are policies that are capable of affecting both shares.
+Trade policies in particular have a direct effect on the tradability of a good.
+
+To sum up, there are reasons to expect a domestic FPI reaction to price fluctuations on the international agricultural commodity markets which depends non-monotonically on income.
+Specifically, the simple model sketched above predicts an inverse-U shaped relationship between income and price transmission.
+Factors related to trade openness, consumption, geography and the quality of infrastructure also, arguably, affect price transmission.
 % !Rnw root = master.Rnw
@@
 \subsection{Data\label{sec:data}}
-
-<<echo=FALSE>>=
-load("infl_0708_1011.Rda")
-load("ts_dat.Rda")
-n.full = nrow(table(ts.dat$id))
-n.sub = nrow(infl)
-infl = infl[infl$iso3c != "BLR", ]
-@
 
@@
 The value of the adjusted series at the break point is thus the same as the one preceding it.} Finally, all the series were normalized with respect to a common base period (Jan 2007).
-In total \Sexpr{n.full} countries have complete price series for the entire ten year period considered whereas, for example, there are \Sexpr{n.sub} countries with complete price series covering the 2005-11 subperiod.
+In total 118 countries have complete price series for the entire ten year period considered whereas, for example, there are 145 countries with complete price series covering the 2005-11 subperiod.
 There is therefore a trade-off between the number of complete price series and the length of the series.
@@
 
-The median country experienced an annualized rate of food and headline inflation of \Sexpr{round(median(infl$fpigr.0708),1)} and \Sexpr{round(median(infl$cpigr.0708),1)} percent, respectively, in 2007-08.
+The median country experienced an annualized rate of food and headline inflation of 9.7 and 7.1 percent, respectively, in 2007-08.
 Nominal food prices are not only driven by exogenous world market shocks; they follow a stochastic trend together with the non-food prices, representing underlying general (core) inflation.
 Dividing with the CPI is a simple way of purging the FPI of underlying domestic inflation.
-RFPI growth, which is approximately the difference between the FPI and CPI inflation rate, was \Sexpr{round(median(infl$rfpigr.0708),1)} percent for the median country in 2007-08.\footnote{A problem with the real FPI is that domestic non-food prices are also affected by international commodity shocks.
+RFPI growth, which is approximately the difference between the FPI and CPI inflation rate, was 2.5 percent for the median country in 2007-08.\footnote{A problem with the real FPI is that domestic non-food prices are also affected by international commodity shocks.
 Low or even negative growth in real domestic food prices in these two periods could therefore reflect a response to the concurrent surge in non-food commodities, oil in particular, rather than a puzzling negative response to an international food price shock.} Domestic price increases were generally lower in 2010-11.
-The median country experienced food and headline inflation rates \Sexpr{round(median(infl$fpigr.1011),1)} and \Sexpr{round(median(infl$cpigr.1011),1)} percent and a modest RFPI growth of \Sexpr{round(median(infl$rfpigr.1011),1)} percent.
+The median country experienced food and headline inflation rates 3.9 and 3.5 percent and a modest RFPI growth of 0.8 percent.
 There is, however, a substantial variation in the inflation rates in both periods.
@@
 
-<<echo=FALSE>>=
-library(dplyr)
-load("lrm_fpi_l1_12.Rda")
-fpi = select(lrm.fpi.l.1_12, -iso3c)
-load("lrm_rfpi_l1_12.Rda")
-rfpi = select(lrm.rfpi.l.1_12, -iso3c)
-fivenum.fpi = data.frame(sapply(fpi, summary))
-fivenum.rfpi = data.frame(sapply(rfpi, summary))
-n = nrow(fpi)
-@
-
 The LRM is a common measure of price transmission.
@@
 Both studies use 12 lags in their country specific ADL models, i.e.
-they assume $p=q=12$.\footnote{\cite{IMF2011} do not include exchange rates in their regressions.} My own estimates based on a sample of \Sexpr{n} countries covering the period 2005:1 to 2014:12 are very similar.
-The average and median FPI LRM from a model with 12 lags of all variables are \Sexpr{round(fivenum.fpi["Mean", "V12"], 2)} and \Sexpr{round(fivenum.fpi["Median", "V12"], 2)}, respectively.
+they assume $p=q=12$.\footnote{\cite{IMF2011} do not include exchange rates in their regressions.} My own estimates based on a sample of 118 countries covering the period 2005:1 to 2014:12 are very similar.
+The average and median FPI LRM from a model with 12 lags of all variables are 0.32 and 0.22, respectively.
 That being said, it is not obvious that so many lags should be included in the models.
 When lag length selection is based on the Akaike Information Criterion (AIC), for example, the optimal number lags in most of the models turn out to be $p=q=1$.\footnote{For each country, I calculated the AIC for each of the $12^2$ combinations of $(p,q)\, \in (1,..,12)$.
-Then I selected the $(p,q)$ combination with the lowest AIC.} But when a common lag length of 1 is used for all countries, average and median FPI transmission LRMs are much lower, namely \Sexpr{round(fivenum.fpi["Mean", "V1"], 2)} and \Sexpr{round(fivenum.fpi["Median", "V1"], 2)}, respectively, whereas the average of the standard errors of the LRM estimates very similar.\footnote{I also tested for residual autocorrelation with the Durbin-Watson test.
+Then I selected the $(p,q)$ combination with the lowest AIC.} But when a common lag length of 1 is used for all countries, average and median FPI transmission LRMs are much lower, namely 0.07 and 0.03, respectively, whereas the average of the standard errors of the LRM estimates very similar.\footnote{I also tested for residual autocorrelation with the Durbin-Watson test.
 There was no evidence of residual autocorrelation in either case.
 Standard errors of the LRMs are based on the so called Bewley transformation of the ADL model instrumented with lags of the dependent variable.} As a compromise between the long lag lengths used in the literature which might cause problems with overfitting and the short optimal lag lengths implied by the AIC criterion which leads to LRM estimates that are lower than commonly found in the literature, the country specific models are based on common lag lengths of $p=q=6$.
-This choice leads to average long run FPI and RFPI transmission rates of \Sexpr{round(fivenum.fpi["Mean", "V6"], 2)} and \Sexpr{round(fivenum.rfpi["Mean", "V6"], 2)}, respectively.
+This choice leads to average long run FPI and RFPI transmission rates of 0.14 and 0.03, respectively.
 One interpretation of the very low RFPI LRM average is that shocks to a country's FPI is passed on almost fully to its non-FPI in the long run, a phenomenon known as propagation in the literature \citep[e.g.][]{pedersen10}.
@@
 Finally, countries with grain state trading enterprises are found to have significantly lower short run (nominal as well as real) food price transmission.
-
-<<echo=false>>=
-load("tp_ctr_mean.Rda")
-load("tp_ctr_mean1.Rda")
-@
-
-\subsection{Estimated income-price transmission turning points}
-Table \ref{tab:tp} summarizes the estimated income turning points derived from the significant coefficients in table \ref{tab:07-08-fc-regressions}-\ref{tab:panel-regressions}.
-The ratios in the last column refer to the number of countries with a per capita GDP less than the estimated turning point shown in the second-last column, relative to the total number of countries in the sample with data on income.
-The average share of countries that are to the left of the turning point, across all specifications, is \Sexpr{round(nc.pc.mean[1],0)} percent.
-There are, however, important differences between the different estimators.
-Panel estimation, in particular, leads to a significantly higher proportion of countries to the left of the turning point than cross section estimation (\Sexpr{round(nc.pc.mean[2],0)} vs.
-\Sexpr{round(nc.pc.mean[3],0)} percent for the non-panel estimators) on average.
-In constant 2011 international dollars, the average income level in 2014 among the group of low-, lower middle- and upper middle income countries was \$1,517, \$5,752 and \$13,719, respectively \citep{worldbank16b}.\footnote{According to the World Bank classification scheme.
-MICs are defined as having a 2011 per capita gross national income of US\$1,026 to \$12,475, where the conversion from local currency to US dollars is based on the Atlas method.} In comparison, the income level leading to the highest predicted price transmission, averaged across all specifications, is \$\Sexpr{format(round(tp.mean[1],0), big.mark= ",", scientific= FALSE)}.
-Countries from the lower end of the high-income spectrum experience the highest short run food price transmission whereas lower middle income countries experience the highest long run food price transmission according to these estimates.

### 2016-06-21 analysis.Rnw
Revised analysis section for 2. round rev. of paper
Initial changes in connection with AE rejection

@@
 <<echo=FALSE>>=
-load("infl_0708_1011.Rda")
-load("ts_dat.Rda")
+str = "C:/Users/nqw235/Google Drive/Research/pov_price_transmission/"
+load(paste0(str, "infl_0708_1011.Rda"))
+load(paste0(str, "ts_dat.Rda"))
 n.full = nrow(table(ts.dat$id))
@@
 library(dplyr)
-load("lrm_fpi_l1_12.Rda")
+load(paste0(str, "lrm_fpi_l1_12.Rda"))
 fpi = select(lrm.fpi.l.1_12, -iso3c)
-load("lrm_rfpi_l1_12.Rda")
+load(paste0(str, "lrm_rfpi_l1_12.Rda"))
 rfpi = select(lrm.rfpi.l.1_12, -iso3c)
@@
 Countries from the lower end of the high-income spectrum experience the highest short run food price transmission whereas lower middle income countries experience the highest long run food price transmission according to these estimates.
+
+\begin{center}
+[Table \ref{tab:tp} around here]
+\par\end{center}

### 2016-06-21 price_data_prep.R

@@ -18,7 +18,7 @@ library(countrycode)
 #####################################################################
 #### Reading and merging the raw price data
 
-str = "C:/Users/nqw235/Google Drev/Research/pov_price_transmission/"
+str = "C:/Users/nqw235/Google Drive/Research/pov_price_transmission/"
 
 ### The ILO data
 temp = paste(str, "laborstaB9.csv", sep = "")

### 2016-06-21 ts_data_prep.R

@@ -14,7 +14,7 @@ library(stringr)
 # er.csv
 # fao_fpi.csv
 
-str = "C:/Users/nqw235/Google Drev/Research/pov_price_transmission/"
+str = "C:/Users/nqw235/Google Drive/Research/pov_price_transmission/"
 
 #####################################################################
 
@@ -94,12 +94,13 @@ yy$ff = as.numeric(zoo(yy$tt))
 yy$tt = NULL
 xxx = inner_join(xx, yy)
 xxx = xxx[order(xxx$iso3c, xxx$ff), ]
+ts.dat = xxx
 
-## Add international food prices
+## Add international oil prices
 xx = ts.dat
 yy =imf.oil
-xx$ff = as.numeric(zoo(xx$tt))
-xx$tt = NULL
+# xx$ff = as.numeric(zoo(xx$tt))
+# xx$tt = NULL
 yy$ff = as.numeric(zoo(yy$tt))
 yy$tt = NULL
 xxx = inner_join(xx, yy)

### 2016-06-21 ts_reg.R

@@ -246,7 +246,7 @@ for (i in 1:max(data$id)) {
 }
 coef.bew.fpi.l6 = as.data.frame(coef.mat)
 coef.bew.fpi.l6$iso3c = names(tbl)
-coef.bew.fpi.l6 = rename(coef.bew.fpi.l6, lrm = V1, lrm.se = V2)
+coef.bew.fpi.l6 = dplyr::rename(coef.bew.fpi.l6, lrm = V1, lrm.se = V2)
 save(coef.bew.fpi.l6, file = "coef_bew_fpi_l6.Rda")
 
 # 2) rfpi
@@ -287,7 +287,7 @@ for (i in 1:max(data$id)) {
 }
 coef.bew.rfpi.l6 = as.data.frame(coef.mat)
 coef.bew.rfpi.l6$iso3c = names(tbl)
-coef.bew.rfpi.l6 = rename(coef.bew.rfpi.l6, lrm = V1, lrm.se = V2)
+coef.bew.rfpi.l6 = dplyr::rename(coef.bew.rfpi.l6, lrm = V1, lrm.se = V2)
 save(coef.bew.rfpi.l6, file = "coef_bew_rfpi_l6.Rda")
 
 ##################################################################

### 2016-08-11 Master.Rnw
Initial changes in connection with AE rejection
JAE sub. vers.

@@
 
-\title{Income and Food Price Transmission\footnote{Revised version of MS 2015310 ``Income and Price Transmission''.
-Submitted to \emph{Agricultural Economics} for 2.
-round review.
-I thank two anonymous reviewers for their valuable comments.}}
+\title{Income and Food Price Transmission}
 
@@
 
-\textbf{Keywords:} price transmission, food crisis, food inflation, determinants, regression analysis, income\newpage
+\textbf{Keywords:} price transmission, food crisis, food inflation, regression analysis, income\newpage
 
@@
 \bibliographystyle{chicago}
-\bibliography{inc_pt}
+\bibliography{poverty_price_transmission}
 
-\SweaveInput{figures1.Rnw}
+\SweaveInput{figures.Rnw}
 \newpage

### 2016-08-11 analysis.Rnw
JAE sub. vers.

@@
-% !Rnw root = master.Rnw
+% !Rnw root = Master.Rnw
 
@@
 \subsubsection{Domestic and international food prices}
-Data on domestic prices are sourced from the International Labor Organization (ILO).\footnote{See \citet{ilo16}} ILO maintains a comprehensive database of domestic food price indexes (FPIs) and consumer price indexes (CPIs) that is used to monitor the cost of living and real wages.
-There are not many studies which draw on this database, however, despite a broad country coverage and long monthly price series.
+Data on domestic prices are sourced from the International Labor Organization (ILO).
+ILO maintains a comprehensive database of domestic food price indexes (FPIs) and consumer price indexes (CPIs) that is used to monitor the cost of living and real wages.
+Not many studies which draw on this database, however, despite a broad country coverage and long monthly price series.
 A major issue with this data is that historical series are not recalculated whenever a new base year is introduced.
@@
 Finally, many of the series contain missing values.
-To ``tidy up'' the raw data I carried out the following steps First, I restricted the sample to the period 2005:1-2014:12.
+To ``tidy up'' the raw data I carried out the following steps: First, I restricted the sample to the period 2005:1-2014:12.
 Secondly, in cases where multiple complete series were available from the same country, a single one was selected.
@@
 
-International food prices are represented by the FAO food price index (FAO FPI) and in some cases its ``real'' version, namely the nominal FAO FPI deflated by the Manufactures Unit Value (MUV) index published by the World Bank.\footnote{see \citet{fao16} and \citet{worldbank16a}.}
+International food prices are represented by the FAO food price index (FAO FPI) and in some cases its ``real'' version, namely the nominal FAO FPI deflated by the Manufactures Unit Value (MUV) index published by the World Bank.
 
@@
 There has been two recent international price spikes; one in 2007-8 and the other in 2010-11.
-The former spike especially is often referred to as ``the food crisis'' in the literature \citep[e.g.][]{Headey2008, Headey2013, timmer10, Ivanic2012, Martin2011}.
+The former spike especially is often referred to as ``the food crisis'' in the literature \citep[e.g.][]{Headey2008, Headey2013, Timmer2010, Ivanic2012, Martin2011}.
 Between January 2007 and June 2008 the FAO FPI rose by 91 points to index 226, corresponding to a 68 percent relative price increase.
@@
 In addition there are two indexes included representing the regulatory performance of the country in question and the quality of its infrastructure.
-These two indexes are abbreviated DTF (Distance to Frontier) and LPI (Logistics Performance Index), respectively.\footnote{The covariates are similar to those of \citet{greb12}.}
+These two indexes are abbreviated DTF (Distance to Frontier) and LPI (Logistics Performance Index), respectively.\footnote{The covariates are similar to those of \citet{Greb2016}.}
 
@@
 Table \ref{tab:07-08-fc-regressions} considers the cross sectional variation in domestic food price changes during the 2007-8 food crises.
-Each of the columns refer to a regression of FPI inflation or real FPI growth on the explanatory variables discussed above as well as the percentage change in a country's domestic currency/US dollar exchange rate during the food crisis.
-All variables except for GDP and the dummies have been standardized in order to facilitate comparisons of effect sizes.
+Each of the columns refer to a regression of FPI inflation or real FPI growth on all or a subset of the explanatory variables discussed above as well as the percentage change in a country's domestic currency/US dollar exchange rate during the food crisis.
+All variables except for GDP and the dummies have been standardized to facilitate comparisons of effect sizes.
 
@@
 A concern is that some of the variables are highly correlated which leads to large standard errors and insignificant effects.
-Therefore, the subset of covariates included in columns (2) and (4) are those which minimizes the Akaike information criterion (AIC) as determined by stepwise regression.\footnote{This was carried out using the \emph{MASS} package in \textsf{R} \citep{venables02}} Stepwise regressions have been criticized for leading to models where the remaining terms appear to be more important than they actually are.
-However, as can be seen, no substantial changes result from this exercise.
+This problem is amplified by the loss of observations due to missing data on some of these variables for some countries.
+Therefore, columns (2) and (4) consider only the variable of primary interest, namely income.
+As can be seen, the increased number of observations and the reduced number of explanatory variables decreases $R^2$ substantially whereas the significance of the income effects remains or increases.
 
@@
 ``Export'' = 1 if the country Restricted or banned export, ``Price'' = 1 if the country administered price control or restricted private trade, ``Stock'' = 1 if the country released stocks (public or imported) at subsidized prices'', ``Tariff'' = 1 if the country reduced tariffs or customs fees on imports and ``Tax'' = 1 if the country suspended or reduced VAT or other taxes, all as of 1.
-December 2008.\label{int_note}} Some of the coefficients to these intervention dummies are substantial but they are all statistically insignificant.
+December 2008.\label{int_note}} Some of the coefficients to these intervention dummies are substantial but they are mainly statistically insignificant.
 Table \ref{tab:1011-regressions} in the appendix contains results related to the 2010-11 international price spike.
-Again, estimated effects are similar to those in  table \ref{tab:07-08-fc-regressions} but most are statistically insignificant.
-In fact,  the cereal consumption share is the only robustly significant predictor of domestic food inflation in 2010-11.
+Again, estimated effects are similar to those in  table \ref{tab:07-08-fc-regressions}.
+The main difference is that cereal consumption is a significant predictor of food inflation in this case.
 
@@
 
-Prior to estimation I tested each of the series for unit roots and cointegration with international prices.\footnote{I tested for stationarity with the KPSS test \citep{kwiatkowski92} and for unit roots with the augmented Dickey-Fuller test \citep{dickey79}.
-I used the Phillips and Ouliaris ``Pz'' test \citep{phillips90} to test for cointegration.
-These tests are implemented in \textsf{R} via the packages \emph{tseries} \citep{trapletti15} and \emph{urca} \citep{pfaff08}, respectively.} Results indicated that the domestic FPI and real FPI series, are nonstationary in levels but stationary in first differences.
+Prior to estimation I tested each of the series for unit roots and cointegration with international prices.\footnote{I tested for stationarity with the KPSS test \citep{Kwiatkowski1992} and for unit roots with the augmented Dickey-Fuller test \citep{Said1984}.
+I used the Phillips and Ouliaris ``Pz'' test \citep{Phillips1990} to test for cointegration.
+These tests are implemented in \textsf{R} via the packages \emph{tseries} \citep{Trapletti2007} and \emph{urca} \citep{Pfaff2005}, respectively.} Results indicated that the domestic FPI and real FPI series, are nonstationary in levels but stationary in first differences.
 Domestic price levels were also found not to be cointegrated with world prices.
@@
 When lag length selection is based on the Akaike Information Criterion (AIC), for example, the optimal number lags in most of the models turn out to be $p=q=1$.\footnote{For each country, I calculated the AIC for each of the $12^2$ combinations of $(p,q)\, \in (1,..,12)$.
-Then I selected the $(p,q)$ combination with the lowest AIC.} But when a common lag length of 1 is used for all countries, average and median FPI transmission LRMs are much lower, namely \Sexpr{round(fivenum.fpi["Mean", "V1"], 2)} and \Sexpr{round(fivenum.fpi["Median", "V1"], 2)}, respectively, whereas the average of the standard errors of the LRM estimates very similar.\footnote{I also tested for residual autocorrelation with the Durbin-Watson test.
+Then I selected the $(p,q)$ combination with the lowest AIC.} But when a common lag length of 1 is used for all countries, average and median FPI transmission LRMs are much lower, namely \Sexpr{round(fivenum.fpi["Mean", "V1"], 2)} and \Sexpr{round(fivenum.fpi["Median", "V1"], 2)}, respectively, whereas the average of the standard errors of the LRM estimates are very similar.\footnote{I also tested for residual autocorrelation with the Durbin-Watson test.
 There was no evidence of residual autocorrelation in either case.
@@
 This choice leads to average long run FPI and RFPI transmission rates of \Sexpr{round(fivenum.fpi["Mean", "V6"], 2)} and \Sexpr{round(fivenum.rfpi["Mean", "V6"], 2)}, respectively.
-One interpretation of the very low RFPI LRM average is that shocks to a country's FPI is passed on almost fully to its non-FPI in the long run, a phenomenon known as propagation in the literature \citep[e.g.][]{pedersen10}.
+One interpretation of the very low RFPI LRM average is that shocks to a country's FPI is passed on almost fully to its non-FPI in the long run, a phenomenon known as propagation in the literature \citep[e.g.][]{Pedersen2010}.
 
 The long run transmission estimates, $\hat{\theta}$, are regressed, on the same explanatory variables as in table \ref{tab:07-08-fc-regressions}.
-However, since the dependent variable is an estimate rather than an observed value and the model now represents an estimated dependent variable (EDV) regression \citep[see][]{hanushek74, lewis05}.
-A common approach in this situation is to use Weighted Least Squares (WLS) estimation with weights $w_{i}$ that are inversely proportional to the standard error of $\theta_{i}$ \citep{saxonhouse76}.
+However, since the dependent variable is an estimate rather than an observed value and the model now represents an estimated dependent variable (EDV) regression \citep[see][]{Hanushek1974, Lewis2005}.
+A common approach in this situation is to use Weighted Least Squares (WLS) estimation with weights $w_{i}$ that are inversely proportional to the standard error of $\theta_{i}$ \citep{Saxonhouse1976}.
 This way more precise estimates are given a higher weight in the regression.
@@
 The inconsistent OLS standard errors can, however, easily be corrected with a robust standard error estimator.
-\citet{lewis05} uses formal reasoning backed up by a Monte Carlo analysis to argue that OLS with robust standard errors is likely to produce more reliable estimates than WLS based on inverse standard error weights, and that the latter is likely to produce downward biased standard errors.
+\citet{Lewis2005} uses formal reasoning backed up by a Monte Carlo analysis to argue that OLS with robust standard errors is likely to produce more reliable estimates than WLS based on inverse standard error weights, and that the latter is likely to produce downward biased standard errors.
 
 Table \ref{tab:fpi-pt-regressions} presents estimates of FPI transmission determinants based on OLS [columns (1)-(2)] as well as WLS [columns (3)-(4)], both with robust standard errors.
-Again, all variables except per capita GDP and the dummies have been standardized and, as in table \ref{tab:fpi-pt-regressions}, column (1) and (3) represent regressions that are based on the full set of covariates whereas columns (2) and (4) represent regressions with the lowest AIC.
+Again, all variables except per capita GDP and the dummies have been standardized and, as in table \ref{tab:fpi-pt-regressions}, column (1) and (3) represent regressions that are based on the full set of covariates whereas columns (2) and (4) represent regressions based solely on income.
 
@@
 The signs of the coefficients are similar to those in table \ref{tab:fpi-pt-regressions} but most effects are statistically insignificant.
-Similarly, table \ref{tab:fpi-pt1-regressions} and \ref{tab:rfpi-pt1-regressions} in the appendix contain results based on FPI and real FPI transmission rates calculated from time series models with a single lag.
-These also point to only a few significant predictors of price transmission.
-A plausible reason is that the LRM estimates based on models with a single lag are relatively more noisy and therefore more difficult to explain.
 
@@
 It is well known that dynamic panel models such as equation (\ref{eq:panel_model}) suffers from (Nickell) bias due to correlation between the lagged dependent variable and the error term.
-However, as demonstrated by \citet{nickell81}, the bias of the within estimator decreases at a rate $1/T$, i.e.
+However, as demonstrated by \citet{Nickell1981}, the bias of the within estimator decreases at a rate $1/T$, i.e.
 endogeneity is mainly a problem in short panel datasets with only a few time periods $T$.
@@
 <<echo=false>>=
-load("tp_ctr_mean.Rda")
-load("tp_ctr_mean1.Rda")
+str = "C:/Users/nqw235/Google Drive/Research/pov_price_transmission/"
+load(paste0(str,"tp_ctr_mean.Rda"))
+load(paste0(str,"tp_ctr_mean1.Rda"))
 @
@@
 \Sexpr{round(nc.pc.mean[3],0)} percent for the non-panel estimators) on average.
-In constant 2011 international dollars, the average income level in 2014 among the group of low-, lower middle- and upper middle income countries was \$1,517, \$5,752 and \$13,719, respectively \citep{worldbank16b}.\footnote{According to the World Bank classification scheme.
+According to the World Bank, the average income level in 2014 among the group of low-, lower middle- and upper middle income countries was \$1,517, \$5,752 and \$13,719, respectively, when measured in constant 2011 international dollars.\footnote{According to the World Bank classification scheme.
 MICs are defined as having a 2011 per capita gross national income of US\$1,026 to \$12,475, where the conversion from local currency to US dollars is based on the Atlas method.} In comparison, the income level leading to the highest predicted price transmission, averaged across all specifications, is \$\Sexpr{format(round(tp.mean[1],0), big.mark= ",", scientific= FALSE)}.

### 2016-08-11 appendix.Rnw
JAE sub. vers.

@@
 \hline
-{\small{}STE} & {\small{}State trading enterprise dummy. 1 if country has a grain state trading enterprise} & {\small{}\citet{greb12}} \tabularnewline
+{\small{}STE} & {\small{}State trading enterprise dummy. 1 if country has a grain state trading enterprise} & {\small{}\citet{Greb2016}} \tabularnewline
 \hline
@@
 library(stargazer)
-reg.list = readRDS("int_reg_m.rds")
-se.list = readRDS("int_reg_se.rds")
+str = "C:/Users/nqw235/Google Drive/Research/pov_price_transmission/"
+reg.list = readRDS(paste0(str,"int_reg_m.rds"))
+se.list = readRDS(paste0(str,"int_reg_se.rds"))
 
 stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
-          title='Determinants of domestic food price changes 2007-8', 
+          title='Determinants of domestic food price changes 2007:1-2008:6', 
           label = "tab:int-regressions",
@@
                              "$Tax$"),
-          notes=c("*, ** and *** signify significance at the 10, 5 and 1 percent",
-                  "levels, respectively. Robust standard errors in parentheses;", 
-                  "samples exclude outliers defined as observations $\\pm$ 3.5 stan-", 
-                  "dard deviations from the mean of the dependent variable.", 
-                  "A description of the variables and their sources can be found", 
-                  "in table \\ref{tab:cs_data} and section \\ref{sec:analysis}. Footnote \\ref{int_note} contains a description", 
-                  "of the 5 intervention dummies."),
+          notes=c("*, ** and *** signify significance at the 10, 5 and 1",
+                  "percent levels, respectively. Robust standard errors in;", 
+                  "parentheses samples exclude outliers defined as obser-", 
+                  "vations $\\pm$ 3.5 standard deviations from the mean of the", 
+                  "dependent variable. A description of the variables and", 
+                  "their sources can be found in table \\ref{tab:cs_data} and section \\ref{sec:analysis}."),
           notes.append=FALSE,
@@
 # library(stargazer)
-reg.list = readRDS("fc_reg1011_m.rds")
-se.list = readRDS("fc_reg1011_se.rds")
+reg.list = readRDS(paste0(str,"fc_reg1011_m.rds"))
+se.list = readRDS(paste0(str,"fc_reg1011_se.rds"))
 
 stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
-          title='Determinants of domestic food price changes 2010-11', 
+          title='Determinants of domestic food price changes 2010:1-2010:12', 
           label = "tab:1011-regressions",
-          dep.var.labels=c("FPI inflation", "RFPI grwth"),
+          dep.var.labels=c("FPI inflation", "RFPI growth"),
           no.space=TRUE,
@@
                              "$LPI$"),
-          notes=c("*, ** and *** signify significance at the 10, 5 and 1 percent",
-                  "levels, respectively. Robust standard errors in parentheses;", 
-                  "samples exclude outliers defined as observations $\\pm$ 3.5 stan-", 
-                  "dard deviations from the mean of the dependent variable.", 
-                  "A description of the variables and their sources can be found", 
-                  "in table \\ref{tab:cs_data} and section \\ref{sec:analysis}."),
+          notes=c("*, ** and *** signify significance at the 10, 5 and 1",
+                  "percent levels, respectively. Robust standard errors in;", 
+                  "parentheses samples exclude outliers defined as obser-", 
+                  "vations $\\pm$ 3.5 standard deviations from the mean of the", 
+                  "dependent variable. A description of the variables and", 
+                  "their sources can be found in table \\ref{tab:cs_data} and section \\ref{sec:analysis}."),
           notes.append=FALSE,
@@
 <<results=tex,echo=F>>=
-reg.list = readRDS("rfpi_pt_reg_m.rds")
-se.list = readRDS("rfpi_pt_reg_se.rds")
+reg.list = readRDS(paste0(str,"rfpi_pt_reg_m.rds"))
+se.list = readRDS(paste0(str,"rfpi_pt_reg_se.rds"))
 
@@
                              "$LPI$"),
-          notes=c("*, ** and *** signify significance at the 10, 5 and 1 percent",
-                  "levels, respectively. Robust standard errors in parentheses;", 
-                  "samples exclude outliers defined as observations $\\pm$ 3.5 stan-", 
-                  "dard deviations from the mean of the dependent variable.", 
-                  "A description of the variables and their sources can be found", 
-                  "in table \\ref{tab:cs_data} and section \\ref{sec:analysis}."),
-          notes.append=FALSE,
-          notes.align="l")
-@
-
-<<results=tex,echo=F>>=
-reg.list = readRDS("fpi_pt_reg_m1.rds")
-se.list = readRDS("fpi_pt_reg_se1.rds")
-
-stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
-          title='Determinants of long run FPI transmission rates', 
-          label = "tab:fpi-pt1-regressions", 
-          dep.var.labels=c("FPI LRM (single lag)"), 
-          column.labels = c("OLS", "WLS"),
-          column.separate = c(2, 2),
-          no.space=TRUE,
-          se = list(se.list[[1]], se.list[[2]], se.list[[3]], se.list[[4]]), 
-          omit.stat = c("adj.rsq","f", "ser"), 
-          font.size = "small", 
-          covariate.labels=c("$GDP$",
-                             "$GDP^{2}$", 
-                             "$CCONS$", 
-                             "$CIMP$", 
-                             "$CEXP$",
-                             "$LLDC$", 
-                             "$STE$",
-                             "$TO$", 
-                             "$DTF$", 
-                             "$LPI$"),
-          notes=c("*, ** and *** signify significance at the 10, 5 and 1 percent",
-                  "levels, respectively. Robust standard errors in parentheses;", 
-                  "samples exclude outliers defined as observations $\\pm$ 3.5 stan-", 
-                  "dard deviations from the mean of the dependent variable.", 
-                  "A description of the variables and their sources can be found", 
-                  "in table \\ref{tab:cs_data} and section \\ref{sec:analysis}."),
-          notes.append=FALSE,
-          notes.align="l")
-@
-
-<<results=tex,echo=F>>=
-reg.list = readRDS("rfpi_pt_reg_m1.rds")
-se.list = readRDS("rfpi_pt_reg_se1.rds")
-
-stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
-          title='Determinants of long run real FPI transmission rates', 
-          label = "tab:rfpi-pt1-regressions", 
-          dep.var.labels=c("RFPI LRM (single lag)"), 
-          column.labels = c("OLS", "WLS"),
-          column.separate = c(2, 2),
-          no.space=TRUE,
-          se = list(se.list[[1]], se.list[[2]], se.list[[3]], se.list[[4]]), 
-          omit.stat = c("adj.rsq","f", "ser"), 
-          font.size = "small", 
-          covariate.labels=c("$GDP$",
-                             "$GDP^{2}$", 
-                             "$CCONS$", 
-                             "$CIMP$", 
-                             "$CEXP$",
-                             "$LLDC$", 
-                             "$STE$",
-                             "$TO$", 
-                             "$DTF$", 
-                             "$LPI$"),
-         notes=c("*, ** and *** signify significance at the 10, 5 and 1 percent",
-                  "levels, respectively. Robust standard errors in parentheses;", 
-                  "samples exclude outliers defined as observations $\\pm$ 3.5 stan-", 
-                  "dard deviations from the mean of the dependent variable.", 
-                  "A description of the variables and their sources can be found", 
-                  "in table \\ref{tab:cs_data} and section \\ref{sec:analysis}."),
+          notes=c("*, ** and *** signify significance at the 10, 5 and 1",
+                  "percent levels, respectively. Robust standard errors in;", 
+                  "parentheses samples exclude outliers defined as obser-", 
+                  "vations $\\pm$ 3.5 standard deviations from the mean of the", 
+                  "dependent variable. A description of the variables and", 
+                  "their sources can be found in table \\ref{tab:cs_data} and section \\ref{sec:analysis}."),
           notes.append=FALSE,

### 2016-08-11 conclusion.Rnw
JAE sub. vers.

@@
 I conclude that countries from the middle of the income spectrum generally have the highest food price transmission rates once we control for factors that are commonly believed to affect price transmission.
-This is a worrying finding in a way, given the fact that nearly three quarters of the world's poor now live in middle-income countries \citep{worldbank16, Sumner2012, Kanbur2012}.
+This is a worrying finding in a way, given the fact that nearly three quarters of the world's poor now live in middle-income countries \citep{Worldbank2016, Sumner2012, Kanbur2012}.
 

### 2016-08-11 cs_infl_regressions.R
JAE sub. vers.

@@ -9,6 +9,8 @@ library(MASS)
 ### Cross section inflation regressions
 ##################################################################
 
+str = "C:/Users/nqw235/Google Drive/Research/pov_price_transmission/"
+setwd(str)
 load("infl_ctr_dat_0708.Rda")
 load("infl_ctr_dat_1011.Rda")
 
@@ -34,6 +36,8 @@ data1 = na.omit(dplyr::select(data, fpigr.0708, rfpigr.0708, ergr.0708,
 fc.reg0708.m = list()
 fc.reg0708.se = list()
 
+#m = lars(x = scale(data$fpigr.0708), y = scale(data$cpigr.0708), type = "lasso")
+
 m = lm(scale(fpigr.0708)~scale(ergr.0708) +
          log(gdp.const.int)+
          I(log(gdp.const.int)^2)+
@@ -51,24 +55,33 @@ rse = sqrt(diag(cov))
 fc.reg0708.m[[1]] = m
 fc.reg0708.se[[1]] = rse
 
-m = lm(scale(fpigr.0708)~scale(ergr.0708)+
+m = lm(scale(fpigr.0708)~
          log(gdp.const.int)+
-         I(log(gdp.const.int)^2)+
-         scale(cshare)+
-         scale(cir)+
-         scale(cer) +
-         lldc+
-         ste +
-         scale(to)+
-         scale(dtf)+
-         scale(lpi.infra),
-       data = filter(data1, fpigr.0708<3.5*sd(fpigr.0708)))
-
-step = stepAIC(m, direction="both")
-cov = vcovHC(step, type = "HC1")
+         I(log(gdp.const.int)^2),
+       data = filter(data, fpigr.0708<3.5*sd(fpigr.0708)))
+cov = vcovHC(m, type = "HC1")
 rse = sqrt(diag(cov))
-fc.reg0708.m[[2]] = step
+fc.reg0708.m[[2]] = m
 fc.reg0708.se[[2]] = rse
+
+# m = lm(scale(fpigr.0708)~scale(ergr.0708)+
+#          log(gdp.const.int)+
+#          I(log(gdp.const.int)^2)+
+#          scale(cshare)+
+#          scale(cir)+
+#          scale(cer) +
+#          lldc+
+#          ste +
+#          scale(to)+
+#          scale(dtf)+
+#          scale(lpi.infra),
[…]

### 2016-08-11 cs_pt_reg.R

@@ -9,6 +9,9 @@ library(MASS)
 ### Cross section price transmission regressions
 ##################################################################
 
+str = "C:/Users/nqw235/Google Drive/Research/pov_price_transmission/"
+setwd(str)
+
 load("infl_ctr_dat_0708.Rda")
 load("infl_ctr_dat_1011.Rda")
 load("coef_bew_fpi_l1.Rda")
@@ -21,8 +24,8 @@ load("coef_bew_rfpi_l6.Rda")
 ## FPI LR multiplier regressions
 fpi.pt.reg.m = list()
 fpi.pt.reg.se = list()
-fpi.bptest.stat = list()
-fpi.bptest.p = list()
+# fpi.bptest.stat = list()
+# fpi.bptest.p = list()
 
 # data = merge(ctr.dat.0708, coef.bew.fpi.l1)
 data = merge(ctr.dat.0708, coef.bew.fpi.l6)
@@ -45,33 +48,39 @@ m = lm(scale(lrm1)~log(gdp.const.int)+
        data = filter(data1, lrm1<3.5*sd(lrm1)))
 cov = vcovHC(m, type = "HC1")
 rse = sqrt(diag(cov))
-
 fpi.pt.reg.m[[1]] = m
 fpi.pt.reg.se[[1]] = rse
-
-fpi.bptest.stat[[1]] = bptest(m)[["statistic"]]
-fpi.bptest.p[[1]] = bptest(m)[["p.value"]]
+# fpi.bptest.stat[[1]] = bptest(m)[["statistic"]]
+# fpi.bptest.p[[1]] = bptest(m)[["p.value"]]
 
 m = lm(scale(lrm1)~log(gdp.const.int)+
-         I(log(gdp.const.int)^2)+
-         scale(cshare)+
-         scale(cir)+
-         scale(cer) +
-         lldc+
-         ste +
-         scale(to)+
-         scale(dtf)+
-         scale(lpi.infra),
-       data = filter(data1, lrm1<3.5*sd(lrm1)))
-
-step = stepAIC(m, direction="both")
-cov = vcovHC(step, type = "HC1")
+         I(log(gdp.const.int)^2),
+       data = filter(data, lrm1<3.5*sd(lrm1)))
+cov = vcovHC(m, type = "HC1")
 rse = sqrt(diag(cov))
-fpi.pt.reg.m[[2]] = step
[…]

### 2016-08-11 data_prep.R

@@ -24,7 +24,7 @@ library(tempdisagg)
 # gdp_05.cs
 # cons_05.csv
 
-str = "C:/Users/nqw235/Google Drev/Research/pov_price_transmission/"
+str = "C:/Users/nqw235/Google Drive/Research/pov_price_transmission/"
 
 # FAO country list
 temp = paste(str, "fao_country_codes_list.csv", sep = "")

### 2016-08-11 figures.Rnw
Initial changes in connection with AE rejection
JAE sub. vers.

--- 2015-12-22 figures.Rnw
@@
 
-\SweaveInput{data_prep.Rnw}
-
 \section{Figures}
-<<pt, echo=FALSE, include=FALSE, fig=TRUE, height=6, width=8>>=
-a = 1.3
-b = 1.2
-par(mfrow=c(1,2))
-curve(1-a^(-x), 0, 10, lty=1, xlab = "Income", ylab = "", ylim = c(0,1))
-curve(b^(-x), add = T, col = "blue", lty=2)
-title(main = "Shares", cex.main = 1)
-legend("topright", c(expression(beta(y)), expression(gamma(y))),lty = c(2,1), col = c("blue", "black"))
-curve((1-a^(-x))*b^(-x), 0, 10, xlab = "Income", ylab = "")
-title(main = "Price transmission", cex.main = 1)
-legend("topright", c(expression(theta(y))), lty = 1, col = "black")
-@
-
-<<gamma, echo=FALSE, include=FALSE, fig=TRUE, height=6, width=8>>=
-library(ggplot2)
-library("foreign")
-
-dat = read.csv("dist_trade.csv", sep = ";")
-dat$gamma = (dat$vop_covx+dat$vop_covm)/dat$vop_covt*100
-ggplot(dat, aes(log(gdppcp00), gamma)) + 
-  geom_point() + 
-  geom_smooth() +
-  theme_bw() +
-  xlab("Log of GDP") +
-  ylab(expression(paste("Estimated ", gamma))) +
-  theme(panel.grid.major = element_line(colour = "grey"))
-@
-
 <<infl, echo=FALSE, include=FALSE, fig=TRUE, height=6, width=8>>=
 library(gridExtra)
+library(ggplot2)
 
-p1 = ggplot(aes(y = value, x = factor(time, labels = c("2007:1-2008:6", "2010:1-2010:12")), 
-                fill = factor(variable, labels = c("FPI inflation", "CPI inflation", "Real FPI growth"))), data = infl_long1) + 
+load(paste0(str, "infl_l1.Rda"))
+load(paste0(str, "infl_0708_1011.Rda"))
+
+p1 = ggplot(aes(y = value, x = factor(period, labels = c("2007:1-2008:6", "2010:1-2010:12")),
+                fill = factor(variable, labels = c("FPI inflation", "CPI inflation", "Real FPI growth"))), data = infl.l1) + 
   geom_boxplot(outlier.shape = 1) +
@@
 
-p2 = ggplot(aes(x=foodinfl78, y=foodinfl1011), data = infl) + 
+p2 = ggplot(aes(x=fpigr.0708, y=fpigr.1011), data = infl) + 
   geom_point(shape=1) +
@@
   xlab("FPI inflation 2007-08") +
-  geom_text(aes(label=ifelse(foodinfl78>40,as.character(iso3c),'')),hjust=1.2,just=-1.2, 
+  geom_text(aes(label=ifelse(fpigr.0708>40,as.character(iso3c),'')),hjust=1.2,vjust=-1.2, 
             size=3) +
-  geom_text(aes(label=ifelse(foodinfl1011>20,as.character(iso3c),'')),hjust=1.2,just=-1.2,
+  geom_text(aes(label=ifelse(fpigr.1011>20,as.character(iso3c),'')),hjust=1.2,vjust=-1.2,
             size=3)
 
-p3 = ggplot(aes(x=cpiinfl78, y=cpiinfl1011), data = infl) +
+p3 = ggplot(aes(x=cpigr.0708, y=cpigr.1011), data = infl) +
   geom_point(shape=1) +
@@
   xlab("CPI inflation 2007-08") +
-  geom_text(aes(label=ifelse(cpiinfl78>30,as.character(iso3c),'')),hjust=1.2,just=-1.2, 
+  geom_text(aes(label=ifelse(cpigr.0708>30,as.character(iso3c),'')),hjust=1.2,vjust=-1.2, 
             size=3) +
-  geom_text(aes(label=ifelse(cpiinfl1011>15,as.character(iso3c),'')),hjust=1.2,just=-1.2,
+  geom_text(aes(label=ifelse(cpigr.1011>15,as.character(iso3c),'')),hjust=1.2,vjust=-1.2,
             size=3)
 
-p4 = ggplot(aes(x=rfpigr78, y=rfpigr1011), data = infl) + 
+p4 = ggplot(aes(x=rfpigr.0708, y=rfpigr.1011), data = infl) + 
   geom_point(shape=1) +
@@
   xlab("Real FPI Growth 2007-08") +
-  geom_text(aes(label=ifelse(rfpigr78>12,as.character(iso3c),'')),hjust=1.2,just=-1.2, 
+  geom_text(aes(label=ifelse(rfpigr.0708>12,as.character(iso3c),'')),hjust=1.2,vjust=-1.2, 
             size=3) +
-  geom_text(aes(label=ifelse(rfpigr1011>10,as.character(iso3c),'')),hjust=1.2,just=-1.2,
+  geom_text(aes(label=ifelse(rfpigr.1011>10,as.character(iso3c),'')),hjust=1.2,vjust=-1.2,
             size=3)
@@
 @
-
-\begin{figure}[H]
-\begin{center}
-\includegraphics[width=1\textwidth]{Master-pt.pdf}
-\caption{The effect of income on price transmission}
-\label{fig:pt}
-\end{center}
-\end{figure}
-
-\begin{figure}[H]
-\begin{center}
-\includegraphics[width=1\textwidth]{Master-gamma.pdf}
-\caption{Estimates of $\gamma$. Source: \cite{Anderson2008}. Own calculations.}
-\label{fig:gamma}
-\end{center}
-\end{figure}
 

### 2016-08-11 intro.Rnw
JAE sub. vers.

@@
 In response to the substantial increase in international food prices during the first decade of the new millennium, a large number of recent studies have quantified global and local poverty impacts of international price shocks.\footnote{E.g.
-\cite{Attanasio2013,Balagtas2014,Coxhead2012,Ferreira2013,Hoyos2011,Ivanic2012,Janvry2010,Mghenyi2011,Vu2011,Wodon2010}.
-In addition, a supplement to volume 39 of \emph{Agricultural Economics} is entirely devoted to this topic \citep[see][]{masters08}.} A major determinant of a country's vulnerability to such shocks is the responsiveness of its domestic food prices \citep[e.g.][]{Hoyos2011}.
+\cite{Attanasio2013,Balagtas2014,Coxhead2012,Ferreira2013,Hatzenbuehler2016, Hoyos2011,Ivanic2012,Janvry2010,Masters2008, Mghenyi2011,Vu2011,Wodon2010}.} A major determinant of a country's vulnerability to such shocks is the responsiveness of its domestic food prices \citep[e.g.][]{Hoyos2011}.
 But what are the main determinants of the domestic price responsiveness to international shocks?
 Are they related to trade, consumption, geography or something else?
-As the title of the paper suggests, a country's income level seems to be an important determinant but other factors also play a role.
-In particular, countries which control their international trade through an agricultural state trading enterprise are found to experience lower food price transmission on average in the short run as well as in the long run.
+As the title of the paper suggests, a country's income level is an important determinant but other factors also seem to play a role.
 
-The literature on domestic price impacts of international commodity price shocks is vast and the topic has experienced a revival in the wake of the 2007-08 food crisis \citep[see][]{Fackler2001, kouyate16}.
-However, since the majority of price transmission studies focus on a single country we know little about the determinants, other than distance and borders, of price transmission differences across countries.
+The literature on domestic price impacts of international commodity price shocks is vast and the topic has experienced a revival in the wake of the 2007-08 food crisis \citep[see][]{Fackler2001, Kouyate2016}.
+However, we know little about the determinants, other than distance and borders, of price transmission differences across countries.
 Moreover, most price transmission studies focus on individual commodity prices rather than food prices in general.
@@
 
-Regarding the choice between individual or aggregate prices, each approach has its strengths and weaknesses \citep{headey10}.
-Commodity specific analyses have a stronger microeconomic foundation in the law of one price but results do not generalize to other food goods or even to the same food good traded in other domestic markets.
+Regarding the choice between individual or aggregate prices, each approach has its strengths and weaknesses \citep{Headey2010}.
+Commodity specific analyses have a stronger microeconomic foundation in the ``law of one price'' but results do not generalize to other food goods or even to the same food good traded in other domestic markets.
 Transmission estimates based on aggregate price data, on the other hand, give a more comprehensive picture and therefore better reflect domestic welfare costs of international food price fluctuations.
-A potential weakness though of the latter approach, is that a country's food price index might not reflect the composition of poor people's food baskets which is often based mainly on grain.
+A potential weakness though of the latter approach, is that a country's food price index might not reflect the composition of poor people's food baskets.
 Another weakness is that the composition of a country's food price index differs substantially from that of the international food price indexes and as well as from other countries' price indexes, which makes it more difficult to compare transmission rates across countries.
@@
 
-The empirical analysis is based on insights derived from a simple economic model that is more suitable for aggregate price data than the law of one price workhorse model.
+The empirical analysis is based on insights derived from a simple economic model that is more suitable for aggregate price data than the ``law of one price'' workhorse model.
 A key prediction is that a country's income level should have an independent effect on price transmission, in addition to factors related to trade openness, geography and so on.

### 2016-08-11 model.Rnw
JAE sub. vers.

@@
 Two studies quantify the determinants of price transmission and market integration based on individual agricultural commodity prices rather than food price indexes.
-\citet{kouyate16} carry out a meta-regression analysis of the effect of distance on agricultural commodity price transmission for a large number of market pairs.
+\citet{Kouyate2016} carry out a meta-regression analysis of the effect of distance on agricultural commodity price transmission for a large number of market pairs.
 Their results show that geographical distance and separation by borders reduces the likelihood of price cointegration and weakens the speed of speed of price transmission significantly.
@@
 
-Finally, \citet{greb12} consider 8 variables covering geographical, infrastructural, institutional and commodity specific factors which are believed to influence agricultural price transmission.
+Finally, \citet{Greb2016} consider 8 variables covering geographical, infrastructural, institutional and commodity specific factors which are believed to influence agricultural price transmission.
 The authors regress their own price transmission estimates, as well as estimates obtained from the literature, on these country characteristics.

### 2016-08-11 panel_reg.R

@@ -8,6 +8,8 @@ library(plm)
 ### Panel price transmission regressions
 ##################################################################
 
+str = "C:/Users/nqw235/Google Drive/Research/pov_price_transmission/"
+setwd(str)
 load("infl_ctr_dat_0708.Rda")
 load("ts_dat.Rda")
 df = merge(ts.dat, ctr.dat.0708, by = "iso3c")
@@ -49,44 +51,65 @@ m = plm(diff(log(adj.fpi)) ~
         effect = "individual",
         model = "within")
 
-cov = vcovHC(m, method="arellano", type="HC1", cluster="group")
+cov = vcovHC(m, method="arellano", type="HC1")
 rse = sqrt(diag(cov))
-
 panel.reg.m[[1]] = m
 panel.reg.se[[1]] = rse
 
-m = plm(diff(diff(log(adj.fpi))) ~ 
-          lag(diff(diff(log(adj.fpi))),1:p) +
-          lag(diff(diff(log(fao.fpi))), 0:1) +
-          lag(diff(diff(log(er))), 0:1) +
-          I(lag(diff(diff(log(fao.fpi))),0)*log(gdp.const.int)) +
-          I(lag(diff(diff(log(fao.fpi))),0)*I(log(gdp.const.int)^2)) + 
-          I(lag(diff(diff(log(fao.fpi))),0)*cshare) +
-          I(lag(diff(diff(log(fao.fpi))),0)*cir) +
-          I(lag(diff(diff(log(fao.fpi))),0)*cer) +
-          I(lag(diff(diff(log(fao.fpi))),0)*lldc) +
-          I(lag(diff(diff(log(fao.fpi))),0)*ste) +
-          I(lag(diff(diff(log(fao.fpi))),0)*to) +
-          I(lag(diff(diff(log(fao.fpi))),0)*dtf) +
-          I(lag(diff(diff(log(fao.fpi))),0)*lpi.infra) -1 |
-          lag(diff(diff(log(fao.fpi))), 0:1) +
-          lag(diff(diff(log(er))), 0:1) +
-          lag(diff(diff(log(adj.fpi))),2:(p+1)) +
-          I(lag(diff(diff(log(fao.fpi))),0)*log(gdp.const.int)) +
-          I(lag(diff(diff(log(fao.fpi))),0)*I(log(gdp.const.int)^2)) +
[…]

### 2016-08-11 tables.Rnw
JAE sub. vers.

@@
 <<results=tex,echo=F>>=
+str = "C:/Users/nqw235/Google Drive/Research/pov_price_transmission/"
 library(stargazer)
-reg.list = readRDS("fc_reg0708_m.rds")
-se.list = readRDS("fc_reg0708_se.rds")
+reg.list = readRDS(paste0(str,"fc_reg0708_m.rds"))
+se.list = readRDS(paste0(str,"fc_reg0708_se.rds"))
 
 stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
-          title='Determinants of domestic food price changes 2007-8', 
+          title='Determinants of domestic food price changes 2007:1-2008:6', 
           label = "tab:07-08-fc-regressions", 
@@
 library(stargazer)
-reg.list = readRDS("fpi_pt_reg_m.rds")
-se.list = readRDS("fpi_pt_reg_se.rds")
-bp.stat.list = readRDS("fpi_bptest_stat.rds")
-bp.p.list = readRDS("fpi_bptest_p.rds")
+reg.list = readRDS(paste0(str,"fpi_pt_reg_m.rds"))
+se.list = readRDS(paste0(str,"fpi_pt_reg_se.rds"))
+bp.stat.list = readRDS(paste0(str,"fpi_bptest_stat.rds"))
+bp.p.list = readRDS(paste0(str,"fpi_bptest_p.rds"))
 
@@
 library(plm)
-reg.list = readRDS("panel_reg_m.rds")
-se.list = readRDS("panel_reg_se.rds")
+reg.list = readRDS(paste0(str,"panel_reg_m.rds"))
+se.list = readRDS(paste0(str,"panel_reg_se.rds"))
 
-stargazer(reg.list[[1]], reg.list[[3]],
-          title='Determinants of short run FPI transmission rates', 
+stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
+          title='Determinants of monthly FPI inflation and RFPI growth', 
           label = "tab:panel-regressions", 
           dep.var.labels=c("FPI inflation", "RFPI growth"),
-          single.row = TRUE,
-          se = list(se.list[[1]], se.list[[3]]),
+          no.space = TRUE,
+          se = list(se.list[[1]], se.list[[2]], se.list[[3]], se.list[[4]]),
           omit = c("adj.fpi", "rfpi"),
@@
           notes=c("*, ** and *** signify significance at the 10, 5 and 1",
-                  "percent levels, respectively. Robust clustered standard", 
-                  "errors in parentheses. A description of the variables and", 
+                  "percent levels, respectively. Robust standard errors", 
+                  "in parentheses. A description of the variables and", 
                   "their sources can be found in table \\ref{tab:cs_data} and section \\ref{sec:analysis}."),
@@
 <<echo=FALSE>>=
-load("infl_ctr_dat_0708.Rda")
+load(paste0(str,"infl_ctr_dat_0708.Rda"))
 dat = ctr.dat.0708
-r1 = readRDS("fc_reg0708_m.rds")
-r2 = readRDS("fpi_pt_reg_m.rds")
-r3 = readRDS("panel_reg_m.rds")
+r1 = readRDS(paste0(str,"fc_reg0708_m.rds"))
+r2 = readRDS(paste0(str,"fpi_pt_reg_m.rds"))
+r3 = readRDS(paste0(str,"panel_reg_m.rds"))
 
-tp = matrix(NA, nrow=6, ncol=1)
+tp = matrix(NA, nrow=9, ncol=1)
 tp[1,1] = -coef(r1[[1]])["log(gdp.const.int)"]/(2*coef(r1[[1]])["I(log(gdp.const.int)^2)"])
-tp[2,1] = -coef(r1[[3]])["log(gdp.const.int)"]/(2*coef(r1[[3]])["I(log(gdp.const.int)^2)"])
-tp[3,1] = -coef(r2[[1]])["log(gdp.const.int)"]/(2*coef(r2[[1]])["I(log(gdp.const.int)^2)"])
-tp[4,1] = -coef(r2[[3]])["log(gdp.const.int)"]/(2*coef(r2[[3]])["I(log(gdp.const.int)^2)"])
-tp[5,1] = -coef(r3[[1]])["I(lag(diff(log(fao.fpi)), 1) * log(gdp.const.int))"]/
+tp[2,1] = -coef(r1[[2]])["log(gdp.const.int)"]/(2*coef(r1[[2]])["I(log(gdp.const.int)^2)"])
+tp[3,1] = -coef(r1[[3]])["log(gdp.const.int)"]/(2*coef(r1[[3]])["I(log(gdp.const.int)^2)"])
+tp[4,1] = -coef(r1[[4]])["log(gdp.const.int)"]/(2*coef(r1[[4]])["I(log(gdp.const.int)^2)"])
+tp[5,1] = -coef(r2[[1]])["log(gdp.const.int)"]/(2*coef(r2[[1]])["I(log(gdp.const.int)^2)"])
+tp[6,1] = -coef(r2[[2]])["log(gdp.const.int)"]/(2*coef(r2[[2]])["I(log(gdp.const.int)^2)"])
+tp[7,1] = -coef(r2[[3]])["log(gdp.const.int)"]/(2*coef(r2[[3]])["I(log(gdp.const.int)^2)"])
+tp[8,1] = -coef(r2[[4]])["log(gdp.const.int)"]/(2*coef(r2[[4]])["I(log(gdp.const.int)^2)"])
+tp[9,1] = -coef(r3[[1]])["I(lag(diff(log(fao.fpi)), 1) * log(gdp.const.int))"]/
   (2*coef(r3[[1]])["I(lag(diff(log(fao.fpi)), 1) * I(log(gdp.const.int)^2))"])
-tp[6,1] = -coef(r3[[3]])["I(lag(diff(log(fao.fpi)), 1) * log(gdp.const.int))"]/
-  (2*coef(r3[[3]])["I(lag(diff(log(fao.fpi)), 1) * I(log(gdp.const.int)^2))"])
 
@@
 
-nc = matrix(NA, nrow=6, ncol=1)
+nc = matrix(NA, nrow=9, ncol=1)
 nc[1,1] = nrow(dat[which(dat$gdp.const.int< tp[1,1] & !is.na(dat$gdp.const.int)), ])
@@
 nc[6,1] = nrow(dat[which(dat$gdp.const.int< tp[6,1] & !is.na(dat$gdp.const.int)), ])
+nc[7,1] = nrow(dat[which(dat$gdp.const.int< tp[7,1] & !is.na(dat$gdp.const.int)), ])
+nc[8,1] = nrow(dat[which(dat$gdp.const.int< tp[8,1] & !is.na(dat$gdp.const.int)), ])
+nc[9,1] = nrow(dat[which(dat$gdp.const.int< tp[9,1] & !is.na(dat$gdp.const.int)), ])
 
 nc.pc.mean =  matrix(NA, nrow=3, ncol=1)
-nc.pc.mean[1] = mean(nc[1:5])/n.gdp
-nc.pc.mean[2] = mean(nc[5])/n.gdp
-nc.pc.mean[3] = mean(nc[1:4])/n.gdp
+nc.pc.mean[1] = mean(nc[1:9])/n.gdp
+nc.pc.mean[2] = mean(nc[9])/n.gdp
+nc.pc.mean[3] = mean(nc[1:8])/n.gdp
 nc.pc.mean = nc.pc.mean*100
@@
 tp.mean =  matrix(NA, nrow=3, ncol=1)
-tp.mean[1] = mean(tp[1:5])
-tp.mean[2] = mean(tp[5])
-tp.mean[3] = mean(tp[1:4])
+tp.mean[1] = mean(tp[1:9])
+tp.mean[2] = mean(tp[9])
+tp.mean[3] = mean(tp[1:8])
 save(nc.pc.mean, file = "tp_ctr_mean.Rda")
@@
 \small
-\begin{tabular}{|c|c|c|c|c|c|}
+\begin{tabular}{|c|c|c|c|c|c|c|}
 \hline 
-Table & Estimator & Dependent variable & TP (\$, PPP) & Left of TP\tabularnewline
+Table & Column & Estimator & Dependent variable & TP (\$, PPP) & Left of TP\tabularnewline
 \hline 
 \hline
-\ref{tab:07-08-fc-regressions} & OLS & FPI infl. 07-08  &  \Sexpr{round(tp[1,1],0)} & \Sexpr{nc[1,1]}/\Sexpr{n.gdp} \tabularnewline
+\ref{tab:07-08-fc-regressions} & (1) & OLS & FPI infl. 07-08  &  \Sexpr{round(tp[1,1],0)} & \Sexpr{nc[1,1]}/\Sexpr{n.gdp} \tabularnewline
 \hline 
-\ref{tab:07-08-fc-regressions} & OLS & RFPI growth 07-08  &  \Sexpr{round(tp[2,1],0)} & \Sexpr{nc[2,1]}/\Sexpr{n.gdp} \tabularnewline
+\ref{tab:07-08-fc-regressions} & (2) & OLS & FPI infl. 07-08  &  \Sexpr{round(tp[2,1],0)} & \Sexpr{nc[2,1]}/\Sexpr{n.gdp} \tabularnewline
 \hline 
-\ref{tab:fpi-pt-regressions} & OLS & FPI LRM  &  \Sexpr{round(tp[3,1],0)} & \Sexpr{nc[3,1]}/\Sexpr{n.gdp} \tabularnewline
+\ref{tab:07-08-fc-regressions} & (3) & OLS & RFPI growth 07-08  &  \Sexpr{round(tp[3,1],0)} & \Sexpr{nc[3,1]}/\Sexpr{n.gdp} \tabularnewline
 \hline 
-\ref{tab:fpi-pt-regressions} & WLS & FPI LRM  &  \Sexpr{round(tp[4,1],0)} & \Sexpr{nc[4,1]}/\Sexpr{n.gdp} \tabularnewline
+\ref{tab:07-08-fc-regressions} & (4) & OLS & RFPI growth 07-08  &  \Sexpr{round(tp[4,1],0)} & \Sexpr{nc[4,1]}/\Sexpr{n.gdp} \tabularnewline
 \hline 
-\ref{tab:panel-regressions} & FE & $\Delta \log(FPI)_{t}$  &  \Sexpr{round(tp[5,1],0)} & \Sexpr{nc[5,1]}/\Sexpr{n.gdp} \tabularnewline
+\ref{tab:fpi-pt-regressions} & (1) & OLS & FPI LRM  &  \Sexpr{round(tp[5,1],0)} & \Sexpr{nc[5,1]}/\Sexpr{n.gdp} \tabularnewline
+\hline 
+\ref{tab:fpi-pt-regressions} & (2) & WLS & FPI LRM  &  \Sexpr{round(tp[6,1],0)} & \Sexpr{nc[6,1]}/\Sexpr{n.gdp} \tabularnewline
+\hline 
+\ref{tab:fpi-pt-regressions} & (3) & OLS & FPI LRM  &  \Sexpr{round(tp[7,1],0)} & \Sexpr{nc[7,1]}/\Sexpr{n.gdp} \tabularnewline
+\hline 
+\ref{tab:fpi-pt-regressions} & (4) & WLS & FPI LRM  &  \Sexpr{round(tp[8,1],0)} & \Sexpr{nc[8,1]}/\Sexpr{n.gdp} \tabularnewline
+\hline 
+\ref{tab:panel-regressions} & (1) & FE & $\Delta \log(FPI)_{t}$  &  \Sexpr{round(tp[9,1],0)} & \Sexpr{nc[9,1]}/\Sexpr{n.gdp} \tabularnewline
 \hline

### 2016-08-11 price_data_prep.R

@@ -481,9 +481,9 @@ infl = merge(infl_0708, infl_1011)
 infl.l = reshape(infl,varying=2:7, direction="long", idvar="iso3c", sep=".", timevar="period")
 infl.l1 = melt(infl.l, id = c("iso3c", "period"))
 
-# save(infl, file = "infl_0708_1011.Rda")
-# save(infl.l, file = "infl_l.Rda")
-# save(infl.l1, file = "infl_l1.Rda")
+# save(infl, file = paste0(str,"infl_0708_1011.Rda"))
+# save(infl.l, file = paste0(str,"infl_l.Rda"))
+# save(infl.l1, file = paste0(str,"infl_l1.Rda"))
 # save(infl_0708, file = "infl_0708.Rda")
 # save(infl_1011, file = "infl_1011.Rda")
 # save(ilo_0515, file = "ilo_0515.Rda")

### 2016-08-11 ts_reg.R
JAE sub. vers.

@@ -8,6 +8,8 @@ library(lmtest)
 ### Time series regressions
 ##################################################################
 
+str = "C:/Users/nqw235/Google Drive/Research/pov_price_transmission/"
+setwd(str)
 load("ts_dat.Rda")
 data = ts.dat
 tbl = table(data$id)
@@ -203,7 +205,7 @@ for (i in 1:max(data$id)) {
 }
 coef.bew.rfpi.l1 = as.data.frame(coef.mat)
 coef.bew.rfpi.l1$iso3c = names(tbl)
-coef.bew.rfpi.l1 = rename(coef.bew.rfpi.l1, lrm = V1, lrm.se = V2)
+coef.bew.rfpi.l1 = dplyr::rename(coef.bew.rfpi.l1, lrm = V1, lrm.se = V2)
 save(coef.bew.rfpi.l1, file = "coef_bew_rfpi_l1.Rda")
 
 ## Common lag length = 6
@@ -302,7 +304,7 @@ imp.se.mat = matrix(NA, nrow = nrow(tbl), ncol = 12)
 
 for (i in 1:max(data$id)) {
   dat = filter(data, id==i)
-  dat = select(dat, cpi.brk:rfpi, fao.fpi:er)
+  dat = dplyr::select(dat, cpi.brk:rfpi, fao.fpi:er)
   ts.dat = ts(dat, start=c(2005, 1), end=c(2014, 12), frequency=12)
   
   if(is.na(dat$er)[1] == FALSE & sd(dat$er, na.rm = TRUE) != 0) {
@@ -352,7 +354,7 @@ imp.se.mat = matrix(NA, nrow = nrow(tbl), ncol = 12)
 
 for (i in 1:max(data$id)) {
   dat = filter(data, id==i)
-  dat = select(dat, cpi.brk:rfpi, r.fao.fpi:er)
+  dat = dplyr::select(dat, cpi.brk:rfpi, r.fao.fpi:er)
   ts.dat = ts(dat, start=c(2005, 1), end=c(2014, 12), frequency=12)
   
   if(is.na(dat$er)[1] == FALSE) {
